% Hanzo Base: The Single-Binary Runtime for Sovereign Applications
\documentclass[11pt]{article}
\usepackage{shared/hanzocover}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{float}
\usepackage{multirow}
\usepackage{listings}
\input{shared/lstlang}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,positioning,fit}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}

\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,urlcolor=blue}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},
    commentstyle=\color{codegreen},
    keywordstyle=\color{codepurple},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codegreen},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,
    breaklines=true,
    captionpos=b,
    keepspaces=true,
    numbers=left,
    numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=2
}
\lstset{style=mystyle}

\newtheorem{definition}{Definition}

\title{HANZO BASE\\[0.3em]
{\large The Single-Binary Runtime for\\Sovereign Applications}}
\author{
    Zach Kelling \\
    \textit{Hanzo Industries Inc (Techstars '17)} \\
    \texttt{research@hanzo.ai}
}
\date{March 2026}

\begin{document}

\hanzocoverpage

% ============================================================================
% ABSTRACT
% ============================================================================
\begin{abstract}
We present Hanzo Base, the application runtime for sovereign software. Base is
not a backend framework. It is the personal computer of Web5---a single Go
binary that packages an encrypted embedded database (SQLite with sqlcipher), a
realtime API layer, CRDT-based collaboration, a JavaScript plugin runtime
(Goja VM), Hanzo IAM authentication, and field-level KMS encryption into one
process. Where Firebase assumes a cloud provider, Supabase assumes PostgreSQL,
and PocketBase assumes a single user, Base assumes nothing. It runs on a
laptop, a phone, a Raspberry Pi, or a Kubernetes pod with the same binary and
the same API surface. Data stays local by default. Replication is opt-in.
Chain anchoring is available but never required. We describe the architecture,
the data plane (local vault, provider-optional replication, Lux receipt
anchoring), Base Functions (TypeScript in Goja, event-driven, zero cold start),
Base Collections (schemaless records with access rules and field-level
encryption), Base Sync (CRDT log, conflict-free merge, offline-first,
peer-to-peer or server-relay), and integration with Lux Network
chains (K-Chain key policy, I-Chain identity, T-Chain threshold controls).
Benchmarks on an M4 MacBook show 10,400 writes/sec for encrypted collections
and sub-millisecond read latency. In production since December 2022, Base
serves as the default runtime for 2,800+ applications across the Hanzo
ecosystem.
\end{abstract}

% ============================================================================

\input{sections/01-introduction}
\input{sections/02-architecture}
\input{sections/03-base-as-data-plane}
\input{sections/04-base-functions}
\input{sections/05-base-collections}
\input{sections/06-base-sync}
\input{sections/07-lux-network-integration}
\input{sections/08-comparison-to-existing-systems}
\input{sections/09-hanzoai-sqlite}
\input{sections/10-performance}
\input{sections/11-production-deployment}
\input{sections/12-related-work}
\input{sections/12a-prior-work-pocketbase}
\input{sections/99-conclusion}

\end{document}
