\section{Integration with Hanzo and Lux Infrastructure}

\label{sec:integration}
% ============================================================================

\subsection{Hanzo ACI Verifiable Computation}

FHE computations on Hanzo ACI nodes produce verifiable proofs:
\begin{enumerate}[leftmargin=1.4em]
    \item The GPU node commits to the encrypted input, model weights, and
    encrypted output via a Merkle hash.
    \item A SNARK proof attests that the committed encrypted output is the
    correct evaluation of the committed model on the committed encrypted input.
    \item The proof is submitted to the ACI consensus layer for verification
    and on-chain attestation.
\end{enumerate}

\subsection{Lux TFHE Integration}

The Lux blockchain provides native TFHE operations~\cite{luxtfhe} for on-chain
encrypted computation. Hanzo FHE Inference uses Lux TFHE for:
\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Key management}: FHE public keys are registered on-chain,
    enabling any ACI node to serve encrypted inference without key exchange.
    \item \textbf{Result verification}: Encrypted inference results are
    verified on-chain by re-executing a subset of operations.
    \item \textbf{Access control}: Decryption keys are managed via
    threshold secret sharing on-chain (\S\ref{sec:threshold}), enabling
    multi-party access policies.
\end{itemize}

\subsection{Hanzo KMS Integration}

FHE key material is managed through Hanzo KMS~\cite{hanzokms}:
\begin{itemize}[leftmargin=1.1em]
    \item Secret keys are stored encrypted at rest with AES-256-GCM envelope
    encryption.
    \item Key rotation follows the KMS per-environment rotation policy.
    \item All key access events are logged to the KMS audit trail.
\end{itemize}

% ============================================================================
