\section{Conclusion}

\label{sec:conclusion}
% ============================================================================

Hanzo FHE Inference enables practical confidential AI-as-a-service by combining
CKKS for neural networks, TFHE for decision trees, and GPU acceleration through
the Candle runtime. Production deployments demonstrate sub-2-second encrypted
inference with negligible accuracy loss, making FHE viable for regulated
industries requiring data privacy guarantees stronger than any hardware-based
approach. Integration with Hanzo ACI and Lux TFHE provides verifiable encrypted
computation with on-chain attestation.

\label{sec:threshold}

% ============================================================================
% BIBLIOGRAPHY
% ============================================================================
\begin{thebibliography}{20}

\bibitem{hanzocandle}
{Hanzo AI Research}.
\newblock Hanzo Candle: Rust-native {ML} framework for production inference.
\newblock \emph{Hanzo AI Research}, 2024.

\bibitem{hanzoaci}
{Hanzo AI Research}.
\newblock Hanzo {ACI}: {AI} Chain Infrastructure for decentralized compute markets.
\newblock \emph{Hanzo AI Research}, December 2024.

\bibitem{hanzojin}
{Hanzo AI Research}.
\newblock Jin: Unified multimodal {AI} architecture.
\newblock \emph{Hanzo AI Research}, 2025.

\bibitem{hanzokms}
{Hanzo AI Research}.
\newblock Hanzo {KMS}: Secrets management with per-environment rotation and audit trail.
\newblock \emph{Hanzo AI Research}, September 2021.

\bibitem{hanzopq}
{Hanzo AI Research}.
\newblock Hanzo post-quantum cryptography: Migration path for {AI} infrastructure.
\newblock \emph{Hanzo AI Research}, April 2026.

\bibitem{luxtfhe}
{Lux Partners}.
\newblock Lux {TFHE}: On-chain fully homomorphic encryption.
\newblock \emph{Lux Network Technical Report}, 2024.

\bibitem{ckks2017}
J.~H.~Cheon, A.~Kim, M.~Kim, and Y.~Song.
\newblock Homomorphic encryption for arithmetic of approximate numbers.
\newblock In \emph{Proc.\ ASIACRYPT}, 2017.

\bibitem{tfhe2020}
I.~Chillotti, N.~Gama, M.~Georgieva, and M.~Izabach\`ene.
\newblock {TFHE}: Fast fully homomorphic encryption over the torus.
\newblock \emph{J.\ Cryptology}, 33(1):34--91, 2020.

\bibitem{fhesurvey2024}
C.~Mouchet, J.~Bossuat, J.-P.~Bossuat, and J.-P.~Hubaux.
\newblock A survey of {FHE} for machine learning.
\newblock \emph{ACM Computing Surveys}, 2024.

\bibitem{sgxattacks}
J.~Van~Bulck, M.~Minkin, O.~Weisse, D.~Genkin, B.~Gras, F.~Piessens, M.~Silberstein, T.~F.~Wenisch, Y.~Yarom, and R.~Strackx.
\newblock Foreshadow: Extracting the keys to the {Intel SGX} kingdom.
\newblock In \emph{Proc.\ USENIX Security}, 2018.

\bibitem{lweestimator}
M.~Albrecht, R.~Player, and S.~Scott.
\newblock On the concrete hardness of {Learning with Errors}.
\newblock \emph{J.\ Mathematical Cryptology}, 9(3):169--203, 2015.

\bibitem{crypten}
B.~Knott, S.~Venkataraman, A.~Hannun, S.~Sengupta, M.~Ibrahim, and L.~van~der~Maaten.
\newblock {CrypTen}: Secure multi-party computation meets machine learning.
\newblock In \emph{Proc.\ NeurIPS}, 2021.

\bibitem{tenseal}
A.~Benaissa, B.~Retiat, B.~Cebere, and A.~E.~Belfedhal.
\newblock {TenSEAL}: A library for encrypted tensor operations using homomorphic encryption.
\newblock In \emph{ICLR Workshop on DPML}, 2021.

\bibitem{concreteml}
{Zama}.
\newblock Concrete {ML}: Privacy-preserving machine learning using {FHE}.
\newblock \emph{Zama Documentation}, 2023.

\end{thebibliography}

