\section{ML-DSA for Digital Signatures}

\label{sec:mldsa}
% ============================================================================

ML-DSA (FIPS 204) replaces Ed25519 and ECDSA for digital signatures. It is
based on the Module Learning with Errors and Module Short Integer Solution
(MLWE/MSIS) problems.

\begin{definition}[ML-DSA]
ML-DSA consists of three algorithms:
\begin{itemize}[leftmargin=1.1em]
    \item $\texttt{KeyGen}() \to (vk, sk)$: Generate verification and
    signing keys.
    \item $\texttt{Sign}(sk, m) \to \sigma$: Produce signature $\sigma$ on
    message $m$.
    \item $\texttt{Verify}(vk, m, \sigma) \to \{0, 1\}$: Verify signature
    validity.
\end{itemize}
\end{definition}

We deploy ML-DSA-65 (NIST Security Level 3):

\begin{table}[H]
\centering
\small
\begin{tabular}{lrrr}
\toprule
\textbf{Parameter} & \textbf{ML-DSA-65} & \textbf{Ed25519} & \textbf{Ratio} \\
\midrule
Public key (bytes) & 1,952 & 32 & 61$\times$ \\
Signature (bytes) & 3,309 & 64 & 52$\times$ \\
Sign ($\mu$s) & 120 & 22 & 5.5$\times$ \\
Verify ($\mu$s) & 45 & 18 & 2.5$\times$ \\
\bottomrule
\end{tabular}
\caption{ML-DSA-65 vs Ed25519 parameter comparison.}
\label{tab:mldsa}
\end{table}

\subsection{IAM JWT Migration}

Hanzo IAM~\cite{hanzoiam} JWT tokens currently use Ed25519 signatures.
The migration proceeds in three phases:

\begin{enumerate}[leftmargin=1.4em]
    \item \textbf{Phase 1 (current)}: IAM signs with both Ed25519 and ML-DSA-65.
    Tokens carry dual signatures. Verifiers accept either.
    \item \textbf{Phase 2 (Q3 2026)}: New tokens are signed with ML-DSA-65 only.
    Verifiers accept both algorithms (backward compatibility for tokens issued
    in Phase 1).
    \item \textbf{Phase 3 (Q4 2026)}: Ed25519 verification is removed. All
    tokens are ML-DSA-65.
\end{enumerate}

\begin{remark}
The larger ML-DSA-65 signature (3,309 bytes vs.\ 64 bytes) increases JWT size
from $\sim$500 bytes to $\sim$3,800 bytes. This remains well within HTTP header
limits and does not impact API performance measurably.
\end{remark}

\subsection{API Request Signing}

External API requests to Hanzo Gateway use HMAC-SHA256 for authentication
(quantum-safe at 128-bit security). Internal service-to-service requests use
mTLS with ML-DSA-65 certificates.

\subsection{Agent Credential Chains}

Agent credentials~\cite{hanzoagentsdk} use certificate chains signed with
Ed25519. Migration replaces the signature algorithm while preserving the chain
structure:
\begin{equation}
\text{Root}_{ML\text{-}DSA} \to \text{Org}_{ML\text{-}DSA} \to
\text{Agent}_{ML\text{-}DSA}.
\end{equation}

During the hybrid period, cross-signed certificates bridge old and new root
keys.

% ============================================================================
