\section{Conclusion}

\label{sec:conclusion}
% ============================================================================

Hanzo post-quantum migration replaces vulnerable classical cryptography across
all infrastructure: ML-KEM-768 for key exchange (TLS, KMS), ML-DSA-65 for
digital signatures (IAM, agent credentials, API signing), and SLH-DSA-192s for
root certificates. The hybrid transition strategy ensures security if either
classical or post-quantum schemes are broken. Performance overhead is bounded
at 1.2ms for cold TLS handshakes and 0.045ms for signature verification.
Migration began in January 2026 with inter-service TLS and completes by Q4
2026 with full ML-DSA JWT signing.

% ============================================================================
% BIBLIOGRAPHY
% ============================================================================
\begin{thebibliography}{15}

\bibitem{hanzogateway}
{Hanzo AI Research}.
\newblock Hanzo Gateway: {API} gateway with {KrakenD} backend.
\newblock \emph{Hanzo AI Research}, 2024.

\bibitem{hanzoiam}
{Hanzo AI Research}.
\newblock Hanzo {IAM}: Identity and access management for {AI} platforms.
\newblock \emph{Hanzo AI Research}, 2024.

\bibitem{hanzokms}
{Hanzo AI Research}.
\newblock Hanzo {KMS}: Secrets management with per-environment rotation and audit trail.
\newblock \emph{Hanzo AI Research}, September 2021.

\bibitem{hanzoagentsdk}
{Hanzo AI Research}.
\newblock Hanzo Agent {SDK}: Framework for building {AI} agents.
\newblock \emph{Hanzo AI Research}, 2025.

\bibitem{hanzothreshold}
{Hanzo AI Research}.
\newblock Hanzo Threshold Signing: {MPC} key management for {AI} infrastructure.
\newblock \emph{Hanzo AI Research}, April 2026.

\bibitem{luxwhitepaper}
{Lux Partners}.
\newblock Lux: A scalable multi-consensus blockchain with post-quantum cryptography.
\newblock \emph{Lux Network Technical Report}, 2024.

\bibitem{nistpq}
{NIST}.
\newblock Post-quantum cryptography standardization.
\newblock \emph{NIST FIPS 203, 204, 205}, August 2024.

\bibitem{shor1994}
P.~Shor.
\newblock Algorithms for quantum computation: Discrete logarithms and factoring.
\newblock In \emph{Proc.\ FOCS}, 1994.

\bibitem{grover1996}
L.~Grover.
\newblock A fast quantum mechanical algorithm for database search.
\newblock In \emph{Proc.\ STOC}, 1996.

\bibitem{quantumtimeline}
M.~Mosca.
\newblock Cybersecurity in an era with quantum computers: Will we be ready?
\newblock \emph{IEEE Security \& Privacy}, 16(5):38--41, 2018.

\bibitem{hybridtls}
D.~Stebila, S.~Fluhrer, and S.~Gueron.
\newblock Hybrid key exchange in {TLS 1.3}.
\newblock \emph{IETF Internet-Draft}, 2024.

\bibitem{chromepq}
{Google Chrome Team}.
\newblock Post-quantum cryptography in {Chrome}.
\newblock \emph{Chromium Blog}, 2024.

\bibitem{signalpq}
{Signal Foundation}.
\newblock {PQXDH}: Post-quantum extended {Diffie-Hellman}.
\newblock \emph{Signal Technical Documentation}, 2023.

\end{thebibliography}

