\documentclass[11pt,twocolumn]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,positioning,fit,backgrounds}
\usepackage[margin=1in]{geometry}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},
    commentstyle=\color{codegreen},
    keywordstyle=\color{codepurple},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codegreen},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,
    breaklines=true,
    captionpos=b,
    keepspaces=true,
    numbers=left,
    numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=2
}
\lstset{style=mystyle}

\title{Jin: Unified Multimodal AI Architecture for Vision-Language, Audio, and Cross-Modal Understanding}
\author{Zach Kelling\thanks{zach@hanzo.ai} \\ \textit{Hanzo Industries \quad Lux Industries \quad Zoo Labs Foundation} \\ \texttt{research@hanzo.ai}}
\date{2025}

\begin{document}

\maketitle

\begin{abstract}
We present Jin, a hypermodal AI architecture that unifies text, vision, audio, and extensible custom modalities through a joint embedding predictive framework. Jin combines three key innovations: (1) modality-specific encoders projecting heterogeneous inputs into a shared latent space with learnable modality embeddings; (2) a diffusion transformer backbone with Mixture of Experts (MoE) routing for efficient cross-modal reasoning; (3) hierarchical zooming JEPA (z-JEPA) for multi-scale visual understanding. Our experiments demonstrate that Jin achieves 82.1\% accuracy on ImageNet with zero-shot transfer, outperforms single-modality baselines by 15\% on AudioCaps retrieval, and enables novel cross-modal generation tasks. The architecture supports runtime modality extension without retraining the core model, enabling deployment flexibility for diverse applications.
\end{abstract}

\section{Introduction}

Human cognition seamlessly integrates information across sensory modalities---we understand speech by combining audio with visual lip movements, interpret scenes by relating objects to their sounds, and communicate using language grounded in physical experience. Current AI systems largely process modalities in isolation, with multimodal models treating cross-modal understanding as a downstream fusion task.

Jin (Japanese: \begin{CJK}{UTF8}{min}神\end{CJK}, ``spirit'' or ``god'') introduces a hypermodal architecture where all modalities share a unified latent space from the ground up. The key insight is that joint embedding predictive architectures (JEPA)~\cite{jepa2022} can be extended beyond single modalities: by training encoders to predict masked representations across modalities, the model learns modality-agnostic features suitable for any downstream task.

\paragraph{Contributions.} This paper makes the following contributions:
\begin{itemize}
    \item A joint embedding space architecture with modality-specific encoders and learnable modality tokens enabling seamless cross-modal fusion.
    \item A diffusion transformer backbone with Mixture of Experts providing efficient routing for modality-specific and cross-modal computation.
    \item z-JEPA, a hierarchical zooming extension of JEPA for multi-scale visual understanding with adaptive computation.
    \item Runtime modality extension enabling new modalities without core model retraining.
    \item Comprehensive benchmarks demonstrating state-of-the-art performance on multimodal understanding tasks.
\end{itemize}

\section{Background}

\subsection{Joint Embedding Predictive Architectures}

JEPA~\cite{jepa2022} learns representations by predicting masked patch embeddings rather than reconstructing raw pixels. Given an image $x$, a context encoder $f_\theta$ produces embeddings for visible patches, and a predictor $g_\phi$ predicts embeddings for masked patches. The target encoder $f_{\bar{\theta}}$ (EMA of context encoder) provides prediction targets:

\begin{equation}
    \mathcal{L}_{\text{JEPA}} = \|g_\phi(f_\theta(x_{\text{vis}})) - f_{\bar{\theta}}(x_{\text{mask}})\|^2
\end{equation}

JEPA avoids reconstruction collapse by operating in embedding space, enabling learning of abstract features without pixel-level supervision.

\subsection{Mixture of Experts}

Sparse MoE~\cite{moe2017} routes tokens through a subset of expert networks:

\begin{equation}
    y = \sum_{i=1}^{K} g_i(x) \cdot E_i(x)
\end{equation}

where $g_i(x)$ is the gating weight for expert $E_i$, with typically $K \ll N$ experts active per token. This enables parameter scaling without proportional compute increase.

\subsection{Diffusion Models}

Diffusion models~\cite{ddpm2020} learn to reverse a noise process:

\begin{equation}
    p_\theta(x_{t-1}|x_t) = \mathcal{N}(x_{t-1}; \mu_\theta(x_t, t), \Sigma_\theta(x_t, t))
\end{equation}

Recent work shows diffusion transformers (DiT)~\cite{dit2023} can generate high-quality images with better scaling properties than U-Nets.

\section{Jin Architecture}

\subsection{Overview}

\begin{figure*}[t]
\centering
\begin{tikzpicture}[
    node distance=0.8cm,
    encoder/.style={rectangle, draw, minimum width=1.8cm, minimum height=1cm, align=center, fill=blue!10},
    joint/.style={rectangle, draw, minimum width=10cm, minimum height=1.2cm, align=center, fill=green!10},
    transformer/.style={rectangle, draw, minimum width=10cm, minimum height=1.5cm, align=center, fill=yellow!10},
    decoder/.style={rectangle, draw, minimum width=1.8cm, minimum height=1cm, align=center, fill=red!10},
    arrow/.style={->, thick}
]
% Inputs
\node[encoder] (text_enc) {Text\\Encoder};
\node[encoder, right=0.3cm of text_enc] (vision_enc) {Vision\\Encoder};
\node[encoder, right=0.3cm of vision_enc] (audio_enc) {Audio\\Encoder};
\node[encoder, right=0.3cm of audio_enc] (custom_enc) {Custom\\Encoder};

% Joint space
\node[joint, below=1cm of vision_enc] (joint) {Joint Embedding Space + Modality Tokens};

% Transformer
\node[transformer, below=0.8cm of joint] (transformer) {Diffusion Transformer with MoE};

% Decoders
\node[decoder, below=1cm of text_enc] (text_dec) at (text_enc |- transformer.south) {Text\\Output};
\node[decoder, below=1cm of vision_enc] (vision_dec) at (vision_enc |- transformer.south) {Vision\\Output};
\node[decoder, below=1cm of audio_enc] (audio_dec) at (audio_enc |- transformer.south) {Audio\\Output};
\node[decoder, below=1cm of custom_enc] (custom_dec) at (custom_enc |- transformer.south) {Custom\\Output};

% Arrows
\draw[arrow] (text_enc) -- (joint);
\draw[arrow] (vision_enc) -- (joint);
\draw[arrow] (audio_enc) -- (joint);
\draw[arrow] (custom_enc) -- (joint);
\draw[arrow] (joint) -- (transformer);
\draw[arrow] (transformer.south -| text_dec) -- (text_dec);
\draw[arrow] (transformer.south -| vision_dec) -- (vision_dec);
\draw[arrow] (transformer.south -| audio_dec) -- (audio_dec);
\draw[arrow] (transformer.south -| custom_dec) -- (custom_dec);

% Labels
\node[above=0.2cm of text_enc] {\small Tokens};
\node[above=0.2cm of vision_enc] {\small Patches};
\node[above=0.2cm of audio_enc] {\small Mel-Spec};
\node[above=0.2cm of custom_enc] {\small Features};
\end{tikzpicture}
\caption{Jin architecture: modality-specific encoders project into a joint embedding space, processed by a diffusion transformer with MoE, and decoded to target modalities.}
\label{fig:architecture}
\end{figure*}

Figure~\ref{fig:architecture} shows the Jin architecture. The system processes inputs through three stages:

\begin{enumerate}
    \item \textbf{Encoding}: Modality-specific encoders project inputs to the joint embedding space.
    \item \textbf{Processing}: A diffusion transformer with MoE processes the joint representation.
    \item \textbf{Decoding}: Modality-specific decoders generate outputs in target modalities.
\end{enumerate}

\subsection{Joint Embedding Space}

\begin{lstlisting}[language=Python]
class JointEmbeddingSpace(nn.Module):
    def __init__(self, config: JinConfig):
        self.encoders = nn.ModuleDict({
            "text": TextEncoder(config),
            "vision": VisionEncoder(config),
            "audio": AudioEncoder(config)
        })
        self.modality_embed = nn.Embedding(
            len(config.modalities), config.embedding_dim)
        self.fusion = nn.MultiheadAttention(
            config.embedding_dim, config.num_heads)

    def encode(self, inputs: Dict[str, Tensor]) -> Tensor:
        encoded = []
        for i, (mod, data) in enumerate(inputs.items()):
            embeds = self.encoders[mod].encode(data)
            mod_emb = self.modality_embed(
                torch.full((B, 1), i, device=device))
            embeds = embeds + mod_emb
            encoded.append(embeds)

        joint = torch.cat(encoded, dim=1)
        fused, _ = self.fusion(joint, joint, joint)
        return fused
\end{lstlisting}

\paragraph{Modality Tokens.} Each modality receives a learnable embedding added to all its tokens, enabling the transformer to distinguish modality origins without explicit conditioning.

\paragraph{Cross-Modal Fusion.} Self-attention over concatenated modality tokens enables cross-modal information flow from the first layer, unlike late-fusion approaches.

\subsection{Modality Encoders}

\paragraph{Text Encoder.} Standard token embedding with learned positional encodings:

\begin{lstlisting}[language=Python]
class TextEncoder(nn.Module):
    def encode(self, x: Tensor) -> Tensor:
        embeds = self.embed(x)  # (B, T, D)
        positions = self.pos_embed[:, :x.shape[1]]
        return embeds + positions
\end{lstlisting}

\paragraph{Vision Encoder.} Patch embedding following ViT~\cite{vit2020}:

\begin{lstlisting}[language=Python]
class VisionEncoder(nn.Module):
    def __init__(self, config, patch_size=14):
        self.proj = nn.Conv2d(3, config.embedding_dim,
                              kernel_size=patch_size,
                              stride=patch_size)

    def encode(self, x: Tensor) -> Tensor:
        patches = self.proj(x)  # (B, D, H/P, W/P)
        patches = rearrange(patches, 'b d h w -> b (h w) d')
        return patches + self.pos_embed[:, :patches.shape[1]]
\end{lstlisting}

\paragraph{Audio Encoder.} Projects mel-spectrogram frames:

\begin{lstlisting}[language=Python]
class AudioEncoder(nn.Module):
    def encode(self, x: Tensor) -> Tensor:
        # x: (B, T, 80) mel bins
        return self.proj(x) + self.pos_embed[:, :x.shape[1]]
\end{lstlisting}

\subsection{Diffusion Transformer with MoE}

\begin{lstlisting}[language=Python]
class DiffusionTransformerBlock(nn.Module):
    def __init__(self, config):
        self.norm1 = nn.LayerNorm(config.hidden_dim)
        self.attn = nn.MultiheadAttention(
            config.hidden_dim, config.num_heads)
        self.norm2 = nn.LayerNorm(config.hidden_dim)
        self.moe = UnifiedMoE(config)
        if config.use_diffusion:
            self.time_embed = nn.Sequential(
                nn.Linear(1, config.hidden_dim),
                nn.SiLU(),
                nn.Linear(config.hidden_dim,
                          config.hidden_dim))

    def forward(self, x, timestep=None):
        x = x + self.attn(self.norm1(x))[0]
        x = x + self.moe(self.norm2(x))
        if timestep is not None:
            x = x + self.time_embed(timestep).unsqueeze(1)
        return x
\end{lstlisting}

\paragraph{MoE Routing.} Each token is routed to top-$k$ experts:

\begin{lstlisting}[language=Python]
class UnifiedMoE(nn.Module):
    def forward(self, x: Tensor) -> Tensor:
        scores = self.router(x)  # (B, N, num_experts)
        weights, indices = torch.topk(
            F.softmax(scores, dim=-1),
            self.experts_per_token, dim=-1)

        output = torch.zeros_like(x)
        for i in range(self.experts_per_token):
            expert_idx = indices[:, :, i]
            expert_weight = weights[:, :, i].unsqueeze(-1)
            for e in range(self.num_experts):
                mask = (expert_idx == e)
                if mask.any():
                    expert_out = self.experts[e](x[mask])
                    output[mask] += expert_weight[mask] * expert_out
        return output
\end{lstlisting}

\subsection{z-JEPA: Hierarchical Zooming}

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    node distance=0.6cm,
    level/.style={rectangle, draw, minimum width=2cm, minimum height=0.7cm, align=center},
    arrow/.style={->, thick}
]
\node[level, fill=blue!30] (l0) {Level 0\\224x224};
\node[level, below=of l0, fill=blue!25] (l1) {Level 1\\112x112};
\node[level, below=of l1, fill=blue!20] (l2) {Level 2\\56x56};
\node[level, below=of l2, fill=blue!15] (l3) {Level 3\\28x28};

\node[right=1.5cm of l0] (d0) {\small Global context};
\node[right=1.5cm of l1] (d1) {\small Object-level};
\node[right=1.5cm of l2] (d2) {\small Part-level};
\node[right=1.5cm of l3] (d3) {\small Fine detail};

\draw[arrow, bend right=20] (l0) to node[left] {\tiny predict} (l1);
\draw[arrow, bend right=20] (l1) to node[left] {\tiny predict} (l2);
\draw[arrow, bend right=20] (l2) to node[left] {\tiny predict} (l3);
\draw[arrow, bend left=20, dashed] (l3) to node[right] {\tiny feedback} (l2);
\draw[arrow, bend left=20, dashed] (l2) to node[right] {\tiny feedback} (l1);
\end{tikzpicture}
\caption{z-JEPA hierarchical levels with bidirectional prediction.}
\label{fig:zjepa}
\end{figure}

z-JEPA extends JEPA with hierarchical zoom levels (Figure~\ref{fig:zjepa}):

\begin{lstlisting}[language=Python]
@dataclass
class ZJEPAConfig:
    num_z_levels: int = 4
    zoom_factors: List[float] = [1.0, 0.5, 0.25, 0.125]
    predict_fine_from_coarse: bool = True
    predict_coarse_from_fine: bool = True
    bidirectional_prediction: bool = True
\end{lstlisting}

\paragraph{Cross-Scale Prediction.} The model learns to predict fine-grained features from coarse-grained context and vice versa:

\begin{equation}
    \mathcal{L}_{\text{cross}} = \sum_{i=0}^{L-1} \left( \|g_i(z_i) - z_{i+1}\|^2 + \|h_{i+1}(z_{i+1}) - z_i\|^2 \right)
\end{equation}

\paragraph{Adaptive Zoom.} An attention network identifies regions of interest for detailed processing:

\begin{lstlisting}[language=Python]
class AdaptiveZoomModule(nn.Module):
    def forward(self, x, global_features):
        roi_coords = self.roi_network(global_features)
        importance = self.importance_scorer(global_features)

        crops = [x]  # Level 0: full image
        for level in range(1, self.num_levels):
            crop = extract_foveal_region(
                x, roi_coords, level)
            crops.append(crop)
        return crops, importance
\end{lstlisting}

\subsection{Modality Extension}

New modalities can be added without retraining:

\begin{lstlisting}[language=Python]
class CustomModalityEncoder(nn.Module, ModalityEncoder):
    def __init__(self, input_dim, config):
        self.proj = nn.Linear(input_dim,
                              config.embedding_dim)

    def encode(self, x: Tensor) -> Tensor:
        return self.proj(x)

# Add at runtime
custom_encoder = CustomModalityEncoder(128, model.config)
model.add_modality("proprioception", custom_encoder)
\end{lstlisting}

The modality embedding table expands dynamically, preserving existing embeddings.

\section{Training}

\subsection{Pre-training Objectives}

\paragraph{Within-Modality JEPA.} Standard masked prediction within each modality.

\paragraph{Cross-Modal JEPA.} Predict masked regions in one modality given context from another:

\begin{equation}
    \mathcal{L}_{\text{cross-modal}} = \mathbb{E}_{x_v, x_a} \left[ \|g(f(x_v)) - f(x_a^{\text{mask}})\|^2 \right]
\end{equation}

\paragraph{Contrastive Alignment.} InfoNCE loss aligns corresponding cross-modal pairs:

\begin{equation}
    \mathcal{L}_{\text{align}} = -\log \frac{\exp(z_v \cdot z_a / \tau)}{\sum_j \exp(z_v \cdot z_a^j / \tau)}
\end{equation}

\paragraph{Total Loss.}
\begin{equation}
    \mathcal{L} = \alpha \mathcal{L}_{\text{JEPA}} + \beta \mathcal{L}_{\text{cross-modal}} + \gamma \mathcal{L}_{\text{align}}
\end{equation}

\subsection{Curriculum Learning}

Training proceeds in stages:
\begin{enumerate}
    \item Single-modality JEPA (epochs 1-50)
    \item Cross-modal prediction (epochs 51-100)
    \item Full multimodal with z-JEPA (epochs 101-150)
\end{enumerate}

\subsection{Scaling}

\begin{table}[h]
\centering
\caption{Jin model configurations.}
\label{tab:scale}
\begin{tabular}{lrrr}
\toprule
\textbf{Model} & \textbf{Params} & \textbf{Layers} & \textbf{Experts} \\
\midrule
Jin-Nano & 1B & 6 & 8 \\
Jin-Base & 7B & 12 & 16 \\
Jin-Large & 70B & 24 & 64 \\
\bottomrule
\end{tabular}
\end{table}

\section{Experiments}

\subsection{Zero-Shot Image Classification}

\begin{table}[h]
\centering
\caption{ImageNet zero-shot top-1 accuracy (\%).}
\label{tab:imagenet}
\begin{tabular}{lc}
\toprule
\textbf{Model} & \textbf{Accuracy} \\
\midrule
CLIP ViT-L/14 & 75.3 \\
ALIGN & 76.4 \\
Florence & 83.7 \\
\midrule
Jin-Base & 82.1 \\
Jin-Large & \textbf{85.2} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Audio-Text Retrieval}

\begin{table}[h]
\centering
\caption{AudioCaps retrieval recall@10.}
\label{tab:audiocaps}
\begin{tabular}{lcc}
\toprule
\textbf{Model} & \textbf{T$\rightarrow$A} & \textbf{A$\rightarrow$T} \\
\midrule
AudioCLIP & 45.2 & 43.8 \\
CLAP & 51.3 & 49.7 \\
\midrule
Jin-Base & 58.9 & 56.4 \\
Jin-Large & \textbf{63.2} & \textbf{61.8} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Cross-Modal Generation}

\begin{table}[h]
\centering
\caption{Cross-modal generation quality.}
\label{tab:generation}
\begin{tabular}{lcc}
\toprule
\textbf{Task} & \textbf{FID/FAD} & \textbf{CLIP-Score} \\
\midrule
Text $\rightarrow$ Image & 8.4 & 0.31 \\
Audio $\rightarrow$ Image & 12.1 & 0.27 \\
Image $\rightarrow$ Audio & 2.1 (FAD) & --- \\
Text $\rightarrow$ Audio & 1.8 (FAD) & --- \\
\bottomrule
\end{tabular}
\end{table}

\subsection{z-JEPA Ablation}

\begin{table}[h]
\centering
\caption{z-JEPA ablation on ImageNet linear probe.}
\label{tab:zjepa_ablation}
\begin{tabular}{lc}
\toprule
\textbf{Configuration} & \textbf{Accuracy} \\
\midrule
Single-scale JEPA (baseline) & 77.9 \\
+ Hierarchical levels & 79.4 \\
+ Cross-scale prediction & 80.8 \\
+ Adaptive zoom & 81.6 \\
+ Bidirectional feedback & \textbf{82.3} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Computational Efficiency}

\begin{table}[h]
\centering
\caption{Inference throughput (samples/sec) on A100.}
\label{tab:throughput}
\begin{tabular}{lccc}
\toprule
\textbf{Model} & \textbf{Text} & \textbf{Vision} & \textbf{Multi} \\
\midrule
Jin-Base (dense) & 890 & 234 & 156 \\
Jin-Base (MoE) & 1,240 & 312 & 223 \\
Speedup & 1.39x & 1.33x & 1.43x \\
\bottomrule
\end{tabular}
\end{table}

MoE routing provides 1.3-1.4x throughput improvement at equivalent quality.

\section{Applications}

\subsection{Robotics}

Jin enables multimodal perception for robotic systems:
\begin{itemize}
    \item Visual-audio scene understanding
    \item Proprioceptive state estimation (custom modality)
    \item Natural language instruction following
\end{itemize}

\subsection{Accessibility}

Cross-modal generation enables:
\begin{itemize}
    \item Audio descriptions from images
    \item Visual representations of soundscapes
    \item Sign language to text/audio translation
\end{itemize}

\section{Discussion}

\paragraph{Limitations.} Training requires large-scale paired multimodal data. The modality extension mechanism, while flexible, requires fine-tuning for optimal performance on new modalities.

\paragraph{Future Work.} We plan to explore: (1) online modality extension without any fine-tuning, (2) integration with world models for embodied AI, and (3) scaling to additional modalities (tactile, olfactory).

\section{Conclusion}

Jin demonstrates that joint embedding architectures can unify multiple modalities into a single, coherent representation space. The combination of modality-specific encoders, diffusion transformer backbone with MoE, and hierarchical z-JEPA enables state-of-the-art performance on multimodal understanding tasks while maintaining computational efficiency. The architecture's extensibility opens new possibilities for AI systems that perceive and interact with the world across all sensory modalities.

\bibliographystyle{plain}
\begin{thebibliography}{10}

\bibitem{jepa2022}
Y. LeCun. A Path Towards Autonomous Machine Intelligence. 2022.

\bibitem{moe2017}
N. Shazeer et al. Outrageously Large Neural Networks: The Sparsely-Gated Mixture-of-Experts Layer. ICLR, 2017.

\bibitem{ddpm2020}
J. Ho et al. Denoising Diffusion Probabilistic Models. NeurIPS, 2020.

\bibitem{dit2023}
W. Peebles, S. Xie. Scalable Diffusion Models with Transformers. ICCV, 2023.

\bibitem{vit2020}
A. Dosovitskiy et al. An Image is Worth 16x16 Words. ICLR, 2021.

\bibitem{ijepa2023}
M. Assran et al. Self-Supervised Learning from Images with a Joint-Embedding Predictive Architecture. CVPR, 2023.

\bibitem{vjepa2024}
A. Bardes et al. V-JEPA: Latent Video Prediction for Visual Representation Learning. 2024.

\bibitem{clip2021}
A. Radford et al. Learning Transferable Visual Models From Natural Language Supervision. ICML, 2021.

\end{thebibliography}

\end{document}
