\documentclass[11pt,twocolumn]{article}
\usepackage{shared/hanzocover}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{listings}
\input{shared/lstlang}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,positioning,fit}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},
    commentstyle=\color{codegreen},
    keywordstyle=\color{codepurple},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codegreen},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,
    breaklines=true,
    captionpos=b,
    keepspaces=true,
    numbers=left,
    numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=2
}
\lstset{style=mystyle}

\title{Hanzo Vector: High-Performance Vector Search Engine for AI Applications}
\author{Zach Kelling\thanks{zach@hanzo.ai} \\ \textit{Hanzo Industries} \\ \texttt{research@hanzo.ai}}
\date{March 2024}

\begin{document}

\hanzocoverpage

\begin{abstract}
We present Hanzo Vector, a high-performance vector search engine implemented in Rust, designed for retrieval-augmented generation (RAG) pipelines and semantic search in AI applications. Vector implements Hierarchical Navigable Small World (HNSW) graphs with configurable distance metrics (cosine, Euclidean, dot product), scalar and product quantization for memory efficiency, and filtered search with pre- and post-filtering strategies. The engine supports collections of up to 100 million vectors with 99\% recall at 1ms p95 latency for 1536-dimensional embeddings. Key innovations include: (1) a hybrid quantization scheme that selects scalar or product quantization per segment based on dimensionality, (2) filtered HNSW traversal that prunes the search graph using metadata predicates during navigation, and (3) a write-ahead log with snapshot isolation for concurrent reads and writes.
\end{abstract}

\section{Introduction}

The emergence of embedding-based retrieval as a core component of AI applications---from RAG~\cite{lewis2020} to semantic search to recommendation systems---has created demand for specialized vector databases. Existing solutions (Pinecone, Weaviate, Qdrant~\cite{qdrant2021}) provide managed services or general-purpose engines, but AI platform operators need an embeddable engine that can be deployed as a sidecar, integrated into existing infrastructure, and tuned for specific workload characteristics.

Hanzo Vector is designed as an embeddable Rust library with an optional gRPC server mode. It prioritizes three properties: (1) predictable latency under concurrent load, (2) memory efficiency through adaptive quantization, and (3) filtered search performance that does not degrade with restrictive filters.

\paragraph{Contributions.}
\begin{itemize}
    \item An HNSW implementation with segment-level adaptive quantization selecting between scalar and product quantization based on dimensionality and recall requirements.
    \item Filtered HNSW traversal integrating metadata predicates into graph navigation, avoiding the recall degradation of post-filtering.
    \item A WAL-based storage engine with snapshot isolation enabling concurrent reads and writes without lock contention.
    \item Benchmarks demonstrating 99\% recall at 1ms p95 latency on 10M vectors with 1536 dimensions.
\end{itemize}

\section{Architecture}

\subsection{System Design}

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    node distance=0.7cm,
    box/.style={rectangle, draw, minimum width=2cm, minimum height=0.5cm, align=center, font=\scriptsize},
    arrow/.style={->, thick}
]
\node[box] (api) {gRPC / HTTP\\API};
\node[box, below=of api] (engine) {Query\\Engine};
\node[box, below left=0.6cm and 0.3cm of engine] (hnsw) {HNSW\\Index};
\node[box, below right=0.6cm and 0.3cm of engine] (filter) {Filter\\Index};
\node[box, below=1.2cm of engine] (store) {Segment\\Store};
\node[box, below=of store] (wal) {WAL +\\Snapshots};
\node[box, right=1.5cm of api] (quant) {Quantizer\\(SQ/PQ)};

\draw[arrow] (api) -- (engine);
\draw[arrow] (engine) -- (hnsw);
\draw[arrow] (engine) -- (filter);
\draw[arrow] (hnsw) -- (store);
\draw[arrow] (filter) -- (store);
\draw[arrow] (store) -- (wal);
\draw[arrow] (quant) -- (hnsw);
\end{tikzpicture}
\caption{Vector search engine architecture.}
\label{fig:vector}
\end{figure}

Vector organizes data into collections, each containing segments (Figure~\ref{fig:vector}). Each segment maintains an HNSW index for vector similarity search and an inverted index for metadata filtering. The query engine merges results across segments.

\subsection{HNSW Index}

The HNSW algorithm~\cite{malkov2018} constructs a multi-layer navigable graph where each layer is a subset of the layer below. Search begins at the top layer and greedily navigates to the nearest neighbor, then descends to the next layer for refinement.

Vector's HNSW implementation uses the following parameters:
\begin{itemize}
    \item $M = 16$: Maximum connections per node (default).
    \item $ef_{\text{construction}} = 200$: Search width during index building.
    \item $ef_{\text{search}}$: Configurable per query (default: 128).
\end{itemize}

The expected search complexity is $O(M \cdot \log N)$ where $N$ is the number of vectors.

\subsection{Quantization}

Vector implements two quantization strategies:

\paragraph{Scalar Quantization (SQ).} Maps each float32 dimension to uint8:
\begin{equation}
    q_i = \text{round}\left(\frac{x_i - \min_i}{\max_i - \min_i} \times 255\right)
\end{equation}
reducing memory by 4x with typical recall loss under 1\%.

\paragraph{Product Quantization (PQ).} Divides the vector into $m$ subspaces, each quantized to $k$ centroids via k-means:
\begin{equation}
    \text{PQ}(\mathbf{x}) = (c_1(\mathbf{x}^1), c_2(\mathbf{x}^2), \ldots, c_m(\mathbf{x}^m))
\end{equation}
reducing memory by up to 32x for high-dimensional vectors.

The adaptive quantizer selects SQ for dimensions $\leq 256$ and PQ for higher dimensions, based on empirical recall-memory tradeoff analysis.

\section{Key Features}

\subsection{Filtered Search}

Metadata filters are integrated into HNSW traversal rather than applied as post-processing:

\begin{lstlisting}[language=Rust]
pub struct SearchRequest {
    pub vector: Vec<f32>,
    pub top_k: usize,
    pub filter: Option<Filter>,
    pub ef: Option<usize>,
}

pub enum Filter {
    Eq(String, Value),
    Range(String, Value, Value),
    And(Vec<Filter>),
    Or(Vec<Filter>),
    Not(Box<Filter>),
}
\end{lstlisting}

During HNSW traversal, candidate nodes are checked against the filter predicate before being added to the result set. If the filter is highly selective (estimated selectivity $< 0.1$), the engine falls back to a pre-filter strategy that scans the metadata index first and restricts HNSW search to matching IDs.

\subsection{Multi-Tenancy}

Collections support a \texttt{tenant\_id} field that partitions the HNSW index into per-tenant subgraphs. Queries are automatically scoped to the authenticated tenant's partition, providing isolation without separate collections.

\subsection{Streaming Upserts}

The WAL-based storage engine enables concurrent reads and writes. New vectors are appended to the WAL and become searchable within 10ms (the flush interval). Background compaction merges WAL entries into immutable segments with optimized HNSW indexes.

\section{Implementation}

Vector is implemented in 28,000 lines of Rust. The HNSW index uses \texttt{unsafe} Rust for SIMD-accelerated distance computations (AVX2 on x86, NEON on ARM). The gRPC API uses \texttt{tonic}; the HTTP API uses \texttt{axum}.

\begin{table}[h]
\centering
\caption{Search performance (10M vectors, 1536-dim).}
\label{tab:search}
\begin{tabular}{lrrr}
\toprule
\textbf{Config} & \textbf{Recall@10} & \textbf{p50 (ms)} & \textbf{p95 (ms)} \\
\midrule
Full precision & 99.5\% & 0.6 & 1.0 \\
SQ (uint8) & 98.8\% & 0.3 & 0.6 \\
PQ (m=96, k=256) & 97.2\% & 0.2 & 0.4 \\
\midrule
+ Filter (10\%) & 99.1\% & 0.8 & 1.4 \\
+ Filter (1\%) & 98.5\% & 1.2 & 2.1 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Memory Usage.} Full precision: 61GB for 10M$\times$1536. SQ: 15GB. PQ: 2.4GB.

\section{Security}

\paragraph{Authentication.} All API calls require a valid Hanzo IAM token. Collection-level ACLs restrict read/write access per tenant.

\paragraph{Encryption at Rest.} Segment files are encrypted with per-collection keys from Hanzo KMS. The WAL is encrypted with a separate key rotated daily.

\paragraph{Network Security.} gRPC endpoints require mTLS. The HTTP API supports TLS~1.3.

\paragraph{Input Validation.} Vector dimensions are validated against the collection schema. Metadata values are type-checked. Malformed requests are rejected before reaching the index.

\section{Conclusion}

Hanzo Vector provides an embeddable, high-performance vector search engine designed for AI workloads. Adaptive quantization, filtered HNSW traversal, and WAL-based storage enable 99\% recall at sub-millisecond latency on collections of tens of millions of vectors. The Rust implementation ensures predictable performance without garbage collection pauses, making Vector suitable for latency-sensitive RAG pipelines.

\bibliographystyle{plain}
\begin{thebibliography}{10}

\bibitem{malkov2018}
Y. Malkov and D. Yashunin. Efficient and robust approximate nearest neighbor search using Hierarchical Navigable Small World graphs. IEEE TPAMI, 2018.

\bibitem{lewis2020}
P. Lewis et al. Retrieval-Augmented Generation for Knowledge-Intensive NLP Tasks. NeurIPS, 2020.

\bibitem{qdrant2021}
Qdrant. Vector Search Engine. 2021.

\bibitem{jegou2011}
H. J\'{e}gou et al. Product Quantization for Nearest Neighbor Search. IEEE TPAMI, 2011.

\end{thebibliography}

\end{document}
