% Hanzo Platform Integration with Digital Securities
\documentclass[11pt]{article}
\usepackage{shared/hanzocover}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx,hyperref,booktabs,enumitem,tikz}
\usepackage{float}

\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,urlcolor=blue}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{definition}{Definition}[section]

\title{Hanzo Platform Integration\\[0.3em]
{\large with Digital Securities}}
\author{
    Zach Kelling \\
    \textit{Hanzo Industries Inc (Techstars '17)} \\
    \texttt{research@hanzo.ai}
}
\date{2025}

\begin{document}
\hanzocoverpage

% ============================================================================
% ABSTRACT
% ============================================================================
\begin{abstract}
We describe how the Hanzo platform infrastructure---Base, IAM, KMS,
Gateway---powers the ArcaLabs digital securities stack. Digital securities
require investor onboarding with KYC/AML verification, trade execution
through a regulated alternative trading system (ATS), institutional MPC
custody, and comprehensive reporting. Rather than building these
capabilities from scratch, ArcaLabs integrates Hanzo's existing platform
services to provide identity management, secret management, API routing,
and data persistence. We detail the full architecture: investor
onboarding via Simplici KYC integrated through Hanzo IAM, trade execution
through the Liquidity ATS with Hanzo Gateway routing, MPC custody with
keys managed by Hanzo KMS, and reporting pipelines built on Hanzo Base.
This integration demonstrates the composability of the Hanzo platform for
regulated financial applications.
\end{abstract}

% ============================================================================
\section{Introduction}
\label{sec:intro}
% ============================================================================

The issuance and trading of digital securities requires infrastructure
that spans identity management, cryptographic key management, API
gateway services, data persistence, and regulatory reporting. Building
each component from scratch would duplicate capabilities that already
exist in production within the Hanzo platform.

Hanzo has operated as an AI-powered commerce and infrastructure platform
since its founding (Techstars '17), serving thousands of applications
through its unified service stack. The platform provides five core
services relevant to digital securities:

\begin{enumerate}[leftmargin=1.4em]
    \item \textbf{Hanzo IAM}: Identity and access management with OIDC,
    OAuth 2.0, and org-scoped permissions.
    \item \textbf{Hanzo KMS}: Secrets and key management with
    envelope encryption.
    \item \textbf{Hanzo Base}: Application runtime with embedded
    SQLite (local dev) or PostgreSQL (production).
    \item \textbf{Hanzo Gateway}: API gateway with rate limiting,
    request routing, and TLS termination.
    \item \textbf{Hanzo Webhooks}: Event delivery with guaranteed
    at-least-once semantics.
\end{enumerate}

ArcaLabs consumes these services to build its digital securities
platform without reimplementing foundational infrastructure.

% ============================================================================
\section{Architecture Overview}
\label{sec:architecture}
% ============================================================================

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
    node distance=1.2cm and 2cm,
    box/.style={draw, rounded corners, minimum width=3cm,
                minimum height=0.8cm, font=\small},
    hanzo/.style={box, fill=black!8},
    arca/.style={box, fill=white},
    ext/.style={box, fill=black!3, dashed},
    arr/.style={->, >=stealth, thick}
]
    % Hanzo layer
    \node[hanzo] (iam) {Hanzo IAM};
    \node[hanzo, right=of iam] (kms) {Hanzo KMS};
    \node[hanzo, below=of iam] (base) {Hanzo Base};
    \node[hanzo, right=of base] (gw) {Hanzo Gateway};
    \node[hanzo, below=of gw] (hooks) {Hanzo Webhooks};

    % ArcaLabs layer
    \node[arca, below=2cm of iam] (onboard) {Investor Onboarding};
    \node[arca, right=of onboard] (trade) {Trade Engine};
    \node[arca, below=of onboard] (custody) {MPC Custody};
    \node[arca, right=of custody] (report) {Reporting};

    % External
    \node[ext, left=of onboard] (simplici) {Simplici KYC};
    \node[ext, right=of trade] (ats) {Liquidity ATS};
    \node[ext, below=of custody] (lux) {Lux C-Chain};

    % Arrows
    \draw[arr] (onboard) -- (iam);
    \draw[arr] (onboard) -- (simplici);
    \draw[arr] (custody) -- (kms);
    \draw[arr] (trade) -- (gw);
    \draw[arr] (trade) -- (ats);
    \draw[arr] (report) -- (base);
    \draw[arr] (report) -- (hooks);
    \draw[arr] (custody) -- (lux);

    % Labels
    \node[above=0.3cm of iam, font=\small\bfseries, xshift=1.5cm]
        {Hanzo Platform};
    \node[above=0.3cm of onboard, font=\small\bfseries, xshift=1.5cm]
        {ArcaLabs Services};
\end{tikzpicture}
\caption{ArcaLabs digital securities architecture on Hanzo platform}
\label{fig:architecture}
\end{figure}

% ============================================================================
\section{Investor Onboarding via Hanzo IAM}
\label{sec:onboarding}
% ============================================================================

Investor onboarding combines Hanzo IAM identity management with Simplici
KYC/AML verification.

\subsection{Identity Creation}

\begin{enumerate}[leftmargin=1.4em]
    \item Investor registers via ArcaLabs platform. Registration request
    routed through Hanzo Gateway.
    \item Hanzo IAM creates an identity record with OIDC subject
    identifier. The investor receives OAuth 2.0 credentials.
    \item IAM assigns the investor to the ArcaLabs organization with
    the \texttt{investor} role.
    \item All subsequent API calls include the IAM bearer token. Gateway
    extracts the org from the JWT \texttt{owner} claim and scopes all
    data queries to that org.
\end{enumerate}

\subsection{KYC/AML via Simplici}

\begin{enumerate}[leftmargin=1.4em]
    \item ArcaLabs onboarding service redirects the investor to
    Simplici's verification flow.
    \item Simplici performs: identity document verification, liveness
    detection, OFAC/sanctions screening, PEP screening, and accredited
    investor verification.
    \item Upon completion, Simplici returns a signed attestation to
    ArcaLabs via webhook.
    \item ArcaLabs compliance service validates the attestation and
    updates the investor's IAM profile with: KYC status, accreditation
    status, accreditation expiry, jurisdiction.
    \item The compliance service writes the investor to the on-chain
    \texttt{ComplianceRegistry} whitelist on Lux C-Chain.
\end{enumerate}

\subsection{Ongoing Monitoring}

Simplici performs continuous monitoring:
\begin{itemize}[leftmargin=1.4em]
    \item Daily sanctions list re-screening.
    \item Adverse media monitoring.
    \item Annual accreditation re-verification trigger.
    \item Alerts delivered to ArcaLabs via Hanzo Webhooks.
\end{itemize}

% ============================================================================
\section{Trade Execution via Liquidity ATS}
\label{sec:trade}
% ============================================================================

Secondary trading of digital securities occurs through the Liquidity
ATS (Alternative Trading System), a SEC-registered trading venue.

\subsection{Order Flow}

\begin{enumerate}[leftmargin=1.4em]
    \item Investor submits order via ArcaLabs trading interface.
    \item Request authenticated by Hanzo IAM, routed through Hanzo
    Gateway to the ArcaLabs trade engine.
    \item Trade engine validates: investor whitelist status, lockup
    compliance, accreditation validity.
    \item Approved order forwarded to Liquidity ATS matching engine.
    \item ATS matches buy and sell orders. Matched trades return to
    ArcaLabs for settlement.
    \item Settlement: atomic swap of security tokens and USDC on Lux
    C-Chain via MPC-signed transaction.
\end{enumerate}

\subsection{Gateway Configuration}

Hanzo Gateway handles the routing between ArcaLabs services and the
Liquidity ATS:

\begin{itemize}[leftmargin=1.4em]
    \item \textbf{Rate limiting}: Per-investor and per-org request limits
    to prevent abuse.
    \item \textbf{Authentication}: IAM token validation at the gateway
    layer. Invalid tokens rejected before reaching backend services.
    \item \textbf{TLS}: End-to-end encryption. Mutual TLS between
    Gateway and ATS.
    \item \textbf{Audit logging}: Every request logged with investor
    identity, timestamp, and response code.
\end{itemize}

% ============================================================================
\section{MPC Custody via Hanzo KMS}
\label{sec:custody}
% ============================================================================

MPC key material is managed through Hanzo KMS (kms.hanzo.ai):

\subsection{Key Hierarchy}

\begin{enumerate}[leftmargin=1.4em]
    \item \textbf{KMS Master Key}: Root encryption key managed by Hanzo
    KMS. Never exported.
    \item \textbf{MPC Share Encryption Keys}: Per-share-holder encryption
    keys derived from the KMS master key via envelope encryption.
    \item \textbf{MPC Key Shares}: The actual CGGMP21 key shares,
    encrypted at rest with their respective share encryption keys.
\end{enumerate}

\subsection{Operational Flow}

\begin{enumerate}[leftmargin=1.4em]
    \item \textbf{DKG}: During distributed key generation, each MPC
    node requests a share encryption key from KMS. The generated key
    share is immediately encrypted and stored.
    \item \textbf{Signing}: When a signing session begins, the MPC node
    requests decryption of its key share from KMS. KMS validates the
    request against IAM identity and returns the decrypted share. The
    share exists in memory only during the signing session.
    \item \textbf{Rotation}: During key resharing, new shares are
    encrypted with fresh KMS-derived keys. Old encrypted shares are
    securely deleted.
    \item \textbf{Audit}: Every KMS access is logged: who requested
    what key, when, and for what purpose.
\end{enumerate}

% ============================================================================
\section{Data Persistence via Hanzo Base}
\label{sec:base}
% ============================================================================

Hanzo Base provides the data layer for ArcaLabs services:

\begin{description}[leftmargin=1.4em]
    \item[Investor records] KYC status, accreditation, jurisdiction,
    wallet addresses. Linked to IAM identity.
    \item[Trade records] Order submissions, matches, settlements.
    Immutable audit trail.
    \item[Compliance events] Whitelist changes, restriction violations,
    corporate actions. Indexed for regulatory reporting.
    \item[NAV history] Daily NAV calculations for ArCoin and other
    fund products.
    \item[MPC session logs] Signing requests, policy decisions,
    completion status. Required for SEC 17a-4 compliance.
\end{description}

Base configuration:
\begin{itemize}[leftmargin=1.4em]
    \item \textbf{Local development}: SQLite embedded in Base runtime.
    Zero external dependencies.
    \item \textbf{Production}: PostgreSQL backend with Base runtime
    providing the application layer, authentication (via IAM), and
    real-time subscriptions.
    \item \textbf{Replication}: PostgreSQL streaming replication for
    high availability.
\end{itemize}

% ============================================================================
\section{Reporting Pipeline}
\label{sec:reporting}
% ============================================================================

ArcaLabs' reporting requirements span regulatory filings, investor
communications, and internal audit:

\subsection{Regulatory Reports}

\begin{itemize}[leftmargin=1.4em]
    \item \textbf{Form D filings}: Generated from issuance records in
    Base. Auto-populated with investor counts, amounts raised,
    exemption type.
    \item \textbf{Blue sky filings}: State-level notice filings based
    on investor jurisdiction data.
    \item \textbf{ATS Form ATS-N}: Quarterly filing for the Liquidity
    ATS, generated from trade records.
    \item \textbf{FINRA trade reports}: Real-time trade reporting
    generated from settlement events.
\end{itemize}

\subsection{Investor Communications}

\begin{itemize}[leftmargin=1.4em]
    \item \textbf{Account statements}: Monthly statements generated from
    Base records, delivered via Hanzo Webhooks to investor email.
    \item \textbf{Tax documents}: K-1 (fund interests) and 1099
    (interest/dividends) generated annually.
    \item \textbf{NAV reports}: Daily NAV publications for fund products.
\end{itemize}

\subsection{Event Pipeline}

\begin{enumerate}[leftmargin=1.4em]
    \item On-chain events (transfers, compliance decisions, corporate
    actions) are indexed by an event listener.
    \item Events written to Base with chain metadata (block number,
    tx hash, timestamp).
    \item Hanzo Webhooks delivers events to configured endpoints
    (reporting systems, compliance dashboards, investor portals).
    \item Events are immutable in Base---no deletion, only append.
\end{enumerate}

% ============================================================================
\section{Full Architecture Diagram}
\label{sec:fullarch}
% ============================================================================

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
    scale=0.85, transform shape,
    node distance=0.9cm and 1.5cm,
    box/.style={draw, rounded corners, minimum width=2.4cm,
                minimum height=0.7cm, font=\footnotesize, align=center},
    hanzo/.style={box, fill=black!8},
    arca/.style={box, fill=white},
    chain/.style={box, fill=black!3, dashed},
    arr/.style={->, >=stealth}
]
    % Top: Investor + Merchant
    \node[arca] (inv) {Investor};
    \node[arca, right=3cm of inv] (merch) {Issuer};

    % Gateway
    \node[hanzo, below=of inv, xshift=1.5cm] (gw) {Hanzo Gateway};

    % IAM + KMS
    \node[hanzo, below left=of gw] (iam) {Hanzo IAM};
    \node[hanzo, below right=of gw] (kms) {Hanzo KMS};

    % ArcaLabs services
    \node[arca, below=1.5cm of iam] (onboard) {Onboarding};
    \node[arca, right=of onboard] (trade) {Trade Engine};
    \node[arca, right=of trade] (custody) {MPC Custody};

    % External
    \node[chain, below left=of onboard] (simplici) {Simplici};
    \node[chain, below=of trade] (ats) {Liquidity ATS};

    % Base + Webhooks
    \node[hanzo, below=1.5cm of custody] (base) {Hanzo Base};
    \node[hanzo, left=of base] (hooks) {Webhooks};

    % Chain
    \node[chain, below=of base] (lux) {Lux C-Chain};

    % Arrows
    \draw[arr] (inv) -- (gw);
    \draw[arr] (merch) -- (gw);
    \draw[arr] (gw) -- (iam);
    \draw[arr] (gw) -- (kms);
    \draw[arr] (iam) -- (onboard);
    \draw[arr] (onboard) -- (simplici);
    \draw[arr] (onboard) -- (trade);
    \draw[arr] (trade) -- (ats);
    \draw[arr] (trade) -- (custody);
    \draw[arr] (custody) -- (kms);
    \draw[arr] (custody) -- (lux);
    \draw[arr] (trade) -- (base);
    \draw[arr] (base) -- (hooks);
    \draw[arr] (hooks) -- (lux);
\end{tikzpicture}
\caption{Complete system architecture: Hanzo platform + ArcaLabs services + external dependencies}
\label{fig:fullarch}
\end{figure}

% ============================================================================
\section{Conclusion}
\label{sec:conclusion}
% ============================================================================

The Hanzo platform provides the foundational infrastructure for
ArcaLabs' digital securities operations. IAM handles investor identity
and access control with org-scoped permissions. KMS secures MPC key
material with envelope encryption and comprehensive audit logging. Base
provides the data persistence layer with production-grade PostgreSQL and
development-friendly SQLite. Gateway routes and rate-limits all traffic
with IAM token validation. Webhooks delivers lifecycle events to
reporting and compliance systems.

This integration demonstrates that regulated financial applications can
be built by composing existing platform services rather than
reimplementing infrastructure. The Hanzo platform's service boundaries
align with the separation of concerns required by securities
regulation: identity (IAM), custody (KMS + MPC), data (Base), and
communication (Gateway + Webhooks).

\begin{thebibliography}{99}
\bibitem{hanzoplatform}
Hanzo Industries, ``Hanzo Platform: AI-Powered Infrastructure,''
Product Documentation, 2024.

\bibitem{cggmp21}
R.~Canetti, R.~Gennaro, S.~Goldfeder, N.~Makriyannis, U.~Peled,
``UC Non-Interactive, Proactive, Threshold ECDSA with Identifiable
Aborts,'' \textit{ACM CCS}, 2020.

\bibitem{luxwhitepaper}
Lux Partners, ``Lux Network: A Scalable Multi-Chain Architecture,''
Technical Report, 2022.

\bibitem{simplici2024}
Simplici Inc., ``Broker-Dealer Integrated KYC/AML Platform,''
Product Documentation, 2024.

\bibitem{sec17a4}
SEC, ``Records to Be Preserved by Certain Exchange Members, Brokers and
Dealers,'' 17 CFR \S~240.17a-4.

\bibitem{oidc}
N.~Sakimura et al., ``OpenID Connect Core 1.0,'' OpenID Foundation,
2014.

\bibitem{oauth2}
D.~Hardt, ``The OAuth 2.0 Authorization Framework,'' RFC 6749, 2012.
\end{thebibliography}

\end{document}
