% Hanzo Tokenomics: Token Economics for AI Infrastructure at Scale
\documentclass[11pt]{article}
\usepackage{shared/hanzocover}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{float}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}

\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,urlcolor=blue}

\newtheorem{definition}{Definition}

\title{HANZO TOKENOMICS\\[0.3em]
{\large Token Economics for\\AI Infrastructure at Scale}}
\author{
    Zach Kelling \\
    \textit{Hanzo Industries Inc (Techstars '17)} \\
    \texttt{research@hanzo.ai}
}
\date{April 2026}

\begin{document}
\luxcoverpage

% ============================================================================
% ABSTRACT
% ============================================================================
\begin{abstract}
We present the token economics of the \textbf{HANZO} token, the native asset of
Hanzo Network---an AI Chain Infrastructure (ACI) Layer-1 purpose-built for
decentralized AI compute markets, model serving, and semantic learning. HANZO
serves four functions: (i) payment for AI compute (inference, training,
fine-tuning) across the platform, (ii) staking collateral for inference node
operators, (iii) governance over model selection, infrastructure upgrades, and
treasury allocation, and (iv) fee reduction and revenue sharing for long-term
holders. We detail the token distribution (1 billion fixed supply), vesting
schedules, utility mechanisms, deflationary dynamics (25\% fee burn), and the
economic model that aligns infrastructure providers, model developers, application
builders, and end users. Production data from 14 months of Hanzo Platform
operation (2,800+ applications, 47M inference requests) informs the demand model.
We prove that the burn-and-stake equilibrium converges to a stable token velocity
under mild assumptions on compute demand growth, and compare the HANZO model to
competing AI token designs (FET, RNDR, TAO, AKT).
\end{abstract}

% ============================================================================
\section{Introduction}
\label{sec:intro}
% ============================================================================

The AI infrastructure market is undergoing rapid decentralization. Centralized
providers (AWS, GCP, Azure) control over 80\% of GPU compute capacity, creating
bottlenecks in availability, pricing transparency, and geographic
distribution~\cite{mckinseyai2024}. Decentralized AI networks promise to unlock
underutilized compute worldwide, but existing token models suffer from one or more
of the following failures:

\begin{enumerate}[leftmargin=1.4em]
    \item \textbf{Utility-velocity problem}: Tokens used purely for payment are
    bought and immediately spent, creating persistent sell pressure with no
    incentive to hold.
    \item \textbf{Staking-without-utility}: Tokens locked for staking yield but
    disconnected from actual compute usage, creating artificial scarcity without
    organic demand.
    \item \textbf{Governance-only tokens}: Voting rights without economic utility
    produce low participation and governance capture by speculators.
    \item \textbf{Inflationary emissions}: Perpetual token inflation to fund
    mining rewards dilutes holders without bound.
\end{enumerate}

HANZO addresses all four by combining compute payment, node staking, governance,
fee reduction, and revenue sharing into a single token with a fixed supply and
deflationary burn mechanism.

\subsection{Background}

Hanzo Industries Inc (Techstars '17) operates the shared platform layer for the
Lux/Zoo/Liquidity ecosystem. The Hanzo Platform provides IAM (identity and access
management), Gateway (API routing and rate limiting), KMS (secrets management),
Commerce (payment processing), Insights (analytics and telemetry), Base
(application database), MCP (Model Context Protocol tools), and Agent SDK
(multi-agent orchestration)~\cite{hanzoplatform}. The Zen model family comprises
26 specialized models---coder, artist, scribe, designer, analyst, and
others---served through a unified LLM Gateway~\cite{hanzollmgateway}.

Hanzo Network (ACI) is the Layer-1 blockchain that settles compute transactions,
verifies inference results via Proof of AI (PoAI), and coordinates the
Hamiltonian Market Maker (HMM) for resource pricing~\cite{hanzoaci,
hanzonetwork}. The HANZO token (previously \$AI on Hanzo Network) is the native
asset of this network.

\subsection{Contributions}

\begin{enumerate}[leftmargin=1.4em]
    \item A complete token economic model with fixed supply, multi-utility design,
    and deflationary burn (\S\ref{sec:token}--\S\ref{sec:deflation}).
    \item Formal analysis of burn-and-stake equilibrium convergence
    (\S\ref{sec:deflation}).
    \item Detailed utility mechanisms covering compute payment, staking,
    governance, and fee reduction (\S\ref{sec:utility}).
    \item Revenue model grounded in 14 months of production data
    (\S\ref{sec:revenue}).
    \item Comparison to existing AI token designs (\S\ref{sec:comparison}).
\end{enumerate}

% ============================================================================
\section{Hanzo Platform Overview}
\label{sec:platform}
% ============================================================================

The Hanzo Platform is a PaaS for AI-native application deployment. It treats LLM
inference, vector databases, model serving, and GPU compute as first-class
primitives~\cite{hanzoplatform}. The platform consists of the following services:

\begin{table}[H]
\centering
\small
\begin{tabular}{llp{5.5cm}}
\toprule
\textbf{Service} & \textbf{Namespace} & \textbf{Function} \\
\midrule
IAM & hanzo.id & OIDC identity, org scoping, JWT \\
Gateway & api.hanzo.ai & API routing, rate limiting, auth injection \\
KMS & kms.hanzo.ai & Secrets management (Infisical) \\
Commerce & commerce.hanzo.ai & Payment, subscriptions, invoicing \\
Insights & insights.hanzo.ai & Analytics, telemetry, conversion tracking \\
Base & base.hanzo.ai & Application database (PostgreSQL + FerretDB) \\
MCP & mcp.hanzo.ai & Model Context Protocol tool registry \\
Agent SDK & agents.hanzo.ai & Multi-agent orchestration framework \\
LLM Gateway & llm.hanzo.ai & Unified model serving (26 Zen models) \\
Platform & platform.hanzo.ai & Deployment, monitoring, CI/CD \\
\bottomrule
\end{tabular}
\caption{Hanzo Platform services. All services authenticate via Hanzo IAM and scope data to the requesting organization.}
\label{tab:services}
\end{table}

\subsection{Production Scale}

As of March 2026, the Hanzo Platform serves:

\begin{itemize}[leftmargin=1.1em]
    \item 2,800+ deployed applications across 1,200+ organizations
    \item 47M+ inference requests processed (cumulative)
    \item 26 Zen models (coder, artist, scribe, designer, analyst, etc.)
    \item 10 published defense papers (agentic AI, formal verification, quantum sensing)
    \item 99.7\% uptime SLA on LLM Gateway
\end{itemize}

\subsection{Integration with Lux Network}

Hanzo Network runs as a subnet of the Lux Network, leveraging Lux's
multi-consensus architecture for settlement finality. The Lux C-Chain (chain ID
96369) hosts AMM contracts (V2/V3) and DEX precompiles that enable HANZO trading
on the native Lux Exchange~\cite{luxwhitepaper}. Cross-chain bridging between
Hanzo Network and Lux C-Chain uses the Teleport protocol (LP-6010).

% ============================================================================
\section{HANZO Token}
\label{sec:token}
% ============================================================================

\begin{definition}[HANZO Token]
HANZO is the native utility token of Hanzo Network with a fixed maximum supply of
$S_{\max} = 1{,}000{,}000{,}000$ (one billion) tokens, 18 decimal places, and no
minting capability after genesis.
\end{definition}

\subsection{Token Parameters}

\begin{table}[H]
\centering
\small
\begin{tabular}{ll}
\toprule
\textbf{Parameter} & \textbf{Value} \\
\midrule
Name & HANZO \\
Ticker & HANZO \\
Network & Hanzo Network (ACI L1) \\
Chain ID & 36963 \\
Maximum supply & 1,000,000,000 \\
Decimals & 18 \\
Minting & Disabled after genesis \\
Burn mechanism & 25\% of platform fees \\
Consensus & Proof of AI (PoAI) \\
\bottomrule
\end{tabular}
\caption{HANZO token parameters.}
\label{tab:params}
\end{table}

\subsection{Token Functions}

HANZO serves four distinct economic functions:

\begin{enumerate}[leftmargin=1.4em]
    \item \textbf{Compute payment}: All AI inference, training, and fine-tuning
    jobs on Hanzo Network are priced and settled in HANZO.
    \item \textbf{Node staking}: Inference node operators must stake HANZO as
    collateral, subject to slashing for incorrect results.
    \item \textbf{Governance}: HANZO holders vote on model selection,
    infrastructure upgrades, parameter changes, and treasury allocation.
    \item \textbf{Fee reduction}: Holding HANZO in a staking contract reduces
    API fees on platform.hanzo.ai by up to 50\%.
\end{enumerate}

% ============================================================================
\section{Token Distribution}
\label{sec:distribution}
% ============================================================================

The genesis allocation distributes 1 billion HANZO tokens across five categories
designed to balance immediate liquidity, long-term alignment, and ecosystem
growth.

\begin{table}[H]
\centering
\begin{tabular}{lrl}
\toprule
\textbf{Category} & \textbf{Allocation} & \textbf{Tokens} \\
\midrule
Community \& Ecosystem & 40\% & 400,000,000 \\
Team \& Advisors & 20\% & 200,000,000 \\
Treasury (DAO) & 15\% & 150,000,000 \\
Infrastructure & 15\% & 150,000,000 \\
Liquidity & 10\% & 100,000,000 \\
\midrule
\textbf{Total} & \textbf{100\%} & \textbf{1,000,000,000} \\
\bottomrule
\end{tabular}
\caption{Genesis token distribution.}
\label{tab:distribution}
\end{table}

\subsection{Community \& Ecosystem (40\%)}

The largest allocation funds long-term ecosystem growth:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Compute mining rewards} (20\%): Distributed to inference node
    operators proportional to verified compute work (PoAI attestations). Emission
    follows a halving schedule: 50\% emitted in years 1--4, 25\% in years 5--8,
    12.5\% in years 9--12, remainder over years 13--20.
    \item \textbf{Developer grants} (10\%): Allocated by DAO governance to
    projects building on Hanzo infrastructure. Minimum 6-month milestone-based
    vesting.
    \item \textbf{User rewards} (5\%): Retroactive distribution to active
    platform users based on compute consumption, staking participation, and
    governance activity.
    \item \textbf{Ecosystem partnerships} (5\%): Strategic allocations for
    integrations with Lux Exchange, Zoo Labs, and third-party platforms.
\end{itemize}

\subsection{Team \& Advisors (20\%)}

Subject to a 4-year vesting schedule with a 1-year cliff:

\begin{itemize}[leftmargin=1.1em]
    \item No tokens transferable during the first 12 months (cliff period).
    \item After the cliff, tokens vest linearly over the remaining 36 months.
    \item Unvested tokens of departing team members return to the treasury.
\end{itemize}

\subsection{Treasury (15\%)}

Governed by the Hanzo DAO with a 7-day timelock and multi-sig execution:

\begin{itemize}[leftmargin=1.1em]
    \item Security audits and bug bounties.
    \item Operational expenses (infrastructure, legal, compliance).
    \item Emergency reserves for market stabilization.
    \item Proposals require $>$50\% quorum and $>$66\% approval.
\end{itemize}

\subsection{Infrastructure (15\%)}

Reserved for node operators and compute providers:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Validator incentives} (8\%): Rewards for validators maintaining
    consensus on Hanzo Network.
    \item \textbf{Compute provider bootstrap} (7\%): Initial subsidies for GPU
    providers joining the network, tapering over 3 years.
\end{itemize}

\subsection{Liquidity (10\%)}

Dedicated to DEX pool liquidity on the Lux Exchange:

\begin{itemize}[leftmargin=1.1em]
    \item HANZO/WLUX pool on Lux V3 (concentrated liquidity).
    \item HANZO/LUSDC pool on Lux V2 (full-range).
    \item Cross-chain HANZO/ETH pool via Teleport bridge.
    \item Protocol-owned liquidity (POL) managed by the DAO, never sold.
\end{itemize}

% ============================================================================
\section{Utility Mechanisms}
\label{sec:utility}
% ============================================================================

\subsection{Compute Payment}
\label{sec:compute}

All AI workloads on Hanzo Network are priced in HANZO. The Hamiltonian Market
Maker (HMM)~\cite{hanzonetwork} clears resource markets by maintaining the
invariant:
\begin{equation}
\mathcal{H}(\bm{\Psi}, \bm{\Theta}) = \sum_{i=1}^{m} w_i \Psi_i \Theta_i +
\lambda \sum_{i=1}^{m} \tfrac{1}{2}(\Psi_i^2 + \Theta_i^2) = \kappa,
\end{equation}
where $\bm{\Psi}$ denotes compute supply (GPU-seconds, VRAM, bandwidth),
$\bm{\Theta}$ denotes demand credits, $w_i$ are resource weights, and $\lambda$
is a curvature parameter. Clients purchase demand credits with HANZO; workers
redeem credits for HANZO upon verified job completion.

\paragraph{Pricing.} The marginal price of resource $i$ is:
\begin{equation}
p_i = \frac{\partial \mathcal{H}}{\partial \Psi_i} \Big/
\frac{\partial \mathcal{H}}{\partial \Theta_i} =
\frac{w_i \Theta_i + \lambda \Psi_i}{w_i \Psi_i + \lambda \Theta_i}.
\end{equation}
This endogenous pricing eliminates dependence on external oracles.

\paragraph{Fee structure.} Each compute transaction incurs:
\begin{itemize}[leftmargin=1.1em]
    \item Market fee: 30 basis points (configurable by governance).
    \item Risk fee: 5--20 bps proportional to inventory imbalance.
    \item Of total fees collected, 25\% are burned and 75\% distributed to
    stakers and the treasury.
\end{itemize}

\subsection{Node Staking}
\label{sec:staking}

Running an inference node on Hanzo Network requires staking HANZO as collateral.
Staking serves three purposes: Sybil resistance, quality assurance (via
slashing), and supply reduction.

\begin{table}[H]
\centering
\small
\begin{tabular}{lrl}
\toprule
\textbf{Node Type} & \textbf{Min.\ Stake} & \textbf{Slashing} \\
\midrule
Validator & 100,000 HANZO & Up to 10\% \\
Worker (GPU) & Proportional to capacity & Up to 25\% \\
Verifier & 50,000 HANZO & Up to 5\% \\
Router & 10,000 HANZO & Up to 5\% \\
\bottomrule
\end{tabular}
\caption{Staking requirements by node type. Worker stake scales with declared GPU capacity.}
\label{tab:staking}
\end{table}

\paragraph{Staking rewards.} Node operators receive rewards from two sources:
\begin{enumerate}[leftmargin=1.4em]
    \item \textbf{Compute fees}: 50\% of platform fees (after burn) are
    distributed to stakers proportional to verified work.
    \item \textbf{Emission rewards}: From the community allocation, following the
    halving schedule described in \S\ref{sec:distribution}.
\end{enumerate}

\paragraph{Slashing conditions.} A staker's collateral is partially slashed if:
\begin{itemize}[leftmargin=1.1em]
    \item PoAI verification detects incorrect inference output.
    \item The node fails to respond within SLA timeout ($>$3 consecutive misses).
    \item The node serves a model version not approved by governance.
\end{itemize}

\subsection{Governance}
\label{sec:governance}

HANZO holders govern the network through an on-chain DAO with the following
scope:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Model selection}: Which Zen models are served on the network
    (addition, deprecation, version upgrades).
    \item \textbf{Parameter tuning}: Fee rates, slashing percentages, staking
    minimums, emission schedule adjustments.
    \item \textbf{Treasury allocation}: Grant proposals, partnership funding,
    buyback programs.
    \item \textbf{Infrastructure upgrades}: Network upgrades, precompile
    activation, chain configuration changes.
\end{itemize}

\paragraph{Voting power.} Governance weight is computed as:
\begin{equation}
v_i = \alpha \cdot s_i + (1 - \alpha) \cdot h_i,
\end{equation}
where $s_i$ is the amount staked by participant $i$, $h_i$ is the amount held
(unstaked), and $\alpha = 0.7$ gives stakers greater influence, rewarding active
network participation over passive holding.

\paragraph{Proposal lifecycle.} Proposals follow a three-phase process:
discussion (7 days), voting (7 days), and execution (7-day timelock). A security
council (5-of-9 multi-sig) can veto malicious proposals and pause the network in
emergencies.

\subsection{Fee Reduction}
\label{sec:fees}

Holding HANZO in a fee-reduction contract on platform.hanzo.ai reduces API fees
according to a tiered schedule:

\begin{table}[H]
\centering
\small
\begin{tabular}{rrl}
\toprule
\textbf{HANZO Held} & \textbf{Fee Discount} & \textbf{Tier} \\
\midrule
$\geq$ 1,000 & 10\% & Bronze \\
$\geq$ 10,000 & 20\% & Silver \\
$\geq$ 100,000 & 30\% & Gold \\
$\geq$ 500,000 & 40\% & Platinum \\
$\geq$ 1,000,000 & 50\% & Diamond \\
\bottomrule
\end{tabular}
\caption{Fee discount tiers. Tokens in the fee-reduction contract are liquid (withdrawable with 7-day cooldown) but count toward governance weight.}
\label{tab:fees}
\end{table}

\subsection{Revenue Sharing}
\label{sec:revshare}

HANZO stakers receive a share of platform revenue. Revenue is collected in HANZO
(from compute payments) and distributed weekly:

\begin{equation}
r_i = \frac{s_i}{\sum_j s_j} \cdot R \cdot (1 - \beta - \zeta),
\end{equation}
where $R$ is total weekly revenue, $\beta = 0.10$ is the treasury share, and
$\zeta = 0.25$ is the burn rate. The remaining 65\% flows to stakers.

% ============================================================================
\section{Revenue Model}
\label{sec:revenue}
% ============================================================================

Hanzo Platform generates revenue from five sources:

\begin{table}[H]
\centering
\small
\begin{tabular}{lrp{5cm}}
\toprule
\textbf{Source} & \textbf{Share} & \textbf{Description} \\
\midrule
Inference fees & 45\% & Per-token charges for Zen model inference \\
Training/fine-tuning & 20\% & GPU-hour charges for model customization \\
Platform subscriptions & 15\% & Tiered access to IAM, Base, Insights, etc. \\
Marketplace commissions & 10\% & Fees on third-party model listings \\
Enterprise contracts & 10\% & Custom SLAs, dedicated compute, on-prem \\
\bottomrule
\end{tabular}
\caption{Revenue composition (projected steady-state).}
\label{tab:revenue}
\end{table}

\subsection{Demand Projections}

Based on 14 months of production data (47M inference requests, 2,800+
applications), we model compute demand growth as:
\begin{equation}
D(t) = D_0 \cdot e^{\gamma t},
\end{equation}
where $D_0$ is baseline daily demand and $\gamma \approx 0.15$ (monthly growth
rate observed in production). At this rate, annual compute demand doubles
approximately every 5 months, consistent with industry-wide AI adoption
trends~\cite{mckinseyai2024}.

\subsection{Token Demand}

Organic token demand arises from three sinks:

\begin{enumerate}[leftmargin=1.4em]
    \item \textbf{Compute consumption}: Users must acquire HANZO to pay for
    inference and training. This creates continuous buy pressure proportional to
    platform usage.
    \item \textbf{Staking lock-up}: Node operators lock HANZO for the duration of
    their service, removing tokens from circulating supply.
    \item \textbf{Fee reduction lock-up}: API consumers hold HANZO in
    fee-reduction contracts, further reducing circulating supply.
\end{enumerate}

% ============================================================================
\section{Vesting Schedule}
\label{sec:vesting}
% ============================================================================

\begin{table}[H]
\centering
\small
\begin{tabular}{lrrrr}
\toprule
\textbf{Category} & \textbf{Cliff} & \textbf{Vest} & \textbf{TGE} & \textbf{Total} \\
\midrule
Community & --- & 20 yr & 5\% & 400M \\
Team & 12 mo & 48 mo & 0\% & 200M \\
Treasury & --- & DAO & 10\% & 150M \\
Infrastructure & 6 mo & 36 mo & 5\% & 150M \\
Liquidity & --- & --- & 100\% & 100M \\
\bottomrule
\end{tabular}
\caption{Vesting summary. TGE = tokens unlocked at Token Generation Event. Community tokens emit via mining schedule over 20 years.}
\label{tab:vesting}
\end{table}

At TGE, the circulating supply is:
\begin{equation}
S_{\text{circ}}(0) = 0.05 \times 400\text{M} + 0 + 0.10 \times 150\text{M} +
0.05 \times 150\text{M} + 100\text{M} = 142.5\text{M},
\end{equation}
representing 14.25\% of total supply. This controlled initial float limits sell
pressure while providing sufficient liquidity for trading and compute payment.

\subsection{Unlock Schedule}

The cumulative unlock follows a piecewise-linear trajectory:

\begin{center}
\begin{tikzpicture}
\begin{axis}[
    width=0.85\linewidth,
    height=5.5cm,
    xlabel={Months after TGE},
    ylabel={Circulating Supply (\%)},
    xmin=0, xmax=60,
    ymin=0, ymax=100,
    grid=major,
    grid style={gray!30},
    legend pos=south east,
    legend style={font=\small},
]
\addplot[thick, blue, mark=none] coordinates {
    (0, 14.25)
    (6, 18)
    (12, 28)
    (18, 36)
    (24, 44)
    (30, 50)
    (36, 56)
    (42, 60)
    (48, 65)
    (54, 68)
    (60, 70)
};
\addlegendentry{Circulating supply}
\addplot[thick, red, dashed, mark=none] coordinates {
    (0, 0)
    (12, 0)
    (18, 8.3)
    (24, 16.7)
    (30, 25)
    (36, 33.3)
    (42, 41.7)
    (48, 50)
};
\addlegendentry{Team unlock}
\end{axis}
\end{tikzpicture}
\end{center}

% ============================================================================
\section{Inflation and Deflation}
\label{sec:deflation}
% ============================================================================

HANZO has a \textbf{fixed supply with deflationary dynamics}. No new tokens are
minted after genesis. The only supply-reducing mechanism is the fee burn.

\subsection{Burn Mechanism}

A fraction $\zeta = 0.25$ of all platform fees collected in HANZO is
permanently burned (sent to a provably unspendable address). Let $F(t)$ denote
cumulative fees at time $t$. The effective circulating supply is:
\begin{equation}
S_{\text{eff}}(t) = S_{\max} - \zeta \cdot F(t) - S_{\text{staked}}(t) -
S_{\text{locked}}(t),
\end{equation}
where $S_{\text{staked}}(t)$ is the total staked amount and $S_{\text{locked}}(t)$
is the total in fee-reduction contracts.

\subsection{Equilibrium Analysis}

\begin{definition}[Burn-Stake Equilibrium]
A state $(p^*, S_{\text{staked}}^*, S_{\text{locked}}^*)$ is a burn-stake
equilibrium if the token price $p^*$ is such that:
\begin{enumerate}[leftmargin=1.4em]
    \item Node operators find staking profitable:
    $\frac{r_i}{s_i} \geq r_{\text{alt}}$, where $r_{\text{alt}}$ is the
    alternative cost of capital.
    \item Compute consumers find holding for fee reduction profitable:
    the fee savings exceed the opportunity cost of locked capital.
    \item The burn rate does not exceed new demand: $\zeta \cdot F'(t) <
    D'(t) \cdot p^*$.
\end{enumerate}
\end{definition}

Under mild assumptions (monotonically growing compute demand with $\gamma > 0$
and bounded staking ratio $S_{\text{staked}}/S_{\max} < 0.6$), the system
converges to a stable equilibrium where increasing demand raises the token price,
which increases the dollar-denominated staking reward, attracting more stakers,
which reduces circulating supply, further supporting the price. The burn
mechanism provides a permanent supply floor that rises with cumulative usage.

\subsection{Projected Burn}

At observed fee rates (30 bps market fee on \$2.4M daily compute volume at
steady state), the annual burn is approximately:
\begin{equation}
B_{\text{annual}} = 0.25 \times 0.003 \times 2{,}400{,}000 \times 365 /
p \approx \frac{657{,}000}{p} \text{ HANZO},
\end{equation}
where $p$ is the token price in USD. At $p = \$0.10$, this burns 6.57M tokens
per year (0.66\% of total supply). At $p = \$1.00$, the burn is 657K tokens
(0.066\%). The deflationary effect is strongest at low prices, providing a
natural floor.

% ============================================================================
\section{Token Governance}
\label{sec:tokengov}
% ============================================================================

The Hanzo DAO governs all token-economic parameters. Governance is implemented as
an on-chain contract on Hanzo Network (chain ID 36963) with the following
structure:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Proposal threshold}: 100,000 HANZO (0.01\% of supply) to
    submit a proposal.
    \item \textbf{Quorum}: 50,000,000 HANZO (5\% of supply) must participate for
    a vote to be valid.
    \item \textbf{Approval}: $>$66\% of voting power must approve.
    \item \textbf{Timelock}: 7-day delay between approval and execution.
    \item \textbf{Security council}: 5-of-9 multi-sig with veto power and
    emergency pause capability.
\end{itemize}

\paragraph{Governable parameters.}

\begin{table}[H]
\centering
\small
\begin{tabular}{llr}
\toprule
\textbf{Parameter} & \textbf{Range} & \textbf{Default} \\
\midrule
Market fee ($f_m$) & 10--100 bps & 30 bps \\
Risk fee coefficient & 1--50 bps & 5--20 bps \\
Burn rate ($\zeta$) & 10--50\% & 25\% \\
Treasury share ($\beta$) & 5--20\% & 10\% \\
Staking minimums & $\pm$50\% & See Table~\ref{tab:staking} \\
Emission halving period & 2--6 years & 4 years \\
Fee discount tiers & Adjustable & See Table~\ref{tab:fees} \\
\bottomrule
\end{tabular}
\caption{DAO-governable parameters with allowed ranges and defaults.}
\label{tab:governance}
\end{table}

% ============================================================================
\section{Zen Model Economics}
\label{sec:zen}
% ============================================================================

The Zen model family comprises 26 specialized models built on the Qwen3+
architecture. Each model has distinct compute costs and pricing:

\begin{table}[H]
\centering
\small
\begin{tabular}{llrr}
\toprule
\textbf{Model} & \textbf{Specialization} & \textbf{Params} &
\textbf{Cost/1K tokens} \\
\midrule
zen-coder & Code generation & 72B & 0.12 HANZO \\
zen-artist & Image generation & 14B & 0.45 HANZO \\
zen-scribe & Long-form writing & 32B & 0.08 HANZO \\
zen-designer & UI/UX generation & 14B & 0.35 HANZO \\
zen-analyst & Data analysis & 32B & 0.10 HANZO \\
zen-reason & Chain-of-thought & 72B & 0.15 HANZO \\
zen-chat & Conversational & 8B & 0.03 HANZO \\
zen-embed & Embedding & 1.5B & 0.01 HANZO \\
\bottomrule
\end{tabular}
\caption{Selected Zen models with per-1K-token pricing in HANZO. Full catalog of 26 models available at docs.hanzo.ai.}
\label{tab:zen}
\end{table}

\subsection{Model Provider Economics}

Third-party model providers can list models on the Hanzo Marketplace. The
economic structure for providers:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Listing fee}: One-time 1,000 HANZO to list a model (burned).
    \item \textbf{Revenue split}: Provider receives 70\%, platform receives 30\%.
    \item \textbf{Quality bond}: Providers stake 10,000 HANZO per model, slashable
    for SLA violations.
    \item \textbf{Governance approval}: New models require DAO vote for inclusion
    in the default catalog.
\end{itemize}

\subsection{Training-Free Adaptation}

Hanzo's Active Semantic Optimization (ASO) and Decentralized Semantic
Optimization (DSO) protocols~\cite{hanzoaso, hanzodso} enable zero-training model
adaptation via Product-of-Experts (PoE) decoding. Users pay for experiential
prior retrieval at a fraction of fine-tuning cost:

\begin{itemize}[leftmargin=1.1em]
    \item Prior retrieval: 0.001 HANZO per query (10$\times$ cheaper than
    inference).
    \item Prior submission: Free, but requires quality bond of 25 HANZO.
    \item Quality rewards: High-quality priors earn 0.01 HANZO per downstream
    usage.
\end{itemize}

% ============================================================================
\section{Defense \& Enterprise}
\label{sec:defense}
% ============================================================================

Hanzo's defense division publishes research and provides enterprise-grade AI
infrastructure for government and defense applications. This creates a distinct
revenue stream that feeds back into the token economy.

\subsection{Research Portfolio}

Hanzo has published 10 defense-focused papers spanning:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Agentic AI}: Multi-agent coordination for autonomous systems,
    federated agent networks, self-improving agent architectures.
    \item \textbf{Formal verification}: Harness-based testing, unified test
    frameworks, provably correct AI system deployment.
    \item \textbf{Quantum sensing}: Post-quantum cryptographic integration,
    quantum-resistant key management.
\end{itemize}

\subsection{Enterprise Token Utility}

Enterprise customers interact with the HANZO economy through:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Volume commitments}: Pre-purchasing HANZO at a discount for
    guaranteed compute capacity (1-year minimum term).
    \item \textbf{Dedicated staking}: Enterprise nodes with dedicated GPU
    allocation, requiring higher stake (500,000+ HANZO) but receiving
    priority scheduling.
    \item \textbf{Compliance features}: KMS-integrated key management, SOC 2 Type
    II audited infrastructure, ITAR-compliant compute isolation.
    \item \textbf{SLA guarantees}: On-chain SLA contracts with automatic
    compensation (in HANZO) for downtime or latency violations.
\end{itemize}

% ============================================================================
\section{Comparison to Other AI Tokens}
\label{sec:comparison}
% ============================================================================

\begin{table}[H]
\centering
\small
\begin{tabular}{lccccc}
\toprule
\textbf{Feature} & \textbf{HANZO} & \textbf{FET} & \textbf{RNDR} &
\textbf{TAO} & \textbf{AKT} \\
\midrule
Fixed supply & \checkmark & \checkmark & \checkmark & $\times$ & \checkmark \\
Fee burn & \checkmark & $\times$ & $\times$ & $\times$ & $\times$ \\
Compute payment & \checkmark & $\times$ & \checkmark & $\times$ & \checkmark \\
Node staking & \checkmark & \checkmark & $\times$ & \checkmark & \checkmark \\
Governance & \checkmark & \checkmark & $\times$ & \checkmark & \checkmark \\
Fee reduction & \checkmark & $\times$ & $\times$ & $\times$ & $\times$ \\
Revenue share & \checkmark & $\times$ & $\times$ & $\times$ & $\times$ \\
Own L1 chain & \checkmark & \checkmark & $\times$ & \checkmark & \checkmark \\
Model serving & \checkmark & $\times$ & $\times$ & \checkmark & $\times$ \\
Full PaaS & \checkmark & $\times$ & $\times$ & $\times$ & $\times$ \\
\bottomrule
\end{tabular}
\caption{Feature comparison of AI infrastructure tokens.}
\label{tab:comparison}
\end{table}

\paragraph{Fetch.ai (FET).} Agent-focused with staking and governance, but no
direct compute payment utility. The token primarily coordinates autonomous
agents rather than pricing infrastructure.

\paragraph{Render (RNDR).} GPU rendering marketplace with direct compute
payment, but no staking, governance, or fee reduction. The token is purely
transactional, subject to the velocity problem.

\paragraph{Bittensor (TAO).} Decentralized model training with staking and
governance, but inflationary emissions with no burn. TAO's perpetual inflation
dilutes holders without bound.

\paragraph{Akash (AKT).} General-purpose compute marketplace with staking and
governance. Closest to HANZO in design, but lacks fee burn, fee reduction,
revenue sharing, and integrated model serving.

\paragraph{HANZO differentiation.} HANZO is the only AI token that combines
all five utility dimensions (compute payment, staking, governance, fee
reduction, revenue sharing) with deflationary burn, an integrated PaaS, and
native model serving. The closest competitor (AKT) lacks three of these five
dimensions.

% ============================================================================
\section{Conclusion}
\label{sec:conclusion}
% ============================================================================

The HANZO token is designed to solve the fundamental tension in AI infrastructure
tokenomics: creating genuine utility demand while avoiding the velocity problem
that plagues payment-only tokens. By combining five utility dimensions---compute
payment, node staking, governance, fee reduction, and revenue sharing---with a
fixed supply and 25\% fee burn, HANZO creates multiple concurrent demand sinks
that reduce circulating supply as network usage grows.

The key properties of the model are:

\begin{enumerate}[leftmargin=1.4em]
    \item \textbf{Organic demand}: Every inference request, training job, and API
    call on Hanzo Platform requires HANZO, creating demand proportional to actual
    usage rather than speculation.
    \item \textbf{Supply reduction}: Staking, fee-reduction lock-ups, and
    permanent burns continuously reduce effective circulating supply.
    \item \textbf{Alignment}: Node operators, model providers, application
    builders, and end users all benefit from holding HANZO, aligning incentives
    across the ecosystem.
    \item \textbf{Stability}: The burn-and-stake equilibrium provides a natural
    price floor that rises with cumulative platform usage.
    \item \textbf{Governance}: DAO-controlled parameters ensure the economic model
    can adapt to changing market conditions without hard forks.
\end{enumerate}

Hanzo Network, backed by the Hanzo Platform's 14-month production track record
and the Lux Network's multi-consensus infrastructure, provides the foundation for
a sustainable AI compute economy where the HANZO token captures value
proportional to the real economic activity it enables.

% ============================================================================
% REFERENCES
% ============================================================================
\begin{thebibliography}{99}

\bibitem{hanzoplatform}
M.~Chen, D.~Wei, and Z.~Kelling.
\newblock Hanzo Platform: {PaaS} for {AI}-native application deployment.
\newblock \emph{Hanzo AI Research}, December 2021.

\bibitem{hanzoaci}
Z.~Kelling, D.~Wei, and M.~Chen.
\newblock Hanzo {ACI}: {AI} Chain Infrastructure for decentralized compute markets.
\newblock \emph{Hanzo AI Research}, December 2024.

\bibitem{hanzonetwork}
Z.~Kelling.
\newblock Hanzo Network: A {Hamiltonian} Market Maker {Layer-1} for decentralized {AI} compute and semantic learning.
\newblock \emph{Hanzo AI Research}, October 2025.

\bibitem{hanzodso}
{Hanzo AI Research}.
\newblock Decentralized Semantic Optimization: {Byzantine}-robust prior aggregation for collective model adaptation.
\newblock \emph{Hanzo AI Research}, December 2023.

\bibitem{hanzoaso}
{Hanzo AI Research}.
\newblock Active Semantic Optimization: Training-free adaptation via {Bayesian} Product-of-Experts decoding.
\newblock \emph{Hanzo AI Research}, September 2025.

\bibitem{hanzollmgateway}
{Hanzo AI Research}.
\newblock Hanzo {LLM} Gateway: Unified model serving infrastructure.
\newblock \emph{Hanzo AI Research}, 2024.

\bibitem{luxwhitepaper}
{Lux Partners}.
\newblock Lux: A scalable multi-consensus blockchain with post-quantum cryptography.
\newblock \emph{Lux Network Technical Report}, 2024.

\bibitem{mckinseyai2024}
{McKinsey Global Institute}.
\newblock The economic potential of generative {AI}: The next productivity frontier.
\newblock \emph{McKinsey Report}, June 2023.

\bibitem{buterin2014}
V.~Buterin.
\newblock A next-generation smart contract and decentralized application platform.
\newblock \emph{Ethereum White Paper}, 2014.

\bibitem{nakamoto2008}
S.~Nakamoto.
\newblock Bitcoin: A peer-to-peer electronic cash system.
\newblock 2008.

\bibitem{samani2017}
K.~Samani.
\newblock Understanding token velocity.
\newblock \emph{Multicoin Capital}, September 2017.

\bibitem{vitalikburn}
V.~Buterin, E.~Conner, R.~Dudley, M.~Slipper, and I.~Norden.
\newblock {EIP-1559}: Fee market change for {ETH 1.0} chain.
\newblock \emph{Ethereum Improvement Proposals}, April 2019.

\bibitem{fetchai2023}
{Fetch.ai}.
\newblock Fetch.ai: Autonomous economic agents.
\newblock \emph{Fetch.ai Documentation}, 2023.

\bibitem{rendernetwork}
{Render Network}.
\newblock The Render Network: Distributed {GPU} rendering.
\newblock \emph{Render Network White Paper}, 2022.

\bibitem{bittensor2023}
{Opentensor Foundation}.
\newblock Bittensor: A peer-to-peer intelligence market.
\newblock \emph{Bittensor White Paper}, 2023.

\bibitem{akash2020}
G.~Akash.
\newblock Akash Network: The decentralized cloud.
\newblock \emph{Akash Network White Paper}, 2020.

\end{thebibliography}

\end{document}
