\section{Investor Onboarding via Hanzo IAM}

\label{sec:onboarding}
% ============================================================================

Investor onboarding combines Hanzo IAM identity management with Simplici
KYC/AML verification.

\subsection{Identity Creation}

\begin{enumerate}[leftmargin=1.4em]
    \item Investor registers via ArcaLabs platform. Registration request
    routed through Hanzo Gateway.
    \item Hanzo IAM creates an identity record with OIDC subject
    identifier. The investor receives OAuth 2.0 credentials.
    \item IAM assigns the investor to the ArcaLabs organization with
    the \texttt{investor} role.
    \item All subsequent API calls include the IAM bearer token. Gateway
    extracts the org from the JWT \texttt{owner} claim and scopes all
    data queries to that org.
\end{enumerate}

\subsection{KYC/AML via Simplici}

\begin{enumerate}[leftmargin=1.4em]
    \item ArcaLabs onboarding service redirects the investor to
    Simplici's verification flow.
    \item Simplici performs: identity document verification, liveness
    detection, OFAC/sanctions screening, PEP screening, and accredited
    investor verification.
    \item Upon completion, Simplici returns a signed attestation to
    ArcaLabs via webhook.
    \item ArcaLabs compliance service validates the attestation and
    updates the investor's IAM profile with: KYC status, accreditation
    status, accreditation expiry, jurisdiction.
    \item The compliance service writes the investor to the on-chain
    \texttt{ComplianceRegistry} whitelist on Lux C-Chain.
\end{enumerate}

\subsection{Ongoing Monitoring}

Simplici performs continuous monitoring:
\begin{itemize}[leftmargin=1.4em]
    \item Daily sanctions list re-screening.
    \item Adverse media monitoring.
    \item Annual accreditation re-verification trigger.
    \item Alerts delivered to ArcaLabs via Hanzo Webhooks.
\end{itemize}

% ============================================================================
