\section{Trade Execution via the regulated ATS}

\label{sec:trade}
% ============================================================================

Secondary trading of digital securities occurs through the regulated platform
ATS (Alternative Trading System), a SEC-registered trading venue.

\subsection{Order Flow}

\begin{enumerate}[leftmargin=1.4em]
    \item Investor submits order via ArcaLabs trading interface.
    \item Request authenticated by Hanzo IAM, routed through Hanzo
    Gateway to the ArcaLabs trade engine.
    \item Trade engine validates: investor whitelist status, lockup
    compliance, accreditation validity.
    \item Approved order forwarded to the regulated ATS matching engine.
    \item ATS matches buy and sell orders. Matched trades return to
    ArcaLabs for settlement.
    \item Settlement: atomic swap of security tokens and USDC on Lux
    C-Chain via MPC-signed transaction.
\end{enumerate}

\subsection{Gateway Configuration}

Hanzo Gateway handles the routing between ArcaLabs services and the
the regulated ATS:

\begin{itemize}[leftmargin=1.4em]
    \item \textbf{Rate limiting}: Per-investor and per-org request limits
    to prevent abuse.
    \item \textbf{Authentication}: IAM token validation at the gateway
    layer. Invalid tokens rejected before reaching backend services.
    \item \textbf{TLS}: End-to-end encryption. Mutual TLS between
    Gateway and ATS.
    \item \textbf{Audit logging}: Every request logged with investor
    identity, timestamp, and response code.
\end{itemize}

% ============================================================================
