\section{Reporting Pipeline}

\label{sec:reporting}
% ============================================================================

ArcaLabs' reporting requirements span regulatory filings, investor
communications, and internal audit:

\subsection{Regulatory Reports}

\begin{itemize}[leftmargin=1.4em]
    \item \textbf{Form D filings}: Generated from issuance records in
    Base. Auto-populated with investor counts, amounts raised,
    exemption type.
    \item \textbf{Blue sky filings}: State-level notice filings based
    on investor jurisdiction data.
    \item \textbf{ATS Form ATS-N}: Quarterly filing for the regulated platform
    ATS, generated from trade records.
    \item \textbf{FINRA trade reports}: Real-time trade reporting
    generated from settlement events.
\end{itemize}

\subsection{Investor Communications}

\begin{itemize}[leftmargin=1.4em]
    \item \textbf{Account statements}: Monthly statements generated from
    Base records, delivered via Hanzo Webhooks to investor email.
    \item \textbf{Tax documents}: K-1 (fund interests) and 1099
    (interest/dividends) generated annually.
    \item \textbf{NAV reports}: Daily NAV publications for fund products.
\end{itemize}

\subsection{Event Pipeline}

\begin{enumerate}[leftmargin=1.4em]
    \item On-chain events (transfers, compliance decisions, corporate
    actions) are indexed by an event listener.
    \item Events written to Base with chain metadata (block number,
    tx hash, timestamp).
    \item Hanzo Webhooks delivers events to configured endpoints
    (reporting systems, compliance dashboards, investor portals).
    \item Events are immutable in Base---no deletion, only append.
\end{enumerate}

% ============================================================================
