\section{Conclusion}

\label{sec:conclusion}
% ============================================================================

The Hanzo platform provides the foundational infrastructure for
ArcaLabs' digital securities operations. IAM handles investor identity
and access control with org-scoped permissions. KMS secures MPC key
material with envelope encryption and comprehensive audit logging. Base
provides the data persistence layer with production-grade PostgreSQL and
development-friendly SQLite. Gateway routes and rate-limits all traffic
with IAM token validation. Webhooks delivers lifecycle events to
reporting and compliance systems.

This integration demonstrates that regulated financial applications can
be built by composing existing platform services rather than
reimplementing infrastructure. The Hanzo platform's service boundaries
align with the separation of concerns required by securities
regulation: identity (IAM), custody (KMS + MPC), data (Base), and
communication (Gateway + Webhooks).

\begin{thebibliography}{99}
\bibitem{hanzoplatform}
Hanzo Industries, ``Hanzo Platform: AI-Powered Infrastructure,''
Product Documentation, 2024.

\bibitem{cggmp21}
R.~Canetti, R.~Gennaro, S.~Goldfeder, N.~Makriyannis, U.~Peled,
``UC Non-Interactive, Proactive, Threshold ECDSA with Identifiable
Aborts,'' \textit{ACM CCS}, 2020.

\bibitem{luxwhitepaper}
Lux Partners, ``Lux Network: A Scalable Multi-Chain Architecture,''
Technical Report, 2022.

\bibitem{simplici2024}
Simplici Inc., ``Broker-Dealer Integrated KYC/AML Platform,''
Product Documentation, 2024.

\bibitem{sec17a4}
SEC, ``Records to Be Preserved by Certain Exchange Members, Brokers and
Dealers,'' 17 CFR \S~240.17a-4.

\bibitem{oidc}
N.~Sakimura et al., ``OpenID Connect Core 1.0,'' OpenID Foundation,
2014.

\bibitem{oauth2}
D.~Hardt, ``The OAuth 2.0 Authorization Framework,'' RFC 6749, 2012.
\end{thebibliography}

