\section{Introduction}


The convergence of artificial intelligence and decentralized systems presents unprecedented opportunities for democratizing compute access while maintaining security and privacy guarantees. However, existing platforms face critical challenges: (1) centralized control over AI inference resources, (2) lack of quantum-resistant security in production systems, (3) fragmented runtime environments requiring manual orchestration, and (4) absence of fair resource pricing mechanisms.

Hanzo Network addresses these challenges through a novel integration of blockchain consensus, confidential computing, and adaptive AI routing. The system was first deployed in September 2025 with over 4,500 commits representing continuous development and refinement.

\subsection{Key Contributions}

\begin{enumerate}
\item \textbf{HLLM Routing Framework}: Physics-inspired model selection combining Hidden Markov Models for regime detection with Hamiltonian mechanics for price dynamics and BitDelta quantization for efficient adaptation.

\item \textbf{Production PQC Implementation}: NIST-compliant ML-KEM (FIPS 203) and ML-DSA (FIPS 204) with hybrid classical/quantum modes, supporting GPU Confidential Computing on H100/Blackwell architectures.

\item \textbf{Unified Runtime Abstraction}: Protocol-agnostic interface for Docker, Kubernetes, WASM, and Firecracker with blockchain-integrated compute metering and automatic billing.

\item \textbf{Lux Consensus Integration}: High-throughput consensus (9M+ blocks/second) with sled database backend and native compute scheduling daemon.

\item \textbf{Native Qwen3 Support}: First-class integration of Qwen3-Embedding-8B (4096-dim) and Qwen3-Reranker-4B models with GPU-accelerated inference via mistral.rs.
\end{enumerate}

