\section{Discussion}


\subsection{Design Trade-offs}

\subsubsection{PQC Key Sizes}

ML-KEM/ML-DSA keys are $5-10\times$ larger than classical counterparts, increasing storage and bandwidth. However, this is acceptable given:
\begin{itemize}
\item Keys are generated infrequently
\item Storage is cheap ($<$\$0.01/GB/month)
\item Quantum threat timeline (5-15 years) justifies overhead
\end{itemize}

\subsubsection{Consensus Throughput vs. Decentralization}

Local throughput (9M blocks/sec) drops to 3,500 in geographically distributed setting. This is expected due to:
\begin{itemize}
\item Network latency dominates consensus time
\item Wave consensus requires $k=20$ samples per decision
\item Trade-off favors safety over speed
\end{itemize}

\subsubsection{HLLM Complexity}

Physics-inspired routing adds 2.8ms overhead vs. simple random selection. Benefits justify cost:
\begin{itemize}
\item 15-30\% cost reduction via optimal model selection
\item Improved user experience through regime-aware routing
\item Adaptive learning without retraining models
\end{itemize}

\subsection{Future Work}

\subsubsection{SLH-DSA Integration}

SPHINCS+ (SLH-DSA) provides hash-based signatures without lattice assumptions. Future versions will support hybrid ML-DSA/SLH-DSA for algorithm agility.

\subsubsection{Hardware Acceleration}

FPGA/ASIC implementations of ML-KEM could reduce latency to $<$10μs, enabling PQC in ultra-low-latency applications.

\subsubsection{Federated Learning}

Extending HLLM to support decentralized model training across nodes while preserving privacy via secure aggregation.

\subsubsection{Zero-Knowledge Proofs}

zk-SNARKs for verifiable compute execution, enabling trustless verification of inference results without revealing model weights.

\subsection{Limitations}

\begin{enumerate}
\item \textbf{Quantum timeline uncertainty}: PQC overhead may be premature if quantum threat is $>$20 years away.

\item \textbf{HLLM tuning}: Regime detection thresholds require per-deployment tuning; no universal defaults.

\item \textbf{Cross-runtime compatibility}: Not all workloads portable across Docker/K8s/WASM without modification.

\item \textbf{Single-chain scaling}: Current architecture limited to single consensus chain; sharding not yet implemented.
\end{enumerate}

