\section*{Acknowledgments}


We thank the Open Quantum Safe project for liboqs implementations, the Lux Network team for consensus engine foundations, and the Qwen team for state-of-the-art embedding models. Special thanks to early adopters providing feedback during beta testing.

\bibliographystyle{plain}
\begin{thebibliography}{99}

\bibitem{bitdelta}
Y. Zhou et al., ``BitDelta: Efficient Training-Free Acceleration of Large Language Models,''
\textit{arXiv preprint arXiv:2405.04144}, 2024.

\bibitem{liboqs}
D. Stebila and M. Mosca, ``Post-quantum Key Exchange for the Internet and the Open Quantum Safe Project,''
\textit{International Conference on Selected Areas in Cryptography}, pp. 1-24, 2016.

\bibitem{fips203}
National Institute of Standards and Technology,
``FIPS 203: Module-Lattice-Based Key-Encapsulation Mechanism Standard,''
\url{https://csrc.nist.gov/pubs/fips/203/final}, 2024.

\bibitem{fips204}
National Institute of Standards and Technology,
``FIPS 204: Module-Lattice-Based Digital Signature Standard,''
\url{https://csrc.nist.gov/pubs/fips/204/final}, 2024.

\bibitem{sp80056c}
National Institute of Standards and Technology,
``SP 800-56C Rev. 2: Recommendation for Key-Derivation Methods in Key-Establishment Schemes,''
\url{https://csrc.nist.gov/pubs/sp/800/56/c/r2/final}, 2020.

\bibitem{friston}
K. Friston, ``The Free-Energy Principle: A Unified Brain Theory?''
\textit{Nature Reviews Neuroscience}, vol. 11, no. 2, pp. 127-138, 2010.

\bibitem{snow}
T. Rocket et al., ``Scalable and Probabilistic Leaderless BFT Consensus through Metastability,''
\textit{arXiv preprint arXiv:1906.08936}, 2019.

\bibitem{golem}
Golem Network, ``Golem: Decentralized Computation Platform,''
\url{https://golem.network/}, 2020.

\bibitem{render}
Render Network, ``The Decentralized GPU Rendering Network,''
\url{https://rendernetwork.com/}, 2023.

\bibitem{akash}
Akash Network, ``The Supercloud for AI,''
\url{https://akash.network/}, 2024.

\bibitem{awspqtls}
Amazon Web Services, ``Introducing Post-Quantum TLS,''
AWS Security Blog, 2023.

\bibitem{anyscale}
Anyscale, ``Ray: A Distributed Framework for Emerging AI Applications,''
\url{https://www.anyscale.com/}, 2023.

\bibitem{baseten}
Baseten, ``Fast Model Inference at Any Scale,''
\url{https://www.baseten.co/}, 2024.

\bibitem{azurecc}
Microsoft Azure, ``Azure Confidential Computing,''
\url{https://azure.microsoft.com/en-us/solutions/confidential-compute/}, 2024.

\bibitem{h100cc}
NVIDIA, ``H100 Confidential Computing: Secure AI for the Data Center,''
NVIDIA Technical White Paper, 2024.

\bibitem{qwen3}
Alibaba Cloud, ``Qwen3: Large Language and Multimodal Models,''
\url{https://qwen.io/}, 2025.

\bibitem{luxconsensus}
Lux Partners Inc., ``Lux Consensus: A Multi-Consensus Blockchain Platform,''
Technical Documentation, 2024.

\bibitem{sled}
Spacejam, ``Sled: Modern Embedded Database,''
\url{https://github.com/spacejam/sled}, 2023.

\bibitem{mistrals}
Mistral AI, ``mistral.rs: High-Performance Inference Engine,''
\url{https://github.com/EricLBuehler/mistral.rs}, 2024.

\bibitem{hamiltonianhmm}
S. Särkkä and Á. F. García-Fernández, ``Temporal Parallelization of Bayesian Smoothers,''
\textit{IEEE Transactions on Automatic Control}, vol. 66, no. 1, pp. 299-306, 2020.

\end{thebibliography}

