\section{Evaluation}


\subsection{Experimental Setup}

We evaluated Verus on production workloads from three merchants:

\begin{itemize}
\item \textbf{Small}: 10K events/day, 5 materialized views.
\item \textbf{Medium}: 1M events/day, 15 materialized views.
\item \textbf{Large}: 100M events/day, 30 materialized views.
\end{itemize}

Infrastructure:
\begin{itemize}
\item Stream processing: 3-node Flink cluster.
\item State store: 3-node Redis cluster.
\item Columnar storage: S3 + 10-node Presto cluster.
\end{itemize}

\subsection{Ingestion Throughput}

\begin{table}[h]
\centering
\caption{Sustained ingestion throughput}
\label{tab:throughput}
\begin{tabular}{lr}
\hline
Configuration & Events/second \\
\hline
Single partition & 8,500 \\
10 partitions & 52,000 \\
100 partitions & 485,000 \\
\hline
\end{tabular}
\end{table}

Throughput scales linearly with partitions. The large merchant's peak load (50K events/second) is well within capacity.

\subsection{Query Latency}

\begin{table}[h]
\centering
\caption{Query latency by type (ms)}
\label{tab:query_latency}
\begin{tabular}{lrrr}
\hline
Query Type & p50 & p95 & p99 \\
\hline
Dashboard (Redis) & 18 & 45 & 82 \\
Dashboard (merged) & 125 & 280 & 520 \\
Ad-hoc (simple) & 850 & 2,100 & 4,500 \\
Ad-hoc (complex) & 3,200 & 8,500 & 15,000 \\
Cohort analysis & 5,500 & 12,000 & 25,000 \\
\hline
\end{tabular}
\end{table}

Dashboard queries served from Redis meet interactive requirements (< 100ms p99). Ad-hoc queries trade latency for flexibility.

\subsection{Freshness}

We measured end-to-end latency from event occurrence to query visibility:

\begin{table}[h]
\centering
\caption{Data freshness (seconds)}
\label{tab:freshness}
\begin{tabular}{lrrr}
\hline
Metric & p50 & p95 & p99 \\
\hline
Materialized views & 1.2 & 2.8 & 5.1 \\
Columnar storage & 45 & 120 & 180 \\
\hline
\end{tabular}
\end{table}

Materialized views update within seconds. Columnar storage has higher latency due to micro-batching for write efficiency.

\subsection{Cost Analysis}

\begin{table}[h]
\centering
\caption{Monthly infrastructure cost (USD)}
\label{tab:cost}
\begin{tabular}{lrrr}
\hline
Component & Small & Medium & Large \\
\hline
Kafka & 50 & 150 & 800 \\
Stream processing & 100 & 400 & 2,500 \\
Redis & 75 & 250 & 1,200 \\
Storage (S3) & 5 & 50 & 500 \\
Query engine & 200 & 600 & 3,000 \\
\hline
Total & 430 & 1,450 & 8,000 \\
\hline
\end{tabular}
\end{table}

Cost scales sub-linearly with volume due to shared infrastructure and compression benefits at scale.

\subsection{Comparison with Alternatives}

We compared Verus against pure-batch (BigQuery) and pure-stream (custom Flink) approaches:

\begin{table}[h]
\centering
\caption{Approach comparison}
\label{tab:comparison}
\begin{tabular}{lccc}
\hline
Metric & BigQuery & Flink-only & Verus \\
\hline
Dashboard latency & 2--5s & 20ms & 20ms \\
Ad-hoc latency & 2--10s & N/A & 1--15s \\
Freshness & 1--4 hours & < 1s & < 5s \\
Query flexibility & High & Low & High \\
Operational complexity & Low & High & Medium \\
\hline
\end{tabular}
\end{table}

Verus achieves the freshness of stream processing with the flexibility of batch analytics.

