\section{Conclusion}


Verus demonstrates that commerce analytics can achieve sub-second freshness without sacrificing query flexibility. The hybrid architecture—stream processing for materialized views, columnar storage for ad-hoc queries, intelligent routing between them—provides the best of both approaches.

Production deployments show the system handles 50,000+ events per second while serving dashboard queries in under 100ms. Merchants gain real-time visibility into their business without the operational complexity of maintaining separate batch and stream systems.

Future work will extend Verus with machine learning integration (anomaly detection models, forecasting) and support for streaming SQL standards (KSQL, Flink SQL).

\begin{thebibliography}{9}

\bibitem{marz2015}
N. Marz and J. Warren, \textit{Big Data: Principles and Best Practices of Scalable Real-Time Data Systems}, Manning, 2015.

\bibitem{druid2014}
F. Yang et al., ``Druid: A Real-time Analytical Data Store,'' \textit{SIGMOD}, 2014.

\bibitem{clickhouse2016}
A. Milovidov, ``ClickHouse: New Open Source Columnar Database,'' \textit{HighLoad++}, 2016.

\bibitem{kafka2011}
J. Kreps, N. Narkhede, and J. Rao, ``Kafka: A Distributed Messaging System for Log Processing,'' \textit{NetDB Workshop}, 2011.

\bibitem{flink2015}
P. Carbone et al., ``Apache Flink: Stream and Batch Processing in a Single Engine,'' \textit{IEEE Data Engineering Bulletin}, vol. 38, no. 4, 2015.

\bibitem{parquet2013}
J. LeFevre et al., ``Parquet,'' \textit{Apache Software Foundation}, 2013.

\end{thebibliography}

