\section{Loop 0: Telemetry Pipeline}

\label{sec:loop0}

\subsection{Formal Definition}

The telemetry pipeline is the foundation on which all other loops depend. Its
design must satisfy three requirements: (1) zero latency overhead on the
critical path, (2) append-only storage for tamper-evidence and crash safety,
and (3) sufficient structural richness to support all downstream analyses.

\begin{definition}[ToolInvocation Record]
\label{def:tool_invocation}
A \textsc{ToolInvocation} record $\tau$ is a tuple:
\begin{equation}
\tau = \bigl( \text{id},\; \text{session\_id},\; \text{ts},\; \text{tool},\;
              \text{args},\; \text{outcome},\; \text{duration\_ms},\; \text{failure\_mode} \bigr),
\end{equation}
where:
\begin{itemize}[leftmargin=1.5em, nosep]
    \item $\text{id} \in \{0,1\}^{256}$: Globally unique record identifier (SHA-256 of content).
    \item $\text{session\_id} \in \mathbb{N}$: Monotonically increasing session counter.
    \item $\text{ts} \in \mathbb{R}_{>0}$: Unix timestamp with millisecond precision.
    \item $\text{tool} \in \mathcal{V}_T$: Tool identifier from the tool vocabulary.
    \item $\text{args} \in \mathcal{A}$: Serialized argument record (JSON, size-bounded).
    \item $\text{outcome} \in \{\textsc{success}, \textsc{failure}, \textsc{timeout}, \textsc{cancelled}\}$: Categorical outcome.
    \item $\text{duration\_ms} \in \mathbb{N}$: Wall-clock execution time.
    \item $\text{failure\_mode} \in \mathcal{V}_F \cup \{\bot\}$: Categorical failure mode or null on success.
\end{itemize}
\end{definition}

\begin{remark}
The content-addressed identifier $\text{id}$ enables deduplication and serves
as an integrity check: any modification to a record is detectable by recomputing
the hash. This property is important for Loop~4's cross-session aggregation,
which must be robust to storage corruption.
\end{remark}

\subsection{Storage Architecture}

Records are appended to a JSONL (JSON Lines) file, one record per line. JSONL
is preferred over binary formats for three reasons:

\begin{enumerate}[leftmargin=1.5em]
    \item \textbf{Crash safety:} A partial write affects only the last line.
          All preceding lines remain valid and parseable.
    \item \textbf{Tier portability:} The same file format works identically
          on a local laptop, a remote container, and an EC2 instance. NATS
          JetStream can replay from JSONL without schema migration.
    \item \textbf{Zero write amplification:} Append-only writes never
          trigger compaction, B-tree rebalancing, or WAL flushing. The
          telemetry path adds at most one \texttt{write(2)} syscall per
          tool invocation.
\end{enumerate}

Records are simultaneously published to a NATS JetStream subject
\texttt{telemetry.{session\_id}.{tool}} for real-time consumption by Loops~1
and~2. The NATS layer provides durability (disk-backed JetStream), replay
capability (for recovering Loop~1 analysis after agent restart), and
subject-based filtering (Loop~2 subscribes only to its trigger subjects).

\begin{definition}[Telemetry Log]
\label{def:telemetry_log}
The telemetry log $\mathcal{T}^{(s)}$ at session $s$ is the ordered sequence:
\begin{equation}
\mathcal{T}^{(s)} = \bigl( \tau_1, \tau_2, \ldots, \tau_{|\mathcal{T}^{(s)}|} \bigr),
\quad \tau_i.\text{ts} \leq \tau_{i+1}.\text{ts},
\end{equation}
stored as an append-only JSONL file. The \emph{session slice}
$\mathcal{T}^{(s)}_{\text{cur}}$ is the subsequence of records with
$\text{session\_id} = s$.
\end{definition}

\subsection{Failure Mode Taxonomy}

The failure mode vocabulary $\mathcal{V}_F$ is a hierarchical taxonomy with
three levels:

\begin{table}[H]
\centering
\caption{Failure Mode Taxonomy (Level-1 categories with examples).}
\label{tab:failure_modes}
\begin{tabular}{llp{6cm}}
\toprule
Code & Category & Example Instances \\
\midrule
\texttt{PERM} & Permission error & File not writable, socket permission denied \\
\texttt{NOTFOUND} & Resource not found & File missing, URL 404, package not installed \\
\texttt{TIMEOUT} & Execution timeout & Command exceeded wall-clock limit \\
\texttt{SYNTAX} & Syntax/parse error & Malformed JSON argument, invalid regex \\
\texttt{RUNTIME} & Runtime exception & Uncaught exception, segfault, OOM \\
\texttt{TOOL\_GAP} & Missing capability & No tool for requested operation \\
\texttt{ARGS} & Argument error & Wrong type, out of range, missing required field \\
\texttt{CONFLICT} & State conflict & Git merge conflict, lock contention \\
\texttt{NETWORK} & Network failure & DNS failure, connection refused, TLS error \\
\bottomrule
\end{tabular}
\end{table}

Level-2 and Level-3 sub-codes refine each category. The full taxonomy is
defined in the SDK and is extensible: new failure modes can be registered by
tool authors without modifying the core taxonomy.

% ============================================================================
% 5. LOOP 1: BUILD IT NOW
% ============================================================================
