\section{Agent Hosting}

\label{sec:agents}
% ============================================================================

Hanzo AI Cloud hosts autonomous agents---persistent AI processes with memory,
tools, and goals. Each agent runs inside a \textbf{capsule}: an isolated
execution environment with its own data, identity, and resource budget.

\subsection{Capsule Architecture}

\begin{definition}[Agent Capsule]
A capsule is a tuple $(\text{vault}, \text{did}, \text{caps}, \text{budget},
\text{mem})$ where:
\begin{itemize}[leftmargin=1.1em]
    \item $\text{vault}$: A private Base instance (encrypted SQLite) containing
    the agent's persistent state.
    \item $\text{did}$: A decentralized identifier on the Lux I-Chain,
    uniquely identifying the agent.
    \item $\text{caps}$: A capability set defining what the agent can access
    (collections, APIs, other agents, network).
    \item $\text{budget}$: A HANZO token budget limiting the agent's compute
    and API spending.
    \item $\text{mem}$: Encrypted working memory (conversation history,
    scratchpad, retrieved context) stored in the vault.
\end{itemize}
\end{definition}

\subsection{Isolation Guarantees}

Capsule isolation is enforced at three levels:

\begin{enumerate}[leftmargin=1.4em]
    \item \textbf{Storage isolation}: Each capsule's vault is a separate
    encrypted SQLite database with a unique encryption key. The platform
    cannot read the vault contents without the agent's key.
    \item \textbf{Compute isolation}: Each capsule runs in a separate Goja VM
    instance (for tool execution) and a separate inference context (for model
    calls). No shared state between capsules.
    \item \textbf{Network isolation}: Capsule network access is restricted to
    the capabilities declared in $\text{caps}$. An agent without the
    \texttt{network} capability cannot make outbound requests.
\end{enumerate}

\subsection{Agent Memory}

Agent memory is structured as three tiers:

\begin{table}[H]
\centering
\small
\begin{tabular}{llp{5cm}}
\toprule
\textbf{Tier} & \textbf{Storage} & \textbf{Behavior} \\
\midrule
Working & In-process & Current conversation context. Lost on capsule
restart. \\
Episodic & Vault (encrypted) & Past interactions, indexed by time and topic.
Persists across restarts. Searchable via vector similarity. \\
Semantic & Vault (encrypted) & Distilled knowledge extracted from episodic
memory. Compact, structured, versioned. \\
\bottomrule
\end{tabular}
\caption{Agent memory tiers. All persistent memory is encrypted in the
capsule vault.}
\label{tab:memory}
\end{table}

Episodic-to-semantic distillation runs as a scheduled Base Function inside the
capsule. The agent periodically reviews its episodic memory and extracts
durable facts, preferences, and learned patterns into the semantic tier. This
is analogous to sleep consolidation in biological memory systems.

\subsection{Budget Enforcement}

Each capsule has a HANZO token budget that limits:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Inference tokens}: Maximum tokens generated per period.
    \item \textbf{API calls}: Maximum outbound API calls per period.
    \item \textbf{Storage}: Maximum vault size in bytes.
    \item \textbf{Compute}: Maximum CPU-seconds for function execution.
\end{itemize}

Budget is enforced at the platform level. When a capsule exhausts its budget,
it is suspended (not terminated). The capsule's vault and memory persist.
The owner can replenish the budget to resume the agent.

% ============================================================================
