\section{Edge Inference}

\label{sec:edge}
% ============================================================================

Not all inference should happen in the cloud. For latency-sensitive
applications (real-time object detection, voice assistants, on-device
classification), Hanzo AI Cloud supports edge inference via Hanzo
Edge~\cite{hanzoedge}.

\subsection{On-Device Runtime}

Hanzo Edge is a Candle-based inference runtime for mobile and IoT devices. It
supports:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Apple Neural Engine}: CoreML integration for ANE-accelerated
    inference on iPhone/iPad/Mac.
    \item \textbf{GPU compute}: Metal (Apple), Vulkan (Android/Linux) for
    GPU-accelerated inference.
    \item \textbf{CPU fallback}: NEON (ARM) and AVX-512 (x86) SIMD-optimized
    CPU inference for devices without GPU access.
    \item \textbf{Dynamic quantization}: INT8/INT4 quantization applied at
    model load time, adapting to available memory.
\end{itemize}

\subsection{CRDT Result Sync}

Edge inference results sync to the user's Base instance via the CRDT protocol
described in the Base runtime paper~\cite{hanzobaseruntime}. A mobile device
runs inference locally, writes the result to its local Base vault, and the
result syncs to the user's other devices and cloud Base instance on next
connection. This enables offline inference with eventual consistency.

\subsection{Model Distribution}

Edge models are distributed from the Hanzo model registry via a CDN-backed
download with integrity verification:

\begin{equation}
\text{verify}(\text{model}_{\text{local}}) :
\text{SHA3-256}(\text{model}_{\text{local}}) \stackrel{?}{=}
\text{model\_id}_{\text{A-Chain}}
\end{equation}

The device downloads the model once, verifies its hash against the A-Chain
registry entry, and caches it locally. Model updates are delta-compressed
(binary diff of quantized weights) to minimize bandwidth.

% ============================================================================
