\section{Comparison to Existing Platforms}

\label{sec:comparison}
% ============================================================================

\begin{table}[H]
\centering
\small
\begin{tabular}{lccccc}
\toprule
\textbf{Feature} & \textbf{Hanzo} & \textbf{OpenAI} & \textbf{Anthropic} &
\textbf{Replicate} & \textbf{Together} \\
\midrule
Confidential inference & \checkmark & $\times$ & $\times$ & $\times$ & $\times$ \\
FHE support & \checkmark & $\times$ & $\times$ & $\times$ & $\times$ \\
Agent capsules & \checkmark & $\times$ & $\times$ & $\times$ & $\times$ \\
Verifiable receipts & \checkmark & $\times$ & $\times$ & $\times$ & $\times$ \\
Multi-provider & \checkmark & $\times$ & $\times$ & \checkmark & \checkmark \\
Open model registry & \checkmark & $\times$ & $\times$ & \checkmark & \checkmark \\
Edge inference & \checkmark & $\times$ & $\times$ & $\times$ & $\times$ \\
On-chain SLA & \checkmark & $\times$ & $\times$ & $\times$ & $\times$ \\
Self-hostable & \checkmark & $\times$ & $\times$ & $\times$ & $\times$ \\
User-controlled keys & \checkmark & $\times$ & $\times$ & $\times$ & $\times$ \\
\bottomrule
\end{tabular}
\caption{Platform comparison. Hanzo AI Cloud is the only platform offering
confidential inference with verifiable attestation.}
\label{tab:comparison}
\end{table}

\paragraph{OpenAI.} Market leader in model quality. No confidential inference.
No verifiable outputs. No self-hosting. Data retention policy permits 30-day
logging. Closed-source models with no weight access.

\paragraph{Anthropic.} Strong safety focus. No confidential inference. No
agent isolation beyond conversation-level context windows. No self-hosting.
Constitutional AI provides behavioral guarantees, not cryptographic ones.

\paragraph{Replicate.} Open model marketplace. No confidential inference.
Shared GPU instances with no tenant isolation. No verifiable outputs. Good
developer experience for model deployment, but no sovereignty guarantees.

\paragraph{Together.} Open model serving with competitive pricing. No
confidential inference. No agent hosting. No verifiable outputs. Together
is a cheaper alternative to OpenAI for open models, not a fundamentally
different architecture.

\paragraph{Hanzo differentiation.} Hanzo AI Cloud is the only platform where
the provider mathematically cannot see the user's data (FHE), the agent's
memory is cryptographically private (capsule vault), and every inference result
is provably attributable (A-Chain receipts). This is not a feature advantage.
It is an architectural category difference.

% ============================================================================
