\section{Byzantine-Robust Aggregation}

\label{sec:aggregation}

\subsection{Threat Model}

\begin{assumption}[Threat Model]
\label{ass:threat}
Of $n$ total participants, at most $f < n/3$ are Byzantine. Byzantine nodes may:
\begin{enumerate}[leftmargin=1.5em]
    \item Submit arbitrary prior data (random noise, targeted attacks).
    \item Collude to coordinate attacks.
    \item Create Sybil identities (bounded by stake requirements).
    \item Selectively withhold or delay messages.
\end{enumerate}
Honest nodes follow the protocol and have priors with quality $q_i > 0.5 + \delta$
for some $\delta > 0$.
\end{assumption}

\subsection{Weighted Median Aggregation}

For each schema element $s \in \mathcal{S}$ and each token position $v$ within
$s$, the aggregate is computed as:

\begin{equation}
    \tilde{\Delta}^{(s)}(v) = \wmed\left(\{\Delta_i^{(s)}(v)\}_{i=1}^n; \; \{w_i\}_{i=1}^n\right),
    \label{eq:wmed}
\end{equation}
where the weight for participant $i$ is:
\begin{equation}
    w_i = S_i \cdot q_i \cdot \mathbb{I}[q_i \geq q_{\min}],
    \label{eq:weights}
\end{equation}
with $S_i$ the stake and $q_i$ the validated quality score.

\begin{theorem}[Byzantine Robustness of Weighted Median]
\label{thm:byz-robust}
Under Assumption~\ref{ass:threat}, if the total honest stake exceeds twice the
total Byzantine stake:
\begin{equation}
    W_H = \sum_{i \in \text{honest}} w_i > 2 \sum_{j \in \text{Byzantine}} w_j = 2W_B,
\end{equation}
then the weighted median (Eq.~\ref{eq:wmed}) equals a value submitted by an
honest node for every schema element and token position.
\end{theorem}

\begin{proof}
Consider any schema element $s$ and token position $v$. Let
$\{\Delta_i^{(s)}(v)\}_{i \in H}$ be the honest submissions and
$\{\Delta_j^{(s)}(v)\}_{j \in B}$ be the Byzantine submissions, with
corresponding weights $\{w_i\}_{i \in H}$ and $\{w_j\}_{j \in B}$.

The total weight is $W = W_H + W_B$. The weighted median selects the value $x$
such that the total weight of values $\leq x$ is $\geq W/2$. For the median
to equal a Byzantine value $\Delta_j^{(s)}(v)$ for $j \in B$, the Byzantine
nodes would need to shift the weight balance so that strictly more than $W/2$
weight lies on one side. This requires:
\begin{equation}
    W_B > W/2 - W_H/2 = (W_H + W_B)/2 - W_H/2 = W_B/2,
\end{equation}
which is trivially true, but for the median to be \emph{moved away from all
honest values}, we need $W_B \geq W_H/2$. Since $W_H > 2W_B$ implies
$W_B < W_H/2$, this is impossible. Therefore, the median must lie within the
range of honest submissions.

Since honest submissions are drawn from $\{-\alpha_i, 0, +\alpha_i\}$ (1-bit
quantized), the median equals one of the honest values.
\end{proof}

\subsection{Coordinate-Wise Aggregation}

For multi-dimensional prior factors, we apply the weighted median coordinate-wise:

\begin{algorithm}[H]
\caption{Byzantine-Robust Prior Aggregation}
\label{alg:aggregation}
\begin{algorithmic}[1]
\Require Submissions $\{E_i\}_{i=1}^n$, weights $\{w_i\}_{i=1}^n$, schema $\mathcal{S}$
\Ensure Aggregated prior $\tilde{E}$
\For{each schema element $s \in \mathcal{S}$}
    \For{each token position $v$ in $s$}
        \State $\tilde{\Delta}^{(s)}(v) \gets \wmed(\{\Delta_i^{(s)}(v)\}_{i=1}^n; \{w_i\})$
    \EndFor
    \State $\tilde{\alpha}^{(s)} \gets \wmed(\{\alpha_i^{(s)}\}_{i=1}^n; \{w_i\})$
\EndFor
\State $\tilde{q} \gets \wmed(\{q_i\}_{i=1}^n; \{w_i\})$
\State $\tilde{E} \gets \text{Assemble}(\{\tilde{\Delta}^{(s)}, \tilde{\alpha}^{(s)}\}_{s \in \mathcal{S}}, \tilde{q})$
\State \Return $\tilde{E}$
\end{algorithmic}
\end{algorithm}

\subsection{Aggregation Quality Guarantee}

\begin{theorem}[Quality Preservation]
\label{thm:quality}
Let $q^* = \med(\{q_i\}_{i \in H})$ be the median quality of honest
submissions. Under Assumption~\ref{ass:threat}, the aggregated prior quality
satisfies:
\begin{equation}
    \tilde{q} \geq q^* - O\left(\frac{1}{\sqrt{|H|}}\right),
\end{equation}
where $|H| = n - f$ is the number of honest participants.
\end{theorem}

\begin{proof}
By Theorem~\ref{thm:byz-robust}, the weighted median quality $\tilde{q}$ lies
within the range of honest quality scores. The median of $|H|$ honest scores
concentrates around the population median at rate $O(1/\sqrt{|H|})$ by the
central limit theorem applied to order statistics~\cite{david2003order}. Since
$\tilde{q}$ is bounded below by the minimum honest quality and the weighted
median is within the honest range, the result follows.
\end{proof}

\subsection{Attack Analysis}

\paragraph{Random Noise Attack.}
Byzantine nodes submit random $\Delta_j^{(s)}(v) \sim \text{Uniform}(\{-1, +1\})$.
Since honest nodes have correlated submissions (reflecting genuine patterns),
the weighted median filters out uncorrelated noise. The expected degradation is
$O(f/n)$ in prior quality.

\paragraph{Targeted Poisoning.}
Byzantine nodes identify high-value schema elements and submit adversarial
factors designed to invert the correct recommendation. By
Theorem~\ref{thm:byz-robust}, this requires controlling $> W_H / 2$ total
weight, which is impossible under the stake and quality constraints.

\paragraph{Gradient Mimicry.}
Sophisticated adversaries might produce plausible-looking priors that subtly
degrade performance. Quality verification via holdout evaluation
(Section~\ref{sec:protocol}, Phase 3) detects such attacks with probability
$\geq 1 - (1-\delta)^k$ after $k$ validation samples.

% ============================================================================
% 6. SYBIL RESISTANCE
% ============================================================================
