\section{Gossip Protocol}

\label{sec:gossip}

\subsection{Protocol Design}

DSO uses an epidemic gossip protocol~\cite{demers1987epidemic} for propagating
high-quality priors across the network. The protocol is designed to be:
\begin{itemize}[leftmargin=1.5em]
    \item \textbf{Pull-based:} Nodes request priors they lack, avoiding
          unsolicited data transfer.
    \item \textbf{Quality-filtered:} Only priors with $q_i \geq q_{\text{fetch}}$
          (default: 0.5) are propagated.
    \item \textbf{Stake-weighted:} Propagation priority is proportional to
          quality score times stake.
\end{itemize}

\begin{algorithm}[H]
\caption{DSO Gossip Protocol (per node)}
\label{alg:gossip}
\begin{algorithmic}[1]
\Require Local replica $L$, registry connection, peer set $\mathcal{P}$
\State Subscribe to registry events: $\texttt{NewSubmission}$, $\texttt{QualityUpdate}$
\Loop
    \State $\texttt{events} \gets \text{Registry.PollEvents}()$
    \For{each $\texttt{NewSubmission}(\texttt{id}, q, \texttt{cid})$ in events}
        \If{$q \geq q_{\text{fetch}}$ \textbf{and} $\texttt{id} \notin L$}
            \State $E \gets \text{FetchFromPeersOrIPFS}(\texttt{cid}, \mathcal{P})$
            \State Verify: $\text{MerkleRoot}(E) = \text{Registry.GetMerkleRoot}(\texttt{id})$
            \If{verification passes}
                \State $L.\text{Insert}(E, \texttt{id})$
                \State Announce availability to peers
            \EndIf
        \EndIf
    \EndFor
    \For{each $\texttt{QualityUpdate}(\texttt{id}, q_{\text{new}})$ in events}
        \If{$q_{\text{new}} < q_{\min}$ \textbf{and} $\texttt{id} \in L$}
            \State $L.\text{Remove}(\texttt{id})$
            \Comment{Garbage collect low-quality}
        \EndIf
    \EndFor
    \State Sleep($\Delta t_{\text{gossip}}$)
    \Comment{Default: 30s}
\EndLoop
\end{algorithmic}
\end{algorithm}

\subsection{Propagation Analysis}

\begin{proposition}[Propagation Time]
\label{prop:propagation}
In a network of $n$ nodes with random peer connections (average degree $d$),
a new prior with quality $q \geq q_{\text{fetch}}$ reaches all honest nodes
in expected time:
\begin{equation}
    T_{\text{prop}} = O\left(\frac{\log n}{d} \cdot \Delta t_{\text{gossip}}\right).
\end{equation}
\end{proposition}

\begin{proof}
Standard epidemic gossip analysis~\cite{demers1987epidemic}. Each gossip round,
each informed node tells $d$ peers. The number of informed nodes doubles each
round until saturation, giving $O(\log n)$ rounds. With gossip interval
$\Delta t_{\text{gossip}}$ and average degree $d$, the effective round time is
$\Delta t_{\text{gossip}} / d$.
\end{proof}

For typical parameters ($n = 1000$, $d = 10$, $\Delta t = 30$s), propagation
time is approximately $O(\log(1000) / 10 \times 30) \approx 21$s.

\subsection{Bandwidth Efficiency}

The 1-bit compression from ASO is critical for gossip efficiency:

\begin{table}[h]
\centering
\begin{tabular}{lcc}
\toprule
Representation & Size per Prior & Bandwidth (100 priors) \\
\midrule
Full precision (FP32) & 128 KB & 12.5 MB \\
16-bit quantized & 64 KB & 6.25 MB \\
1-bit compressed & 4.3 KB & 430 KB \\
1-bit + schema sparse & 1.2 KB & 120 KB \\
\bottomrule
\end{tabular}
\caption{Prior data sizes under different representations. 1-bit compression
with schema-sparse encoding reduces bandwidth by $100\times$ compared to
full precision.}
\label{tab:bandwidth}
\end{table}

\subsection{Incentive-Compatible Propagation}

Nodes that propagate high-quality priors earn fee rebates:
\begin{equation}
    \text{rebate}_i = \gamma_{\text{prop}} \sum_{j: \text{fetched from } i} q_j \cdot \text{size}_j,
    \label{eq:rebate}
\end{equation}
where $\gamma_{\text{prop}}$ is the propagation reward rate (default: 0.001 \$AI
per quality-byte). This incentivizes nodes to maintain high availability and
serve popular priors.

% ============================================================================
% 8. CONVERGENCE ANALYSIS
% ============================================================================
