\section{Reproducibility Details}

\label{app:reproducibility}

\begin{itemize}[leftmargin=1.5em]
    \item \textbf{Environment:} Docker containers with Python 3.11, Go 1.21,
          LanceDB 0.3.0, IPFS Kubo 0.25.0.
    \item \textbf{Network simulation:} Custom simulator with configurable
          latency, bandwidth, and packet loss.
    \item \textbf{Byzantine behavior:} Random noise ($\Delta \sim \text{Unif}(\{-1,+1\})$)
          for all reported experiments. Targeted attacks evaluated in
          Appendix~\ref{app:security}.
    \item \textbf{Seeds:} Fixed seeds (42, 123, 456) for three runs; results
          report mean $\pm$ std (std $< 0.3\%$ in all cases).
    \item \textbf{Hardware:} 10$\times$ M1 MacBook Pro (simulated agents),
          1$\times$ 4-core VM for registry and IPFS.
\end{itemize}

% ============================================================================
% REFERENCES
% ============================================================================
\bibliographystyle{plainnat}
\begin{thebibliography}{30}

\bibitem[Baker et~al.(2020)]{baker2020emergent}
Baker, B., Kanitscheider, I., Marber, T., et~al.
\newblock Emergent tool use from multi-agent autocurricula.
\newblock \emph{ICLR}, 2020.

\bibitem[Benet(2014)]{benet2014ipfs}
Benet, J.
\newblock IPFS --- Content addressed, versioned, P2P file system.
\newblock \emph{arXiv preprint arXiv:1407.3561}, 2014.

\bibitem[Bittensor(2023)]{bittensor2023}
Bittensor.
\newblock Bittensor: A peer-to-peer intelligence market.
\newblock Whitepaper, 2023.

\bibitem[Blanchard et~al.(2017)]{blanchard2017machine}
Blanchard, P., El~Mhamdi, E.~M., Guerraoui, R., and Stainer, J.
\newblock Machine learning with adversaries: Byzantine tolerant gradient
  descent.
\newblock \emph{NeurIPS}, 2017.

\bibitem[Christianos et~al.(2021)]{christianos2021scaling}
Christianos, F., Papoudakis, G., Rahman, M.~A., and Albrecht, S.~V.
\newblock Scaling multi-agent reinforcement learning with selective parameter
  sharing.
\newblock \emph{ICML}, 2021.

\bibitem[David \& Nagaraja(2003)]{david2003order}
David, H.~A. and Nagaraja, H.~N.
\newblock \emph{Order Statistics}.
\newblock Wiley, 3rd edition, 2003.

\bibitem[Demers et~al.(1987)]{demers1987epidemic}
Demers, A., Greene, D., Hauser, C., et~al.
\newblock Epidemic algorithms for replicated database maintenance.
\newblock \emph{PODC}, 1987.

\bibitem[FLock(2024)]{flock2024}
FLock.
\newblock FLock: Federated learning on blockchain.
\newblock \url{https://flock.io}, 2024.

\bibitem[Hanzo AI Research(2026)]{hanzo2026aso}
Hanzo AI Research.
\newblock Active Semantic Optimization: Training-free adaptation via Bayesian
  Product-of-Experts decoding.
\newblock Technical report, Hanzo AI, 2026.

\bibitem[Hyeon-Woo et~al.(2022)]{hyeon2022fedpara}
Hyeon-Woo, N., Ye-Bin, M., and Tae-Hyun, O.
\newblock FedPara: Low-rank Hadamard product for communication-efficient
  federated learning.
\newblock \emph{ICLR}, 2022.

\bibitem[Jim{\'e}nez et~al.(2024)]{jimenez2024swebench}
Jim{\'e}nez, C.~E., Yang, J., Wettig, A., et~al.
\newblock SWE-bench: Can language models resolve real-world {GitHub} issues?
\newblock \emph{ICLR}, 2024.

\bibitem[Koloskova et~al.(2020)]{koloskova2020unified}
Koloskova, A., Stich, S.~U., and Jaggi, M.
\newblock Decentralized stochastic optimization and gossip algorithms with
  compressed communication.
\newblock \emph{ICML}, 2020.

\bibitem[Lamport et~al.(1982)]{lamport1982byzantine}
Lamport, L., Shostak, R., and Pease, M.
\newblock The Byzantine generals problem.
\newblock \emph{ACM TOPLAS}, 4(3):382--401, 1982.

\bibitem[LanceDB(2024)]{lancedb2024}
LanceDB.
\newblock LanceDB: Developer-friendly serverless vector database.
\newblock \url{https://lancedb.github.io/lancedb/}, 2024.

\bibitem[Li et~al.(2020)]{li2020federated}
Li, T., Sahu, A.~K., Zaheer, M., Sanjabi, M., Talwalkar, A., and Smith, V.
\newblock Federated optimization in heterogeneous networks.
\newblock \emph{MLSys}, 2020.

\bibitem[McMahan et~al.(2017)]{mcmahan2017communication}
McMahan, B., Moore, E., Ramage, D., Hampson, S., and Arcas, B.~A.~y.
\newblock Communication-efficient learning of deep networks from decentralized
  data.
\newblock \emph{AISTATS}, 2017.

\bibitem[Mhamdi et~al.(2018)]{mhamdi2018hidden}
El~Mhamdi, E.~M., Guerraoui, R., and Rouault, S.
\newblock The hidden vulnerability of distributed learning in {B}yzantium.
\newblock \emph{ICML}, 2018.

\bibitem[Minsker(2015)]{minsker2015geometric}
Minsker, S.
\newblock Geometric median and robust estimation in {Banach} spaces.
\newblock \emph{Bernoulli}, 21(4):2308--2335, 2015.

\bibitem[Pease et~al.(1980)]{pease1980reaching}
Pease, M., Shostak, R., and Lamport, L.
\newblock Reaching agreement in the presence of faults.
\newblock \emph{JACM}, 27(2):228--234, 1980.

\bibitem[Warnat-Herresthal et~al.(2021)]{warnat2021swarm}
Warnat-Herresthal, S., Schultze, H., Shastry, K.~L., et~al.
\newblock Swarm Learning for decentralized and confidential clinical machine
  learning.
\newblock \emph{Nature}, 594:265--270, 2021.

\bibitem[Williams(2019)]{williams2019arweave}
Williams, S.
\newblock Arweave: A protocol for economically sustainable information
  permanence.
\newblock Yellowpaper, 2019.

\bibitem[Yin et~al.(2018)]{yin2018byzantine}
Yin, D., Chen, Y., Kannan, R., and Bartlett, P.
\newblock Byzantine-robust distributed learning: Towards optimal statistical
  rates.
\newblock \emph{ICML}, 2018.

\bibitem[Zhang et~al.(2024)]{zhang2024fedlora}
Zhang, Y., Chen, X., Wang, Y., et~al.
\newblock FedLoRA: Communication-efficient federated learning with low-rank
  adaptation.
\newblock \emph{arXiv preprint arXiv:2402.11846}, 2024.

\end{thebibliography}

