\section{Architecture}
\label{sec:arch}

\subsection{AI Chain = A-Chain, Reframed}

The Hanzo AI Chain is the Lux A-Chain (LP-134), the unified global
attestation ledger, viewed as the AI Virtual Machine. There is exactly one
chain; the term ``AI Chain'' is the public-facing identity of A-Chain when
it is acting as the open AI protocol for the Hanzo Network. This avoids the
proliferation of ledgers that has plagued every prior decentralized AI
proposal: a single canonical chain hosts attestations for value-settlement
chains (P/Q/Z lanes), and it hosts the AI registry, agent registry,
compute markets, and AI coin economics described in this paper.

\begin{figure}[h]
\centering
\begin{tikzpicture}[
    node distance=0.9cm and 1.1cm,
    box/.style={rectangle, draw, minimum width=2.4cm, minimum height=0.7cm,
                font=\small, rounded corners=1pt},
    arrow/.style={->, thick}
]
\node[box] (achain) {A-Chain (= AI Chain / AIVM)};
\node[box, below left=of achain] (lux) {Lux Quasar 3.0 (LP-105)};
\node[box, below right=of achain] (gpu) {QuasarGPU (LP-132)};
\node[box, below=of achain] (stm) {QuasarSTM (LP-010)};
\node[box, above=of achain] (hmm) {HMM Registry};
\node[box, left=of hmm] (mkts) {Compute Markets};
\node[box, right=of hmm] (agents) {Agent Registry};
\draw[arrow] (achain) -- (lux);
\draw[arrow] (achain) -- (gpu);
\draw[arrow] (achain) -- (stm);
\draw[arrow] (hmm) -- (achain);
\draw[arrow] (mkts) -- (achain);
\draw[arrow] (agents) -- (achain);
\end{tikzpicture}
\caption{The AI Chain is the AIVM persona of A-Chain. It is settled by
Quasar 3.0, executed by QuasarGPU, and scheduled by QuasarSTM. On top of
that single ledger sit the HMM registry, the agent registry, and the
compute markets.}
\label{fig:arch}
\end{figure}

\subsection{Settlement: Lux Quasar 3.0 (LP-105)}

Every block on the AI Chain is finalized by a single \emph{QuasarCert}
that aggregates three independent post-quantum verification lanes:

\begin{itemize}[leftmargin=1.2em]
  \item \textbf{P-lane} — BLS12-381 aggregate signature (classical fast path);
  \item \textbf{Q-lane} — Ringtail two-round Ring-LWE threshold signature;
  \item \textbf{Z-lane} — a Z-Chain Groth16 SNARK rolling up $N$ ML-DSA-65
        validator signatures into a single 192-byte proof plus a 32-byte
        public-inputs hash.
\end{itemize}

Lane independence guarantees that an adversary must simultaneously break
the discrete logarithm on BLS12-381, Module-LWE for Ringtail, and
Module-LWE/MSIS for ML-DSA to forge a certificate. Breaking any one or any
two lanes leaves the third intact. See LP-105 for the formal definition,
the certificate-subject binding, and the security proofs.

For the AI Chain, this means every inference receipt and every market
state transition inherits the same three-lane post-quantum envelope used
to settle LUX value transfers. There is no separate, weaker security tier
for ``AI things.''

\subsection{Execution: QuasarGPU (LP-132)}

QuasarGPU is the GPU-resident execution substrate that verifies QuasarCerts
and runs the AIVM. The same hardware that batches BLS aggregate verify
($\sim$875\,$\mu$s for $n{=}100$), Ringtail batch verify ($O(1)$ amortized
after DKG), and the single Groth16 pairing check (200--500\,$\mu$s) also
hosts the inference kernels invoked by AIVM opcodes. Co-location is
deliberate: a GPU that has already loaded a Hanzo model into HBM can
service the next inference job without paging weights from main memory,
and it can produce the TEE quote that anchors the result without leaving
the trusted boundary.

QuasarGPU exposes a small set of primitives the AIVM relies on:

\begin{itemize}[leftmargin=1.2em]
  \item \texttt{ATTEST(quote)} — produce a vendor-rooted attestation
        quote (Intel TDX, AMD SEV-SNP, NVIDIA H100/H200/B100 confidential
        compute, ARM CCA) covering the loaded model commitment and
        runtime version;
  \item \texttt{INFER(model\_root, input)} — run forward pass under the
        attested boundary, returning $(\text{output}, \text{quote})$;
  \item \texttt{VERIFY(quasarcert)} — verify a QuasarCert on GPU
        (used by light clients and by cross-chain bridges).
\end{itemize}

\subsection{Scheduling: QuasarSTM Lanes (LP-010)}

QuasarSTM (LP-010) is the lane-based software transactional memory model
used by the Quasar pipeline. The AI Chain reuses the same lane
abstraction: independent compute jobs that do not touch overlapping state
run in parallel lanes; only state-touching transitions (operator stake
updates, market settlement, model registry mutations) serialize. Because
inference is read-mostly with respect to the chain (model commitments are
read; only the receipt is written), the AI Chain achieves throughput
limited by GPU capacity, not by chain serialization.

\subsection{The AIVM Opcode Surface}

The AIVM is a superset of the EVM with tensor-aware opcodes. Standard
EVM opcodes execute unchanged. AI-specific operations are exposed as
precompiles addressable from any EVM contract:

\begin{table}[h]
\centering
\small
\begin{tabular}{lll}
\toprule
\textbf{Address} & \textbf{Name} & \textbf{Purpose} \\
\midrule
\texttt{0x0a01} & \texttt{HMM\_LOOKUP}    & Resolve model\_root → metadata \\
\texttt{0x0a02} & \texttt{INFER\_REQUEST} & Submit inference job to market \\
\texttt{0x0a03} & \texttt{INFER\_VERIFY}  & Verify TEE quote + receipt \\
\texttt{0x0a04} & \texttt{TRAIN\_BOUNTY}  & Post training bounty \\
\texttt{0x0a05} & \texttt{GPU\_LEASE}     & Lease GPU pool capacity \\
\texttt{0x0a06} & \texttt{AGENT\_AUTH}    & Authorize agent action under cap \\
\texttt{0x0a07} & \texttt{F\_CALL}        & Cross-chain call to F-Chain (FHE) \\
\texttt{0x0a08} & \texttt{M\_CUSTODY}     & Cross-chain call to M-Chain (MPC) \\
\bottomrule
\end{tabular}
\caption{AIVM precompile addresses. EVM contracts may call any of these
to interact with the AI Chain registries and markets without leaving the
solidity programming model.}
\end{table}

\subsection{Block Time and Throughput}

The AI Chain targets the same 500\,ms block time and $\geq$100\,K
TPS-per-chain envelope as the rest of Quasar 3.0 (LP-105 §8). Inference
throughput is bounded by aggregate operator GPU capacity, not by the
chain itself; in practice each block can record thousands of inference
receipts because each receipt is a fixed-size commitment, not a
re-execution.
