\section{Compute Markets}
\label{sec:markets}

The AI Chain hosts three composable compute markets. All three settle in
AI, all three rely on TEE attestation for correctness, and all three
share the operator stake and slashing mechanisms described in
\S\ref{sec:aicoin}.

\subsection{Inference Market}

\paragraph{Listing.}
A compute operator with $\geq$10{,}000 AI staked may list capacity for
any registered HMM. A listing pins (model\_root, max\_throughput,
max\_latency, geographic region, supported precisions). Listings are
indexed in an on-chain order book keyed by model\_root.

\paragraph{Job submission.}
A caller submits an inference job by invoking the
\texttt{INFER\_REQUEST} precompile with (model\_root, input\_root,
max\_payment, deadline). The chain matches the job to a listing using
the operator's posted price function and the caller's
\texttt{max\_payment} ceiling. The matched operator escrows the
caller's payment and starts the job.

\paragraph{Receipt.}
On completion the operator emits a receipt
$(\text{output\_root}, \text{tee\_quote}, \text{operator\_sig})$. The
\texttt{tee\_quote} covers the model\_root, the runtime version, the
input\_root, and the output\_root; this is what makes the result
verifiable. The receipt is recorded in the block; payment is released
from escrow; the operator's stake is unaffected.

\paragraph{Throughput.}
Per-token pricing scales linearly with output length. Operators can
batch many concurrent jobs on the same loaded model in HBM, so per-job
fixed cost is small. Targets at activation: $\leq$200\,ms first-token
latency on 7B-class models, $\leq$2\,s on 70B-class.

\subsection{Training Market}

\paragraph{Bounty.}
A model owner posts a bounty by invoking \texttt{TRAIN\_BOUNTY}
with (parent\_model\_root, dataset\_root, target\_metric,
target\_threshold, payout, deadline). The chain holds the payout in
escrow.

\paragraph{Trainer.}
Any operator may attempt the bounty. The training run is conducted on
the operator's GPUs; the operator must produce a TEE attestation
covering the training environment. On submission the operator emits a
candidate (new\_model\_root, training\_log\_root, tee\_quote).

\paragraph{Verification.}
Verifiers, drawn from the staked operator set by VRF, replay a sampled
fraction $\alpha$ of the training steps inside their own TEEs and
compare the resulting state. A successful replay releases the bounty;
a failed replay slashes the candidate operator. This is the same
verification-by-sampling pattern Hanzo ACI describes for general
compute, specialized to training (\texttt{\textasciitilde/work/hanzo/papers/hanzo-aci}).

\paragraph{Settlement.}
On success the new model is registered as a child of the parent in the
HMM provenance DAG (\S\ref{sec:hmm}). The training\_log\_root is
preserved in CAS, allowing third-party post-hoc audits.

\subsection{GPU Pool Market}

\paragraph{Capacity rental.}
Where the inference market sells per-token throughput against a
specific model, the GPU pool market sells raw GPU-seconds against
any workload. An operator lists pool capacity by GPU SKU (e.g.\ H100,
H200, B100, B200, MI300X, TPUv5) and time window. A renter posts a
job that includes a workload commitment and a TEE-verifiable runtime
image.

\paragraph{Use cases.}
Pool rental is targeted at workloads that do not fit the inference or
training markets: research experiments, custom kernels, fine-tuning
runs that bring their own data and infra, benchmarking, ML systems
research. The chain provides only the rental contract, the TEE
attestation, and the settlement; the workload itself is opaque to the
chain.

\paragraph{Pricing.}
Pool listings post a per-GPU-second price. Renters pay up front in AI
into escrow; payment is released to the operator at the end of the
window minus any documented downtime. Disputes are resolved by TEE
log evidence.

\subsection{TEE Attestation Across All Three Markets}

All three markets rely on the A-Chain attestation primitive (LP-134).
A market job is honest iff (a) the TEE quote validates against the
correct vendor root, (b) the quoted measurement matches the agreed
runtime image, and (c) the quoted input/output commitments match what
the chain recorded. Any divergence is a slashable offense.

\begin{algorithm}
\caption{Honest-job check (executed by any verifier)}
\begin{algorithmic}[1]
\Require receipt $r$, vendor root $V$, expected runtime image $I$
\State $q \gets r.\text{tee\_quote}$
\If{not VerifyQuote($q$, $V$)} \Return \textsc{SlashOperator} \EndIf
\If{$q.\text{measurement} \neq I.\text{measurement}$}
  \Return \textsc{SlashOperator}
\EndIf
\If{$q.\text{inputs} \neq r.\text{input\_root}$}
  \Return \textsc{SlashOperator}
\EndIf
\If{$q.\text{outputs} \neq r.\text{output\_root}$}
  \Return \textsc{SlashOperator}
\EndIf
\State \Return \textsc{Honest}
\end{algorithmic}
\end{algorithm}

\subsection{Spot vs.\ Reserved Capacity}

Operators may sell capacity in two pricing modes per market:

\begin{itemize}[leftmargin=1.2em]
  \item \emph{Spot}: dynamic price set per epoch, first-come-first-served,
        cancellable on either side with a small penalty. Targets latency-
        tolerant workloads at minimum cost.
  \item \emph{Reserved}: a forward contract paid up front for a guaranteed
        capacity window. Targets predictable workloads (production
        inference, scheduled training).
\end{itemize}

The two modes share the same settlement, attestation, and slashing
machinery; they differ only in payment timing and cancellation policy.

\subsection{Why Markets, Not a Single Centralized Scheduler}

The three-market structure is not a UX choice; it is a property
required for the chain to function as the open AI protocol. A single
centralized scheduler — even a benevolent one — would be a trust
assumption equivalent to an OpenAI or an Anthropic. Markets, by
contrast, expose the price function explicitly, let any operator
participate, and let any caller verify the result. The chain mediates,
it does not adjudicate.
