\section{Cross-Chain Integration}
\label{sec:xchain}

The AI Chain is one chain in the Lux multi-chain layout (LP-134, the
P/Q/Z/A/B/M/F pipeline). Each peer chain has a single, well-defined
purpose; AI Chain composes with them through small, typed precompiles
rather than redefining their work.

\subsection{C-Chain (EVM)}

The Lux C-Chain runs the general-purpose EVM. The AI Chain runs the
AIVM, which is a strict superset of the EVM (\S\ref{sec:arch}). Any
solidity contract that runs on C-Chain runs on AI Chain unmodified;
contracts on AI Chain may additionally call the AIVM precompiles.

The two chains exchange value and messages through B-Chain bridges.
A C-Chain ERC-20 contract holding USDC can settle the receivable side
of a Hanzo inference contract that lives on AI Chain; the bridge
message is what makes the cross-chain transfer atomic with respect to
the AI receipt.

\subsection{F-Chain (FHE Confidential Inference)}

F-Chain is the Lux fully homomorphic encryption chain (LP-013) and the
home of confidential ERC-20 patterns (LP-067). For inference workloads
where the input data must remain encrypted end-to-end, the AI Chain
delegates execution to F-Chain through the \texttt{F\_CALL} precompile.

The flow:

\begin{enumerate}[leftmargin=1.4em]
  \item Caller encrypts input under CKKS to the operator's public key
        on F-Chain.
  \item Caller invokes \texttt{F\_CALL}(model\_root, fhe\_input,
        max\_payment) on AI Chain.
  \item AI Chain emits a B-Chain bridge message routing the FHE
        ciphertext and the model commitment to F-Chain.
  \item F-Chain operator runs the homomorphic forward pass under TEE,
        emits an encrypted output and a TEE quote.
  \item Bridge returns (encrypted\_output, tee\_quote) to AI Chain;
        the receipt is recorded.
  \item Caller decrypts the output under their secret key.
\end{enumerate}

The provider sees only ciphertext throughout. The receipt on AI Chain
proves which model produced the result without revealing inputs or
outputs. This is the same primitive Hanzo AI Cloud describes for
confidential inference (\texttt{\textasciitilde/work/hanzo/papers/hanzo-ai-cloud}),
now anchored on chain.

\subsection{M-Chain (Threshold Custody)}

M-Chain is the Lux MPC chain. Model owners with high-value or
regulated weights may custody their model decryption keys on M-Chain
under threshold MPC: a quorum of custodians must sign to release a
decryption shard. The AI Chain references the custody record through
the \texttt{M\_CUSTODY} precompile; an operator that wants to load
encrypted weights must request and receive the threshold-signed
release.

This composes with HMMs in two useful ways. First, a model owner can
publish weights publicly under a permissive license while reserving a
specialized fine-tune under threshold custody. Second, regulated
domains (medical, financial) can satisfy data-residency requirements
by anchoring decryption authority to a custody quorum that includes
in-jurisdiction parties.

\subsection{B-Chain (Bridges)}

B-Chain (LP-134 §5) is the chain of bridges. All cross-chain messages
in the AI Chain workflows above flow over B-Chain. The bridge is itself
QuasarCert-secured: a B-Chain message attesting that ``AI Chain block
$h$ recorded receipt $r$'' is signed by the same three-lane consensus
that finalized the underlying block. There is no weaker bridge security
tier.

\subsection{P-Chain, Q-Chain, Z-Chain (Consensus Substrate)}

P/Q/Z are the consensus chains that produce the lanes of every
QuasarCert (LP-105). The AI Chain is a consumer of those certificates,
not a producer of them, in the same way every other Lux chain is.
This is why AI Chain inherits Quasar 3.0 security without modification:
the lanes are produced once, on the consensus substrate, and reused by
every chain that asks for finality.

\subsection{Topology Summary}

\begin{figure}[h]
\centering
\begin{tikzpicture}[
    node distance=0.55cm and 1.0cm,
    box/.style={rectangle, draw, minimum width=1.7cm, minimum height=0.55cm,
                font=\footnotesize, rounded corners=1pt},
    arrow/.style={<->, thick}
]
\node[box] (a) {A (= AI)};
\node[box, above=of a]      (c) {C (EVM)};
\node[box, below=of a]      (b) {B (Bridges)};
\node[box, left=of a]       (f) {F (FHE)};
\node[box, right=of a]      (m) {M (MPC)};
\node[box, below left=of b] (p) {P (val)};
\node[box, below=of b]      (q) {Q (cer)};
\node[box, below right=of b](z) {Z (snark)};
\draw[arrow] (c) -- (a);
\draw[arrow] (a) -- (b);
\draw[arrow] (a) -- (f);
\draw[arrow] (a) -- (m);
\draw[arrow] (b) -- (p);
\draw[arrow] (b) -- (q);
\draw[arrow] (b) -- (z);
\end{tikzpicture}
\caption{The AI Chain (A) sits at the center: it composes with the EVM
(C) above, with bridges (B) below, with FHE (F) on the left, and with
MPC (M) on the right. P/Q/Z produce the lanes that finalize every
chain.}
\end{figure}
