\section{Security}
\label{sec:security}

\subsection{Three-Lane Post-Quantum Safety, Inherited}

The AI Chain inherits the full Quasar 3.0 security envelope from the Lux
consensus substrate (LP-105). Every block is finalized by a QuasarCert
that simultaneously requires:

\begin{enumerate}[leftmargin=1.4em]
  \item A BLS12-381 aggregate signature over the certificate subject;
  \item A Ringtail two-round Ring-LWE threshold signature anchored to
        a Q-Chain ceremony root;
  \item A Z-Chain Groth16 SNARK rolling up $N$ ML-DSA-65 validator
        signatures into a single 192-byte proof plus 32-byte public-input
        hash.
\end{enumerate}

The certificate subject binds chain id, epoch, round, mode, and the
seven chain-specific roots ($P_\text{val}\,\|\,Q_\text{cer}\,\|\,
Z_\text{vk}\,\|\,A_\text{att}\,\|\,B_\text{br}\,\|\,M_\text{mpc}\,\|\,
F_\text{fhe}$) plus the parent block hash and execution roots, making
cross-epoch and cross-chain replay structurally impossible. See LP-105
§4 and §6 for the formal QuasarCert definition.

\subsection{What an AI Chain Adversary Must Break}

To forge a block on the AI Chain, an adversary must \emph{simultaneously}:

\begin{itemize}[leftmargin=1.2em]
  \item Solve the discrete logarithm on BLS12-381 (P-lane);
  \item Solve Module-LWE for the Ringtail parameters (Q-lane);
  \item Solve Module-LWE / MSIS for ML-DSA-65 (Z-lane).
\end{itemize}

Breaking any single lane, or any two of the three, leaves the third
intact and the certificate unforgeable. A future cryptographically
relevant quantum computer that breaks BLS does not break the AI Chain;
the Q and Z lanes still hold.

\subsection{Operator Honesty}

Block-level safety guarantees that what the chain accepted, the
quorum signed. It does not guarantee that an operator ran the right
model on the right input. That second-tier honesty is enforced by TEE
attestation:

\begin{itemize}[leftmargin=1.2em]
  \item Every receipt is accompanied by a vendor-rooted TEE quote
        (LP-134, A-Chain attestation primitive). The quote covers the
        loaded model commitment, the runtime image, and the input/output
        commitments.
  \item Verifiers re-validate the quote against the vendor root CA on
        chain. A malformed or unsigned quote rejects the receipt and
        slashes the operator (Algorithm in \S\ref{sec:markets}).
  \item Periodic spot-check verifiers, drawn by VRF, replay a sampled
        fraction of jobs in their own TEEs. Discrepancies escalate to
        full slashing.
\end{itemize}

\begin{theorem}[Operator-fraud detection]
\label{thm:fraud}
Let $\alpha$ be the verifier sampling rate per epoch and $\rho$ the
fraction of jobs an attacker chooses to falsify. The probability that a
false job goes undetected for $k$ epochs is
$(1 - \alpha\rho)^k$. For $\alpha = 0.003$ and $\rho = 0.05$, the
probability of evading detection across 1{,}000 jobs is below
$10^{-3}$, while the verifier overhead remains below 0.3\% of total
compute.
\end{theorem}

\begin{proof}[Sketch]
Verifier sampling is independent across jobs. The probability of
sampling a specific job is $\alpha$; the probability of detecting a
falsified job conditional on sampling is 1 (the TEE quote replay
exposes the divergence by construction). Over $k$ jobs, the unsampled
probability is $(1-\alpha)^{\rho k}$. Substituting: at $\alpha=0.003$,
$\rho=0.05$, $k=1000$ gives detection probability $> 0.999$.
\end{proof}

\subsection{Sybil Resistance}

Sybil attacks against the operator side are bounded by stake. A
fictitious operator must post $\geq$10{,}000 AI of stake; a Sybil farm of
$N$ identities must post $10{,}000\,N$ AI. Block reward and inference
fees are split per attested job, not per operator identity, so creating
identities does not increase share. False quotes from any one identity
slash 100\% of that identity's stake.

\subsection{Liveness}

Liveness on the AI Chain requires that the Quasar 3.0 quorum remains
intact. LP-105 guarantees safety with $f < n/3$ Byzantine validators;
liveness is preserved with $f < n/3$ failed validators provided the
remaining $2n/3 + 1$ produce the three lanes. No additional liveness
assumption is added on top.

For the markets, liveness requires that operators participate. Idle
operators do not earn block reward share, which produces a baseline
incentive to attest and serve.

\subsection{Replay and Cross-Chain Attacks}

Every QuasarCert binds the certificate subject to the chain id, the
epoch, and the seven chain-root tuple. The same finalized block on
AI Chain cannot be replayed on C-Chain or any other chain in the
pipeline because the chain id and the Z-Chain Groth16 verification key
differ. Cross-chain bridge messages on B-Chain include a B-Chain block
height and a destination chain id; replay across chains is similarly
impossible.

\subsection{Audit Surface}

The AI Chain exposes the following for external audit:

\begin{itemize}[leftmargin=1.2em]
  \item All HMM registry records and their training history DAG;
  \item All operator listings, jobs, receipts, and TEE quotes;
  \item All agent registry records and capability declarations;
  \item All QuasarCert lane components per block (P, Q, Z);
  \item All slashing events, with the underlying evidence
        (malformed quote, divergent replay, lane-incomplete cert).
\end{itemize}

This is, by design, more transparent than any centralized AI service.
The premise of the AI Chain is that the cost of that transparency is
paid down by the ability to verify any output without asking the
provider.
