\section{Conclusion}

The Hanzo AI Chain is the AI Virtual Machine: A-Chain (LP-134), the
unified Lux attestation chain, viewed as the open AI protocol. There
is one ledger, one consensus (Quasar 3.0, LP-105), one execution
substrate (QuasarGPU, LP-132), and one scheduling model (QuasarSTM,
LP-010). On top of that single ledger sit the Hanzo model registry,
the agent registry, and three composable compute markets, all
denominated in the native AI coin and all secured by three-lane
post-quantum certificates.

The economic claim is simple: a content-addressed registry of models,
a verifiable receipt for every inference call, a market for compute
priced in a coin that burns under load, and an agent layer with
durable identity together produce an open AI protocol that does not
require a hyperscaler. The technical claim is that this protocol can be
delivered without a separate chain and without a weaker security tier:
A-Chain's three-lane envelope and QuasarGPU's TEE primitives are exactly
the primitives an open AI protocol needs, and they are already in
production on the Lux primary network.

Activation is 25 December 2025. The spec is frozen as of 15 December
2025. From that boundary forward, AI is the native coin of the Hanzo
Network, HMMs are first-class on-chain objects, agents have persistent
identity and capability, and any operator with a TEE-attestable GPU can
sell compute into a market that any caller can price-discover and any
auditor can verify.

\paragraph{Cross-references.}
This paper is part of the Hanzo papers collection
(\texttt{\textasciitilde/work/hanzo/papers}). Related work in the same
collection: \emph{Hanzo ACI: AI Chain Infrastructure}
(\texttt{hanzo-aci/}); \emph{Hanzo AI Cloud: Confidential Inference,
Agent Hosting, and Model Serving} (\texttt{hanzo-ai-cloud/});
\emph{Hanzo Agent SDK} (\texttt{hanzo-agent-sdk/}); \emph{Hanzo Cloud
Infrastructure} (\texttt{hanzo-cloud-infrastructure/}).

Related Lux papers (\texttt{\textasciitilde/work/lux/papers}):
\emph{LP-010: QuasarSTM} (\texttt{lp-010-quasar-stm-3/});
\emph{LP-105: Quasar 3.0 Consensus}
(\texttt{lp-105-quasar-consensus/});
\emph{Lux A-Chain Attestation} (\texttt{lux-achain-attestation/});
\emph{Lux Decentralized Cloud} (\texttt{lux-decentralized-cloud/}).

\bibliographystyle{plain}
\begin{thebibliography}{9}

\bibitem{lp010}
Lux Industries.
\textit{LP-010: QuasarSTM 3.0 — Lane-based Software Transactional
Memory.}
Lux Network Technical Specifications, December 2025.

\bibitem{lp105}
Z. Kelling.
\textit{LP-105: Quasar 3.0 Consensus — Three-Lane Post-Quantum
Certificates.}
Lux Network Technical Specifications, December 2025.

\bibitem{lp132}
Lux Industries.
\textit{LP-132: QuasarGPU — GPU-Resident Verification and Execution
Substrate.}
Lux Network Technical Specifications, December 2025.

\bibitem{lp134}
Lux Industries.
\textit{LP-134: A-Chain — Unified Attestation for the P/Q/Z/A/B/M/F
Pipeline.}
Lux Network Technical Specifications, December 2025.

\bibitem{lp013}
Lux Industries.
\textit{LP-013: F-Chain — Fully Homomorphic Encryption Substrate.}
Lux Network Technical Specifications, 2025.

\bibitem{lp067}
Lux Industries.
\textit{LP-067: Confidential ERC-20 on F-Chain.}
Lux Network Technical Specifications, 2025.

\bibitem{lp136}
Lux Industries.
\textit{LP-136: Lux Decentralized Cloud — Compute Settlement Layer.}
Lux Network Technical Specifications, 2025.

\bibitem{aci}
Hanzo AI Research.
\textit{Hanzo ACI: AI Chain Infrastructure for Decentralized Compute
Markets.}
Hanzo Papers, 2025.

\bibitem{aicloud}
Z. Kelling.
\textit{Hanzo AI Cloud: Confidential Inference, Agent Hosting, and
Model Serving.}
Hanzo Papers, 2025.

\bibitem{agentsdk}
A. Worring and Z. Kelling.
\textit{Hanzo Agent SDK: Building AI Agents with Orchestration, Tool
Use, and Memory Systems.}
Hanzo Papers, 2025.

\end{thebibliography}
