\section{Crypto Payment Architecture}

\label{sec:architecture}
% ============================================================================

\subsection{Payment Flow}

\begin{enumerate}[leftmargin=1.4em]
    \item \textbf{Initiation}: Customer selects crypto payment at
    checkout. SDK displays available currencies and chains.
    \item \textbf{Rate lock}: Hanzo locks the exchange rate for 15
    minutes. Rate sourced from aggregated exchange feeds (Binance,
    Coinbase, Kraken) with TWAP smoothing.
    \item \textbf{Address generation}: MPC-derived payment address
    generated for this transaction (unique per payment).
    \item \textbf{Payment}: Customer sends cryptocurrency to the
    generated address.
    \item \textbf{Confirmation}: Chain-specific confirmation monitoring
    (see Section~\ref{sec:multichain}).
    \item \textbf{Settlement}: Received crypto is converted to the
    merchant's settlement currency (default USDC) and transferred to
    the merchant's MPC-managed settlement wallet.
    \item \textbf{Fiat off-ramp}: If configured, USDC is converted to
    fiat and wire-transferred to the merchant's bank account within 24
    hours.
\end{enumerate}

\subsection{MPC Wallet Management}

Each merchant's payment infrastructure is backed by an MPC key group
using CGGMP21 threshold ECDSA~\cite{cggmp21}:

\begin{itemize}[leftmargin=1.4em]
    \item \textbf{2-of-3 signing}: Hanzo Operations + Merchant Co-Signer
    + Disaster Recovery.
    \item \textbf{Single DKG ceremony}: One key generation produces
    signing capability for all supported chains (secp256k1 for EVM and
    Bitcoin, ed25519 for Solana via separate ceremony).
    \item \textbf{Payment addresses}: Derived from the MPC public key
    using chain-specific address encoding (Ethereum \texttt{keccak256},
    Bitcoin \texttt{bech32}, Solana \texttt{base58}).
    \item \textbf{Unique addresses}: Each payment gets a unique derived
    address via BIP-32 HD derivation from the MPC master key.
\end{itemize}

% ============================================================================
