\section{Conclusion}

\label{sec:conclusion}
% ============================================================================

Hanzo Crypto Payment Rails extends the proven checkout.js commerce
platform with multi-chain cryptocurrency acceptance. MPC custody
eliminates the key management burden for merchants, while the
integration with Hanzo Commerce API provides a unified interface for
fiat and crypto payments. Lux Network's sub-second finality makes it
the preferred settlement layer, with Bitcoin, Ethereum, and Solana
supported for customer flexibility. Production deployments demonstrate
that crypto payments can match the merchant experience of traditional
card payments: drop-in SDK integration, instant notification, and
next-day fiat settlement.

\begin{thebibliography}{99}
\bibitem{coinmarket2024}
CoinMarketCap, ``Global Cryptocurrency Market Capitalization,''
December 2024.

\bibitem{cggmp21}
R.~Canetti, R.~Gennaro, S.~Goldfeder, N.~Makriyannis, U.~Peled,
``UC Non-Interactive, Proactive, Threshold ECDSA with Identifiable
Aborts,'' \textit{ACM CCS}, 2020.

\bibitem{luxwhitepaper}
Lux Partners, ``Lux Network: A Scalable Multi-Chain Architecture,''
Technical Report, 2022.

\bibitem{bip32}
P.~Wuille, ``BIP-32: Hierarchical Deterministic Wallets,''
Bitcoin Improvement Proposals, 2012.

\bibitem{circle2024}
Circle, ``USDC: A Digital Dollar for the Global Economy,''
Technical Documentation, 2024.

\bibitem{lightning2024}
Lightning Labs, ``The Lightning Network: Scalable Off-Chain Instant
Payments,'' Technical Report, 2024.

\bibitem{pci2024}
PCI Security Standards Council, ``PCI DSS v4.0,'' 2024.

\bibitem{hanzocommerce}
Hanzo Industries, ``Hanzo Commerce: AI-Powered E-Commerce Platform,''
Product Documentation, 2024.
\end{thebibliography}

