\section*{Acknowledgments}


We thank the Ethereum Foundation for supporting this research, ZKNOX for their optimization contributions, and the broader cryptographic community for their work on post-quantum algorithms. Special recognition goes to the teams behind FALCON and Dilithium for creating practical lattice-based signature schemes.

\begin{thebibliography}{99}

\bibitem{LN16}
P. Longa and M. Naehrig, ``Speeding up the Number Theoretic Transform for Faster Ideal Lattice-Based Cryptography,'' Cryptology ePrint Archive, Report 2016/504, 2016.

\bibitem{RD23}
R. Dubois, ``Speeding up elliptic computations for Ethereum Account Abstraction,'' Cryptology ePrint Archive, Report 2023/939, 2023.

\bibitem{FALCON}
T. Prest et al., ``FALCON: Fast-Fourier Lattice-based Compact Signatures over NTRU,'' NIST Post-Quantum Cryptography Standardization, 2020.

\bibitem{Dilithium}
L. Ducas et al., ``CRYSTALS-Dilithium: A Lattice-Based Digital Signature Scheme,'' IACR Transactions on Cryptographic Hardware and Embedded Systems, 2018.

\bibitem{SPHINCS}
D. J. Bernstein et al., ``SPHINCS+: Stateless hash-based signatures,'' NIST Post-Quantum Cryptography Standardization, 2020.

\bibitem{Willow}
Google Quantum AI, ``Quantum error correction below the surface code threshold,'' Nature, 2024.

\bibitem{EIP616}
G. Colvin, ``EIP-616: SIMD Operations for the EVM,'' Ethereum Improvement Proposals, 2017.

\bibitem{Vitalik}
V. Buterin, ``An Introduction to FFTs and Fast Multiplication of Polynomials,'' 2019. [Online]. Available: \url{https://vitalik.eth.limo/general/2019/05/12/fft.html}

\bibitem{STARK}
E. Ben-Sasson et al., ``Scalable, transparent, and post-quantum secure computational integrity,'' Cryptology ePrint Archive, Report 2018/046, 2018.

\bibitem{NIST}
NIST, ``Post-Quantum Cryptography Standardization,'' 2024. [Online]. Available: \url{https://csrc.nist.gov/projects/post-quantum-cryptography}

\end{thebibliography}

