\section{Conclusion}

\label{sec:conclusion}
% ============================================================================

We have presented three consensus mechanisms for decentralized AI on Hanzo ACI:
Proof of AI Work for probabilistic inference verification with provable
economic security, GPU Cluster Consensus for Byzantine-tolerant distributed
computation, and Inference Attestation for on-chain provenance binding.
Production deployment across 240 GPU nodes demonstrates 99.97\% cheating
detection with only 6.2\% verification overhead, making honest operation the
only rational strategy for profit-maximizing GPU operators.

% ============================================================================
% BIBLIOGRAPHY
% ============================================================================
\begin{thebibliography}{15}

\bibitem{hanzoaci}
{Hanzo AI Research}.
\newblock Hanzo {ACI}: {AI} Chain Infrastructure for decentralized compute markets.
\newblock \emph{Hanzo AI Research}, December 2024.

\bibitem{hanzotokenomics}
{Hanzo AI Research}.
\newblock Hanzo Tokenomics: Token economics for {AI} infrastructure at scale.
\newblock \emph{Hanzo AI Research}, April 2026.

\bibitem{luxwhitepaper}
{Lux Partners}.
\newblock Lux: A scalable multi-consensus blockchain with post-quantum cryptography.
\newblock \emph{Lux Network Technical Report}, 2024.

\bibitem{megatron}
M.~Shoeybi, M.~Patwary, R.~Puri, P.~LeGresley, J.~Casper, and B.~Catanzaro.
\newblock Megatron-{LM}: Training multi-billion parameter language models using model parallelism.
\newblock \emph{arXiv:1909.08053}, 2019.

\bibitem{byzmean}
D.~Yin, Y.~Chen, R.~Kannan, and P.~Bartlett.
\newblock Byzantine-robust distributed learning: Towards optimal statistical rates.
\newblock In \emph{Proc.\ ICML}, 2018.

\bibitem{bittensor}
{Opentensor Foundation}.
\newblock Bittensor: A peer-to-peer intelligence market.
\newblock \emph{Bittensor White Paper}, 2023.

\bibitem{ritual}
{Ritual}.
\newblock Ritual: Decentralized {AI} inference.
\newblock \emph{Ritual Documentation}, 2024.

\bibitem{gensyn}
{Gensyn}.
\newblock Gensyn: A protocol for deep learning computation.
\newblock \emph{Gensyn White Paper}, 2023.

\end{thebibliography}

