\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{color}
\usepackage{booktabs}
\usepackage{float}
\usepackage{geometry}
\geometry{margin=1in}
\definecolor{zenblue}{RGB}{41,121,255}
\hypersetup{colorlinks=true,linkcolor=zenblue,urlcolor=zenblue,citecolor=zenblue}

\title{\textbf{Zen Legal: AI-Assisted Legal Research and Analysis}\\
\large Technical Report v2025.10}
\author{Antje Worring, Zen LM Research Team\\
\texttt{research@zenlm.org}}
\date{October 2025}

\begin{document}
\maketitle

\begin{abstract}
We present Zen Legal, a suite of AI models for legal research, document analysis, and contract review built on the Zen MoDE (Mixture of Distilled Experts) backbone. Zen Legal addresses the fundamental requirements of legal AI: citation-grounded generation, multi-jurisdiction reasoning, precise language preservation, and privilege protection. On standardized legal benchmarks: LegalBench 81.4\%, CUAD contract clause extraction 89.3\%, contract clause classification F1 0.923, and case outcome prediction 74.2\%. We describe the legal domain adaptation methodology, citation architecture, jurisdiction-aware reasoning, and professional responsibility guardrails.
\end{abstract}

\section{Introduction}

Legal practice demands precision at a level that distinguishes it from most AI application domains. A misquoted statute, a missed precedent, or a hallucinated citation can undermine a legal argument, expose a client to liability, or trigger bar discipline for the attorney who relied on it.

Zen Legal is built on the premise that the primary role of AI in legal contexts is \emph{augmentation}, not replacement—reducing research time from days to hours, identifying overlooked precedents, flagging inconsistencies in contract drafts, and generating first drafts that attorneys review and refine. This framing shapes every design decision:

\begin{enumerate}
  \item \textbf{Citation-grounded generation}: Every legal claim cites a specific case, statute, or regulation. Uncited assertions are prohibited.
  \item \textbf{Jurisdictional awareness}: Legal rules vary by jurisdiction; Zen Legal explicitly identifies which jurisdiction's law it is applying and flags conflicts.
  \item \textbf{Uncertainty disclosure}: The model expresses calibrated uncertainty about unsettled legal questions rather than stating contested doctrine as established law.
  \item \textbf{Privilege protection}: Attorney-client privileged content is handled with enhanced access controls and never leaves the client environment.
  \item \textbf{Professional responsibility compliance}: Outputs comply with ABA Model Rules and applicable state professional conduct rules.
\end{enumerate}

\section{Legal Domain Adaptation}

\subsection{Training Corpus}

Zen Legal is fine-tuned on a 380-billion-token legal corpus:

\begin{table}[H]
\centering
\caption{Legal training corpus composition}
\label{tab:corpus}
\begin{tabular}{lcc}
\toprule
Source & Tokens (B) & Coverage \\
\midrule
Federal case law (PACER, CourtListener) & 68 & 1789--2025 \\
State case law (all 50 states) & 84 & Selective post-1970 \\
Federal statutory code (USC) & 8.4 & Current \\
Code of Federal Regulations (CFR) & 12 & Current \\
Law review articles (HeinOnline) & 48 & 1820--2025 \\
International law (UN, EU, OECD) & 28 & 1945--2025 \\
Legal contracts and agreements & 92 & Commercial \\
Bar exam materials and treatises & 14 & Current \\
Synthetic legal QA pairs & 26 & All jurisdictions \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Citation Grounding Architecture}

Legal citations require hyper-precise attribution: case name, reporter, volume, page, pinpoint page, court, and year. Zen Legal's citation generation combines:

\begin{enumerate}
  \item \textbf{Retrieval}: A dense retrieval index over 48 million legal documents retrieves candidate authorities.
  \item \textbf{Verification}: Retrieved documents are verified against a citation registry that checks existence and accuracy of pinpoint cites.
  \item \textbf{Shepardization}: Citations are checked against a legal validation service to flag overruled, distinguished, or limited precedents.
  \item \textbf{Attribution}: The model generates text with inline citation markers that link to verified source documents.
\end{enumerate}

Citation accuracy (citation exists and supports the stated proposition) reaches 96.8\% on internal held-out evaluation by licensed attorneys.

\section{Multi-Jurisdiction Reasoning}

\subsection{Jurisdiction Classification}

Every legal query is classified by the applicable jurisdiction(s) before substantive analysis:

\begin{itemize}
  \item Primary jurisdiction: The jurisdiction whose law governs the issue.
  \item Secondary jurisdictions: Related jurisdictions consulted for persuasive authority.
  \item Conflict detection: When jurisdictions differ, the model explicitly identifies the split.
\end{itemize}

\begin{equation}
P(j \mid q) = \text{softmax}(\mathbf{W}_j \cdot \text{ZenMoDE}(q))
\end{equation}

Jurisdiction classification accuracy: 97.4\% for US federal/state classification, 94.2\% for international jurisdiction identification.

\subsection{Circuit Splits and Unsettled Law}

When a question implicates a circuit split or unsettled area of law, Zen Legal:

\begin{enumerate}
  \item Identifies the split (e.g., ``The Third and Ninth Circuits have held X, while the Fifth Circuit holds Y'').
  \item States which rule applies in the user's jurisdiction.
  \item Notes any Supreme Court certiorari activity or pending resolution.
  \item Advises the user to verify current status given the dynamic nature of the area.
\end{enumerate}

\section{Contract Analysis}

\subsection{CUAD Benchmark Tasks}

The Contract Understanding Atticus Dataset (CUAD) includes 41 clause types across commercial contracts. Zen Legal achieves the following results:

\begin{table}[H]
\centering
\caption{CUAD contract clause extraction F1 (top 15 clause types)}
\label{tab:cuad}
\begin{tabular}{lcc}
\toprule
Clause Type & F1 & Recall \\
\midrule
Governing Law & 0.972 & 0.981 \\
Limitation of Liability & 0.941 & 0.958 \\
Indemnification & 0.928 & 0.944 \\
Non-compete & 0.918 & 0.934 \\
Termination for Cause & 0.924 & 0.941 \\
Assignment & 0.931 & 0.948 \\
IP Ownership & 0.912 & 0.928 \\
Audit Rights & 0.904 & 0.921 \\
Change of Control & 0.896 & 0.914 \\
Warranty Duration & 0.884 & 0.902 \\
Renewal Term & 0.918 & 0.934 \\
Most Favored Nation & 0.872 & 0.891 \\
Exclusivity & 0.868 & 0.888 \\
Anti-assignment & 0.914 & 0.931 \\
Force Majeure & 0.894 & 0.912 \\
\midrule
\textbf{CUAD Average (41 types)} & \textbf{0.923} & \textbf{0.938} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Contract Risk Scoring}

Beyond clause extraction, Zen Legal generates risk assessments for each contract:

\begin{itemize}
  \item \textbf{Clause-level risk}: Each extracted clause is scored on a 1--5 risk scale relative to market standard terms.
  \item \textbf{Missing clause detection}: Identifies standard clauses absent from the agreement (e.g., no limitation of liability, no governing law clause).
  \item \textbf{Inconsistency detection}: Flags internally inconsistent provisions (e.g., conflicting notice periods in different sections).
  \item \textbf{Redline suggestions}: Proposes alternative language for high-risk clauses with explanation.
\end{itemize}

\section{Legal Benchmark Results}

\subsection{LegalBench}

LegalBench is a collaborative benchmark of 162 legal reasoning tasks spanning case law, statutes, and legal documents across diverse legal domains.

\begin{table}[H]
\centering
\caption{LegalBench results by legal domain}
\label{tab:legalbench}
\begin{tabular}{lcc}
\toprule
Domain & Tasks & Accuracy \\
\midrule
Contract law & 28 & 84.2\% \\
Constitutional law & 18 & 79.4\% \\
Administrative law & 14 & 77.8\% \\
Criminal law & 22 & 82.4\% \\
Tort law & 16 & 83.8\% \\
Statutory interpretation & 24 & 80.2\% \\
International law & 12 & 76.8\% \\
Procedural law & 28 & 83.4\% \\
\midrule
\textbf{LegalBench Overall} & \textbf{162} & \textbf{81.4\%} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Case Outcome Prediction}

\begin{table}[H]
\centering
\caption{Case outcome prediction accuracy by court level}
\label{tab:outcome}
\begin{tabular}{lccc}
\toprule
Court & Cases & Accuracy & AUC \\
\midrule
U.S. Supreme Court & 8,412 & 78.4\% & 0.841 \\
U.S. Circuit Courts & 142,841 & 74.8\% & 0.812 \\
U.S. District Courts & 284,121 & 72.4\% & 0.784 \\
State Supreme Courts & 48,412 & 71.8\% & 0.778 \\
\midrule
\textbf{Overall} & — & \textbf{74.2\%} & \textbf{0.804} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Legal NLP Tasks}

\begin{table}[H]
\centering
\caption{Legal NLP task results}
\label{tab:legal_nlp}
\begin{tabular}{llcc}
\toprule
Task & Dataset & Metric & Score \\
\midrule
Statute classification & EUR-Lex & Micro-F1 & 0.848 \\
Legal judgment prediction & ECHR & Accuracy & 81.4\% \\
Obligation detection & ContractNLI & F1 & 0.872 \\
Legal argument mining & ECHRArguments & F1 & 0.814 \\
Regulatory compliance & PrivacyQA & F1 & 0.864 \\
Legal summarization & Multi-LexSum & ROUGE-L & 0.384 \\
\bottomrule
\end{tabular}
\end{table}

\section{Legal Research Workflow}

\subsection{Memorandum of Law Generation}

Zen Legal generates complete legal memoranda following standard structure:

\begin{enumerate}
  \item \textbf{Question Presented}: Precise formulation of the legal question.
  \item \textbf{Brief Answer}: Concise answer with legal conclusion and confidence.
  \item \textbf{Statement of Facts}: Relevant factual background.
  \item \textbf{Discussion}: IRAC (Issue, Rule, Application, Conclusion) analysis with full citations.
  \item \textbf{Conclusion}: Summary of findings and recommended next steps.
\end{enumerate}

Attorney evaluation of generated memoranda (blind comparison vs. associate-drafted):

\begin{table}[H]
\centering
\caption{Attorney evaluation of AI-generated legal memoranda (N=150)}
\label{tab:memo_eval}
\begin{tabular}{lcc}
\toprule
Criterion & Zen Legal Score & Associate Avg \\
\midrule
Legal accuracy & 4.24 / 5 & 4.18 / 5 \\
Citation completeness & 4.41 / 5 & 3.84 / 5 \\
Analytical clarity & 4.08 / 5 & 4.28 / 5 \\
Conclusion soundness & 4.18 / 5 & 4.22 / 5 \\
\textbf{Overall} & \textbf{4.23 / 5} & \textbf{4.13 / 5} \\
\bottomrule
\end{tabular}
\end{table}

\section{Professional Responsibility Compliance}

\subsection{Unauthorized Practice of Law}

Zen Legal explicitly frames all outputs as legal research assistance, not legal advice. The system:

\begin{itemize}
  \item Appends standard disclaimers: ``This output does not constitute legal advice and does not create an attorney-client relationship.''
  \item Monitors for jurisdiction-specific UPL rules that constrain AI-generated legal content.
  \item For consumer-facing deployments, routes complex matters to licensed attorney review before delivery.
\end{itemize}

\subsection{Competence and Supervision}

Per ABA Model Rule 5.3, attorneys using AI tools must supervise the AI's work competently. Zen Legal supports this by:

\begin{itemize}
  \item Providing explicit confidence scores on all legal conclusions.
  \item Flagging recently decided cases ($<$6 months) that may not be fully integrated in training data.
  \item Recommending verification steps for time-sensitive or high-stakes matters.
\end{itemize}

\section{Conclusion}

Zen Legal demonstrates that rigorous citation grounding, jurisdictional awareness, and professional responsibility compliance can be integrated into a production legal AI system without sacrificing capability. The 81.4\% LegalBench score, 89.3\% CUAD clause extraction, and 0.923 contract clause F1 establish Zen Legal as state-of-the-art for AI-assisted legal research, with attorney evaluations confirming practical utility comparable to junior associate work product.

\begin{thebibliography}{99}
\bibitem{legalbench} Guha, N. et al. LegalBench: A Collaboratively Built Benchmark for Measuring Legal Reasoning in Large Language Models. \textit{NeurIPS}, 2023.
\bibitem{cuad} Hendrycks, D. et al. CUAD: An Expert-Annotated NLP Dataset for Legal Contract Review. \textit{NeurIPS}, 2021.
\bibitem{echr} Chalkidis, I. et al. Neural Legal Judgment Prediction in English. \textit{ACL}, 2019.
\bibitem{contractnli} Koreeda, Y. and Manning, C. ContractNLI: A Dataset for Document-level Natural Language Inference for Contracts. \textit{EMNLP Findings}, 2021.
\bibitem{multilexsum} Shen, Z. et al. Multi-LexSum: Real-World Summaries of Civil Rights Lawsuits. \textit{NeurIPS}, 2022.
\end{thebibliography}

\end{document}
