\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{color}
\usepackage{booktabs}
\usepackage{float}
\usepackage{geometry}
\geometry{margin=1in}
\definecolor{zenblue}{RGB}{41,121,255}
\definecolor{zengreen}{RGB}{52,199,89}
\hypersetup{colorlinks=true,linkcolor=zenblue,urlcolor=zenblue,citecolor=zenblue}

\title{\textbf{Zen-Pro: A High-Performance Language Model for Complex Reasoning}\\
\large Technical Report v2025.02}
\author{Antje Worring, Zach Kelling, Zen LM Research Team\\
\texttt{research@zenlm.org}}
\date{February 2025}

\begin{document}
\maketitle

\begin{abstract}
We present \textbf{Zen-Pro}, a 72 billion parameter dense transformer model designed for complex
reasoning, professional scientific applications, and advanced code generation. Zen-Pro extends the
Zen MoDE (Mixture of Distilled Experts) architecture to the 72B parameter scale, trained on
4.5 trillion tokens with a curriculum emphasizing mathematics, science, and multi-step reasoning.
Post-training combines extended chain-of-thought (CoT) supervised fine-tuning with reinforcement
learning from verifiable rewards (RLVR) on formally checkable problem domains. Zen-Pro achieves
frontier-level performance: MMLU 89.2\%, MATH 84.1\%, HumanEval 91.3\%, and GPQA 71.8\%,
competitive with the strongest models at this capability tier while supporting a 128K token
context window for long-form professional workflows.
\end{abstract}

\tableofcontents
\newpage

%% ─────────────────────────────────────────────────────────────────────────────
\section{Introduction}

The transition from 7B to 70B$+$ parameter models represents more than a quantitative scaling
step. Empirical research consistently shows qualitative capability emergence at larger scales,
including multi-step mathematical reasoning, scientific hypothesis evaluation, and nuanced
instruction disambiguation \cite{wei2022emergent, srivastava2022beyond}. Zen-Pro is our 72B
offering in this tier, designed to serve professional and enterprise use cases where accuracy,
reasoning depth, and reliability are paramount.

Zen-Pro makes the following contributions:

\begin{itemize}
  \item A 72B dense transformer trained on 4.5T tokens with a curriculum that front-loads
        high-quality mathematical, scientific, and code data in the final training phase.
  \item Extended post-training with long chain-of-thought (CoT) fine-tuning and reinforcement
        learning from verifiable rewards (RLVR) using formal verification on math and code tasks.
  \item A 128K token context window via extended RoPE, validated on real-world long-document
        tasks including legal review, codebase analysis, and scientific literature synthesis.
  \item Frontier-tier benchmark performance that closes the gap with models of significantly
        larger total parameter counts.
\end{itemize}

Zen-Pro serves as the flagship high-performance model in the Zen family for deployments requiring
production-grade reasoning without the infrastructure overhead of the Zen-Max MoE tier.

%% ─────────────────────────────────────────────────────────────────────────────
\section{Architecture}

\subsection{Model Hyperparameters}

Zen-Pro scales the Zen MoDE architecture to 72B parameters. Table~\ref{tab:arch} lists the
key architectural hyperparameters.

\begin{table}[H]
\centering
\caption{Zen-Pro architecture hyperparameters.}
\label{tab:arch}
\begin{tabular}{lc}
\toprule
\textbf{Hyperparameter} & \textbf{Value} \\
\midrule
Parameters (total)           & 72.7B \\
Layers                       & 80 \\
Attention heads              & 64 \\
KV heads (GQA)               & 8 \\
Hidden dimension             & 8192 \\
FFN intermediate dimension   & 29{,}568 \\
Vocabulary size              & 151{,}936 \\
Context length (training)    & 131{,}072 \\
Position encoding            & RoPE ($\theta = 5{,}000{,}000$) \\
Activation function          & SiLU \\
Normalization                & RMSNorm \\
Tied embeddings              & No \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Grouped Query Attention at Scale}

At 72B parameters and 80 layers, memory bandwidth is the dominant inference bottleneck.
Zen-Pro uses GQA with 64 query heads and 8 KV heads, achieving an 8$\times$ reduction in
KV cache size relative to full multi-head attention (MHA). This enables:

\begin{equation}
  \text{KV cache size} = 2 \cdot L \cdot n_{\text{kv}} \cdot d_k \cdot S \cdot \text{dtype\_bytes}
\end{equation}

where $L=80$ layers, $n_{\text{kv}}=8$ KV heads, $d_k = 128$ head dimension, and $S$ is sequence
length. For a 128K-token sequence in BF16, this yields approximately 26.8 GB of KV cache,
feasible on 4$\times$ A100-80GB with tensor parallelism.

\subsection{Sliding Window Attention for Long Context}

For the final 20 layers (layers 60--79), Zen-Pro employs sliding window attention (SWA)
\cite{beltagy2020longformer} with a window size of 16K tokens in addition to full global
attention on every 4th layer. This hybrid pattern maintains linear scaling in memory for
long sequences while preserving global information flow:

\begin{equation}
  \text{Attention scope} =
  \begin{cases}
    \text{full }[0, S] & \text{if } l \bmod 4 = 0 \\
    \text{window }[i-w, i] & \text{otherwise}
  \end{cases}
\end{equation}

where $w = 16{,}384$ is the sliding window size and $i$ is the current token position.

\subsection{Extended Context via RoPE Scaling}

To support 128K tokens, we use a higher RoPE base frequency of $\theta = 5{,}000{,}000$ combined
with position interpolation during the long-context fine-tuning phase. Pretraining is conducted
at 8K context, with a two-stage context extension:

\begin{enumerate}
  \item Extend to 32K at reduced learning rate ($5 \times 10^{-5}$) for 10B tokens.
  \item Extend to 128K at further reduced learning rate ($1 \times 10^{-5}$) for 5B tokens.
\end{enumerate}

This staged extension preserves short-context performance while enabling long-context
generalization.

%% ─────────────────────────────────────────────────────────────────────────────
\section{Training Methodology}

\subsection{Pretraining Data Curriculum}

Zen-Pro trains on 4.5T tokens with a data curriculum that emphasizes reasoning-intensive sources.
The final 10\% of training uses a concentrated curriculum of the highest-quality sources.

\begin{table}[H]
\centering
\caption{Pretraining data composition (4.5T tokens total).}
\label{tab:data}
\begin{tabular}{lcc}
\toprule
\textbf{Domain} & \textbf{Tokens (B)} & \textbf{Fraction} \\
\midrule
Web text (quality-filtered) & 1{,}800 & 40.0\% \\
Books and long-form         &   675 & 15.0\% \\
Code (all languages)        &   675 & 15.0\% \\
Scientific articles         &   450 & 10.0\% \\
Mathematics (formal+informal) & 360 &  8.0\% \\
Multilingual web            &   315 &  7.0\% \\
Curated reasoning chains    &   225 &  5.0\% \\
\midrule
Total                       & 4{,}500 & 100.0\% \\
\bottomrule
\end{tabular}
\end{table}

Curated reasoning chains include synthetic chain-of-thought solutions to mathematical competition
problems (AMC, AIME, Olympiad), formal theorem proving datasets (Lean, Isabelle), and
verified code problem solutions with test suite passes.

\subsection{Long Chain-of-Thought Fine-Tuning}

After base pretraining, we apply a specialized CoT-SFT phase on 2.5M long-form reasoning
trajectories. Unlike standard SFT which targets short responses, CoT-SFT trains the model to
generate step-by-step reasoning of 500--5{,}000 tokens before producing a final answer:

\begin{align}
  \mathcal{L}_{\text{CoT-SFT}} &= -\sum_{t=1}^{T_r} \log p(r_t \mid x, r_{<t})
                                   - \lambda \sum_{t=1}^{T_a} \log p(a_t \mid x, r, a_{<t})
\end{align}

where $r$ is the reasoning chain, $a$ is the final answer, and $\lambda = 0.5$ down-weights
the answer loss to encourage richer intermediate reasoning.

\subsection{Reinforcement Learning from Verifiable Rewards}

RLVR \cite{lightman2023lets, cobbe2021gsm8k} replaces human preference labels with
programmatically verifiable correctness signals on three domains:

\begin{itemize}
  \item \textbf{Mathematics}: symbolic equivalence checking of numeric and algebraic answers.
  \item \textbf{Code}: unit test execution with pass@k evaluation.
  \item \textbf{Logic}: SAT/SMT solver verification of formal claims.
\end{itemize}

The reward function is:

\begin{equation}
  r(y, y^*) = \mathbb{1}[\text{verify}(y) = y^*] - \alpha \cdot \frac{|y|}{L_{\max}}
\end{equation}

where the first term is 1 for correct verified answers and the second term penalizes unnecessary
verbosity with coefficient $\alpha = 0.02$ and maximum allowed length $L_{\max}$.

Group Relative Policy Optimization (GRPO) \cite{shao2024deepseekmath} is used as the RL
algorithm, sampling $G=8$ responses per prompt and computing advantages within the group to
reduce variance without a separate value network.

%% ─────────────────────────────────────────────────────────────────────────────
\section{Evaluation}

\subsection{Standard Benchmarks}

\begin{table}[H]
\centering
\caption{Zen-Pro benchmark results versus comparable frontier models.}
\label{tab:benchmarks}
\begin{tabular}{lcccc}
\toprule
\textbf{Benchmark} & \textbf{Zen-Pro (72B)} & \textbf{Competitor A (70B)} & \textbf{Competitor B (72B)} & \textbf{Competitor C (8$\times$7B)} \\
\midrule
MMLU (5-shot)          & \textbf{89.2} & 88.1 & 87.6 & 82.4 \\
MMLU-Pro               & \textbf{72.4} & 71.1 & 70.8 & 61.3 \\
ARC-Challenge          & 72.8 & 73.1 & 71.4 & 66.4 \\
HellaSwag              & 88.4 & 87.9 & 88.1 & 86.7 \\
WinoGrande             & 82.3 & 81.8 & 81.4 & 78.2 \\
\midrule
MATH (4-shot, CoT)     & \textbf{84.1} & 80.3 & 78.9 & 62.1 \\
GSM8K (8-shot, CoT)    & \textbf{93.4} & 92.1 & 91.8 & 87.3 \\
GPQA Diamond           & \textbf{71.8} & 68.4 & 66.9 & 51.2 \\
\midrule
HumanEval (pass@1)     & \textbf{91.3} & 88.7 & 87.4 & 74.8 \\
MBPP (pass@1)          & \textbf{84.6} & 82.3 & 80.9 & 71.4 \\
\midrule
MT-Bench (1-10)        & \textbf{9.1}  &  8.9 &  8.7 &  8.4 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Scientific Reasoning}

Zen-Pro shows strong performance on scientific reasoning benchmarks, reflecting the emphasis on
scientific literature and formal reasoning in the training curriculum.

\begin{table}[H]
\centering
\caption{Scientific reasoning benchmark results.}
\label{tab:science}
\begin{tabular}{lcc}
\toprule
\textbf{Benchmark} & \textbf{Zen-Pro (72B)} & \textbf{Prior Best (Open)} \\
\midrule
GPQA (Graduate-level)    & 71.8 & 68.4 \\
GPQA (Diamond subset)    & 64.3 & 61.7 \\
SciQ                     & 97.1 & 96.8 \\
ARC-Challenge            & 72.8 & 73.1 \\
MMLU-STEM subjects       & 91.4 & 89.2 \\
MedQA (USMLE)            & 82.6 & 80.1 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Long-Context Evaluation}

\begin{table}[H]
\centering
\caption{RULER long-context benchmark scores at multiple context lengths.}
\label{tab:longctx}
\begin{tabular}{lccccc}
\toprule
\textbf{Model} & \textbf{8K} & \textbf{16K} & \textbf{32K} & \textbf{64K} & \textbf{128K} \\
\midrule
Zen-Pro (72B)      & 96.2 & 94.8 & 92.1 & 88.3 & 82.7 \\
Competitor A (70B) & 95.8 & 92.3 & 86.4 & 71.2 & 52.3 \\
Competitor B (72B) & 94.9 & 91.7 & 85.1 & 68.4 & 48.7 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Coding Benchmarks}

\begin{table}[H]
\centering
\caption{Code generation benchmark results.}
\label{tab:code}
\begin{tabular}{lcc}
\toprule
\textbf{Benchmark} & \textbf{Zen-Pro (72B)} & \textbf{Competitor A (70B)} \\
\midrule
HumanEval (pass@1)         & 91.3 & 88.7 \\
HumanEval$+$ (pass@1)      & 87.1 & 84.2 \\
MBPP (pass@1)              & 84.6 & 82.3 \\
MBPP$+$ (pass@1)           & 79.3 & 76.8 \\
SWE-bench Verified (pass@1)& 38.7 & 34.1 \\
LiveCodeBench (3-month)    & 62.4 & 58.3 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Inference Efficiency}

\begin{table}[H]
\centering
\caption{Inference throughput on 4$\times$ A100-80GB (tensor parallel, BF16).}
\label{tab:inference}
\begin{tabular}{lcc}
\toprule
\textbf{Metric} & \textbf{8K context} & \textbf{128K context} \\
\midrule
Throughput (tok/s, batch=1)  & 42  & 18 \\
Throughput (tok/s, batch=8)  & 310 & 124 \\
TTFT P50 (ms, 4K prompt)     & 210 & 1{,}840 \\
GPU memory (weights+KV)      & 148 GB & 174 GB \\
\bottomrule
\end{tabular}
\end{table}

%% ─────────────────────────────────────────────────────────────────────────────
\section{Ablation Studies}

\subsection{Effect of RLVR}

Table~\ref{tab:ablation} compares Zen-Pro with RLVR against the SFT-only checkpoint.

\begin{table}[H]
\centering
\caption{Ablation: effect of RLVR post-training stage.}
\label{tab:ablation}
\begin{tabular}{lccc}
\toprule
\textbf{Model} & \textbf{MATH} & \textbf{HumanEval} & \textbf{GPQA} \\
\midrule
Zen-Pro SFT only   & 77.3 & 86.2 & 63.4 \\
Zen-Pro + RLVR     & 84.1 & 91.3 & 71.8 \\
\midrule
Improvement ($\Delta$) & $+$6.8 & $+$5.1 & $+$8.4 \\
\bottomrule
\end{tabular}
\end{table}

RLVR produces the largest gains on GPQA, reflecting that formal scientific reasoning benefits
most from verifiable reward signals where human preference annotation is difficult to obtain at
scale and quality.

\subsection{Effect of Long CoT Training}

\begin{table}[H]
\centering
\caption{Ablation: standard SFT vs.\ long chain-of-thought SFT.}
\label{tab:cot_ablation}
\begin{tabular}{lcc}
\toprule
\textbf{Model} & \textbf{MATH} & \textbf{AIME 2024} \\
\midrule
Standard SFT      & 71.4 & 23.3\% \\
Long CoT SFT      & 78.9 & 36.7\% \\
Long CoT + RLVR   & 84.1 & 53.3\% \\
\bottomrule
\end{tabular}
\end{table}

%% ─────────────────────────────────────────────────────────────────────────────
\section{Related Work}

Large-scale dense transformers at the 70B tier have been explored extensively following the
LLaMA 2 \cite{touvron2023llama2} release, with subsequent models pushing reasoning capability
through data quality improvements and specialized post-training. Reinforcement learning with
verifiable rewards traces to outcome-based reward modeling \cite{cobbe2021gsm8k} and was
prominently applied in mathematical reasoning work \cite{lightman2023lets, shao2024deepseekmath}.
Our GRPO implementation follows the group relative policy gradient approach, adapted for a mix
of math, code, and logic verification environments.

Long-context scaling via RoPE interpolation has been studied in multiple works
\cite{peng2023yarn, chen2023extending}, and the hybrid sliding-window / full-attention pattern
draws inspiration from Longformer \cite{beltagy2020longformer}. Our two-stage context extension
is a pragmatic variant of these approaches optimized for training stability.

%% ─────────────────────────────────────────────────────────────────────────────
\section{Limitations}

Zen-Pro at 72B requires significant compute for inference: at minimum 2$\times$ A100-80GB in
BF16 without quantization, limiting accessibility compared to smaller models. While RLVR
substantially improves mathematical and code reasoning, commonsense reasoning and open-ended
creative tasks rely entirely on SFT signal, leaving room for improvement. Long-context
performance degrades for retrieval tasks beyond 64K tokens, and generation quality on very
long outputs (>8K tokens) has not been systematically evaluated.

%% ─────────────────────────────────────────────────────────────────────────────
\section{Conclusion}

Zen-Pro represents the Zen family's high-performance tier, demonstrating that rigorous data
curation combined with verifiable-reward reinforcement learning yields frontier-level reasoning
capability at the 72B scale. The model achieves MMLU 89.2\%, MATH 84.1\%, HumanEval 91.3\%,
and GPQA 71.8\%, competitive with the strongest open-weight models while maintaining a 128K
context window for professional workflows. The RLVR training protocol generalizes well across
verifiable domains and is expected to underpin future Zen family post-training pipelines.

%% ─────────────────────────────────────────────────────────────────────────────
\begin{thebibliography}{99}

\bibitem{wei2022emergent}
J.~Wei et al., ``Emergent Abilities of Large Language Models,'' \textit{TMLR}, 2022.

\bibitem{srivastava2022beyond}
A.~Srivastava et al., ``Beyond the Imitation Game: Quantifying and Extrapolating the Capabilities
of Language Models,'' \textit{arXiv:2206.04615}, 2022.

\bibitem{beltagy2020longformer}
I.~Beltagy, M.~E.~Peters, and A.~Cohan, ``Longformer: The Long-Document Transformer,''
\textit{arXiv:2004.05150}, 2020.

\bibitem{lightman2023lets}
H.~Lightman et al., ``Let's Verify Step by Step,'' \textit{arXiv:2305.20050}, 2023.

\bibitem{cobbe2021gsm8k}
K.~Cobbe et al., ``Training Verifiers to Solve Math Word Problems,'' \textit{arXiv:2110.14168}, 2021.

\bibitem{shao2024deepseekmath}
Z.~Shao et al., ``DeepSeekMath: Pushing the Limits of Mathematical Reasoning in Open Language
Models,'' \textit{arXiv:2402.03300}, 2024.

\bibitem{peng2023yarn}
B.~Peng et al., ``YaRN: Efficient Context Window Extension of Large Language Models,''
\textit{arXiv:2309.00071}, 2023.

\bibitem{chen2023extending}
S.~Chen et al., ``Extending Context Window of Large Language Models via Positional Interpolation,''
\textit{arXiv:2306.15595}, 2023.

\bibitem{touvron2023llama2}
H.~Touvron et al., ``Llama 2: Open Foundation and Fine-Tuned Chat Models,''
\textit{arXiv:2307.09288}, 2023.

\bibitem{su2021rope}
J.~Su et al., ``RoFormer: Enhanced Transformer with Rotary Position Embedding,''
\textit{arXiv:2104.09864}, 2021.

\bibitem{ainslie2023gqa}
J.~Ainslie et al., ``GQA: Training Generalized Multi-Query Transformer Models,''
\textit{EMNLP}, 2023.

\bibitem{rafailov2023dpo}
R.~Rafailov et al., ``Direct Preference Optimization,'' \textit{NeurIPS}, 2023.

\bibitem{ouyang2022instructgpt}
L.~Ouyang et al., ``Training Language Models to Follow Instructions with Human Feedback,''
\textit{NeurIPS}, 2022.

\end{thebibliography}

\end{document}
