\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{color}
\usepackage{booktabs}
\usepackage{float}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{geometry}
\geometry{margin=1in}
\definecolor{zenblue}{RGB}{41,121,255}
\hypersetup{colorlinks=true,linkcolor=zenblue,urlcolor=zenblue,citecolor=zenblue}

\title{\textbf{Zen Training Methodology: From Pretraining to Deployment}\\
\large Technical Report v2025.03}
\author{Zen LM Research Team\\
\texttt{research@zenlm.org}}
\date{March 2025}

\begin{document}
\maketitle

\begin{abstract}
We present the complete training methodology for the Zen family of large language models,
spanning pretraining on 7 trillion tokens through supervised fine-tuning and reinforcement
learning from human feedback. Our approach introduces several key innovations: a hybrid
data mixing strategy that balances quality and diversity, perplexity-based quality filtering
that removes low-signal content, and a constitutional training regime that instills
safety properties without sacrificing capability. We report detailed training curves,
compute budget analysis, and ablation results demonstrating the contribution of each
methodological component. Models trained under this methodology achieve state-of-the-art
performance on standard benchmarks while maintaining robust safety properties and
deployment reliability. The Zen MoDE (Mixture of Distilled Experts) architecture
enables efficient scaling from 600M to 480B parameters under a unified training pipeline.
\end{abstract}

\section{Introduction}

Training frontier language models requires careful coordination of data curation,
architectural decisions, and optimization strategies across a multi-stage pipeline.
The Zen training methodology addresses the full lifecycle: from raw web data ingestion
through pretraining, instruction tuning, and alignment, to final deployment optimization.

Key challenges we address include:
\begin{itemize}
  \item \textbf{Data quality at scale}: Curating 7T tokens that balance breadth and quality
  \item \textbf{Stable large-scale optimization}: Preventing loss spikes and ensuring smooth convergence
  \item \textbf{Alignment without degradation}: Instilling safety properties while preserving capability
  \item \textbf{Compute efficiency}: Achieving optimal FLOPs allocation across model sizes
\end{itemize}

The Zen MoDE architecture introduces a Mixture of Distilled Experts paradigm, where
each expert specializes in semantic domains while sharing a unified vocabulary and
positional encoding scheme. This architectural choice significantly impacts training
dynamics and data mixing decisions.

\section{Background and Related Work}

\subsection{Scaling Laws}

The Chinchilla scaling laws~\cite{hoffmann2022chinchilla} established that optimal
training requires roughly 20 tokens per parameter. Subsequent work has shown that
inference-optimal training often favors over-training smaller models on more tokens.
The Zen family follows an inference-optimal strategy, training smaller models on
significantly more data than Chinchilla-optimal prescriptions.

\subsection{Data Curation}

Prior work on data curation has emphasized the importance of deduplication, quality
filtering, and domain mixing. C4~\cite{raffel2020c4}, The Pile~\cite{gao2021pile}, and
RedPajama~\cite{together2023redpajama} established key preprocessing pipelines. Zen
extends these with perplexity-based filtering and domain-aware mixing.

\subsection{Multi-Stage Training}

The standard paradigm of pretraining followed by instruction tuning (SFT) and RLHF
was established by InstructGPT~\cite{ouyang2022rlhf}. Zen introduces constitutional
training as an intermediate stage between SFT and RLHF, reducing human annotation
requirements while improving alignment quality.

\section{Data Curation Pipeline}

\subsection{Raw Data Collection}

The Zen pretraining corpus aggregates data from multiple sources:

\begin{table}[H]
\centering
\begin{tabular}{lrrr}
\toprule
\textbf{Source} & \textbf{Raw Size} & \textbf{After Filter} & \textbf{Mix Weight} \\
\midrule
Web (Common Crawl) & 68T tokens & 4.2T & 60\% \\
Books \& Literature & 400B tokens & 380B & 12\% \\
Scientific Papers & 150B tokens & 140B & 8\% \\
Code Repositories & 800B tokens & 650B & 12\% \\
Wikipedia + Encyclopedias & 80B tokens & 78B & 3\% \\
Curated Q\&A & 120B tokens & 110B & 5\% \\
\midrule
\textbf{Total} & \textbf{$\sim$70T} & \textbf{7T} & \textbf{100\%} \\
\bottomrule
\end{tabular}
\caption{Zen pretraining corpus composition after filtering.}
\end{table}

\subsection{Quality Filtering}

Our quality pipeline applies filters in sequence, each removing a fraction of documents:

\begin{algorithm}[H]
\caption{Zen Quality Filtering Pipeline}
\begin{algorithmic}[1]
\REQUIRE Raw document corpus $\mathcal{D}$
\ENSURE Filtered corpus $\mathcal{D}^*$
\STATE $\mathcal{D}_1 \leftarrow \text{LanguageID}(\mathcal{D})$, keep $p(\text{en}) > 0.65$
\STATE $\mathcal{D}_2 \leftarrow \text{URLFilter}(\mathcal{D}_1)$, remove spam/adult domains
\STATE $\mathcal{D}_3 \leftarrow \text{MinHashDedup}(\mathcal{D}_2, \text{ngram}=13, \text{jaccard}>0.8)$
\STATE $\mathcal{D}_4 \leftarrow \text{HeuristicFilter}(\mathcal{D}_3)$:
\STATE \quad Remove docs $< 100$ tokens or $> 100$K tokens
\STATE \quad Remove docs with symbol/word ratio $> 0.1$
\STATE \quad Remove docs with repeated n-gram ratio $> 0.3$
\STATE Train reference LM $\mathcal{M}_\text{ref}$ on high-quality seed corpus (200B tokens)
\STATE $\mathcal{D}_5 \leftarrow \{d \in \mathcal{D}_4 : \text{PPL}_{\mathcal{M}_\text{ref}}(d) < \tau_\text{ppl}\}$
\STATE $\tau_\text{ppl} \leftarrow$ 90th percentile of $\text{PPL}_{\mathcal{M}_\text{ref}}$ on $\mathcal{D}_4$
\STATE $\mathcal{D}^* \leftarrow \text{DomainSample}(\mathcal{D}_5, \text{weights}=W)$
\RETURN $\mathcal{D}^*$
\end{algorithmic}
\end{algorithm}

\subsubsection{Perplexity-Based Quality Filtering}

A key innovation is using a reference language model to score document quality.
We train a 1.3B parameter reference model on a manually curated seed corpus of
200B high-quality tokens. Documents scoring above the 90th percentile perplexity
threshold are removed, eliminating repetitive, incoherent, or template-generated content.

The perplexity filter reduces web content from 68T to 4.2T tokens (6.2\% retention),
while books and scientific papers show 95\%+ retention, confirming the filter's
discriminative power.

\subsection{Hybrid Data Mixing}

Static mixing ratios degrade performance as training progresses because the model
learns different domains at different rates. We employ adaptive mixing via a
domain loss monitor:

\begin{equation}
w_d^{(t+1)} = w_d^{(t)} \cdot \exp\left(\eta \cdot \frac{\mathcal{L}_d^{(t)} - \bar{\mathcal{L}}^{(t)}}{\sigma_\mathcal{L}^{(t)}}\right)
\end{equation}

where $w_d^{(t)}$ is the weight for domain $d$ at step $t$, $\mathcal{L}_d^{(t)}$ is
the per-domain validation loss, and $\eta = 0.01$ is the adaptation rate. Weights
are normalized after each update. This causes the optimizer to allocate more capacity
to domains where the model lags.

\section{Pretraining Setup}

\subsection{Architecture: Zen MoDE}

The Zen MoDE (Mixture of Distilled Experts) architecture replaces dense FFN layers
with a mixture of $E$ expert networks, activated sparsely:

\begin{equation}
\text{MoDE}(x) = \sum_{i=1}^{k} g_i(x) \cdot E_i(x), \quad g(x) = \text{TopK}(\text{softmax}(W_g x), k)
\end{equation}

where $k=2$ experts are selected per token from $E=64$ total experts. Expert
specialization emerges from domain-structured data: we observe that distinct experts
activate preferentially for code, mathematics, and natural language.

\subsection{Model Configurations}

\begin{table}[H]
\centering
\begin{tabular}{lrrrrr}
\toprule
\textbf{Model} & \textbf{Params} & \textbf{Layers} & \textbf{d\_model} & \textbf{Heads} & \textbf{Experts} \\
\midrule
Zen-600M & 600M & 24 & 1024 & 16 & -- \\
Zen-7B & 7B & 32 & 4096 & 32 & -- \\
Zen-32B & 32B & 64 & 7168 & 56 & -- \\
Zen-235B-MoE & 235B & 94 & 7168 & 56 & 128 \\
Zen-480B-MoE & 480B & 128 & 8192 & 64 & 256 \\
\bottomrule
\end{tabular}
\caption{Zen model family configurations.}
\end{table}

\subsection{Optimization}

All models are trained with AdamW ($\beta_1=0.9$, $\beta_2=0.95$, $\epsilon=10^{-8}$,
weight decay $\lambda=0.1$). Learning rate follows a warmup-cosine schedule:

\begin{equation}
\eta_t = \eta_{\max} \cdot \begin{cases} t / T_{\text{warm}} & t < T_{\text{warm}} \\ \frac{1}{2}\left(1 + \cos\left(\pi \cdot \frac{t - T_{\text{warm}}}{T - T_{\text{warm}}}\right)\right) & t \geq T_{\text{warm}} \end{cases}
\end{equation}

with $T_{\text{warm}} = 2000$ steps, $\eta_{\max}$ model-size dependent (see Table 3),
and final learning rate $\eta_{\min} = \eta_{\max}/10$.

\begin{table}[H]
\centering
\begin{tabular}{lrrr}
\toprule
\textbf{Model} & \textbf{Peak LR} & \textbf{Batch Size (tokens)} & \textbf{Train Steps} \\
\midrule
Zen-600M & $6\times10^{-4}$ & 2M & 3.5M \\
Zen-7B & $3\times10^{-4}$ & 4M & 1.75M \\
Zen-32B & $1\times10^{-4}$ & 8M & 875K \\
Zen-235B-MoE & $5\times10^{-5}$ & 16M & 437K \\
Zen-480B-MoE & $3\times10^{-5}$ & 32M & 219K \\
\bottomrule
\end{tabular}
\caption{Training hyperparameters per model size.}
\end{table}

\subsection{Compute Infrastructure}

Training runs on H100 SXM5 80GB clusters with NVLink interconnects. Pipeline
parallelism depth $P$, tensor parallelism $T$, and data parallelism $D$ are
configured per model size:

\begin{table}[H]
\centering
\begin{tabular}{lrrrr}
\toprule
\textbf{Model} & \textbf{GPUs} & \textbf{P} & \textbf{T} & \textbf{D} \\
\midrule
Zen-7B & 512 & 1 & 4 & 128 \\
Zen-32B & 1024 & 4 & 8 & 32 \\
Zen-235B-MoE & 2048 & 8 & 8 & 32 \\
Zen-480B-MoE & 4096 & 16 & 8 & 32 \\
\bottomrule
\end{tabular}
\caption{Parallelism strategies per model. MoE models use expert parallelism with EP=64.}
\end{table}

\section{Supervised Fine-Tuning}

\subsection{SFT Data Construction}

The SFT dataset contains 2.4M instruction-response pairs spanning:
multiturn conversation (35\%), coding tasks (25\%), mathematical reasoning (20\%),
long-form writing (12\%), and factual Q\&A (8\%). All pairs are human-validated
or model-generated with human verification.

\subsection{SFT Training Protocol}

SFT is conducted for 3 epochs with learning rate $5\times10^{-6}$, batch size 256,
and maximum sequence length 32K tokens. Only response tokens contribute to the
cross-entropy loss:

\begin{equation}
\mathcal{L}_\text{SFT} = -\sum_{t \in \text{response}} \log p_\theta(x_t \mid x_{<t})
\end{equation}

We apply NEFTune noise injection~\cite{jain2023neftune} with $\alpha=5$ to improve
generalization on open-ended tasks.

\section{Constitutional Training}

Between SFT and RLHF, we apply a constitutional training stage that teaches the model
to critique and revise its own outputs according to a set of principles:

\begin{enumerate}
  \item Generate response $r_0$ to prompt $p$
  \item Apply critique template: ``Identify ways the response violates principle $c_i$''
  \item Generate revision $r_1$ guided by the critique
  \item Train on $(p, r_1)$ pairs using SFT loss
\end{enumerate}

This reduces human annotation requirements for the subsequent RLHF stage by 40\%
while improving harmlessness scores by 12 points (see Section 7).

\section{RLHF Pipeline}

\subsection{Reward Model Training}

The reward model (RM) is initialized from the Zen-7B SFT checkpoint and trained on
400K human preference pairs. Given two responses $r_a, r_b$ to the same prompt,
the RM is trained to maximize the margin:

\begin{equation}
\mathcal{L}_\text{RM} = -\mathbb{E}_{(p,r_w,r_l)} \left[\log \sigma\left(r_\phi(p, r_w) - r_\phi(p, r_l)\right)\right]
\end{equation}

where $r_w$ is the preferred response and $r_l$ the rejected response.

\subsection{PPO Training}

Policy optimization uses Proximal Policy Optimization with the KL penalty:

\begin{equation}
\mathcal{L}_\text{PPO} = \mathbb{E}\left[r_\phi(p, r) - \beta \cdot \text{KL}\left[\pi_\theta(\cdot|p) \| \pi_{\text{ref}}(\cdot|p)\right]\right]
\end{equation}

$\beta = 0.04$ is tuned to balance helpfulness and proximity to the SFT policy.
We apply a reward clipping of $[-5, 5]$ and run 2 PPO epochs per batch.

\section{Experiments and Results}

\subsection{Training Curves}

Pretraining loss curves exhibit three distinct phases observed across all model sizes:
(1) rapid descent in the first 5\% of training as the model learns basic token statistics,
(2) steady decline through 90\% of training as domain knowledge accumulates,
(3) a slower final phase with diminishing returns signaling data saturation.

Loss spike prevention is achieved via gradient norm clipping at 1.0 and a loss
spike detector that halves the learning rate for 100 steps when $\mathcal{L}_t > 1.5 \cdot \bar{\mathcal{L}}_{t-100:t}$.

\subsection{Ablation Results}

\begin{table}[H]
\centering
\begin{tabular}{lcccc}
\toprule
\textbf{Configuration} & \textbf{MMLU} & \textbf{HumanEval} & \textbf{MATH} & \textbf{MT-Bench} \\
\midrule
Full pipeline & \textbf{85.3} & \textbf{78.2} & \textbf{67.4} & \textbf{8.62} \\
No PPL filter & 83.1 & 76.4 & 64.8 & 8.31 \\
Static mix (no adaptive) & 84.0 & 77.1 & 65.9 & 8.44 \\
No constitutional training & 84.9 & 78.0 & 67.1 & 8.41 \\
No NEFTune & 84.8 & 77.8 & 67.2 & 8.39 \\
\bottomrule
\end{tabular}
\caption{Ablation study on Zen-7B. All numbers are pass@1 or accuracy (\%).}
\end{table}

\subsection{Compute Budget Analysis}

We analyze the FLOPs-to-performance tradeoff by training model variants with
different compute budgets:

\begin{equation}
\text{Performance}(C) \approx A - B \cdot C^{-\alpha}
\end{equation}

where $C$ is the training compute in FLOPs, and we fit $A=89.1$, $B=142.3$,
$\alpha=0.095$ for MMLU performance of Zen-7B. This implies diminishing returns
beyond $3\times10^{23}$ FLOPs for this model size, motivating the switch to larger
models for higher capability targets.

\section{Analysis}

\subsection{Data Quality vs. Quantity}

Our experiments confirm that data quality dominates quantity beyond a threshold.
Doubling corpus size with the PPL filter disabled yields +0.8 MMLU points, while
halving corpus size with tighter filtering ($\tau_\text{ppl}$ at 80th percentile)
yields +1.1 MMLU points.

\subsection{Expert Specialization in MoDE}

Analysis of expert activation patterns in Zen-235B-MoE reveals clear domain
specialization: code-related tokens activate a consistent subset of 8--12 experts,
mathematics activates a partially overlapping set, and natural language distributes
more broadly. This specialization emerges organically from the data without
explicit routing supervision.

\subsection{Scaling Behavior}

Zen models follow a modified scaling law that accounts for MoE efficiency:

\begin{equation}
\mathcal{L}(N_\text{active}, D) = \frac{A}{N_\text{active}^\alpha} + \frac{B}{D^\beta}
\end{equation}

where $N_\text{active}$ is the number of active parameters per token (not total
parameters), $D$ is the dataset size, $\alpha=0.076$, $\beta=0.095$. MoE models
achieve lower loss at fixed active-parameter compute compared to dense models.

\section{Conclusion}

The Zen training methodology demonstrates that careful data curation, adaptive mixing,
constitutional training, and staged alignment produce models that are both capable and
safe. The perplexity-based filtering provides the largest single capability gain
among the methodological components we study. Constitutional training between SFT and
RLHF reduces annotation costs while improving alignment quality, making the pipeline
more scalable.

Future work will explore continual pretraining strategies, domain-specific fine-tuning
at scale, and automated data quality assessment to further reduce human annotation
requirements in the training pipeline.

\bibliographystyle{plain}
\begin{thebibliography}{99}
\bibitem{hoffmann2022chinchilla} Hoffmann et al. (2022). Training Compute-Optimal Large Language Models. \textit{NeurIPS}.
\bibitem{raffel2020c4} Raffel et al. (2020). Exploring the Limits of Transfer Learning with a Unified Text-to-Text Transformer. \textit{JMLR}.
\bibitem{gao2021pile} Gao et al. (2021). The Pile: An 800GB Dataset of Diverse Text for Language Modeling. \textit{arXiv:2101.00027}.
\bibitem{together2023redpajama} Together AI (2023). RedPajama: An Open Source Recipe to Reproduce LLaMA Training Dataset. \textit{GitHub}.
\bibitem{ouyang2022rlhf} Ouyang et al. (2022). Training language models to follow instructions with human feedback. \textit{NeurIPS}.
\bibitem{jain2023neftune} Jain et al. (2023). NEFTune: Noisy Embeddings Improve Instruction Finetuning. \textit{arXiv:2310.05914}.
\end{thebibliography}

\end{document}
