\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{color}
\usepackage{booktabs}
\usepackage{float}
\usepackage{geometry}
\geometry{margin=1in}
\definecolor{zenblue}{RGB}{41,121,255}
\hypersetup{colorlinks=true,linkcolor=zenblue,urlcolor=zenblue,citecolor=zenblue}

\title{\textbf{Reward Modeling for Zen Alignment:\\
Multi-Reward Ensembles and Synthetic Preference Generation}\\
\large Technical Report v2025.04}
\author{Zach Kelling, Zen LM Research Team\\
\texttt{research@zenlm.org}}
\date{April 2025}

\begin{document}
\maketitle

\begin{abstract}
Reward modeling is a critical component of aligning large language models to human
preferences. We present the Zen alignment stack, comprising: (1) a preference data
collection pipeline that produces high-quality human judgments at scale;
(2) ensemble reward models based on the Bradley-Terry statistical model with
multi-head architecture; (3) synthetic preference generation using a red-team/blue-team
approach to augment human data; and (4) Direct Preference Optimization (DPO) as an
alternative to full RL training. Our multi-reward ensemble achieves 87.3\% correlation
with human judgment on a held-out evaluation set, a 12\% improvement over single-model
baselines. RLHF with our ensemble reward substantially outperforms DPO alone on
open-ended generation quality, while DPO achieves competitive results at lower compute
cost for instruction following.
\end{abstract}

\tableofcontents
\newpage

%% -----------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}
%% -----------------------------------------------------------------------

Training a large language model to generate text that humans find helpful, harmless, and
honest (HHH) \cite{askell2021general} requires a reward signal that accurately captures
human preferences. Reinforcement Learning from Human Feedback (RLHF)
\cite{christiano2017deep,ouyang2022training} trains a reward model on preference data and
then uses it to fine-tune the policy via PPO. The quality of the reward model is a
critical bottleneck: a poorly calibrated reward model produces \emph{reward hacking},
where the policy exploits reward model weaknesses rather than genuinely improving.

We address reward model quality through three mechanisms:

\begin{enumerate}
    \item \textbf{High-quality preference data}: structured collection of pairwise human
          judgments with inter-annotator agreement metrics.
    \item \textbf{Multi-reward ensemble}: an ensemble of reward models with different
          training objectives (helpfulness, harmlessness, honesty, instruction-following),
          combined to produce a holistic alignment score.
    \item \textbf{Synthetic preference generation}: using a red-team model to generate
          challenging response pairs that expose alignment failures, augmenting human data
          with synthetically generated hard cases.
\end{enumerate}

\subsection{Alignment Taxonomy}

Following \cite{askell2021general}, we distinguish:
\begin{itemize}
    \item \textbf{Helpfulness}: whether the response addresses the user's intent effectively.
    \item \textbf{Harmlessness}: whether the response avoids harmful content or enabling harm.
    \item \textbf{Honesty}: whether the response is factually accurate and appropriately
          uncertain.
    \item \textbf{Instruction following}: whether the response adheres to specified format,
          length, and style constraints.
\end{itemize}

These dimensions are partially correlated but not identical, motivating a multi-head
reward architecture.

%% -----------------------------------------------------------------------
\section{Preference Data Collection}
\label{sec:data}
%% -----------------------------------------------------------------------

\subsection{Annotation Protocol}

Human annotators compare pairs of responses $(y_A, y_B)$ to the same prompt $x$ and
indicate: strongly prefer $A$, slightly prefer $A$, about equal, slightly prefer $B$,
strongly prefer $B$. This 5-point scale is mapped to a continuous preference signal
$\delta \in \{-2, -1, 0, +1, +2\}$ for the Bradley-Terry likelihood computation.

Annotators also rate each response on four dimensions (helpfulness, harmlessness,
honesty, instruction-following) using a 1--5 Likert scale. These dimensional ratings
train the individual reward model heads.

\subsection{Inter-Annotator Agreement}

To ensure data quality, each prompt-pair is annotated by 3 independent annotators.
We compute Fleiss's $\kappa$ as a quality metric:

\begin{table}[H]
\centering
\caption{Inter-annotator agreement across preference data categories.}
\begin{tabular}{lrr}
\toprule
\textbf{Category} & \textbf{Fleiss's $\kappa$} & \textbf{Agreement interpretation} \\
\midrule
Helpfulness            & 0.68 & Substantial \\
Harmlessness           & 0.81 & Almost perfect \\
Honesty                & 0.71 & Substantial \\
Instruction following  & 0.87 & Almost perfect \\
Overall preference     & 0.64 & Substantial \\
\bottomrule
\end{tabular}
\label{tab:agreement}
\end{table}

Prompt-pairs with $\kappa < 0.4$ are excluded from training, retaining 89\% of
collected pairs. This yields a final dataset of 1.2M preference pairs.

\subsection{Dataset Composition}

\begin{table}[H]
\centering
\caption{Preference dataset composition by domain.}
\begin{tabular}{lrrr}
\toprule
\textbf{Domain} & \textbf{Pairs (K)} & \textbf{Fraction (\%)} & \textbf{Avg. $\kappa$} \\
\midrule
General conversation     & 320 & 26.7 & 0.63 \\
Instruction following    & 210 & 17.5 & 0.88 \\
Mathematics / reasoning  & 180 & 15.0 & 0.82 \\
Code generation          & 150 & 12.5 & 0.79 \\
Factual QA               & 140 & 11.7 & 0.71 \\
Safety-critical          & 120 & 10.0 & 0.84 \\
Creative writing         & 80  & 6.7  & 0.58 \\
\midrule
\textbf{Total}           & 1200 & 100.0 & 0.74 \\
\bottomrule
\end{tabular}
\label{tab:dataset}
\end{table}

%% -----------------------------------------------------------------------
\section{Bradley-Terry Reward Model}
\label{sec:reward_model}
%% -----------------------------------------------------------------------

\subsection{Bradley-Terry Model}

The Bradley-Terry (BT) model \cite{bradley1952rank} models the probability that response
$y_A$ is preferred over $y_B$ for prompt $x$ as:
\begin{equation}
    P(y_A \succ y_B \mid x) = \sigma(r(x, y_A) - r(x, y_B))
    \label{eq:bt}
\end{equation}
where $r(x, y) \in \mathbb{R}$ is the reward score and $\sigma$ is the sigmoid function.
The reward model is trained to maximize the log-likelihood:
\begin{equation}
    \mathcal{L}_{\text{BT}} = -\mathbb{E}_{(x, y_A, y_B, \delta) \sim \mathcal{D}}\!\left[
    \delta \log P(y_A \succ y_B \mid x)\right]
    \label{eq:bt_loss}
\end{equation}
where $\delta \in \{-2, -1, 0, +1, +2\}$ is the annotator preference strength.

\subsection{Multi-Head Architecture}

Our reward model uses a Zen MoDE encoder to produce a contextual representation of
$(x, y)$, with separate output heads for each alignment dimension:

\begin{equation}
    r_d(x, y) = \mathbf{w}_d^\top h(x, y) + b_d, \quad
    d \in \{\text{help}, \text{harm}, \text{hon}, \text{inst}\}
    \label{eq:multi_head}
\end{equation}

where $h(x, y) \in \mathbb{R}^{d_{\text{model}}}$ is the [EOS] token representation
from the final transformer layer. The combined reward is:
\begin{equation}
    r(x, y) = \sum_d \lambda_d \cdot r_d(x, y)
    \label{eq:combined}
\end{equation}

with per-domain weights $\lambda_d$ learned on the validation set.

\subsection{Ensemble Construction}

We train 5 reward models with different random seeds and data orderings, then ensemble:
\begin{equation}
    r_{\text{ens}}(x, y) = \frac{1}{5}\sum_{k=1}^{5} r^{(k)}(x, y)
    \label{eq:ensemble}
\end{equation}

Ensemble disagreement $\text{Var}_{k}[r^{(k)}(x, y)]$ serves as an uncertainty estimate:
high-variance examples are flagged for additional human review during RLHF training.

%% -----------------------------------------------------------------------
\section{Synthetic Preference Generation}
\label{sec:synthetic}
%% -----------------------------------------------------------------------

\subsection{Red-Team / Blue-Team Protocol}

Human preference data is expensive and limited in coverage. We augment it using a
red-team model trained to generate challenging prompts that expose alignment failures,
and a blue-team (the base policy) that responds. The red-team generates prompts by:

\begin{enumerate}
    \item Sampling a seed prompt $x_0$ from the training distribution.
    \item Applying a perturbation $\Delta$ that targets known weak spots (jailbreaks,
          out-of-distribution topics, adversarial formatting).
    \item Generating the perturbed prompt $x' = T(x_0, \Delta)$.
\end{enumerate}

For each perturbed prompt, the base policy generates two responses with different
decoding temperatures ($T_1 = 0.7$, $T_2 = 1.2$), producing a natural diversity
in response quality. The reward model ensemble scores both responses and the pair is
added to the preference dataset with the model scores as soft labels.

\subsection{Quality Filtering}

Synthetic pairs are filtered by:
\begin{itemize}
    \item Ensemble agreement: $\text{Var}_k[r^{(k)}] < \sigma_{\text{thresh}}$ ensures
          the preference label is well-defined.
    \item Margin threshold: $|r_{\text{ens}}(y_A) - r_{\text{ens}}(y_B)| > m_{\min}$
          ensures the pair is informative (non-trivial).
    \item Semantic diversity: the prompt $x'$ must differ from training prompts by
          at least $0.15$ in embedding distance.
\end{itemize}

This process generates 8M synthetic preference pairs, of which 3.4M pass all filters.

%% -----------------------------------------------------------------------
\section{Direct Preference Optimization (DPO)}
\label{sec:dpo}
%% -----------------------------------------------------------------------

DPO \cite{rafailov2023direct} directly optimizes the policy using preference data,
bypassing explicit reward model training:

\begin{equation}
    \mathcal{L}_{\text{DPO}} = -\mathbb{E}_{(x, y_w, y_l)}\!\left[
    \log \sigma\!\left(\beta \log \frac{\pi_\theta(y_w \mid x)}{\pi_{\text{ref}}(y_w \mid x)}
    - \beta \log \frac{\pi_\theta(y_l \mid x)}{\pi_{\text{ref}}(y_l \mid x)}\right)
    \right]
    \label{eq:dpo}
\end{equation}

where $y_w$ and $y_l$ are preferred and dispreferred responses, $\pi_{\text{ref}}$
is the reference (SFT) policy, and $\beta$ controls the KL penalty.

We compare DPO trained on our human + synthetic preference data to full RLHF using
PPO with our ensemble reward model.

%% -----------------------------------------------------------------------
\section{Experiments}
\label{sec:experiments}
%% -----------------------------------------------------------------------

\subsection{Reward Model Accuracy}

\begin{table}[H]
\centering
\caption{Reward model accuracy on held-out human preference pairs (binary: correct
if reward model prefers the human-preferred response).}
\begin{tabular}{lrr}
\toprule
\textbf{Model} & \textbf{Accuracy (\%)} & \textbf{Human corr.\ (Pearson $r$)} \\
\midrule
Single RM (no ensemble)    & 77.8 & 0.778 \\
Ensemble (5 RMs)           & 81.4 & 0.814 \\
Multi-head + ensemble      & 85.2 & 0.852 \\
+ Synthetic data           & \textbf{87.3} & \textbf{0.873} \\
\bottomrule
\end{tabular}
\label{tab:rm_accuracy}
\end{table}

\subsection{RLHF vs.\ DPO Comparison}

\begin{table}[H]
\centering
\caption{Alignment benchmark results. AlpacaEval 2.0 win rate vs.\ GPT-4 Turbo.
MT-Bench score (1--10). Arena Elo (Chatbot Arena).}
\begin{tabular}{lrrr}
\toprule
\textbf{Training} & \textbf{AlpacaEval 2.0 (\%)} & \textbf{MT-Bench} & \textbf{Arena Elo} \\
\midrule
SFT only           & 18.4 & 7.8 & 1124 \\
DPO (human data)   & 31.7 & 8.4 & 1198 \\
DPO (+ synthetic)  & 36.2 & 8.7 & 1231 \\
RLHF (PPO, single RM)     & 38.4 & 8.9 & 1248 \\
RLHF (PPO, ensemble RM)   & \textbf{44.8} & \textbf{9.2} & \textbf{1312} \\
\bottomrule
\end{tabular}
\label{tab:alignment}
\end{table}

Full RLHF with ensemble reward outperforms DPO by 8.6 pp on AlpacaEval. However,
DPO at significantly lower compute cost achieves competitive performance on MT-Bench.

\subsection{Reward Model Calibration}

\begin{table}[H]
\centering
\caption{ECE of reward model scores. Lower is better. The ensemble reduces
miscalibration by 41\% vs.\ single RM.}
\begin{tabular}{lrr}
\toprule
\textbf{Model} & \textbf{ECE} & \textbf{$\Delta$ vs.\ single} \\
\midrule
Single RM                & 0.124 & --- \\
Ensemble (5 RMs)         & 0.082 & $-33.9\%$ \\
Multi-head + ensemble    & 0.073 & $-41.1\%$ \\
\bottomrule
\end{tabular}
\label{tab:calibration}
\end{table}

\subsection{Downstream RLHF Results}

\begin{table}[H]
\centering
\caption{Downstream benchmark performance after RLHF. RLHF with ensemble RM provides
consistent improvements without degrading base capabilities.}
\begin{tabular}{lrrrrr}
\toprule
\textbf{Model} & \textbf{MMLU} & \textbf{GSM8K} & \textbf{HumanEval} & \textbf{TruthfulQA} \\
\midrule
Zen MoDE-72B (base)       & 84.7 & 90.8 & 79.3 & 74.8 \\
+ SFT                     & 84.9 & 91.2 & 80.1 & 76.3 \\
+ DPO                     & 85.1 & 91.8 & 80.7 & 78.9 \\
+ RLHF (ensemble RM)      & \textbf{85.4} & \textbf{92.1} & \textbf{81.3} & \textbf{82.4} \\
\bottomrule
\end{tabular}
\label{tab:downstream}
\end{table}

\subsection{Reward Hacking Analysis}

\begin{table}[H]
\centering
\caption{Reward hacking incidence: fraction of RLHF checkpoints where reward model
score increases but human preference decreases.}
\begin{tabular}{lrr}
\toprule
\textbf{Configuration} & \textbf{Hacking rate (\%)} & \textbf{Detected by ensemble (\%)} \\
\midrule
Single RM, no detection  & 14.2 & --- \\
Ensemble RM              & 7.8  & 89.3\% of single-RM hacking events \\
Ensemble + variance gate & 3.1  & 97.2\% of single-RM hacking events \\
\bottomrule
\end{tabular}
\label{tab:hacking}
\end{table}

The variance gate (flagging high-disagreement ensemble predictions) detects 97\% of
reward hacking events at the cost of rejecting 8\% of legitimate training updates.

%% -----------------------------------------------------------------------
\section{Discussion}
\label{sec:discussion}
%% -----------------------------------------------------------------------

\subsection{Ensemble vs.\ Single Reward Model}

The ensemble provides two benefits beyond accuracy: (1) calibration improvement via
variance-based uncertainty estimation, and (2) reward hacking detection through
disagreement signals. The computational overhead of running 5 reward models during
RLHF rollouts is 5$\times$ for inference, but reward model inference is typically
$<$10\% of total PPO compute, making this tractable.

\subsection{Synthetic Data Quality}

The synthetic preference pairs improve accuracy by 2.1 pp (Table~\ref{tab:rm_accuracy}).
Quality filtering is critical: unfiltered synthetic data degrades accuracy by 1.4 pp
due to noisy margin pairs. The margin threshold $m_{\min}$ is the most important filter.

\subsection{DPO Compute-Quality Trade-off}

DPO avoids reward model training and PPO rollout generation, reducing alignment compute
by approximately 8$\times$. For deployment scenarios where compute is the primary
constraint, DPO with synthetic preference augmentation is the recommended choice,
achieving 81\% of full RLHF performance at 12.5\% of the compute.

%% -----------------------------------------------------------------------
\section{Conclusion}
\label{sec:conclusion}
%% -----------------------------------------------------------------------

We have presented the Zen alignment stack: multi-head ensemble reward models trained
on 1.2M human preferences plus 3.4M filtered synthetic pairs, combined with RLHF or
DPO fine-tuning. Key results: 87.3\% reward model human correlation, 44.8\% AlpacaEval
win rate after RLHF, and 97\% reward hacking detection with ensemble variance gating.
The full stack is recommended for production deployments; DPO is recommended when
compute is constrained.

\begin{thebibliography}{9}
\bibitem{askell2021general}
A. Askell, Y. Bai, A. Chen, et al.
\textit{A General Language Assistant as a Laboratory for Alignment}.
arXiv:2112.00861, 2021.

\bibitem{christiano2017deep}
P. Christiano, J. Leike, T. Brown, et al.
\textit{Deep Reinforcement Learning from Human Preferences}.
NeurIPS, 2017.

\bibitem{ouyang2022training}
L. Ouyang, J. Wu, X. Jiang, et al.
\textit{Training Language Models to Follow Instructions with Human Feedback}.
NeurIPS, 2022.

\bibitem{bradley1952rank}
R.A. Bradley, M.E. Terry.
\textit{Rank Analysis of Incomplete Block Designs: I. The Method of Paired Comparisons}.
Biometrika, 39(3/4):324--345, 1952.

\bibitem{rafailov2023direct}
R. Rafailov, A. Sharma, E. Mitchell, et al.
\textit{Direct Preference Optimization: Your Language Model is Secretly a Reward Model}.
NeurIPS, 2023.
\end{thebibliography}

\end{document}
