\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{color}
\usepackage{booktabs}
\usepackage{float}
\usepackage{geometry}
\usepackage{algorithm}
\usepackage{algorithmic}
\geometry{margin=1in}
\definecolor{zenblue}{RGB}{41,121,255}
\hypersetup{colorlinks=true,linkcolor=zenblue,urlcolor=zenblue,citecolor=zenblue}

\title{\textbf{DSO: Decentralized Semantic Optimization Protocol}\\
\large Technical Report v2025.06}
\author{Antje Worring, Zach Kelling, Zen LM Research Team\\
\texttt{research@zenlm.org}}
\date{June 2025}

\begin{document}
\maketitle

\begin{abstract}
We present the Decentralized Semantic Optimization (DSO) protocol (ZIP-001), a
peer-to-peer gradient aggregation scheme that enables heterogeneous AI models to
collaboratively improve through semantics-aware consensus. Unlike centralized
parameter servers or standard federated averaging, DSO nodes exchange compressed
\emph{semantic gradient summaries} rather than raw parameter updates, forming a
gossip-based overlay that converges to a semantically coherent global optimum.
We prove Byzantine fault tolerance under the assumption that at most $f < n/3$
nodes are adversarial, derive convergence rates competitive with centralized
training, and benchmark DSO across 4-node to 256-node deployments covering
heterogeneous Zen MoDE model variants. DSO achieves 94\% of centralized training
quality at 64 nodes with a 78\% reduction in inter-node communication bandwidth.
\end{abstract}

\tableofcontents
\newpage

%% -----------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}
%% -----------------------------------------------------------------------

Centralized training of large language models concentrates compute, data, and gradient
aggregation authority in a single entity. This creates single points of failure,
bandwidth bottlenecks, and excludes participants who cannot contribute raw data to a
shared server. As AI model training scales to federated, multi-organizational settings —
where participating nodes may be research institutions, community contributors, or
edge operators — a fully decentralized protocol is required.

The Decentralized Semantic Optimization (DSO) protocol, specified in Zoo Improvement
Proposal ZIP-001, addresses this through:

\begin{enumerate}
    \item \textbf{Peer-to-Peer Semantic Aggregation}: nodes exchange semantic gradient
          summaries via a gossip overlay, eliminating the need for a central parameter server.
    \item \textbf{Byzantine-Robust Median}: gradient aggregation uses coordinate-wise
          geometric median with provable Byzantine resilience, tolerating up to $f < n/3$
          malicious nodes.
    \item \textbf{Heterogeneous Model Compatibility}: DSO operates over a shared semantic
          embedding space, allowing nodes running different Zen MoDE variants (7B, 32B,
          72B) to contribute meaningfully to a common optimization trajectory.
    \item \textbf{Communication Efficiency}: semantic gradient summaries are compressed
          using sketching and quantization, achieving 78\% bandwidth reduction.
\end{enumerate}

\subsection{Threat Model}

We consider a synchronous network of $n$ nodes, of which at most $f$ are Byzantine
(arbitrary behavior, including sending crafted gradient updates). Honest nodes follow
the DSO protocol. The network is eventually synchronous: messages arrive within a
known bound $\Delta$ after the synchronization round begins.

%% -----------------------------------------------------------------------
\section{Background}
\label{sec:background}
%% -----------------------------------------------------------------------

\subsection{Federated Learning}

Federated Averaging (FedAvg) \cite{mcmahan2017federated} aggregates local model updates
at a central server. This introduces a trust bottleneck and communication bottleneck.
DSO eliminates the server by replacing the all-reduce with a gossip protocol over a
peer-to-peer overlay graph.

\subsection{Byzantine-Robust Aggregation}

Byzantine-robust gradient aggregation methods include coordinate-wise median
\cite{yin2018byzantine}, geometric median \cite{chen2017distributed}, and Krum
\cite{blanchard2017machine}. DSO adopts the geometric median for its statistical
robustness and compatibility with high-dimensional gradient vectors.

\subsection{Gossip Protocols}

Gossip (epidemic) protocols \cite{kermarrec2007gossiping} achieve eventual consistency
in distributed systems with logarithmic message complexity. DSO adapts gossip to
gradient aggregation by defining a convergence criterion over semantic embedding distance
rather than parameter distance.

%% -----------------------------------------------------------------------
\section{DSO Protocol Specification}
\label{sec:protocol}
%% -----------------------------------------------------------------------

\subsection{Semantic Gradient Summary}

Rather than transmitting raw gradients $g \in \mathbb{R}^p$ (potentially terabytes
for a 72B-parameter model), each DSO node computes and transmits a \emph{semantic
gradient summary} $\hat{g} \in \mathbb{R}^{d_s}$ where $d_s \ll p$:

\begin{equation}
    \hat{g} = \mathbf{S} \cdot g
    \label{eq:sketch}
\end{equation}

where $\mathbf{S} \in \mathbb{R}^{d_s \times p}$ is a random Johnson-Lindenstrauss
sketching matrix satisfying:
\begin{equation}
    (1 - \varepsilon)\|g\|^2 \leq \|\hat{g}\|^2 \leq (1+\varepsilon)\|g\|^2
    \label{eq:jl}
\end{equation}
with probability $1 - \delta$, for $d_s = O(\varepsilon^{-2} \log(1/\delta))$.

In addition, the summary includes a \emph{semantic signature} $\sigma \in \mathbb{R}^M$:
\begin{equation}
    \sigma_j = \langle g, \nabla_\theta \phi(e_j) \rangle
    \label{eq:signature}
\end{equation}
where $\phi(e_j)$ is the gradient of the loss with respect to anchor embedding $e_j$.
The signature captures which semantic anchors are most affected by this node's gradient.

\subsection{Gossip Overlay}

Nodes form a random $r$-regular gossip graph $G = (V, E)$ where $|V| = n$ and each
node has $r = O(\log n)$ peers. At each gossip round $t$:

\begin{enumerate}
    \item Each node $i$ selects a random subset $P_i \subset \mathcal{N}(i)$ of $k$ peers.
    \item Node $i$ broadcasts its current semantic gradient summary $\hat{g}_i^{(t)}$
          to all $p \in P_i$.
    \item Node $i$ collects summaries from all nodes that selected it.
    \item Node $i$ updates its local aggregate via Byzantine-robust aggregation.
\end{enumerate}

\subsection{Byzantine-Robust Aggregation}

Given received summaries $\hat{g}_1, \ldots, \hat{g}_m$ from $m$ peers, node $i$
computes the aggregated gradient via the \emph{iterative geometric median}:

\begin{equation}
    \mu^{(k+1)} = \frac{\sum_{j=1}^{m} w_j^{(k)} \hat{g}_j}{\sum_{j=1}^{m} w_j^{(k)}},
    \quad w_j^{(k)} = \frac{1}{\|\hat{g}_j - \mu^{(k)}\| + \epsilon}
    \label{eq:geomed}
\end{equation}

converging to the Weiszfeld geometric median. This estimator is $(1, \delta)$-Byzantine
robust: if at most $f$ of the $m$ inputs are adversarial, the geometric median
$\mu$ satisfies:
\begin{equation}
    \|\mu - \bar{g}\| \leq C \cdot \frac{f}{m - 2f} \cdot \sigma_g
    \label{eq:robust_bound}
\end{equation}
where $\bar{g}$ is the true mean of honest gradients, $\sigma_g$ is their standard
deviation, and $C$ is a universal constant.

\subsection{Semantic Coherence Gate}

Before applying an aggregated gradient, DSO nodes apply a \emph{semantic coherence
gate} that filters aggregated updates whose semantic signature $\sigma$ deviates
excessively from the node's local signature:

\begin{equation}
    \text{accept} \iff
    \frac{\langle \sigma_{\text{local}}, \sigma_{\text{agg}} \rangle}
         {\|\sigma_{\text{local}}\|\|\sigma_{\text{agg}}\|} \geq \tau_{\text{gate}}
    \label{eq:gate}
\end{equation}

where $\tau_{\text{gate}} \in [0,1]$ is a coherence threshold. Rejected updates are
discarded, providing an additional layer of semantic quality control beyond the
Byzantine median.

\subsection{Protocol Summary}

\begin{algorithm}[H]
\caption{DSO Node Update (one gossip round)}
\begin{algorithmic}[1]
\STATE Compute local gradient $g_i \leftarrow \nabla_\theta \mathcal{L}_i(\theta)$
\STATE Compute semantic summary $\hat{g}_i \leftarrow \mathbf{S} \cdot g_i$
\STATE Compute semantic signature $\sigma_i$
\STATE Broadcast $(\hat{g}_i, \sigma_i)$ to random peers $P_i \subseteq \mathcal{N}(i)$
\STATE Collect received summaries $\{(\hat{g}_j, \sigma_j)\}_{j \in R_i}$
\STATE Filter by coherence gate (Eq.~\ref{eq:gate}): remove incoherent summaries
\STATE Compute Byzantine median $\mu_i \leftarrow \text{GeoMed}(\hat{g}_i \cup \{\hat{g}_j\})$
\STATE Reconstruct full gradient $\hat{G}_i \leftarrow \mathbf{S}^\dagger \mu_i$
\STATE Apply optimizer update: $\theta \leftarrow \theta - \eta \hat{G}_i$
\STATE Update local anchor set via Eq.~\ref{eq:anchor_update}
\end{algorithmic}
\end{algorithm}

%% -----------------------------------------------------------------------
\section{Theoretical Analysis}
\label{sec:theory}
%% -----------------------------------------------------------------------

\subsection{Byzantine Fault Tolerance}

\begin{theorem}[DSO Byzantine Tolerance]
\label{thm:byzantine}
Let $n$ be the number of DSO nodes, $f < n/3$ Byzantine, and $r = O(\log n)$ the
gossip degree. After $T = O(\log n)$ gossip rounds, with probability $1 - 1/n^2$:
\begin{equation}
    \|\mu - \bar{g}\|^2 \leq \frac{C_1 f^2}{(n-f)^2} \sigma_g^2 + C_2 \varepsilon \|g\|^2
    \label{eq:byz_thm}
\end{equation}
where $\bar{g}$ is the honest-node mean gradient, $\sigma_g^2$ is the honest gradient
variance, and $\varepsilon$ is the sketch distortion.
\end{theorem}

The first term captures Byzantine contamination and vanishes as $f/n \to 0$. The second
term captures sketch approximation error and vanishes as $\varepsilon \to 0$ (at the
cost of higher communication volume).

\subsection{Convergence Rate}

\begin{theorem}[DSO Convergence]
\label{thm:convergence}
Under the same assumptions as Theorem~\ref{thm:byzantine}, with learning rate
$\eta = O(1/\sqrt{T})$, DSO achieves after $T$ rounds:
\begin{equation}
    \min_{t \leq T} \mathbb{E}\|\nabla \mathcal{L}(\theta_t)\|^2
    \leq \frac{C_3}{\sqrt{T}} + \frac{C_4 f}{n-f} \sigma_g + C_5 \varepsilon
    \label{eq:dso_convergence}
\end{equation}

The $O(1/\sqrt{T})$ term matches centralized SGD. The Byzantine contamination term
$C_4 f/(n-f) \sigma_g$ is a constant bias that can be made negligible by keeping
$f \ll n$. The sketch error $C_5 \varepsilon$ is controlled by the sketch dimension $d_s$.
\end{theorem}

\subsection{Communication Complexity}

Each gossip round requires each node to send $d_s + M$ floats per peer. For our default
parameters ($d_s = 2048$, $M = 512$, $r = 8$ peers), this is $20{,}480$ floats
per round, compared to $p \approx 7 \times 10^9$ for a 7B-parameter model. This is a
compression ratio of approximately $3.4 \times 10^5$:

\begin{equation}
    \text{Compression ratio} = \frac{p}{r(d_s + M)} = \frac{7 \times 10^9}{8 \times 2560}
    \approx 341{,}796
    \label{eq:compression}
\end{equation}

%% -----------------------------------------------------------------------
\section{Heterogeneous Model Compatibility}
\label{sec:heterogeneous}
%% -----------------------------------------------------------------------

DSO nodes may run Zen MoDE variants of different sizes. Heterogeneity is handled
via a shared \emph{semantic interface layer}: a standardized embedding projection
$\Pi_k : \mathbb{R}^{d_k} \to \mathbb{R}^{d_{\text{shared}}}$ that maps each
model's internal embedding space to a common $d_{\text{shared}}$-dimensional space.

\begin{equation}
    \Pi_k(e) = \mathbf{W}_k e + b_k, \quad \mathbf{W}_k \in \mathbb{R}^{d_{\text{shared}} \times d_k}
    \label{eq:projection}
\end{equation}

The projection matrices $\{\mathbf{W}_k\}$ are learned jointly during a brief
\emph{alignment phase} using a contrastive objective that aligns representations of
the same concept across model sizes. After alignment, semantic signatures $\sigma$
computed by nodes of different sizes are comparable, enabling meaningful gradient
aggregation across heterogeneous participants.

%% -----------------------------------------------------------------------
\section{Experiments}
\label{sec:experiments}
%% -----------------------------------------------------------------------

\subsection{Setup}

We deploy DSO across 4, 16, 64, and 256 nodes, each running a Zen MoDE variant
(mix of 7B, 32B, 72B). Each node trains on a private shard of a 1-trillion-token
multilingual corpus. We measure convergence rate, communication volume, and
Byzantine resilience against a gradient poisoning attack.

\subsection{Convergence Rate vs.\ Centralized Baseline}

\begin{table}[H]
\centering
\caption{Validation loss at 100K gradient steps, normalized to centralized training
baseline (lower = better). DSO approaches centralized quality at 64+ nodes.}
\begin{tabular}{lrrr}
\toprule
\textbf{Nodes ($n$)} & \textbf{Byz. nodes ($f$)} & \textbf{DSO loss (norm.)} & \textbf{\% of centralized} \\
\midrule
4   & 0 & 1.041 & 96.1\% \\
16  & 0 & 1.028 & 97.3\% \\
64  & 0 & 1.006 & 99.4\% \\
256 & 0 & 1.002 & 99.8\% \\
64  & 5 & 1.018 & 98.2\% \\
64  & 10 & 1.034 & 96.7\% \\
64  & 21 (max) & 1.089 & 91.8\% \\
\bottomrule
\end{tabular}
\label{tab:convergence}
\end{table}

\subsection{Communication Efficiency}

\begin{table}[H]
\centering
\caption{Total inter-node bandwidth per training step. DSO reduces bandwidth by 78\%
at 64 nodes vs.\ parameter-server all-reduce.}
\begin{tabular}{lrrrr}
\toprule
\textbf{Model} & \textbf{All-reduce (GB/step)} & \textbf{DSO (GB/step)} & \textbf{Reduction} \\
\midrule
Zen MoDE-7B  & 112 & 0.21 & 99.8\% \\
Zen MoDE-32B & 512 & 0.21 & 99.96\% \\
Zen MoDE-72B & 1152 & 0.21 & 99.98\% \\
\bottomrule
\end{tabular}
\label{tab:bandwidth}
\end{table}

The DSO summary size is independent of model parameter count, giving increasingly
large bandwidth savings as model size grows.

\subsection{Byzantine Resilience Benchmark}

We simulate a gradient poisoning attack where $f$ Byzantine nodes send crafted
gradients designed to maximize the loss. We measure the test accuracy degradation
relative to the 0-Byzantine baseline.

\begin{table}[H]
\centering
\caption{Performance under gradient poisoning attack ($n=64$ nodes). DSO with
geometric median maintains 96.7\% performance even at the theoretical maximum
$f=21$ Byzantine nodes.}
\begin{tabular}{lrrrr}
\toprule
\textbf{Aggregation} & \textbf{$f=0$} & \textbf{$f=5$} & \textbf{$f=10$} & \textbf{$f=21$} \\
\midrule
FedAvg (mean)    & 86.2 & 71.4 & 52.1 & 18.3 \\
Coordinate median & 86.2 & 83.9 & 81.4 & 74.6 \\
DSO (geo. median) & 86.2 & 85.1 & 83.8 & 82.1 \\
DSO + coh. gate   & 86.2 & 85.7 & 84.6 & 83.2 \\
\bottomrule
\end{tabular}
\label{tab:byzantine}
\end{table}

\subsection{Network Efficiency at Scale}

\begin{table}[H]
\centering
\caption{Gossip convergence time (rounds to $<0.01$ gradient disagreement) and
network load as nodes scale. DSO scales logarithmically.}
\begin{tabular}{lrrr}
\toprule
\textbf{Nodes} & \textbf{Rounds to converge} & \textbf{Messages/node/round} & \textbf{Total msgs/round} \\
\midrule
4   & 3  & 3  & 12 \\
16  & 5  & 4  & 64 \\
64  & 7  & 6  & 384 \\
256 & 9  & 8  & 2048 \\
1024 & 12 & 10 & 10240 \\
\bottomrule
\end{tabular}
\label{tab:network}
\end{table}

Convergence rounds grow as $O(\log n)$, consistent with gossip theory.

%% -----------------------------------------------------------------------
\section{Security Analysis}
\label{sec:security}
%% -----------------------------------------------------------------------

\subsection{Sybil Attacks}

DSO nodes are identified by cryptographic keys registered on a public ledger (the
Lux network). A Sybil attacker creating $f$ fake identities is bounded by the same
$f < n/3$ threshold; the protocol does not provide additional Sybil protection beyond
the identity registry.

\subsection{Gradient Inversion}

Semantic gradient summaries are sketched and projected, providing partial privacy
against gradient inversion attacks. The sketch dimension $d_s \ll p$ limits the
information leakable from a single summary; however, aggregating many rounds can
increase leakage. We recommend combining DSO with differential privacy noise
injection for sensitive deployments.

\subsection{Model Poisoning}

The semantic coherence gate (Eq.~\ref{eq:gate}) provides defense against model
poisoning attacks that inject semantically incoherent gradients. In our experiments,
the coherence gate detects and discards 94\% of poisoning gradients while passing
99.7\% of legitimate gradients.

%% -----------------------------------------------------------------------
\section{Conclusion}
\label{sec:conclusion}
%% -----------------------------------------------------------------------

DSO (ZIP-001) provides a Byzantine-robust, bandwidth-efficient, and heterogeneity-aware
decentralized training protocol for the Zen MoDE model family. Key results:

\begin{itemize}
    \item 94--99\% of centralized training quality at 64--256 nodes.
    \item 99.8--99.98\% bandwidth reduction vs.\ all-reduce (model-size independent).
    \item Tolerates up to $f < n/3$ Byzantine nodes with geometric median aggregation.
    \item Semantic coherence gate provides additional defense against model poisoning.
    \item $O(\log n)$ gossip convergence in round complexity.
\end{itemize}

The DSO protocol specification is published at \url{https://zips.zoo.ngo/zip-001}
and the reference implementation is available at \url{https://github.com/hanzoai/dso}.

\begin{thebibliography}{9}
\bibitem{mcmahan2017federated}
H.B. McMahan, E. Moore, D. Ramage, S. Hampson, B. Agüera y Arcas.
\textit{Communication-Efficient Learning of Deep Networks from Decentralized Data}.
AISTATS, 2017.

\bibitem{yin2018byzantine}
D. Yin, Y. Chen, R. Kannan, P. Bartlett.
\textit{Byzantine-Robust Distributed Learning: Towards Optimal Statistical Rates}.
ICML, 2018.

\bibitem{chen2017distributed}
Y. Chen, L. Su, J. Xu.
\textit{Distributed Statistical Machine Learning in Adversarial Settings: Byzantine Gradient Descent}.
POMACS, 2017.

\bibitem{blanchard2017machine}
P. Blanchard, E.M. El Mhamdi, R. Guerraoui, J. Stainer.
\textit{Machine Learning with Adversaries: Byzantine Tolerant Gradient Descent}.
NeurIPS, 2017.

\bibitem{kermarrec2007gossiping}
A.-M. Kermarrec, M. van Steen.
\textit{Gossiping in Distributed Systems}.
Operating Systems Review, 2007.
\end{thebibliography}

\end{document}
