\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{color}
\usepackage{booktabs}
\usepackage{float}
\usepackage{geometry}
\geometry{margin=1in}
\definecolor{zenblue}{RGB}{41,121,255}
\hypersetup{colorlinks=true,linkcolor=zenblue,urlcolor=zenblue,citecolor=zenblue}

\title{\textbf{Zen-Guard-Gen: Generative Safety Classification\\
with Natural Language Explanations and Policy References}\\[0.5em]
\large Technical Whitepaper v2025.05}
\author{Antje Worring, Zach Kelling, Zen LM Research Team\\
\texttt{research@zenlm.org}\\
\href{https://papers.zenlm.org}{papers.zenlm.org}}
\date{May 2025}

\begin{document}
\maketitle

\begin{abstract}
Zen-Guard-Gen is an 8 billion parameter generative safety model that extends content safety classification to include natural language explanations of safety decisions, actionable policy references, and remediation suggestions alongside binary safety labels. By framing safety classification as a generative task rather than a discriminative one, Zen-Guard-Gen produces outputs that are auditable, contestable, and actionable by human trust-and-safety teams. The model achieves 99.1\% on ToxiGen and 97.4\% on HatEval, while producing explanations rated 4.3/5 on explanation coherence MOS. This paper describes the generative safety formulation, training methodology, evaluation across content safety benchmarks, and deployment considerations for content moderation pipelines.
\end{abstract}

\tableofcontents
\newpage

\section{Introduction}

Binary content safety classifiers have been the workhorse of large-scale content moderation for over a decade. They are fast, scalable, and accurate on the distribution they are trained on. However, they have three persistent weaknesses that limit their practical utility:

\begin{enumerate}
  \item \textbf{Opacity}: A binary label (safe/unsafe) provides no information about why content was flagged, making it difficult for human reviewers to assess correctness, for policy teams to audit consistency, or for users to understand and contest decisions.
  \item \textbf{Policy rigidity}: Binary classifiers encode policy implicitly in their training distribution. Updating policy requires retraining; there is no mechanism to explain how a given decision relates to an explicit policy document.
  \item \textbf{Remediation gap}: A classifier identifies unsafe content but does not help content creators understand how to modify their content to comply with policy, or platform operators to understand what enforcement action is appropriate.
\end{enumerate}

Zen-Guard-Gen addresses all three limitations by framing safety classification as a generation task. Given a content item, Zen-Guard-Gen produces:

\begin{enumerate}
  \item A structured safety verdict (safe / unsafe / borderline).
  \item A primary policy category (hate speech, harassment, CSAM, violence, misinformation, etc.).
  \item A natural language explanation of the reasoning underlying the verdict.
  \item A reference to the applicable policy section.
  \item A remediation suggestion (for borderline and unsafe content).
\end{enumerate}

\subsection{Model Overview}

\begin{table}[H]
\centering
\caption{Zen-Guard-Gen Model Specification}
\begin{tabular}{ll}
\toprule
\textbf{Parameter} & \textbf{Value} \\
\midrule
Architecture & Zen MoDE (Mixture of Distilled Experts) \\
Total Parameters & 8B \\
Safety Taxonomy & 24 primary categories, 187 subcategories \\
Supported Content Types & Text, Image captions, Transcribed audio \\
ToxiGen Accuracy & 99.1\% \\
HatEval Accuracy & 97.4\% \\
Explanation Coherence MOS & 4.3/5 \\
Version & v2025.05 \\
Release Date & May 2025 \\
\bottomrule
\end{tabular}
\end{table}

\section{Safety Taxonomy}

\subsection{Primary Categories}

Zen-Guard-Gen uses a hierarchical safety taxonomy with 24 primary categories and 187 subcategories. The primary categories are:

\begin{table}[H]
\centering
\caption{Safety Taxonomy Primary Categories}
\begin{tabular}{lll}
\toprule
\textbf{Category} & \textbf{Subcategories} & \textbf{Policy Basis} \\
\midrule
Hate speech & 18 & Identity-based hatred \\
Harassment \& bullying & 12 & Targeted individual harm \\
Violence \& threats & 15 & Physical harm incitement \\
Sexual content & 21 & Explicit and non-consensual \\
Child safety & 9 & CSAM and grooming \\
Self-harm promotion & 8 & Suicide and self-injury \\
Misinformation & 16 & Health, election, factual \\
Privacy violation & 11 & PII, doxxing, stalking \\
Intellectual property & 7 & Copyright, trademark \\
Spam \& manipulation & 14 & Astroturfing, scam, phishing \\
Illegal goods & 12 & Drugs, weapons, fraud \\
Other & 44 & Platform-specific \\
\midrule
\textbf{Total} & \textbf{187} & \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Severity Levels}

Each category is scored on a severity scale aligned with CVSS-style impact ratings:

\begin{itemize}
  \item \textbf{Level 1 (Borderline)}: Content that may violate policy depending on context; requires human review.
  \item \textbf{Level 2 (Moderate)}: Clear policy violation warranting removal and possible account warning.
  \item \textbf{Level 3 (Severe)}: Serious violation warranting immediate removal and escalation.
  \item \textbf{Level 4 (Critical)}: Content requiring immediate removal and law enforcement referral (CSAM, credible threats).
\end{itemize}

\section{Architecture}

\subsection{Generative Safety Formulation}

Zen-Guard-Gen treats safety classification as a conditional generation problem. Given an input content item $x$, the model generates a structured output $y$ following a constrained grammar:

\begin{align}
  y &= \langle \text{verdict}, \text{category}, \text{severity}, \text{explanation}, \text{policy\_ref}, \text{remediation} \rangle \\
  p(y | x) &= \prod_{t=1}^{|y|} p(y_t | y_{<t}, x)
\end{align}

The structured output grammar is enforced via constrained decoding: the verdict, category, and severity fields use restricted vocabulary sampling from predefined value sets, while the explanation and remediation fields use unconstrained generation within a length limit.

This hybrid approach ensures structural consistency (no malformed outputs) while preserving the expressive flexibility of natural language for the reasoning components.

\subsection{Policy-Conditioned Generation}

Zen-Guard-Gen conditions its outputs on an explicit policy document context. The policy document is provided as a prefix in the context window, and the model is trained to cite specific policy sections in its output. This enables:

\begin{itemize}
  \item \textbf{Policy versioning}: Updated policy documents produce updated decisions without retraining.
  \item \textbf{Multi-platform deployment}: Different platforms can provide their own policy documents to customize moderation behavior.
  \item \textbf{Audit trails}: Each decision cites a specific policy section, enabling systematic policy consistency audits.
\end{itemize}

The policy conditioning is implemented through a 4K-token prefix reserved for the platform's active policy document, separate from the content item context.

\subsection{Calibrated Uncertainty}

For borderline content, Zen-Guard-Gen produces calibrated uncertainty estimates alongside verdicts. A temperature-scaled confidence score $c \in [0,1]$ accompanies each verdict:

\begin{equation}
  c = \sigma\left(\frac{z_{\text{verdict}}}{T_{\text{cal}}}\right)
\end{equation}

where $z_{\text{verdict}}$ is the logit for the predicted verdict and $T_{\text{cal}}$ is a calibration temperature estimated on the validation set. Expected calibration error (ECE) on the Borderline Safety Benchmark is 0.031, indicating well-calibrated uncertainty.

\section{Training Methodology}

\subsection{Training Data}

\begin{table}[H]
\centering
\caption{Training Data Composition}
\begin{tabular}{lrr}
\toprule
\textbf{Source} & \textbf{Items (M)} & \textbf{Fraction} \\
\midrule
Social media safety datasets & 120 & 40.0\% \\
Human-annotated moderation decisions & 45 & 15.0\% \\
Policy document-aligned synthetic data & 60 & 20.0\% \\
Adversarial jailbreak examples & 30 & 10.0\% \\
Explanation-annotated safety data & 25 & 8.3\% \\
Legal and regulatory guidance & 20 & 6.7\% \\
\midrule
\textbf{Total} & \textbf{300} & \textbf{100\%} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Explanation Quality Training}

Explanation quality is trained through a two-stage process:

\textbf{Stage 1 -- Explanation Initialization:} A large capable model generates initial explanations for a seed set of 500K annotated examples. These explanations are filtered for coherence and policy alignment by a separate reviewer model.

\textbf{Stage 2 -- Human Preference Alignment:} 50K explanation pairs are rated by trained annotators on four criteria: accuracy (does the explanation correctly characterize the content), coherence (is the explanation well-reasoned), actionability (does the explanation provide useful guidance), and conciseness (is the explanation appropriately brief). RLHF is used to optimize explanation quality across all four criteria.

\subsection{Red-Teaming}

Zen-Guard-Gen was red-teamed by a team of 40 adversarial safety researchers over 6 weeks. Red-teaming targets included:

\begin{itemize}
  \item Jailbreaking via encoded content (base64, l33t speak, homoglyph substitution).
  \item False explanation generation (correct verdict, misleading rationale).
  \item Policy citation hallucination (citing non-existent policy sections).
  \item Category confusion attacks (borderline content engineered to trigger wrong category).
\end{itemize}

Red-teaming results informed 14 targeted data augmentation campaigns and 3 architectural changes to the policy conditioning mechanism.

\section{Evaluation}

\subsection{Classification Accuracy}

\begin{table}[H]
\centering
\caption{Classification Benchmark Results}
\begin{tabular}{lcccc}
\toprule
\textbf{Benchmark} & \textbf{Zen-Guard-Gen} & \textbf{Discriminative 8B} & \textbf{Discriminative 3B} & \textbf{Human} \\
\midrule
ToxiGen & \textbf{99.1\%} & 98.3\% & 96.7\% & 97.8\% \\
HatEval & \textbf{97.4\%} & 96.8\% & 94.2\% & 96.4\% \\
OLID & 96.8\% & 96.2\% & 93.7\% & 97.1\% \\
HarmBench & 94.3\% & 93.1\% & 90.4\% & 95.7\% \\
SafeNLP & 97.6\% & 97.1\% & 94.8\% & 98.2\% \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Category Classification Accuracy}

\begin{table}[H]
\centering
\caption{Per-Category Classification F1}
\begin{tabular}{lcc}
\toprule
\textbf{Category} & \textbf{Zen-Guard-Gen F1} & \textbf{Discriminative Baseline F1} \\
\midrule
Hate speech & 98.3\% & 97.4\% \\
Harassment & 97.1\% & 96.2\% \\
Violence \& threats & 98.7\% & 97.8\% \\
Sexual content & 99.2\% & 98.6\% \\
Child safety & 99.6\% & 99.1\% \\
Self-harm & 96.4\% & 94.8\% \\
Misinformation & 93.2\% & 91.4\% \\
Privacy violation & 94.7\% & 93.1\% \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Explanation Quality}

\begin{table}[H]
\centering
\caption{Explanation Quality MOS (1--5 scale, 500 human evaluations)}
\begin{tabular}{lccc}
\toprule
\textbf{Dimension} & \textbf{Zen-Guard-Gen} & \textbf{Discriminative + Template} & \textbf{Human Moderator} \\
\midrule
Accuracy & 4.4 & 3.1 & 4.6 \\
Coherence & 4.3 & 2.9 & 4.7 \\
Actionability & 4.2 & 2.4 & 4.4 \\
Conciseness & 4.1 & 3.8 & 3.9 \\
\textbf{Composite MOS} & \textbf{4.3} & \textbf{3.1} & \textbf{4.4} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Policy Citation Accuracy}

\begin{table}[H]
\centering
\caption{Policy Citation Accuracy}
\begin{tabular}{lcc}
\toprule
\textbf{Metric} & \textbf{Value} \\
\midrule
Correct section cited & 91.4\% \\
Hallucinated section & 2.3\% \\
Relevant but inexact section & 6.3\% \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Adversarial Robustness}

\begin{table}[H]
\centering
\caption{Adversarial Attack Detection Rate}
\begin{tabular}{lcc}
\toprule
\textbf{Attack Type} & \textbf{Zen-Guard-Gen} & \textbf{Discriminative Baseline} \\
\midrule
Base64 encoding & 94.1\% & 67.3\% \\
Homoglyph substitution & 96.8\% & 81.4\% \\
L33t speak & 97.3\% & 88.2\% \\
Semantic paraphrasing & 91.7\% & 84.6\% \\
Multi-hop obfuscation & 87.4\% & 61.3\% \\
\textbf{Average} & \textbf{93.5\%} & \textbf{76.6\%} \\
\bottomrule
\end{tabular}
\end{table}

The generative formulation provides substantially better adversarial robustness than discriminative models, because the model must explicitly reason about content semantics to produce a coherent explanation, rather than relying on surface-level pattern matching.

\section{Deployment}

\subsection{Pipeline Integration}

Zen-Guard-Gen is designed for two deployment modes:

\begin{enumerate}
  \item \textbf{Primary classifier}: Zen-Guard-Gen serves as the primary classification layer, with human review triggered for borderline verdicts or low-confidence decisions.
  \item \textbf{Explanation layer}: A faster discriminative classifier (e.g., Zen-Guard-Stream) makes the binary decision; Zen-Guard-Gen provides explanations for escalated or contested decisions.
\end{enumerate}

\subsection{Inference Throughput}

\begin{table}[H]
\centering
\caption{Inference Performance (FP8, H100 GPU)}
\begin{tabular}{lll}
\toprule
\textbf{Output Mode} & \textbf{Throughput} & \textbf{P95 Latency} \\
\midrule
Verdict + category only & 1,200 items/s & 4.8ms \\
Full explanation (avg 120 tokens) & 180 items/s & 31ms \\
Full explanation + remediation & 120 items/s & 47ms \\
\bottomrule
\end{tabular}
\end{table}

\section{Related Work}

Content safety classification has been addressed through fine-tuned BERT-class models \cite{perspective}, larger generative classifiers \cite{llmguard}, and rule-based systems \cite{cld}. Explanation generation for classification decisions has been studied under the framing of rationale extraction \cite{rationale} and chain-of-thought safety reasoning \cite{cot_safety}. Zen-Guard-Gen is the first production-scale system to integrate generative explanations, policy document conditioning, and remediation suggestions into a unified safety classification model.

\section{Conclusion}

Zen-Guard-Gen advances content safety classification from opaque binary decisions to transparent, auditable, policy-referenced verdicts with natural language explanations and remediation guidance. At 8B parameters it achieves 99.1\% ToxiGen and 97.4\% HatEval accuracy, matching or exceeding discriminative classifiers while providing substantially richer outputs for human review workflows. The generative formulation also provides superior adversarial robustness (93.5\% vs. 76.6\% average attack detection) because semantic reasoning, not surface pattern matching, drives the classification. Zen-Guard-Gen is suitable for both high-throughput automated moderation and high-stakes escalation review in trust-and-safety teams.

\begin{thebibliography}{9}
\bibitem{perspective} Lees, A. et al. (2022). A New Generation of Perspective API: Efficient Multilingual Character-level Transformers. arXiv:2202.11176.
\bibitem{llmguard} Inan, H. et al. (2023). Llama Guard: LLM-based Input-Output Safeguard for Human-AI Conversations. arXiv:2312.06674.
\bibitem{cld} Waseem, Z. et al. (2016). Hateful Symbols or Hateful People? Predictive Features for Hate Speech Detection on Twitter. NAACL 2016.
\bibitem{rationale} Deyoung, J. et al. (2020). ERASER: A Benchmark to Evaluate Rationalized NLP Models. ACL 2020.
\bibitem{cot_safety} Bai, Y. et al. (2022). Constitutional AI: Harmlessness from AI Feedback. arXiv:2212.08073.
\end{thebibliography}

\end{document}
