\section{Safety Boundaries}

\label{sec:safety}

Operative implements a defense-in-depth safety system to prevent agents from causing unintended harm.

\subsection{Permission Model}

Safety boundaries are configured through a declarative policy:

\begin{lstlisting}[language=Python, caption=Safety boundary configuration]
safety:
  filesystem:
    allowed_paths:
      - ~/Downloads
      - ~/Documents/agent-workspace
    denied_paths:
      - ~/.ssh
      - ~/.aws
      - /etc
    max_file_size: 100MB

  network:
    allowed_domains:
      - "*.google.com"
      - "*.github.com"
    denied_domains:
      - "*.darkweb.*"
    block_downloads_from: ["*"]

  system:
    allow_app_launch: true
    allow_app_close: false
    allow_system_settings: false
    allow_shell_commands: false
    max_concurrent_windows: 5

  interaction:
    require_confirmation:
      - purchase
      - delete
      - send_email
      - form_submission
    max_actions_per_minute: 60
    idle_timeout: 300
\end{lstlisting}

\subsection{Sandboxing Layers}

\begin{enumerate}[leftmargin=*]
    \item \textbf{Action-level filtering}: Each action is checked against the policy before execution. Denied actions return an error to the planner, which must select an alternative approach.
    \item \textbf{Confirmation gates}: Actions matching \texttt{require\_confirmation} patterns trigger a human-in-the-loop approval step. The agent pauses and presents the proposed action with context to the user.
    \item \textbf{VM isolation}: For maximum security, Operative supports execution within a Docker container or lightweight VM with restricted host access. The screen is shared via VNC or headless framebuffer.
    \item \textbf{Rollback capability}: Reversible actions maintain undo state. If a sequence of actions produces unintended results, the planner can issue rollback commands to restore previous state.
\end{enumerate}

\subsection{Rate Limiting and Circuit Breaking}

To prevent runaway agents, Operative enforces:

\begin{itemize}[leftmargin=*]
    \item \textbf{Action rate limit}: Maximum 60 actions per minute (configurable), preventing rapid uncontrolled clicking.
    \item \textbf{Cost limit}: For actions that incur LLM costs, a per-task budget ceiling.
    \item \textbf{Repetition detection}: If the agent repeats the same action sequence more than 3 times, execution pauses for human review.
    \item \textbf{Error escalation}: After 5 consecutive failed actions, the agent enters supervised mode requiring human approval for each subsequent action.
\end{itemize}

