% Hanzo Crypto Commerce: Blockchain Payment APIs and Smart Contract Infrastructure
\documentclass[11pt,twocolumn]{article}
\usepackage{shared/hanzocover}
\usepackage{amsmath, amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{float}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,urlcolor=blue}

\title{Hanzo Crypto Commerce: Blockchain Payment APIs and Smart Contract
Infrastructure for Enterprise}
\author{
    Zach Kelling \\
    \textit{CTO, Hanzo AI (Techstars '17)} \\
    \texttt{zach@hanzo.ai}
}
\date{2018}

\begin{document}
\hanzocoverpage

\twocolumn[
\begin{@twocolumnfalse}
\maketitle
\begin{abstract}
We describe the Hanzo crypto commerce platform, a unified API layer that enables
any business to accept cryptocurrency payments, deploy smart contracts, and
manage digital wallets without blockchain expertise. The platform provides
zero-volatility settlement---merchants receive fiat in their bank accounts
regardless of the cryptocurrency used by the customer---alongside custodial and
non-custodial wallet options, audited smart contract templates for token sales
and airdrops, and a management dashboard for real-time operational visibility.
Built by a 2017 Techstars graduate, the system is designed on the principle that
the best payment integration is one where the developer writes as little payments
code as possible.
\end{abstract}
\end{@twocolumnfalse}
]

\section{Introduction}
\label{sec:intro}

By 2018, cryptocurrency adoption had reached a critical inflection point:
millions of consumers held digital assets, yet the merchant infrastructure to
accept those assets remained fragmented, technically demanding, and
operationally risky. Price volatility alone deterred most businesses---a Bitcoin
payment worth \$100 at confirmation could be worth \$85 by settlement.

Hanzo addresses this gap with a complete crypto-commerce platform whose design
principle is borrowed from Stripe: \emph{measure success by how little payments
code the developer has to write}. The platform handles wallet management,
transaction signing, settlement, and compliance so that engineering teams can
focus on product.

\section{Architecture}
\label{sec:architecture}

The platform consists of four layers:

\begin{enumerate}[leftmargin=1.1em]
    \item \textbf{Payment Gateway.} Accepts Bitcoin, Ethereum, and ERC-20 tokens.
    Converts to fiat at the moment of confirmation using real-time exchange rates,
    delivering zero-volatility settlement to the merchant's bank account.
    \item \textbf{Wallet Services.} Provides both custodial wallets (for ease of
    use) and non-custodial wallets (for maximum security), selectable per user
    or per transaction.
    \item \textbf{Smart Contract Engine.} Deploys audited, parameterized smart
    contracts from a template library---ICOs, airdrops, collectibles, referral
    programs---without requiring Solidity expertise.
    \item \textbf{Management Dashboard.} A unified interface for tracking
    payments, managing smart contracts, monitoring token sale progress, and
    gaining business insights in real time.
\end{enumerate}

\section{Payment Processing}
\label{sec:payments}

\subsection{Zero-Volatility Settlement}

The core innovation is instant fiat conversion. When a customer pays in
cryptocurrency:

\begin{enumerate}[leftmargin=1.1em]
    \item The Hanzo gateway generates a payment address and locks an exchange
    rate for a configurable confirmation window.
    \item Upon sufficient blockchain confirmations, the gateway executes an
    immediate market sell on partnered exchanges.
    \item The merchant receives settlement in their local currency (USD, EUR,
    GBP) via standard bank transfer, typically within one business day.
\end{enumerate}

The merchant bears zero cryptocurrency price risk. The spread between the locked
rate and the execution rate is absorbed by the platform's liquidity reserves.

\subsection{Multi-Currency Support}

The payment gateway accepts any currency for which Hanzo maintains exchange
liquidity. At launch: BTC, ETH, and major ERC-20 tokens. The unified API
abstracts currency differences---a single integration endpoint handles all
supported assets.

\section{Wallet Solutions}
\label{sec:wallets}

\begin{table}[H]
\centering
\small
\begin{tabular}{lcc}
\toprule
\textbf{Feature} & \textbf{Custodial} & \textbf{Non-Custodial} \\
\midrule
Key management & Hanzo HSM & User device \\
Recovery & Email/phone & Seed phrase \\
Transaction signing & Server-side & Client-side \\
Target user & Consumer apps & Power users \\
\bottomrule
\end{tabular}
\caption{Wallet mode comparison.}
\label{tab:wallets}
\end{table}

Both modes share the same API surface. Developers choose at initialization time
and can offer both options to end users within the same application.

\section{Smart Contract Platform}
\label{sec:contracts}

Hanzo provides audited smart contract templates covering the most common
blockchain business patterns:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Token Sales (ICOs).} Configurable cap, pricing tiers,
    whitelisting, and vesting schedules.
    \item \textbf{Airdrops.} Batch token distribution with Merkle-proof-based
    claiming to minimize gas costs.
    \item \textbf{Collectibles.} ERC-721 non-fungible tokens with metadata
    management and marketplace hooks.
    \item \textbf{Referral Programs.} On-chain referral tracking with automatic
    token rewards.
\end{itemize}

Each template is parameterized via the API---no Solidity required. For custom
requirements, Hanzo's smart contract development service provides bespoke
implementation with third-party audit.

\section{Developer Experience}
\label{sec:dx}

The API design follows three principles:

\begin{enumerate}[leftmargin=1.1em]
    \item \textbf{Composable.} Small, orthogonal endpoints that combine to
    support any payment flow.
    \item \textbf{Multi-platform.} Client libraries for web, iOS, Android,
    and server-side in Node.js, Python, Go, Ruby, and Haskell.
    \item \textbf{Non-breaking.} Versioned API changes allow merchants to
    upgrade on their own schedule.
\end{enumerate}

Two JavaScript libraries provide drop-in frontend integration:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Shop.js.} Pre-built UI components for checkout flows with
    validation, formatting, and browser detection.
    \item \textbf{Coin.js.} A lightweight library for accepting crypto payments
    or launching token sales with minimal frontend code.
\end{itemize}

\section{Management Dashboard}
\label{sec:dashboard}

The Hanzo Dashboard provides:

\begin{itemize}[leftmargin=1.1em]
    \item Real-time charge tracking across payment types, countries, and
    currencies.
    \item Smart contract lifecycle management---deploy, monitor, pause, and
    upgrade.
    \item Token sale progress visualization with investor analytics.
    \item Webhook configuration for integrating payment events into existing
    business systems.
\end{itemize}

\section{Security and Compliance}
\label{sec:security}

All smart contracts undergo formal verification and third-party audit before
deployment. The payment gateway implements:

\begin{itemize}[leftmargin=1.1em]
    \item Hardware security module (HSM) key storage for custodial wallets.
    \item Built-in fraud detection using transaction pattern analysis.
    \item KYC/AML integration for regulated jurisdictions.
    \item PCI-DSS-compliant card data handling for hybrid fiat-crypto flows.
\end{itemize}

\section{Services}
\label{sec:services}

Beyond the self-service platform, Hanzo offers professional services:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{White paper authoring.} Technical writing and token economics
    design for blockchain projects.
    \item \textbf{Security audit.} Smart contract and protocol review.
    \item \textbf{Token distribution management.} End-to-end token sale
    operation.
\end{itemize}

\section{Conclusion}
\label{sec:conclusion}

Hanzo's crypto commerce platform eliminates the three barriers to enterprise
cryptocurrency adoption: technical difficulty, price volatility, and
compliance uncertainty. By providing a Stripe-like developer experience for
blockchain payments, audited smart contract templates, and zero-volatility
settlement, the platform enables any business to participate in the
cryptocurrency economy without specialized blockchain expertise. Hanzo is
dedicated to advancing decentralization by making crypto payments accessible,
opening the door for the world to transact on the blockchain.

\end{document}
