\section{Architecture}


\subsection{Core Principles}

Astle.js adheres to three architectural principles:

\textbf{Principle 1: State as Derived Value.} Application state is computed from an ordered sequence of events. The current state is a pure function of the event history:

\begin{equation}
\text{state}_n = \text{reduce}(\text{state}_0, [\text{event}_1, \ldots, \text{event}_n])
\end{equation}

\textbf{Principle 2: Events are Immutable.} Once recorded, events cannot be modified. Corrections are represented as compensating events, preserving audit history.

\textbf{Principle 3: Optimistic by Default.} User actions produce immediate local state changes. Server confirmation or rejection triggers reconciliation.

\subsection{Event Model}

Events in Astle.js follow a standard schema:

\begin{lstlisting}[language=JavaScript,caption=Event schema]
interface Event {
  id: string;          // UUID
  type: string;        // e.g., "cart/addItem"
  payload: object;     // Event-specific data
  timestamp: number;   // Unix milliseconds
  origin: "local" | "remote";
  status: "pending" | "confirmed" | "rejected";
}
\end{lstlisting}

Events transition through states:
\begin{enumerate}
\item \texttt{pending}: Applied locally, awaiting server confirmation.
\item \texttt{confirmed}: Server acknowledged; event is permanent.
\item \texttt{rejected}: Server rejected; compensating event generated.
\end{enumerate}

\subsection{State Domains}

Astle.js partitions state into domains with independent reducers:

\subsubsection{Cart Domain}

The cart domain manages shopping cart state:

\begin{lstlisting}[language=JavaScript,caption=Cart state shape]
interface CartState {
  items: CartItem[];
  subtotal: number;
  discounts: Discount[];
  total: number;
  itemCount: number;
}
\end{lstlisting}

Cart events include:
\begin{itemize}
\item \texttt{cart/addItem}: Add product to cart.
\item \texttt{cart/updateQuantity}: Change item quantity.
\item \texttt{cart/removeItem}: Remove item from cart.
\item \texttt{cart/applyCoupon}: Apply promotional code.
\item \texttt{cart/clear}: Empty cart contents.
\end{itemize}

\subsubsection{Checkout Domain}

The checkout domain manages the checkout flow:

\begin{lstlisting}[language=JavaScript,caption=Checkout state machine]
type CheckoutStatus =
  | "idle"
  | "collecting_info"
  | "processing_payment"
  | "confirming"
  | "complete"
  | "failed";
\end{lstlisting}

State transitions are guarded by validation:

\begin{equation}
\text{transition}(s, e) = \begin{cases}
s' & \text{if valid}(s, e) \\
s & \text{otherwise}
\end{cases}
\end{equation}

\subsubsection{Customer Domain}

The customer domain tracks authentication and profile state:

\begin{lstlisting}[language=JavaScript,caption=Customer state]
interface CustomerState {
  authenticated: boolean;
  profile: Profile | null;
  addresses: Address[];
  paymentMethods: PaymentMethod[];
}
\end{lstlisting}

\subsection{Reducer Composition}

Domain reducers compose into the root reducer:

\begin{lstlisting}[language=JavaScript,caption=Root reducer composition]
const rootReducer = combineReducers({
  cart: cartReducer,
  checkout: checkoutReducer,
  customer: customerReducer,
  ui: uiReducer,
});

function cartReducer(state = initialCart, event) {
  switch (event.type) {
    case "cart/addItem":
      return addItemToCart(state, event.payload);
    case "cart/updateQuantity":
      return updateCartQuantity(state, event.payload);
    // ...
    default:
      return state;
  }
}
\end{lstlisting}

\subsection{Optimistic Updates}

Astle.js implements optimistic updates through a three-phase protocol:

\textbf{Phase 1: Local Application.} User action generates event with \texttt{status: "pending"}. Reducer applies event to produce optimistic state. UI renders immediately.

\textbf{Phase 2: Server Submission.} Event dispatched to server via WebSocket or HTTP. Client awaits confirmation.

\textbf{Phase 3: Reconciliation.} Server responds with confirmation or rejection.
\begin{itemize}
\item \textbf{Confirmed}: Event status updates; state unchanged.
\item \textbf{Rejected}: Compensating event generated; state rolls back.
\end{itemize}

\begin{lstlisting}[language=JavaScript,caption=Optimistic middleware]
function optimisticMiddleware(store) {
  return next => async event => {
    // Phase 1: Apply locally
    event.status = "pending";
    next(event);

    try {
      // Phase 2: Submit to server
      const response = await api.submit(event);

      // Phase 3a: Confirm
      next({ type: "event/confirm", payload: { id: event.id } });
    } catch (error) {
      // Phase 3b: Reject and compensate
      next({ type: "event/reject", payload: { id: event.id, error } });
      next(compensate(event));
    }
  };
}
\end{lstlisting}

\subsection{Real-Time Synchronization}

Astle.js maintains a persistent WebSocket connection for real-time updates:

\begin{lstlisting}[language=JavaScript,caption=WebSocket synchronization]
class SyncClient {
  constructor(store, url) {
    this.store = store;
    this.socket = new WebSocket(url);
    this.queue = [];

    this.socket.onmessage = (msg) => {
      const event = JSON.parse(msg.data);
      event.origin = "remote";
      this.store.dispatch(event);
    };

    this.socket.onclose = () => {
      this.reconnect();
    };
  }

  send(event) {
    if (this.socket.readyState === WebSocket.OPEN) {
      this.socket.send(JSON.stringify(event));
    } else {
      this.queue.push(event);
    }
  }
}
\end{lstlisting}

Server-originated events (inventory updates, price changes) flow through the same reducer pipeline, ensuring consistent state regardless of event origin.

