\documentclass[11pt,a4paper]{article}
\usepackage{hanzo-defense}

\doctitle{Quantum Key Distribution and Quantum Sensing\\Integration for Post-Quantum Blockchain\\Consensus and Threshold Cryptography}
\docid{HAI-DEF-2026-009}
\docversion{Version 1.0 --- February 27, 2026}
\docdate{February 2026}
\docauthors{Antje Worring, Zach Kelling --- Hanzo Team}
\docorgs{Hanzo Industries \quad Lux Industries \quad Zoo Labs Foundation\\research@hanzo.ai}

\begin{document}
\makecover

\begin{abstract}
We present an architecture for integrating quantum key distribution (QKD) hardware and quantum sensing devices into a production post-quantum blockchain stack. The system extends five existing production components---QuantumVM (Q-Chain), the KEM cryptographic interface, RNS mesh transport, Sovereign KMS, and the Hanzo Zero Trust fabric---with QKD-derived key material and quantum sensor attestations. A novel \emph{triple-hybrid} key derivation combines classical Diffie-Hellman (X25519), post-quantum lattice encapsulation (ML-KEM-768, \fips{203}), and physics-layer QKD shared secrets via HKDF, ensuring link security if \emph{any one} of three independent foundations holds: elliptic curve hardness, Module-LWE hardness, or the laws of quantum mechanics. Quantum sensors---atomic clocks, magnetometers, and gravimeters---feed cryptographically attested measurements into the Q-Chain via a Sensor Oracle, stamped with dual BLS/Ringtail threshold signatures for immutable provenance. We specify ETSI GS QKD 014 and ITU-T Y.3800 compliant interfaces, analyze integration with threshold MPC signing ceremonies, and project a 3,000--5,000 line implementation leveraging existing production interfaces. No competing blockchain system offers production post-quantum consensus, let alone QKD or quantum sensing integration.
\end{abstract}

\tableofcontents
\newpage

% ============================================================================
\section{Introduction}
% ============================================================================

Post-quantum cryptography protects against \emph{algorithmic} quantum attacks---Shor's algorithm breaking RSA and elliptic curves, Grover's algorithm weakening symmetric ciphers. However, algorithmic security rests on computational hardness \emph{assumptions} (Module-LWE, Module-SIS) that, while believed strong, lack the unconditional guarantee of physics-based security.

Quantum key distribution provides information-theoretic security grounded in the laws of quantum mechanics: an eavesdropper disturbing a quantum channel is detectable with probability approaching 1. Quantum sensing exploits quantum coherence and entanglement for measurement precision beyond classical limits.

This paper presents an architecture that layers physics-based security (QKD) atop algorithmic security (PQ crypto) atop classical security (ECDH), creating a triple-hybrid defense-in-depth unique in the blockchain and military networking landscape.

\subsection{Existing Production Foundation}

The architecture builds on five deployed systems:

\begin{enumerate}
\item \textbf{QuantumVM (Q-Chain)}: Production VM with ML-DSA-44/65/87 signing, SLH-DSA stamping, and Quasar dual-threshold consensus (BLS + Ringtail).
\item \textbf{KEM Interface}: Pluggable key encapsulation supporting ML-KEM-768, ML-KEM-1024, X25519, and hybrid modes.
\item \textbf{RNS Mesh Transport}: Medium-agnostic networking with hybrid PQ handshake (X25519 + ML-KEM-768 + ML-DSA-65) over LoRa, serial, TCP, and satellite.
\item \textbf{Sovereign KMS}: Enterprise key management with Shamir sharing, HPKE wrapping, transit engine, and TFHE bridge.
\item \textbf{Zero-Trust Fabric}: Identity-first overlay network with pluggable transport underlays and HSM integration.
\end{enumerate}

\subsection{What This Paper Adds}

\begin{enumerate}
\item \textbf{QKD-KEM Bridge}: ETSI GS QKD 014 client wrapping QKD-derived keys into the existing \texttt{kem.KEM} interface.
\item \textbf{Triple-Hybrid Key Derivation}: $K = \text{HKDF}(\text{X25519\_ss} \| \text{ML-KEM\_ss} \| \text{QKD\_key})$.
\item \textbf{Sensor Oracle}: New transaction type in QuantumVM for quantum sensor attestations with dual-threshold signatures.
\item \textbf{QRNG Integration}: Drop-in replacement of \texttt{crypto/rand} with quantum random number generator hardware.
\item \textbf{QKD Network Manager}: ITU-T Y.3800 compliant key pool orchestration across multiple QKD links.
\end{enumerate}

% ============================================================================
\section{Quantum Key Distribution Integration}
% ============================================================================

\subsection{QKD Protocol Support}

The architecture supports all major QKD protocol families via the ETSI standardized hardware interface:

\begin{table}[H]
\centering
\small
\begin{tabular}{@{}llrrl@{}}
\toprule
\textbf{Protocol} & \textbf{Type} & \textbf{Range} & \textbf{Rate} & \textbf{Vendors} \\
\midrule
BB84 & DV, prepare-measure & $<$100 km & $\sim$1 Mbps & ID Quantique \\
CV-QKD & Continuous-variable & $<$50 km & $\sim$10 Mbps & Toshiba \\
MDI-QKD & Measurement-device-indep. & $<$400 km & $\sim$100 kbps & QuantumCTek \\
TF-QKD & Twin-field & $<$600 km & $\sim$10 kbps & Toshiba, USTC \\
E91 & Entanglement-based & $<$100 km & $\sim$10 kbps & Research \\
\bottomrule
\end{tabular}
\caption{Supported QKD protocols via ETSI hardware interface.}
\label{tab:qkd-protocols}
\end{table}

\subsection{ETSI GS QKD 014 Client}

All QKD hardware interaction uses the ETSI GS QKD 014 REST API, the industry standard key delivery interface:

\begin{lstlisting}[caption={ETSI QKD 014 key request flow.}]
// Request quantum-derived key material
GET /api/v1/keys/{slave_SAE_ID}/status
  -> { "source_KME_ID": "kme-alpha",
       "key_count": 42,
       "max_key_size": 256 }

POST /api/v1/keys/{slave_SAE_ID}/enc_keys
  body: { "number": 1, "size": 256 }
  -> { "keys": [
         { "key_ID": "a1b2c3d4-...",
           "key": "base64(256-bit key)" }
       ]}

// Peer retrieves same key by ID
POST /api/v1/keys/{master_SAE_ID}/dec_keys
  body: { "key_IDs": ["a1b2c3d4-..."] }
  -> { "keys": [
         { "key_ID": "a1b2c3d4-...",
           "key": "base64(same 256-bit key)" }
       ]}
\end{lstlisting}

The ETSI 014 abstraction makes the architecture hardware-agnostic: BB84, CV-QKD, or TF-QKD appliances from any vendor present the same REST interface.

\subsection{QKD-KEM Bridge}

The QKD-KEM bridge implements the existing \texttt{kem.KEM} interface, enabling QKD to slot into any component that uses key encapsulation:

\begin{lstlisting}[language=Go,caption={QKD-KEM bridge implementing existing KEM interface.}]
type QKDKEM struct {
    client *etsi014.Client
    peerID string
}

func (q *QKDKEM) Encapsulate() (ct, ss []byte, err error) {
    // Request key from local QKD node
    key, err := q.client.GetKey(q.peerID, 256)
    if err != nil { return nil, nil, err }
    // "ciphertext" is just the key_ID (peer fetches same key)
    return []byte(key.KeyID), key.Material, nil
}

func (q *QKDKEM) Decapsulate(ct []byte) (ss []byte, err error) {
    keyID := string(ct)
    key, err := q.client.GetKeyByID(q.peerID, keyID)
    return key.Material, err
}
\end{lstlisting}

This plugs directly into the RNS link handshake at \texttt{rns\_link.go} where ML-KEM ciphertexts are exchanged. The QKD-KEM has zero ciphertext overhead---the key is delivered out-of-band by the quantum channel.

% ============================================================================
\section{Triple-Hybrid Key Derivation}
% ============================================================================

\subsection{Security Model}

\begin{definition}[Triple-Hybrid Security]
A key exchange is triple-hybrid secure if the derived session key $K$ is indistinguishable from random under the assumption that \emph{at least one} of the following holds:
\begin{enumerate}
\item The Computational Diffie-Hellman problem on Curve25519 is hard (classical).
\item The Module-LWE problem with ML-KEM-768 parameters is hard (post-quantum algorithmic).
\item The laws of quantum mechanics prevent undetected eavesdropping on the QKD channel (physics-layer).
\end{enumerate}
\end{definition}

\subsection{Protocol}

\begin{algorithm}[H]
\caption{Triple-Hybrid Key Exchange}
\begin{algorithmic}[1]
\State \textbf{Layer 1 --- Classical ECDH}:
\State \quad $\text{ss}_1 \gets \text{X25519.DH}(\text{ek}_A, \text{epk}_B)$ \quad (32 bytes)
\Statex
\State \textbf{Layer 2 --- Post-Quantum KEM}:
\State \quad $(\text{ct}_2, \text{ss}_2) \gets \text{ML-KEM-768.Encaps}(\text{epk}_B^M)$ \quad (32 bytes)
\Statex
\State \textbf{Layer 3 --- Physics-Layer QKD}:
\State \quad $\text{ss}_3 \gets \text{QKD.GetKey}(\text{peer\_ID}, 256)$ \quad (32 bytes)
\State \quad $\text{key\_ID} \gets \text{QKD.KeyID}(\text{ss}_3)$
\Statex
\State \textbf{Combine}: $K \gets \text{HKDF-SHA256}(\text{salt}, \text{ss}_1 \| \text{ss}_2 \| \text{ss}_3)$
\State \textbf{Derive}: $K_{\text{send}}, K_{\text{recv}} \gets \text{HKDF-Expand}(K, \text{``triple-hybrid-v1''})$
\State \textbf{Exchange}: Send $\text{epk}_A^X \| \text{ct}_2 \| \text{key\_ID}$ to peer
\State \textbf{Session}: AES-256-GCM with derived keys
\end{algorithmic}
\end{algorithm}

\begin{theorem}[Triple-Hybrid IND-CCA2]
If any one of X25519, ML-KEM-768, or the QKD key is secure (\ie unknown to the adversary), then the HKDF output $K$ is computationally indistinguishable from a uniformly random key of the same length. This follows from the extraction property of HKDF: given sufficient min-entropy in any one input, the output is pseudorandom regardless of the other inputs.
\end{theorem}

\subsection{Wire Format}

\begin{table}[H]
\centering
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{Component} & \textbf{Classical} & \textbf{PQ Hybrid} & \textbf{Triple Hybrid} \\
\midrule
X25519 ephemeral PK & 32 B & 32 B & 32 B \\
ML-KEM-768 ciphertext & --- & 1,088 B & 1,088 B \\
QKD key\_ID (UUID) & --- & --- & 36 B \\
ML-DSA-65 signature & --- & 3,309 B & 3,309 B \\
\midrule
\textbf{Total per-direction} & \textbf{32 B} & \textbf{$\sim$4.4 KB} & \textbf{$\sim$4.5 KB} \\
\bottomrule
\end{tabular}
\caption{Wire overhead comparison: triple-hybrid adds only 36 bytes (UUID) over PQ hybrid.}
\label{tab:triple-hybrid-wire}
\end{table}

The QKD layer adds only 36 bytes (a UUID key reference) to the handshake---the actual key material is delivered by the quantum channel, not the classical network.

% ============================================================================
\section{Quantum Sensing and Attestation}
% ============================================================================

\subsection{Sensor Types}

\begin{table}[H]
\centering
\small
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Sensor} & \textbf{Technology} & \textbf{Precision} & \textbf{Naval Application} \\
\midrule
Atomic clock & CSAC/optical lattice & $<$1 $\mu$s accuracy & GPS-denied timing \\
Magnetometer & SERF/NV-center & $<$1 fT sensitivity & Submarine detection \\
Gravimeter & Atom interferometry & $<$1 $\mu$Gal & Subsurface mapping \\
Quantum radar & Microwave photons & Enhanced SNR & Target detection \\
Quantum lidar & Entangled photons & Sub-shot-noise & Terrain mapping \\
\bottomrule
\end{tabular}
\caption{Quantum sensor types with naval applications.}
\label{tab:sensors}
\end{table}

\subsection{Sensor Oracle Architecture}

Quantum sensor measurements are ingested into the Q-Chain via a Sensor Oracle transaction type:

\begin{lstlisting}[language=Go,caption={Sensor attestation transaction.}]
type SensorAttestation struct {
    SensorType    SensorType  // AtomicClock, Magnetometer, etc.
    DeviceID      [32]byte    // Hardware attestation identity
    Timestamp     int64       // Sub-nanosecond precision
    Measurement   []byte      // Sensor-specific data format
    Uncertainty   float64     // Measurement uncertainty
    TEESignature  []byte      // TEE-signed device attestation
    MLDSASignature []byte     // ML-DSA-65 over all fields
}
\end{lstlisting}

The QuantumVM's existing \texttt{StampBlock} interface accepts arbitrary block IDs and messages---sensor attestations are stamped with the same Quasar dual-signature (BLS + Ringtail) that certifies consensus blocks.

\subsection{Atomic Clocks for Consensus Timing}

Chip-scale atomic clocks (CSAC) provide GPS-independent timing for DDIL environments:

\begin{itemize}
\item \textbf{Current}: QuantumVM uses \texttt{time.Now()} for block timestamps (OS clock, $\sim$1 ms accuracy).
\item \textbf{With CSAC}: Sub-microsecond timestamp ordering via PTP/1PPS interface from quantum clock.
\item \textbf{Byzantine detection}: Validators with drifted clocks produce timestamps outside the consensus window, enabling automatic identification of faulty or compromised nodes.
\item \textbf{GPS-denied operation}: Aligns with RNS mesh networking for maritime/submarine scenarios where GPS signals are unavailable or jammed.
\end{itemize}

\subsection{Magnetometers for Physical Security}

Quantum magnetometers detecting sub-femtotesla fields enable:

\begin{itemize}
\item \textbf{Tamper detection}: Cryptographically attested magnetic field baseline for data center physical security. Any hardware modification changes the field signature.
\item \textbf{Submarine detection}: Blockchain-attested magnetic anomaly data from distributed sensor arrays.
\item \textbf{Navigation}: Magnetic field map matching for GPS-denied underwater navigation, attested on-chain for integrity.
\end{itemize}

\subsection{Gravimeters for Subsurface Intelligence}

Atom-interferometry gravimeters provide:

\begin{itemize}
\item \textbf{Subsurface mapping}: Detect underground structures, tunnels, and voids.
\item \textbf{Immutable geological records}: Blockchain-attested gravity surveys with provenance chains.
\item \textbf{Anti-spoofing}: Ground-truth gravity data anchored to consensus prevents adversary falsification.
\end{itemize}

% ============================================================================
\section{QRNG Integration}
% ============================================================================

\subsection{Drop-In Entropy Source}

Quantum random number generators (QRNG) provide entropy from quantum vacuum fluctuations or single-photon detection:

\begin{lstlisting}[language=Go,caption={QRNG entropy source replacement.}]
// Current (QuantumSigner, line 104):
crypto.rand.Read(nonce)  // reads /dev/urandom

// With QRNG hardware:
qrng.Read(nonce)  // reads /dev/qrng0
// Fallback: crypto/rand if QRNG unavailable
\end{lstlisting}

\begin{table}[H]
\centering
\begin{tabular}{@{}llrl@{}}
\toprule
\textbf{Device} & \textbf{Interface} & \textbf{Rate} & \textbf{Standard} \\
\midrule
ID Quantique Quantis & USB/PCIe & 4--240 Mbps & NIST SP 800-90B \\
Quside FQRNG & PCIe & 400+ Mbps & NIST SP 800-22 \\
Toshiba QRNG & Integrated & 2 Gbps & MIST/BSI certified \\
\bottomrule
\end{tabular}
\caption{QRNG hardware options.}
\label{tab:qrng}
\end{table}

QRNG integration is the lowest-effort, highest-value component: a single import replacement at \texttt{rand.Read()} call sites across the key generation and nonce generation code paths.

% ============================================================================
\section{Integration with Threshold MPC}
% ============================================================================

\subsection{QKD-Authenticated DKG Ceremony}

Distributed Key Generation requires secure channels between all $n$ parties. QKD provides information-theoretically secure channels for share distribution:

\begin{algorithm}[H]
\caption{QKD-Authenticated Threshold Key Generation}
\begin{algorithmic}[1]
\State \textbf{Setup}: Each pair $(P_i, P_j)$ establishes QKD link
\State \quad $K_{ij} \gets \text{QKD.GetKey}(P_j, 256)$ (symmetric key from quantum channel)
\Statex
\State \textbf{DKG Phase 1}: Commitment broadcast
\For{each party $P_i$}
  \State Sample polynomial $f_i(x)$ of degree $t-1$
  \State Broadcast commitment $C_i = g^{f_i(0)}$ (public, signed ML-DSA-65)
  \For{each party $P_j$}
    \State Encrypt share: $e_{ij} = \text{AES-256-GCM}(K_{ij}, f_i(j))$
    \State Send $e_{ij}$ via classical channel (QKD key protects confidentiality)
  \EndFor
\EndFor
\Statex
\State \textbf{Advantage}: Share distribution is information-theoretically secure
\State Even a quantum computer cannot recover $f_i(j)$ from $e_{ij}$ without $K_{ij}$
\end{algorithmic}
\end{algorithm}

\subsection{QKD-Secured Signing Sessions}

Threshold signing messages between MPC parties use QKD-derived session keys:

\begin{itemize}
\item \textbf{CGGMP21}: Round messages encrypted with QKD keys via ZAP transport.
\item \textbf{FROST}: Commitment and response shares protected by QKD-authenticated channels.
\item \textbf{Ringtail}: Lattice threshold shares doubly protected---PQ algorithmic + QKD physics-layer.
\item \textbf{BLS}: Aggregate signature shares transmitted over QKD-secured links.
\end{itemize}

\subsection{Key Rotation with QKD Refresh}

Ringtail epoch rotation (every 10 minutes) can incorporate QKD key refresh:

\begin{enumerate}
\item At epoch boundary, validators request fresh QKD keys from all peer links.
\item New Ringtail threshold keys generated using QKD-authenticated DKG.
\item Old keys destroyed; forward secrecy guaranteed by both ephemeral PQ keys and QKD one-time keys.
\item Triple protection: even if lattice assumptions fail AND classical crypto breaks, QKD keys remain secure.
\end{enumerate}

% ============================================================================
\section{Integration with Consensus}
% ============================================================================

\subsection{QKD-Authenticated Validator Links}

Quasar consensus samples $k$ peers per round. QKD provides an additional authentication layer:

\begin{itemize}
\item \textbf{Pre-established QKD links}: Validators with fiber-connected QKD appliances establish shared keys.
\item \textbf{Consensus message authentication}: HMAC using QKD-derived key appended alongside BLS signatures.
\item \textbf{Sybil resistance}: QKD links are physical---an adversary cannot create virtual QKD connections without quantum hardware and fiber access.
\item \textbf{Partial deployment}: Validators without QKD hardware continue using PQ-hybrid authentication. The system degrades gracefully.
\end{itemize}

\subsection{Sensor-Attested Block Production}

Blocks can include quantum sensor attestations as special transactions:

\begin{itemize}
\item Atomic clock timestamps replace OS timestamps for sub-microsecond ordering.
\item Magnetometer readings provide physical-layer environmental context.
\item All sensor data receives Quasar dual-signature (BLS + Ringtail) for immutable provenance.
\item Sensor attestation chain provides tamper-evident physical security monitoring.
\end{itemize}

% ============================================================================
\section{KMS Integration}
% ============================================================================

\subsection{QKD Key Import}

The Sovereign KMS Transit Engine supports key import via \texttt{importKeyMaterial()}:

\begin{itemize}
\item QKD-derived keys imported as a new key version type: \texttt{hanzo-qkd:v\{N\}:\{base64\}}.
\item Keys stored with the same per-org isolation as algorithmic keys.
\item QKD keys can serve as root keys for Shamir splitting, providing physics-layer root-of-trust.
\item HPKE wrapping extended to include QKD key material alongside ML-KEM-768.
\end{itemize}

\subsection{Key Hierarchy with QKD Root}

\begin{table}[H]
\centering
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Layer} & \textbf{Key Source} & \textbf{Security Basis} & \textbf{Rotation} \\
\midrule
Root & QKD + Shamir & Physics + threshold & Ceremony \\
Organization & AES-256 (QKD-derived) & Physics-layer & Policy \\
Session & Triple-hybrid KDF & Classical + PQ + physics & Per-session \\
Message & ML-KEM ephemeral & Algorithmic PQ & Per-message \\
\bottomrule
\end{tabular}
\caption{Key hierarchy with QKD-anchored root of trust.}
\label{tab:key-hierarchy-qkd}
\end{table}

% ============================================================================
\section{Zero-Trust Fabric Integration}
% ============================================================================

\subsection{QKD Transport Underlay}

The Hanzo ZT fabric supports pluggable transport underlays. A QKD-authenticated underlay implements the existing \texttt{Transport.Connection} interface:

\begin{itemize}
\item \textbf{Physical link}: Fiber-optic QKD channel between zero-trust nodes.
\item \textbf{Key delivery}: ETSI GS QKD 014 provides symmetric keys.
\item \textbf{Data encryption}: AES-256-GCM using QKD-derived keys on the classical channel.
\item \textbf{Authentication}: QKD keys authenticate zero-trust identity assertions.
\item \textbf{Policy enforcement}: Zero-trust policies applied after QKD authentication---QKD provides the channel, ZT provides the authorization.
\end{itemize}

% ============================================================================
\section{Standards Compliance}
% ============================================================================

\begin{table}[H]
\centering
\small
\begin{tabular}{@{}llp{5cm}@{}}
\toprule
\textbf{Standard} & \textbf{Scope} & \textbf{Compliance} \\
\midrule
ETSI GS QKD 004 & Session-based key delivery & Client implementation \\
ETSI GS QKD 014 & REST key delivery API & Client implementation \\
ITU-T Y.3800 & QKD network overview & Functional architecture \\
ITU-T Y.3802 & QKD functional architecture & Key management layer \\
NIST SP 800-90B & Entropy source validation & QRNG qualification \\
\fips{203} (ML-KEM) & Key encapsulation & Production (existing) \\
\fips{204} (ML-DSA) & Digital signatures & Production (existing) \\
\fips{205} (SLH-DSA) & Hash-based signatures & Production (existing) \\
CNSA 2.0 & NSA algorithm suite & Full compliance \\
\bottomrule
\end{tabular}
\caption{Standards compliance matrix.}
\label{tab:standards}
\end{table}

% ============================================================================
\section{Naval Deployment Scenarios}
% ============================================================================

\subsection{Shore-to-Ship QKD via Submarine Fiber}

\begin{itemize}
\item QKD appliances at shore facility and pier-side ship terminal.
\item Pre-load QKD key pool before departure (keys consumed during deployment).
\item Triple-hybrid key derivation for all ship-to-shore encrypted communications.
\item Key pool size determines operational endurance without QKD refresh.
\end{itemize}

\subsection{Satellite QKD for Global Coverage}

\begin{itemize}
\item Satellite-based QKD (demonstrated by Micius satellite at 1,200 km) for global key distribution.
\item Trusted-node relay architecture for intercontinental links.
\item QKD keys distributed to fleet units via satellite pass windows.
\item Stored keys consumed for consensus authentication between passes.
\end{itemize}

\subsection{Submarine Quantum Sensing Array}

\begin{itemize}
\item Distributed quantum magnetometer array for ASW (anti-submarine warfare).
\item Sensor attestations transmitted via RNS mesh (acoustic/ELF radio).
\item Blockchain-anchored magnetic anomaly data for post-mission analysis.
\item Tamper-evident sensor chain prevents adversary data injection.
\end{itemize}

\subsection{GPS-Denied Navigation}

\begin{itemize}
\item Quantum atomic clocks provide autonomous timing without GPS.
\item Quantum gravimeters enable terrain-referenced navigation underwater.
\item Quantum magnetometers support magnetic-field navigation.
\item All sensor data attested on Q-Chain with dual-threshold signatures.
\end{itemize}

% ============================================================================
\section{Implementation Roadmap}
% ============================================================================

\begin{table}[H]
\centering
\begin{tabular}{@{}llrl@{}}
\toprule
\textbf{Phase} & \textbf{Component} & \textbf{LOC} & \textbf{Timeline} \\
\midrule
1 & QRNG entropy integration & $\sim$200 & Q2 2026 (planned) \\
2 & ETSI GS QKD 014 client & $\sim$800 & Q3 2026 (planned) \\
3 & QKD-KEM bridge & $\sim$400 & Q3 2026 (planned) \\
4 & Triple-hybrid key derivation & $\sim$300 & Q3 2026 (planned) \\
5 & Sensor Oracle (QuantumVM) & $\sim$1,000 & Q4 2026 (planned) \\
6 & QKD Network Manager & $\sim$1,500 & Q1 2027 (planned) \\
7 & Zero-trust QKD underlay & $\sim$800 & Q1 2027 (planned) \\
\midrule
& \textbf{Total} & \textbf{$\sim$5,000} & \\
\bottomrule
\end{tabular}
\caption{Implementation roadmap: 5,000 lines of Go leveraging existing interfaces.}
\label{tab:roadmap}
\end{table}

All components are additive---they implement existing interfaces (\texttt{kem.KEM}, \texttt{Transaction}, \texttt{crypto/rand}, \texttt{Transport.Connection}) without modifying production code.

% ============================================================================
\section{Competitive Analysis}
% ============================================================================

\begin{table}[H]
\centering
\small
\begin{tabular}{@{}lcccc@{}}
\toprule
\textbf{Capability} & \textbf{Ours} & \textbf{QAN} & \textbf{IOTA} & \textbf{Algorand} \\
\midrule
PQ signatures (prod) & \checkmark & Planned & -- & Partial \\
PQ key exchange (prod) & \checkmark & -- & -- & -- \\
Threshold PQ (Ringtail) & \checkmark & -- & -- & -- \\
QKD integration & Design & -- & -- & -- \\
Quantum sensing & Design & -- & -- & -- \\
Triple-hybrid KDF & Design & -- & -- & -- \\
Mesh/DDIL transport & \checkmark & -- & Partial & -- \\
Zero-trust fabric & \checkmark & -- & -- & -- \\
\bottomrule
\end{tabular}
\caption{Competitive landscape: no competitor has PQ consensus, let alone QKD/sensing.}
\label{tab:competitive}
\end{table}

% ============================================================================
\section{Conclusion}
% ============================================================================

We have presented an architecture for integrating quantum key distribution and quantum sensing into a production post-quantum blockchain stack. The triple-hybrid key derivation---combining classical ECDH, lattice-based KEM, and physics-layer QKD---provides defense-in-depth across three independent security foundations.

The architecture is uniquely feasible because the production stack already provides the integration surfaces: a pluggable KEM interface for QKD keys, a QuantumVM for sensor attestations, a mesh transport for DDIL deployment, and a zero-trust fabric for QKD-authenticated channels. The estimated 5,000 lines of new code leverage these existing interfaces without modifying production systems.

No competing blockchain or military networking system offers this combination of production post-quantum cryptography, QKD integration architecture, and quantum sensor attestation---let alone over a delay-tolerant mesh network operating on LoRa, packet radio, and satellite links.

% ============================================================================
\begin{thebibliography}{20}

\bibitem{fips203} NIST, ``ML-KEM Standard,'' FIPS 203, 2024.

\bibitem{fips204} NIST, ``ML-DSA Standard,'' FIPS 204, 2024.

\bibitem{fips205} NIST, ``SLH-DSA Standard,'' FIPS 205, 2024.

\bibitem{cnsa2} NSA, ``CNSA Suite 2.0,'' 2022.

\bibitem{etsi014} ETSI, ``QKD; Protocol and data format of REST-based key delivery API,'' GS QKD 014 V1.1.1, 2019.

\bibitem{etsi004} ETSI, ``QKD; Application Interface,'' GS QKD 004 V2.1.1, 2020.

\bibitem{ituy3800} ITU-T, ``Overview on Networks Supporting Quantum Key Distribution,'' Y.3800, 2019.

\bibitem{ituy3802} ITU-T, ``Quantum Key Distribution Networks -- Functional Architecture,'' Y.3802, 2020.

\bibitem{nist90b} NIST, ``Recommendation for the Entropy Sources Used for Random Bit Generation,'' SP 800-90B, 2018.

\bibitem{bb84} C. Bennett, G. Brassard, ``Quantum Cryptography: Public Key Distribution and Coin Tossing,'' TCS 2014 (orig.\ 1984).

\bibitem{tfqkd} M. Lucamarini \etal, ``Overcoming the Rate-Distance Limit of QKD without Quantum Repeaters,'' Nature 557, 2018.

\bibitem{micius} S.-K. Liao \etal, ``Satellite-to-Ground Quantum Key Distribution,'' Nature 549, 2017.

\bibitem{qkdblockchain} ``Quantum Secured Blockchain Framework,'' Nature Scientific Reports, 2025.

\bibitem{tfqkdbc} ``Protecting Distributed Blockchain with Twin-Field QKD,'' arXiv:2603.14826, 2026.

\bibitem{ringtail} ``Practical Two-Round Threshold Signature from LWE,'' IACR ePrint 2024/1113.

\bibitem{quasar} Hanzo Team, ``Quasar Consensus,'' Lux Industries, 2025.

\bibitem{infleqtion} Infleqtion, ``Tiqker Quantum Atomic Clock,'' CES 2026.

\bibitem{shamir} A. Shamir, ``How to Share a Secret,'' CACM, 1979.

\end{thebibliography}

\end{document}
