\section{Motivation}


Hanzo Base (HIP Base Runtime) packages an encrypted SQLite database per
organization into a single Go binary. This architecture delivers
sub-millisecond CRUD latency, per-tenant isolation by construction, and zero
external database dependencies. It is the right architecture for sovereign
applications.

It has one gap: durability on ephemeral infrastructure.

Kubernetes pods are ephemeral. A node drain, an OOM kill, a failed health
check, or a routine deployment all destroy local state. Current Base
deployments use either \texttt{emptyDir} volumes (state lost on every
restart) or PersistentVolumeClaims (state survives restart but not node
failure, and PVCs introduce scheduling constraints that conflict with
horizontal scaling).

The consequences are concrete:

\begin{enumerate}[nosep]
    \item \textbf{Data loss on restart.} An \texttt{emptyDir} deployment of
    ATS loses all trade records, order history, and portfolio state when the
    pod restarts. This is unacceptable for a regulated securities exchange.

    \item \textbf{Node affinity locks.} A PVC-backed deployment pins a pod to
    one node. If that node fails, the pod cannot reschedule until the volume
    is manually detached or the cloud provider's volume controller completes
    failover---typically 2--10 minutes on GKE.

    \item \textbf{No disaster recovery.} Neither \texttt{emptyDir} nor PVC
    provides off-site backup. A storage controller bug, a zone outage, or
    accidental \texttt{kubectl delete pvc} destroys all state with no
    recovery path.

    \item \textbf{No cross-region replication.} PVCs are zone-scoped. A zone
    failure takes down both the pod and its data. Regional redundancy requires
    an external replication mechanism.
\end{enumerate}

This HIP solves all four problems with a single mechanism: continuous WAL
replication to S3-compatible object storage, encrypted with age before
leaving the pod.

% ============================================================================
% 2. BACKGROUND
% ============================================================================
