\section{Post-Quantum Security}


The Hanzo ecosystem has achieved comprehensive post-quantum (PQ)
coverage across all cryptographic layers. All previously planned
mitigations are now deployed.

\subsection{PQ Scorecard}

\begin{table}[H]
\centering
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Layer} & \textbf{Algorithm} & \textbf{PQ Status} & \textbf{Status} \\
\midrule
Disk encryption     & AES-256 (sqlcipher)          & Safe (128-bit PQ)    & Deployed \\
HKDF derivation     & HKDF-SHA-256                 & Safe (128-bit PQ)    & Deployed \\
Field encryption    & AES-256-GCM                  & Safe (128-bit PQ)    & Deployed \\
S3 backup           & age ML-KEM-768+X25519 (\texttt{age1pq}) & Safe (FIPS 203) & Deployed \\
TLS                 & X25519MLKEM768 (first curve) & Safe (hybrid PQ)     & Deployed \\
JWT signing         & ML-DSA-65 (FIPS 204)         & Safe (Module-LWE+SIS)& Deployed \\
JWKS validation     & ML-DSA-65                    & Safe                 & Deployed \\
MPC transport       & Hybrid KEM (X25519+ML-KEM-768) & Safe              & Deployed \\
MPC key shares      & PQ KEM encryption            & Safe                 & Deployed \\
Master keys         & Cloud HSM (FIPS 140-2 L3)    & Hardware isolation   & Deployed \\
Chain signing       & secp256k1 ECDSA              & \textbf{Not PQ-safe} & EVM constraint \\
\bottomrule
\end{tabular}
\caption{Complete post-quantum scorecard. All layers PQ-safe except EVM chain signing.}
\label{tab:pq-scorecard}
\end{table}

\subsection{S3 Backup Encryption (Deployed)}

All age encryption uses ML-KEM-768 + X25519 hybrid recipients
(\texttt{age1pq} prefix). The file key is derived from both shared
secrets: $\text{file\_key} = \text{HKDF-SHA256}(ss_1 \| ss_2,
\texttt{"age-hybrid-v1"})$. Label enforcement prevents downgrade to
classical-only recipients.

\subsection{TLS Post-Quantum (Deployed)}

All TLS 1.3 connections negotiate X25519MLKEM768 as the first curve.
This provides PQ protection for all data in transit, including S3
uploads and inter-service communication.

\subsection{JWT Signing with ML-DSA-65 (Deployed)}

Hanzo IAM issues JWT tokens signed with ML-DSA-65 (FIPS~204). The
JWKS endpoint exposes the ML-DSA-65 public key. All services validate
JWTs using the PQ-safe signature. EUF-CMA security under Module-LWE
and Module-SIS hardness assumptions.

\subsection{MPC Transport and Key Shares (Deployed)}

MPC node-to-node communication uses hybrid KEM (X25519 + ML-KEM-768).
MPC key shares at rest are encrypted with PQ KEM encryption.

\subsection{Cloud HSM for Master Keys (Deployed)}

Master encryption keys are stored in Cloud HSM (GCP Cloud KMS,
FIPS~140-2 Level~3 Cavium HSMs). The master key never leaves the HSM:
wrap, unwrap, and sign operations execute inside the HSM. Three
keyrings per ecosystem (devnet, testnet, mainnet), 12 keys total.
Mainnet uses HSM protection level.

\subsection{Harvest-Now-Decrypt-Later: Closed}

The combination of ML-KEM-768 hybrid encryption (S3), X25519MLKEM768
TLS (transit), ML-DSA-65 JWT signing (auth), PQ KEM for MPC, and
Cloud HSM for master keys closes the HNDL attack vector completely.

% ============================================================================
% 12. REFERENCES
% ============================================================================
