% Decentralized Semantic Optimization (DSO) core section
\section{Decentralized Semantic Optimization (DSO)}
\subsection{Experience Priors}
Each agent/node maintains an \emph{experience prior} \(E\): token/embedding-level memory distilled from rollouts. Locally, nodes run \textbf{Active Semantic Optimization (ASO)} to extract \emph{semantic advantages} from groups of rollouts (TF-GRPO). Priors are compressed (\S\ref{sec:bitdelta}) and written to the on-chain \emph{ExperienceRegistry} with Merkle proofs.

\subsection{Training-Free GRPO as Bayesian PoE}
For a base model with conditional \(p_\theta(y\mid x)\) and a set of experiences \(\{e_k\}\) mapping to token-level factors \(\phi_k(y\mid x)\), decoding uses a \emph{product-of-experts}:
\begin{equation}\label{eq:poe}
 p(y\mid x,E) \propto p_\theta(y\mid x)\,\prod_{k} \phi_k(y\mid x)^{\alpha_k},\quad \alpha_k\ge 0.
\end{equation}
Here \(\phi_k\) are distilled from group-relative semantic advantage; weights \(\alpha_k\) are learned by introspective calibration without gradient updates to \(\theta\).

\subsection{Distributed Aggregation}
Hanzo Network aggregates \emph{priors, not gradients}. Let node priors be \(\{E_i\}\). We publish hashes and quality scores; the chain computes a byzantine-robust aggregate \(\tilde E = \operatorname{median}_q\{E_i\}\) under a fixed schema (token bins / embedding centroids). Conflicting contributions resolve by stake-weighted quorum plus quality caps.

\section{ExperienceRegistry and P2P Sync}
\textbf{Registry.} On-chain contract stores: content-addressed CID, Merkle root, schema version, quality vector, submitter, slashing bond. \textbf{Storage.} Off-chain IPFS/Arweave; local SQLite+LanceDB with Merkle verification. \textbf{Sync.} Gossip protocol with CRDT merge; priority given to high-quality shards (fee rebates bias peers to propagate them).
