\section{Conclusion}


Hanzo Agent SDK provides a principled framework for multi-agent orchestration with type-safe contexts, hierarchical memory, and semantic routing. Our evaluation demonstrates significant improvements in task completion and token efficiency compared to existing approaches. The SDK is open-source and designed for production deployment with comprehensive observability features.

\paragraph{On-chain identity via the Hanzo AI Chain.}
For agents that need persistent identity, capability-bound authorization, durable spend caps, and on-chain reputation that follows them across jobs, the SDK integrates with the agent registry on the Hanzo AI Chain (AIVM). Each registered agent is a content-addressed record committing to a model set, a Merkle-rooted capability declaration, a per-epoch AI spend cap, and a reputation score updated from TEE-attested job outcomes. Cross-chain actions (settlement on Lux C-Chain, confidential inference on F-Chain, threshold custody on M-Chain) are mediated by the \texttt{AGENT\_AUTH} precompile, which enforces capability scope at the AIVM boundary. See \emph{Hanzo AI Chain (AIVM)} (\texttt{\textasciitilde/work/hanzo/papers/hanzo-ai-chain/}) for the registry schema, the reputation function, and worked examples; the chain activates on 25 December 2025 alongside Lux Quasar 3.0 (LP-105).

\bibliographystyle{plain}
\begin{thebibliography}{10}

\bibitem{langchain2022}
H. Chase. LangChain: Building applications with LLMs through composability. 2022.

\bibitem{autogpt2023}
T. Richards. Auto-GPT: An Autonomous GPT-4 Experiment. 2023.

\bibitem{multiagent2023}
Y. Wu et al. AutoGen: Enabling Next-Gen LLM Applications via Multi-Agent Conversation. arXiv:2308.08155, 2023.

\bibitem{mcp2022}
Anthropic. Model Context Protocol Specification. 2022.

\bibitem{swebench2023}
C. E. Jimenez et al. SWE-bench: Can Language Models Resolve Real-World GitHub Issues? arXiv:2310.06770, 2023.

\bibitem{gaia2023}
G. Mialon et al. GAIA: A Benchmark for General AI Assistants. arXiv:2311.12983, 2023.

\end{thebibliography}

