\section{Background}

\label{sec:background}

\subsection{Classical Federated Learning}

Federated learning~\cite{mcmahan2017communication} trains a shared model
across $N$ participants without centralizing raw data. The canonical
FedAvg algorithm alternates between local training steps and global parameter
averaging:
\begin{equation}
    \bm{\theta}^{t+1} = \sum_{i=1}^{N} \frac{n_i}{n} \bm{\theta}_i^{t+1},
    \quad \bm{\theta}_i^{t+1} = \bm{\theta}^t - \eta \nabla_{\bm{\theta}} \mathcal{L}_i(\bm{\theta}^t),
\end{equation}
where $n_i$ is the local dataset size and $n = \sum_i n_i$.

\paragraph{Communication Compression.}
Communication overhead is the primary bottleneck in federated learning.
\cite{seide20141bit} propose 1-bit SGD, quantizing gradient signs:
$\hat{g} = \sign(g) \cdot \|g\|_1 / d$.
\cite{bernstein2018signsgd} formalize SignSGD and prove convergence under
symmetric gradient noise. 1-bit compression reduces communication by $32\times$
versus full-precision floats.

\paragraph{Byzantine Robustness.}
Byzantine participants~\cite{lamport1982byzantine} can send arbitrary
gradients to poison the aggregate. Krum~\cite{blanchard2017machine} selects the
gradient closest to its $n - f - 2$ nearest neighbors as the aggregate.
\cite{yin2018byzantine} prove that coordinate-wise median and trimmed mean
achieve statistical rates within a factor of $f^2$ of the optimal non-Byzantine
estimator. DeltaSoup inherits this framework and adapts it to semantic
experience embeddings.

\subsection{DSO: Decentralized Semantic Optimization}

DSO~\cite{hanzo2026dso} (ZIP-001) is a protocol for sharing and aggregating
experiential priors across distributed language model agents.

\begin{definition}[Experiential Prior]
\label{def:prior}
An experiential prior $E_m$ consists of:
\begin{itemize}[leftmargin=1.5em]
    \item A natural language pattern description $\text{pattern}_m$ ($\leq 32$ words).
    \item A task embedding $\bm{e}_m \in \mathbb{R}^d$ encoding the domain.
    \item A quality score $q_m \in (0, 1)$ estimating pattern reliability.
    \item A 1-bit quantized expert factor $\widehat{\Delta}_m \in \{-1,+1\}^V$
          with scale $\alpha_m > 0$ for PoE decoding.
    \item Metadata: timestamp, source node, domain tag, iteration count.
\end{itemize}
\end{definition}

The \emph{experience library} $\mathcal{L} = \{E_1, \ldots, E_M\}$ is the
shared artifact. Agents use it for retrieve-and-apply at decode time:
\begin{equation}
    \pi_{\text{ASO}}(y_t \mid x, y_{<t}) \propto \pi_\theta(y_t \mid x, y_{<t})
    \prod_{m \in \text{retrieved}(x)} \phi_m(y_t)^{\eta_m},
    \label{eq:poe}
\end{equation}
where $\eta_m = \log(q_m / (1 - q_m))$ and $\phi_m$ is constructed from
$\widehat{\Delta}_m$.

Unlike classical federated learning, DSO shares no model parameters, no
gradients, and no raw training data. The shared artifact is entirely composed
of compressed semantic patterns in natural language plus corresponding
quantized embedding perturbations.

\subsection{BitDelta Compression}
\label{sec:background-bitdelta}

BitDelta (ZIP-007)~\cite{hanzo2026bitdelta} applies 1-bit quantization to
embedding \emph{deltas} relative to a shared baseline. For experience embedding
$\bm{e} \in \mathbb{R}^d$ and baseline $\bar{\bm{e}} \in \mathbb{R}^d$:

\begin{definition}[BitDelta Quantization]
\label{def:bitdelta}
Given delta $\Delta = \bm{e} - \bar{\bm{e}} \in \mathbb{R}^d$, the BitDelta
quantization is:
\begin{equation}
    \widehat{\Delta} = \alpha \cdot \sign(\Delta),
    \quad \alpha = \frac{1}{d} \|\Delta\|_1 = \frac{1}{d}\sum_{j=1}^d |\Delta_j|,
    \label{eq:bitdelta}
\end{equation}
where $\alpha \in \mathbb{R}_{>0}$ is the per-group scale factor and
$\sign(\Delta) \in \{-1, +1\}^d$ is the elementwise sign vector stored as
packed bits (1 bit per dimension).
\end{definition}

With group-wise scaling (group size $g = 128$), each group uses one float32
scale and $g$ bits, giving:
\begin{equation}
    \text{bits per element} = \frac{32 + g}{g} = 1 + \frac{32}{g}.
\end{equation}
For $g = 128$: $1.25$ bits/element, a $25.6\times$ reduction from float32.
Accounting for metadata overhead, practical compression is ${\approx}10\times$.

\subsection{DeltaSoup Aggregation}
\label{sec:background-deltasoup}

DeltaSoup is a Byzantine-robust aggregation protocol for community delta
contributions. Conceptually, it applies the geometric median~\cite{minsker2015geometric}
rather than the arithmetic mean, so a minority of adversarial contributors
cannot shift the aggregate arbitrarily.

Given $N$ compressed experience batches $\{\hat{B}_i\}_{i=1}^N$ received from
peers, DeltaSoup:
\begin{enumerate}[leftmargin=1.5em]
    \item Decompresses all batches to full-precision embeddings.
    \item Computes pairwise cosine similarity across all experiences.
    \item Filters experiences whose embedding lies more than $2\sigma$ from
          the coordinate-wise median (outlier rejection).
    \item Merges surviving experiences with confidence-weighted averaging.
\end{enumerate}

Formal properties are developed in Section~\ref{sec:deltasoup}.

% ============================================================================
% 3. SYSTEM ARCHITECTURE
% ============================================================================
