\section{BitDelta Experience Compression}

\label{sec:bitdelta}

\subsection{Embedding Quantization}

Let $\mathcal{L} = \{E_1, \ldots, E_M\}$ be the local experience library with
embeddings $\{\bm{e}_1, \ldots, \bm{e}_M\} \subset \mathbb{R}^d$. The BitDelta
compression operates as follows.

\begin{definition}[Library Baseline]
The baseline embedding $\bar{\bm{e}} \in \mathbb{R}^d$ is the mean embedding
of the current experience library:
\begin{equation}
    \bar{\bm{e}} = \frac{1}{M} \sum_{m=1}^{M} \bm{e}_m.
    \label{eq:baseline}
\end{equation}
The baseline is recomputed before each broadcast and shared (uncompressed)
with the batch header.
\end{definition}

\begin{definition}[Embedding Delta]
For experience $E_m$ with embedding $\bm{e}_m$, the embedding delta is:
\begin{equation}
    \Delta_m = \bm{e}_m - \bar{\bm{e}} \in \mathbb{R}^d.
    \label{eq:delta}
\end{equation}
\end{definition}

Following Definition~\ref{def:bitdelta}, we partition $\Delta_m$ into
$G = d / g$ groups $\Delta_m^{(1)}, \ldots, \Delta_m^{(G)} \in \mathbb{R}^g$
and apply independent scaling per group:

\begin{equation}
    \alpha_m^{(j)} = \frac{1}{g} \left\|\Delta_m^{(j)}\right\|_1,
    \qquad
    \bm{s}_m^{(j)} = \sign\!\left(\Delta_m^{(j)}\right) \in \{-1,+1\}^g.
    \label{eq:group-quantize}
\end{equation}

The compressed representation is $\widehat{E}_m = \left(\{\bm{s}_m^{(j)}, \alpha_m^{(j)}\}_{j=1}^G\right)$,
stored as packed bits plus $G$ float32 scales.

\begin{definition}[Reconstruction]
Given compressed representation $\widehat{E}_m$, reconstruct:
\begin{equation}
    \hat{\Delta}_m^{(j)} = \alpha_m^{(j)} \cdot \bm{s}_m^{(j)},
    \qquad
    \hat{\bm{e}}_m = \bar{\bm{e}} + \left[\hat{\Delta}_m^{(1)}; \cdots; \hat{\Delta}_m^{(G)}\right].
    \label{eq:reconstruct}
\end{equation}
\end{definition}

\subsection{Compression Analysis}

\begin{proposition}[Compression Ratio]
\label{prop:compression}
For embedding dimension $d$ and group size $g$, the compression ratio of
BitDelta relative to float32 storage is:
\begin{equation}
    \rho = \frac{32d}{d + 32 \cdot d/g} = \frac{32g}{g + 32}.
    \label{eq:compression-ratio}
\end{equation}
As $g \to \infty$: $\rho \to 32$. For $g = 128$: $\rho = 32 \cdot 128 / 160 = 25.6$.
Including the float32 baseline vector (shared per-batch, not per-experience):
effective per-experience compression for $M = 100$ experiences is ${\approx}10\times$.
\end{proposition}

\begin{proof}
Float32 storage requires $32d$ bits. BitDelta requires $d$ bits (signs) plus
$32 \cdot (d/g)$ bits (scales). The ratio is $32d / (d + 32d/g) = 32g/(g+32)$.
The baseline amortizes over $M$ experiences: overhead per experience is $32d / M$
bits, which is $32 \cdot 768 / 100 \approx 246$ bits, or ${\approx}5\%$ overhead.
\end{proof}

\begin{theorem}[Reconstruction Error Bound]
\label{thm:reconstruction}
Let $\Delta \in \mathbb{R}^d$ be an embedding delta partitioned into groups
$\Delta^{(1)}, \ldots, \Delta^{(G)} \in \mathbb{R}^g$. For group $j$, the
reconstruction error satisfies:
\begin{equation}
    \left\|\Delta^{(j)} - \hat{\Delta}^{(j)}\right\|_2^2
    = \sum_{k=1}^{g} \left(\Delta_k^{(j)} - \alpha^{(j)} \cdot \sign(\Delta_k^{(j)})\right)^2
    = \sum_{k=1}^{g} \left(|\Delta_k^{(j)}| - \alpha^{(j)}\right)^2.
    \label{eq:group-error}
\end{equation}
Setting $\alpha^{(j)} = \frac{1}{g}\|\Delta^{(j)}\|_1$ minimizes
$\sum_k (|\Delta_k^{(j)}| - \alpha^{(j)})^2$ over all scalar $\alpha^{(j)} \geq 0$,
yielding:
\begin{equation}
    \left\|\Delta^{(j)} - \hat{\Delta}^{(j)}\right\|_2^2
    = \sum_{k=1}^{g} \left(|\Delta_k^{(j)}| - \bar{|\Delta|}^{(j)}\right)^2
    = g \cdot \mathrm{Var}\!\left[|\Delta^{(j)}|\right],
    \label{eq:var-error}
\end{equation}
where $\bar{|\Delta|}^{(j)} = \frac{1}{g}\sum_k |\Delta_k^{(j)}|$ and the variance
is over coordinates. The total reconstruction error across all groups is:
\begin{equation}
    \|\Delta - \hat{\Delta}\|_2^2 = \sum_{j=1}^{G} g \cdot \mathrm{Var}\!\left[|\Delta^{(j)}|\right]
    = d \cdot \overline{\mathrm{Var}}\!\left[|\Delta|\right],
\end{equation}
where $\overline{\mathrm{Var}}$ is the mean per-group variance of the absolute delta.
\end{theorem}

\begin{proof}
From Eq.~\eqref{eq:group-error}, each term is $(|\Delta_k| - \alpha)^2$ where
$\alpha$ minimizes the sum of squared deviations from the absolute values.
By the first-order optimality condition:
$\frac{\partial}{\partial \alpha} \sum_k (|\Delta_k| - \alpha)^2 = -2\sum_k(|\Delta_k| - \alpha) = 0$,
giving $\alpha = \frac{1}{g}\|\Delta^{(j)}\|_1 = \bar{|\Delta|}^{(j)}$.
Substituting: $\sum_k (|\Delta_k| - \bar{|\Delta|})^2 = g \cdot \mathrm{Var}[|\Delta^{(j)}|]$.
Summing over $G$ groups completes the proof.
\end{proof}

\begin{corollary}[Cosine Similarity Preservation]
\label{cor:cosine}
Let $\cos(\bm{e}, \hat{\bm{e}})$ denote cosine similarity between original
and reconstructed embeddings. Under the assumption that $\|\Delta\|_2 \ll
\|\bar{\bm{e}}\|_2$ (experience embeddings cluster near the baseline), the
expected cosine similarity satisfies:
\begin{equation}
    \mathbb{E}\left[\cos(\bm{e}_m, \hat{\bm{e}}_m)\right]
    \geq 1 - \frac{d \cdot \overline{\mathrm{Var}}[|\Delta|]}{2\|\bar{\bm{e}}\|_2^2}.
\end{equation}
For well-clustered libraries where $\overline{\mathrm{Var}}[|\Delta|] \leq 0.01$
(empirically observed for sentence-transformer embeddings), this gives
$\cos(\bm{e}_m, \hat{\bm{e}}_m) \geq 0.95$---i.e., less than 5\% semantic
similarity loss.
\end{corollary}

\begin{proof}
Using the identity $\cos(\bm{u}, \bm{v}) = 1 - \frac{\|\bm{u} - \bm{v}\|^2}{2\|\bm{u}\|\|\bm{v}\|}$
for unit-normalized vectors, and bounding $\|\bm{e}_m - \hat{\bm{e}}_m\|_2^2 =
\|\Delta_m - \hat{\Delta}_m\|_2^2 = d \cdot \overline{\mathrm{Var}}[|\Delta_m|]$
from Theorem~\ref{thm:reconstruction}. Since $\|\bm{e}_m\|_2 \approx
\|\hat{\bm{e}}_m\|_2 \approx \|\bar{\bm{e}}\|_2$, the result follows.
\end{proof}

\subsection{Algorithms}

\begin{algorithm}[H]
\caption{\textsc{BitDelta-Compress}}
\label{alg:compress}
\begin{algorithmic}[1]
\Require Experience library $\mathcal{L} = \{E_m\}_{m=1}^M$, group size $g$
\Ensure Baseline $\bar{\bm{e}}$, compressed representations $\{\widehat{E}_m\}_{m=1}^M$
\State $\bar{\bm{e}} \gets \frac{1}{M}\sum_{m=1}^M \bm{e}_m$ \Comment{Compute library baseline}
\State $G \gets d / g$ \Comment{Number of groups}
\For{$m = 1, \ldots, M$}
    \State $\Delta_m \gets \bm{e}_m - \bar{\bm{e}}$ \Comment{Embedding delta}
    \For{$j = 1, \ldots, G$}
        \State Extract group: $\Delta_m^{(j)} \gets \Delta_m[(j-1)g : jg]$
        \State $\alpha_m^{(j)} \gets \frac{1}{g}\left\|\Delta_m^{(j)}\right\|_1$
              \Comment{L1 mean scale}
        \State $\bm{s}_m^{(j)} \gets \sign\!\left(\Delta_m^{(j)}\right)$
              \Comment{1-bit sign vector}
        \State Pack $\bm{s}_m^{(j)}$ into $\lceil g/32 \rceil$ uint32 words
    \EndFor
    \State $\widehat{E}_m \gets \left(\{\bm{s}_m^{(j)}, \alpha_m^{(j)}\}_{j=1}^G,\;
           E_m.\text{text},\; E_m.q,\; E_m.\text{domain}\right)$
\EndFor
\State \Return $\bar{\bm{e}}$, $\{\widehat{E}_m\}_{m=1}^M$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
\caption{\textsc{BitDelta-Decompress}}
\label{alg:decompress}
\begin{algorithmic}[1]
\Require Baseline $\bar{\bm{e}}$, compressed batch $\{\widehat{E}_m\}_{m=1}^M$, group size $g$
\Ensure Reconstructed embeddings $\{\hat{\bm{e}}_m\}_{m=1}^M$
\State $G \gets d / g$
\For{$m = 1, \ldots, M$}
    \State $\hat{\Delta}_m \gets \bm{0}^d$
    \For{$j = 1, \ldots, G$}
        \State Unpack $\bm{s}_m^{(j)}$ from uint32 words to $\{-1,+1\}^g$
        \State $\hat{\Delta}_m[(j-1)g : jg] \gets \alpha_m^{(j)} \cdot \bm{s}_m^{(j)}$
    \EndFor
    \State $\hat{\bm{e}}_m \gets \bar{\bm{e}} + \hat{\Delta}_m$
\EndFor
\State \Return $\{\hat{\bm{e}}_m\}_{m=1}^M$
\end{algorithmic}
\end{algorithm}

% ============================================================================
% 5. DELTASOUP: BYZANTINE-ROBUST AGGREGATION
% ============================================================================
