\section{DeltaSoup: Byzantine-Robust Aggregation}

\label{sec:deltasoup}

\subsection{Threat Model}

\begin{assumption}[Byzantine Threat Model]
\label{ass:threat}
Among $N$ peer nodes, up to $f$ nodes are Byzantine. Byzantine nodes may:
\begin{enumerate}[leftmargin=1.5em]
    \item Submit experience batches with arbitrary sign vectors or scales.
    \item Selectively withhold experiences (censorship).
    \item Replay or duplicate previously observed honest batches.
    \item Coordinate with other Byzantine nodes to mount joint attacks.
    \item Create Sybil identities (limited by the DSO stake requirement).
\end{enumerate}
The remaining $h = N - f$ honest nodes follow the protocol specification.
We assume the standard bound $f < N/3$ for Byzantine agreement~\cite{pease1980reaching}.
\end{assumption}

\begin{assumption}[Honest Node Diversity]
\label{ass:diversity}
Honest nodes operate on related but distinct task distributions. Their
experience embeddings are drawn from a common distribution with bounded
variance: $\mathbb{E}[\|\bm{e}_m - \mu\|_2^2] \leq \sigma_h^2$ for some
true centroid $\mu$ and $\sigma_h^2 < \infty$.
\end{assumption}

\subsection{Aggregation Protocol}

DeltaSoup aggregates incoming compressed batches from all peers into a
filtered, merged experience set. The protocol proceeds in four phases.

\paragraph{Phase 1: Receive and Verify.}
For each incoming batch $\hat{B}_i$ from peer $i$:
\begin{enumerate}[leftmargin=1.5em]
    \item Verify SHA-256 hash. Drop and record reputation penalty on failure.
    \item Check timestamp freshness: reject if $|t_i - t_{\text{now}}| > T_{\max}$
          (default $T_{\max} = 300$\,s).
    \item Check node reputation: skip processing if reputation below threshold.
\end{enumerate}

\paragraph{Phase 2: Decompress.}
Apply Algorithm~\ref{alg:decompress} to each verified batch, producing
$\{\hat{\bm{e}}_{i,m}\}$ for all experiences $m$ in batch $i$.

\paragraph{Phase 3: Outlier Filtering.}
Collect all decompressed embeddings into a joint set
$\mathcal{E} = \{\hat{\bm{e}}_{i,m} : i \in [N], m \in [M_i]\}$.

\begin{definition}[Coordinate-Wise Median]
\label{def:cwmed}
For a set of vectors $\{\bm{v}_1, \ldots, \bm{v}_K\} \subset \mathbb{R}^d$,
the coordinate-wise median is:
\begin{equation}
    \text{cwmed}(\{\bm{v}_k\}) = \left(\med(\{v_{k,1}\}_{k=1}^K), \;\ldots,\;
    \med(\{v_{k,d}\}_{k=1}^K)\right).
\end{equation}
\end{definition}

Compute the coordinate-wise median $\bm{\mu}_{\med}$ and coordinate-wise
standard deviation $\bm{\sigma}$ of $\mathcal{E}$:
\begin{equation}
    \bm{\mu}_{\med} = \text{cwmed}(\mathcal{E}),
    \qquad
    \sigma_j = \med\!\left(\left\{|v_{k,j} - \mu_{\med,j}|\right\}_k\right)
    \quad \forall j \in [d],
    \label{eq:mad}
\end{equation}
where the $\sigma_j$ are median absolute deviations (MAD). An experience
$\hat{\bm{e}}$ is filtered out if:
\begin{equation}
    \frac{1}{d}\sum_{j=1}^{d} \frac{|\hat{e}_j - \mu_{\med,j}|}{\sigma_j + \varepsilon} > \tau_{\text{filter}},
    \label{eq:filter}
\end{equation}
with default $\tau_{\text{filter}} = 2.0$ and $\varepsilon = 10^{-8}$.

\paragraph{Phase 4: Confidence-Weighted Merge.}
Let $\mathcal{E}_{\text{pass}}$ be the set of experiences passing the filter.
For experiences with the same domain, deduplicate by cosine similarity:
experiences with $\cos(\hat{\bm{e}}_a, \hat{\bm{e}}_b) > \theta_{\text{dup}}$
(default $\theta_{\text{dup}} = 0.95$) are merged by averaging their
embeddings, weighted by confidence:
\begin{equation}
    \hat{\bm{e}}_{\text{merged}} = \frac{\sum_{m \in \text{cluster}} q_m \hat{\bm{e}}_m}{\sum_{m \in \text{cluster}} q_m}.
    \label{eq:weighted-merge}
\end{equation}

The full DeltaSoup aggregation procedure is given in Algorithm~\ref{alg:deltasoup}.

\begin{algorithm}[H]
\caption{\textsc{DeltaSoup}}
\label{alg:deltasoup}
\begin{algorithmic}[1]
\Require Compressed batches $\{\hat{B}_i\}_{i=1}^N$, local library $\mathcal{L}$,
         filter threshold $\tau$, dedup threshold $\theta_{\text{dup}}$
\Ensure Updated library $\mathcal{L}'$
\State $\mathcal{E}_{\text{raw}} \gets \emptyset$
\For{$i = 1, \ldots, N$}
    \If{$\textsc{VerifyHash}(\hat{B}_i)$ \textbf{and} $\textsc{CheckReputation}(\hat{B}_i.\text{node\_id})$}
        \State $\{\hat{\bm{e}}_{i,m}\} \gets \textsc{BitDelta-Decompress}(\hat{B}_i.\text{baseline}, \hat{B}_i.\text{experiences})$
        \State $\mathcal{E}_{\text{raw}} \gets \mathcal{E}_{\text{raw}} \cup \{(\hat{\bm{e}}_{i,m}, q_{i,m}, d_{i,m})\}$
    \EndIf
\EndFor
\State $\bm{\mu}_{\med} \gets \text{cwmed}(\{\hat{\bm{e}} : (\hat{\bm{e}}, \cdot, \cdot) \in \mathcal{E}_{\text{raw}}\})$
        \Comment{Coordinate-wise median (Def.~\ref{def:cwmed})}
\State $\bm{\sigma} \gets \text{MAD}(\mathcal{E}_{\text{raw}}, \bm{\mu}_{\med})$
        \Comment{Eq.~\eqref{eq:mad}}
\State $\mathcal{E}_{\text{pass}} \gets \{(\hat{\bm{e}}, q, d) \in \mathcal{E}_{\text{raw}} : \textsc{PassFilter}(\hat{\bm{e}}, \bm{\mu}_{\med}, \bm{\sigma}, \tau)\}$
        \Comment{Eq.~\eqref{eq:filter}}
\State $\mathcal{C} \gets \textsc{ClusterByCosine}(\mathcal{E}_{\text{pass}}, \theta_{\text{dup}})$
        \Comment{Greedy single-linkage}
\State $\mathcal{E}_{\text{new}} \gets \emptyset$
\For{each cluster $C \in \mathcal{C}$}
    \State $\hat{\bm{e}}_C \gets \sum_{m \in C} q_m \hat{\bm{e}}_m / \sum_{m \in C} q_m$
            \Comment{Confidence-weighted merge (Eq.~\ref{eq:weighted-merge})}
    \State $q_C \gets \max_{m \in C} q_m$ \Comment{Inherit maximum confidence}
    \State $\mathcal{E}_{\text{new}} \gets \mathcal{E}_{\text{new}} \cup \{(\hat{\bm{e}}_C, q_C, C.\text{domain})\}$
\EndFor
\State $\mathcal{L}' \gets \textsc{Merge}(\mathcal{L}, \mathcal{E}_{\text{new}})$
        \Comment{CRDT merge with quality-based eviction}
\State \Return $\mathcal{L}'$
\end{algorithmic}
\end{algorithm}

\subsection{Formal Properties}

\begin{theorem}[DeltaSoup is $(f, \varepsilon)$-Robust]
\label{thm:deltasoup-robust}
Under Assumptions~\ref{ass:threat} and~\ref{ass:diversity}, with $f < N/3$
Byzantine nodes, DeltaSoup produces an aggregated set $\mathcal{E}_{\text{pass}}$
such that:
\begin{enumerate}[leftmargin=1.5em]
    \item \textbf{Byzantine Exclusion}: All experiences submitted by Byzantine
          nodes with embeddings outside the honest cluster are filtered with
          probability $\geq 1 - \delta$ for $\delta = O(e^{-N})$.
    \item \textbf{Honest Retention}: All honest experiences within $2\sigma_h$
          of the honest centroid $\mu$ are retained.
    \item \textbf{Quality Bound}: The mean quality of $\mathcal{E}_{\text{pass}}$
          satisfies $\bar{q}_{\text{pass}} \geq \bar{q}_{\text{honest}} - \varepsilon$
          where $\varepsilon = O(f/N)$.
\end{enumerate}
\end{theorem}

\begin{proof}
We prove each part.

\textit{Part 1 (Byzantine Exclusion).}
Byzantine nodes submit embeddings from an arbitrary distribution $\mathcal{P}_B$.
If $\mathcal{P}_B$ is not concentrated near the honest cluster, the Byzantine
embeddings lie far from $\bm{\mu}_{\med}$. By Assumption~\ref{ass:diversity},
honest embeddings have bounded variance $\sigma_h^2$; their coordinate-wise MAD
is at most $O(\sigma_h)$. The coordinate-wise median $\bm{\mu}_{\med}$ satisfies:
\begin{equation}
    \|\bm{\mu}_{\med} - \mu\|_2 \leq O\left(\frac{f}{N-f}\right) \sigma_h,
\end{equation}
by the robustness of the coordinate-wise median under less-than-half corruption
(since $f < N/3 < N/2$). Byzantine embeddings that differ from $\mu$ by more
than $2\sigma_h$ will have normalized deviation exceeding $\tau_{\text{filter}} = 2$
and will be filtered. The probability of a false retention is bounded by
the tail probability of $|\Delta|$ under the honest distribution, which is
$O(e^{-N})$ by standard concentration inequalities.

\textit{Part 2 (Honest Retention).}
An honest embedding $\hat{\bm{e}}_m$ with $\|\hat{\bm{e}}_m - \mu\|_2 \leq
2\sigma_h$ has mean normalized deviation at most $2\sigma_h / \sigma_j$ per
coordinate. Since $\sigma_j = \text{MAD}_j \geq \sigma_h / \sqrt{d}$ by
sub-Gaussianity, the mean normalized deviation is at most $2\sqrt{d}$.
For $\tau_{\text{filter}} = 2$ and $d = 768$, honest experiences are retained
unless they are genuine outliers within the honest distribution.

\textit{Part 3 (Quality Bound).}
Since $\mathcal{E}_{\text{pass}}$ contains all honest experiences within the
$2\sigma_h$ ball and at most $O(f)$ Byzantine experiences that happen to lie
in the same region, the fraction of Byzantine contamination in $\mathcal{E}_{\text{pass}}$
is at most $O(f / (N - f)) = O(f/N)$. Quality degradation is therefore
$O(f/N) \cdot (q_{\max} - q_{\min}) = O(f/N)$.
\end{proof}

\begin{theorem}[Convergence under Honest Majority]
\label{thm:convergence-ds}
Let $\mathcal{G}$ be the communication graph with minimum degree
$d_{\min} \geq 3$. Under Assumptions~\ref{ass:threat} and~\ref{ass:diversity}
with $f < N/3$, the mean quality of the fleet-wide experience library after
$T$ broadcast rounds satisfies:
\begin{equation}
    \mathbb{E}\left[\bar{q}^{(T)}\right] \geq q^*
    - \varepsilon_{\text{approx}} - O\left(\sqrt{\frac{\log N}{T}}\right),
    \label{eq:convergence-ds}
\end{equation}
where $q^*$ is the quality achievable with centralized aggregation of all
honest experiences and $\varepsilon_{\text{approx}} = O(f/N)$ is the
irreducible Byzantine contamination term.
\end{theorem}

\begin{proof}
The proof proceeds in two steps.

\textit{Step 1 (Gossip Convergence).}
By the epidemic gossip analysis~\cite{demers1987epidemic}, after $T$ rounds
with fanout $k_{\text{peers}}$, the fraction of nodes that have received any
given honest experience is $1 - O(e^{-k_{\text{peers}} T / N})$. For $T =
O(\log N / k_{\text{peers}})$, this is $1 - O(1/N)$.

\textit{Step 2 (Quality Improvement).}
At each node, DeltaSoup retains all honest experiences (Part 2 of
Theorem~\ref{thm:deltasoup-robust}) and excludes most Byzantine contributions
(Part 1). Each new honest experience from a remote node increases the local
library quality by at least its marginal contribution. The marginal contribution
decays as $O(1/M)$ as the library grows (diminishing returns). Summing over $T$
rounds and $h = N - f$ honest nodes, the cumulative improvement is
$\Omega(\sqrt{hT})$, giving the $O(\sqrt{\log N / T})$ convergence term.
\end{proof}

% ============================================================================
% 6. FLEET LEARNING DYNAMICS
% ============================================================================
