\section{Privacy and Security}

\label{sec:privacy}

\subsection{What Is and Is Not Shared}

\paragraph{Shared (non-sensitive):}
\begin{itemize}[leftmargin=1.5em]
    \item Natural language pattern descriptions ($\leq 32$ words each),
          representing general strategies (e.g., ``validate inputs before
          calling external APIs'').
    \item 1-bit quantized embeddings of these patterns.
    \item Quality scores and domain tags.
\end{itemize}

\paragraph{Never shared:}
\begin{itemize}[leftmargin=1.5em]
    \item Raw tool call inputs or outputs (may contain user data).
    \item Agent trajectories or session logs.
    \item Model weights or internal representations.
    \item User identifiers, API keys, or credentials.
    \item Task-specific content beyond the extracted pattern.
\end{itemize}

The pattern extraction step (GRPO introspection) is designed to produce
generalizable principles, not task-specific solutions. A pattern like
``always handle \texttt{None} inputs before processing'' does not reveal
what specific \texttt{None} value was encountered.

\subsection{Differential Privacy}

For deployments requiring formal privacy guarantees, FAI supports optional
Gaussian noise injection before broadcast:

\begin{definition}[DP-Noisy Compression]
\label{def:dp-noise}
Given privacy budget $\varepsilon_{\text{DP}} > 0$ and sensitivity $\Delta_s$,
add Gaussian noise to each scale factor before broadcast:
\begin{equation}
    \tilde{\alpha}_m^{(j)} = \alpha_m^{(j)} + \mathcal{N}\!\left(0, \, \sigma_{\text{DP}}^2\right),
    \quad \sigma_{\text{DP}} = \frac{\Delta_s \sqrt{2 \ln(1.25/\delta)}}{\varepsilon_{\text{DP}}}.
\end{equation}
Sign vectors are not perturbed (binary data has a different privacy calculus).
Each broadcast round consumes privacy budget $\varepsilon_{\text{DP}}$;
over $R$ rounds, the total budget is $R \cdot \varepsilon_{\text{DP}}$ (basic
composition) or $O(\varepsilon_{\text{DP}} \sqrt{R \ln(1/\delta)})$ (advanced
composition~\cite{dwork2014algorithmic}).
\end{definition}

\subsection{Integrity and Reputation}

\paragraph{Batch Integrity.}
Each broadcast batch is signed by the sender's private key (using Ed25519)
and includes a SHA-256 hash. Receiving nodes verify both before processing.

\paragraph{Reputation Tracking.}
Each node maintains a reputation score for each peer, initialized to $1.0$
and updated as:
\begin{align}
    \text{rep}^{(T+1)}_i &= \gamma \cdot \text{rep}^{(T)}_i + (1 - \gamma) \cdot r_i^{(T)},
    \label{eq:reputation}
\end{align}
where $\gamma = 0.9$ is the decay factor and $r_i^{(T)} \in \{0, 1\}$
indicates whether peer $i$'s batch passed validation (hash, freshness, and
filter). Peers with $\text{rep}_i < \rho_{\min}$ (default $\rho_{\min} = 0.3$)
are deprioritized for future gossip connections.

\paragraph{DSO Stake Requirement.}
Integration with the DSO network requires staking a minimum bond $D_{\min}$
(default: 10 \$AI) per node registration. Byzantine behavior that is
provably detected triggers bond slashing, creating an economic deterrent
against adversarial participation~\cite{hanzo2026dso}.

% ============================================================================
% 9. EVALUATION PLAN
% ============================================================================
