\section{Conclusion}

\label{sec:conclusion}

We presented \textbf{Federated Agent Intelligence (FAI)}, a system enabling
collective learning across autonomous AI agent fleets without sharing model
weights, raw data, or training compute. The core insight is that frozen
language model agents learn implicitly through experience, and that this
knowledge can be extracted as compressed semantic priors, shared via
decentralized gossip, and aggregated robustly against Byzantine adversaries.

FAI composes three mechanisms: GRPO-based local experience extraction,
BitDelta compression achieving ${\approx}10\times$ bandwidth reduction with
$<5\%$ semantic loss, and DeltaSoup Byzantine-robust aggregation tolerating
$f < N/3$ adversarial contributors. We proved formal guarantees including
convergence rate $O(\sqrt{\log N / T})$, $(f,\varepsilon)$-Byzantine robustness,
and monotonic fleet quality improvement under honest majority.

\paragraph{Key Results.}
\begin{itemize}[leftmargin=1.5em]
    \item \textbf{23.7\% task success rate} at $N=100$ nodes, versus 18.2\%
          isolated (a $+5.5\%$ absolute improvement approaching the centralized
          oracle's $+5.8\%$).
    \item \textbf{Byzantine resilience}: 92.8\% of clean performance retained
          at $f = N/3$ with DeltaSoup; naive averaging collapses to baseline.
    \item \textbf{$10.7\times$ compression}: BitDelta reduces per-experience
          bandwidth from 3,072 bytes to 288 bytes, enabling practical P2P operation.
    \item \textbf{0.957 mean cosine similarity} after compression/decompression
          cycle at group size $g = 128$.
    \item \textbf{Constant per-node bandwidth}: ${\approx}29$\,KB/round
          regardless of fleet size (gossip fanout $k=8$).
\end{itemize}

\paragraph{Limitations.}
DeltaSoup requires $f < N/3$; larger Byzantine fractions degrade performance.
The experience quality metric (Definition~\ref{def:quality-metric}) is a proxy
for downstream task performance and may not perfectly correlate in all domains.
Cross-domain transfer is limited by domain tag filtering.

\paragraph{Broader Impact.}
FAI enables organizations to share the benefits of agent experience without
sharing sensitive data. Any agent operating in a domain benefits from the
collective experience of all agents in that domain, accelerating the
time-to-competence for new deployments and reducing redundant exploration
across the fleet. This democratizes access to high-quality AI agent performance
for smaller organizations that cannot afford to run large independent fleets.

% ============================================================================
% APPENDIX
% ============================================================================
\appendix

