\section{Conclusion}


Hanzo demonstrates that AI-first architecture yields significant improvements across commerce metrics. By designing machine learning into the platform foundation rather than bolting it on, we achieve:
\begin{itemize}
\item 28\% average order value increase.
\item 45\% fraud loss reduction.
\item 18\% inventory turnover improvement.
\end{itemize}

The key insight is architectural: unified event streams, shared feature stores, and consistent model serving create compounding benefits impossible with siloed AI additions.

Future work will extend Hanzo with conversational commerce (chatbots with full platform context), visual search (find similar products from images), and automated merchandising (AI-generated product descriptions and imagery).

\begin{thebibliography}{9}

\bibitem{linden2003}
G. Linden, B. Smith, and J. York, ``Amazon.com Recommendations: Item-to-Item Collaborative Filtering,'' \textit{IEEE Internet Computing}, vol. 7, no. 1, pp. 76--80, 2003.

\bibitem{paypal2013}
PayPal Engineering, ``Fraud Detection with Machine Learning,'' \textit{PayPal Engineering Blog}, 2013.

\bibitem{alibaba2017}
Alibaba Group, ``Artificial Intelligence at Alibaba,'' \textit{Alibaba Tech Blog}, 2017.

\bibitem{youtube2016}
P. Covington, J. Adams, and E. Sargin, ``Deep Neural Networks for YouTube Recommendations,'' \textit{RecSys}, 2016.

\bibitem{liu2009}
T.-Y. Liu, ``Learning to Rank for Information Retrieval,'' \textit{Foundations and Trends in Information Retrieval}, vol. 3, no. 3, pp. 225--331, 2009.

\bibitem{hyndman2011}
R. J. Hyndman and G. Athanasopoulos, \textit{Forecasting: Principles and Practice}, OTexts, 2011.

\bibitem{netflix2009}
J. Bennett and S. Lanning, ``The Netflix Prize,'' \textit{KDD Cup and Workshop}, 2007.

\end{thebibliography}

