\section{KMS Secrets Verification}

\label{sec:kms}
% ============================================================================

Hanzo KMS~\cite{hanzokms} manages 12,000+ secrets with envelope encryption.

\subsection{Envelope Encryption Correctness}

\begin{definition}[Envelope Encryption]
A secret $s$ is encrypted as $\texttt{Enc}_{KEK}(\texttt{DEK}) \|
\texttt{Enc}_{DEK}(s)$ where DEK is a per-secret data encryption key and
KEK is the master key encryption key.
\end{definition}

\begin{theorem}[Encryption Correctness]
For any secret $s$ and keys $(KEK, DEK)$:
\begin{equation}
\texttt{Dec}_{DEK}(\texttt{Enc}_{DEK}(s)) = s,
\end{equation}
and
\begin{equation}
\texttt{Dec}_{KEK}(\texttt{Enc}_{KEK}(DEK)) = DEK.
\end{equation}
Composition: the full decrypt-unwrap pipeline recovers $s$ exactly.
\end{theorem}

\subsection{Key Rotation Safety}

\begin{theorem}[Rotation Safety]
During key rotation from $KEK_{\text{old}}$ to $KEK_{\text{new}}$:
\begin{enumerate}[leftmargin=1.4em]
    \item No secret $s$ is ever stored in plaintext (the DEK is always
    encrypted under at least one KEK).
    \item No secret is inaccessible (either $KEK_{\text{old}}$ or
    $KEK_{\text{new}}$ can decrypt the DEK at every point during rotation).
    \item After rotation completes, $KEK_{\text{old}}$ can be safely destroyed
    without losing access to any secret.
\end{enumerate}
\end{theorem}

The proof models rotation as a state machine with five states (idle, re-wrap
started, re-wrap in progress, re-wrap completed, old key destroyed) and proves
that the safety invariant holds at every state transition.

\subsection{Bug Found: Partial Rotation Failure}

Formalization revealed that if the re-wrap process fails mid-way (e.g.,
database connection lost), some secrets remain wrapped under the old KEK while
others are wrapped under the new KEK. If the old KEK is then destroyed
(automated cleanup), the un-re-wrapped secrets become inaccessible. The fix
adds a rotation completion check before old key destruction.

% ============================================================================
