\section{Conclusion}

\label{sec:conclusion}
% ============================================================================

The Hanzo formal verification suite provides mechanically-checked proofs of
173 critical properties across five infrastructure systems. The formalization
effort discovered 8 bugs undetected by testing, including a permission
escalation via revocation race, a path traversal in route matching, and a
rolling update failure mode leaving zero healthy replicas. All 12,000 lines of
Lean 4 proofs are checked by the Lean kernel with no \texttt{sorry} axioms,
providing mathematical certainty that the specified properties hold. The proof
suite runs in CI in 47 seconds, ensuring that specification changes are
continuously verified.

% ============================================================================
% BIBLIOGRAPHY
% ============================================================================
\begin{thebibliography}{15}

\bibitem{hanzoagentsdk}
{Hanzo AI Research}.
\newblock Hanzo Agent {SDK}: Framework for building {AI} agents.
\newblock \emph{Hanzo AI Research}, 2025.

\bibitem{hanzogateway}
{Hanzo AI Research}.
\newblock Hanzo Gateway: {API} gateway with {KrakenD} backend.
\newblock \emph{Hanzo AI Research}, 2024.

\bibitem{hanzokms}
{Hanzo AI Research}.
\newblock Hanzo {KMS}: Secrets management with per-environment rotation and audit trail.
\newblock \emph{Hanzo AI Research}, September 2021.

\bibitem{hanzoaci}
{Hanzo AI Research}.
\newblock Hanzo {ACI}: {AI} Chain Infrastructure for decentralized compute markets.
\newblock \emph{Hanzo AI Research}, December 2024.

\bibitem{hanzoiam}
{Hanzo AI Research}.
\newblock Hanzo {IAM}: Identity and access management for {AI} platforms.
\newblock \emph{Hanzo AI Research}, 2024.

\bibitem{luxwhitepaper}
{Lux Partners}.
\newblock Lux: A scalable multi-consensus blockchain with post-quantum cryptography.
\newblock \emph{Lux Network Technical Report}, 2024.

\bibitem{lean4}
L.~de~Moura and S.~Ullrich.
\newblock The {Lean 4} theorem prover and programming language.
\newblock In \emph{Proc.\ CADE}, 2021.

\bibitem{sel4}
G.~Klein, K.~Elphinstone, G.~Heiser, et~al.
\newblock {seL4}: Formal verification of an {OS} kernel.
\newblock In \emph{Proc.\ SOSP}, 2009.

\bibitem{compcert}
X.~Leroy.
\newblock Formal verification of a realistic compiler.
\newblock \emph{Commun.\ ACM}, 52(7):107--115, 2009.

\bibitem{ironfleet}
C.~Hawblitzel, J.~Howell, M.~Kapritsos, J.~R.~Lorch, B.~Parno, M.~L.~Roberts,
S.~Setty, and B.~Zill.
\newblock {IronFleet}: Proving practical distributed systems correct.
\newblock In \emph{Proc.\ SOSP}, 2015.

\end{thebibliography}

