\section{SKILL.md Format Specification}

\label{app:skill-format}

\begin{verbatim}
---
name: <skill-identifier>
version: 1.0.0
trigger: >
  <Natural language description of when this skill applies.
   Should be specific enough to avoid false positives.>
scope:
  - developer
  - operator       # optional: role-specific scoping
tags:
  - <tag1>
  - <tag2>
difficulty: beginner | intermediate | advanced
estimated_time: <human-readable duration>
---

# <Skill Title>

## Overview

<One-paragraph description of what this skill accomplishes.>

## When to Use

<Bullet list of triggering conditions.>

## Procedure

1. <Step 1>
2. <Step 2>
   ...

## Common Pitfalls

- <Pitfall 1 and how to avoid it>
- <Pitfall 2 and how to avoid it>

## Examples

### Example 1: <Short description>

<Concrete example with code snippets if applicable.>

## See Also

- <Related skill name>
- <Related extension name>
\end{verbatim}

% ============================================================================
% REFERENCES
% ============================================================================
\bibliographystyle{plainnat}
\begin{thebibliography}{40}

\bibitem[Anthropic(2024a)]{anthropic2024claude}
Anthropic.
\newblock Claude 3.5 technical report.
\newblock Technical report, Anthropic, 2024.

\bibitem[Anthropic(2024b)]{anthropic2024mcp}
Anthropic.
\newblock Model Context Protocol: Open standard for context in AI systems.
\newblock \url{https://modelcontextprotocol.io}, 2024.

\bibitem[Bavishi et~al.(2019)]{bavishi2019autopandas}
Bavishi, R., Lemieux, C., Fox, R., Sen, K., and Stoica, I.
\newblock AutoPandas: Neural-backed generators for program synthesis.
\newblock \textit{Proceedings of the ACM on Programming Languages}, 3(OOPSLA):
  1--27, 2019.

\bibitem[Claessen and Hughes(2000)]{claessen2000quickcheck}
Claessen, K. and Hughes, J.
\newblock QuickCheck: A lightweight tool for random testing of Haskell programs.
\newblock In \textit{Proceedings of the Fifth ACM SIGPLAN International
  Conference on Functional Programming (ICFP)}, pp.\ 268--279, 2000.

\bibitem[Git(2015)]{git2015worktree}
Git.
\newblock git-worktree: Manage multiple working trees.
\newblock \url{https://git-scm.com/docs/git-worktree}, 2015.

\bibitem[Gulwani et~al.(2017)]{gulwani2017program}
Gulwani, S., Polozov, O., and Singh, R.
\newblock Program synthesis.
\newblock \textit{Foundations and Trends in Programming Languages}, 4(1--2):
  1--119, 2017.

\bibitem[Hanzo AI(2026a)]{hanzo2026aso}
Hanzo AI Research.
\newblock Active Semantic Optimization: Training-free adaptation via Bayesian
  Product-of-Experts decoding.
\newblock Technical report, Hanzo AI Inc., February 2026.

\bibitem[Hanzo AI(2026b)]{hanzo2026dso}
Hanzo AI Research.
\newblock Decentralized Semantic Optimization: Byzantine-robust prior
  aggregation for collective model adaptation.
\newblock Technical report, Hanzo AI Inc., February 2026.

\bibitem[Holland(1975)]{holland1975adaptation}
Holland, J.~H.
\newblock \textit{Adaptation in Natural and Artificial Systems}.
\newblock University of Michigan Press, 1975.

\bibitem[Hu et~al.(2024)]{hu2024adas}
Hu, S., Lu, C., and Clune, J.
\newblock Automated design of agentic systems.
\newblock \textit{arXiv preprint arXiv:2408.08435}, 2024.

\bibitem[Koza(1994)]{koza1994genetic}
Koza, J.~R.
\newblock \textit{Genetic Programming II: Automatic Discovery of Reusable
  Programs}.
\newblock MIT Press, 1994.

\bibitem[Madaan et~al.(2023)]{madaan2023selfrefine}
Madaan, A., Tandon, N., Gupta, P., Hallinan, S., Gao, L., Wiegreffe, S.,
  Alon, U., Dziri, N., Prabhumoye, S., Yang, Y., et~al.
\newblock Self-Refine: Iterative refinement with self-feedback.
\newblock In \textit{Advances in Neural Information Processing Systems
  (NeurIPS)}, 2023.

\bibitem[ngrok(2024)]{ngrok2024bmo}
ngrok.
\newblock bmo: An agent that builds its own tools.
\newblock \url{https://github.com/ngrok/bmo}, 2024.

\bibitem[OpenAI(2023)]{openai2023gpt4}
OpenAI.
\newblock GPT-4 technical report.
\newblock Technical report, OpenAI, 2023.

\bibitem[Schmidhuber(1987)]{schmidhuber1987evolutionary}
Schmidhuber, J.
\newblock Evolutionary principles in self-referential learning.
\newblock Diploma thesis, Technische Universit{\"a}t M{\"u}nchen, 1987.

\bibitem[Shao et~al.(2024)]{shao2024deepseekmath}
Shao, Z., Wang, P., Zhu, Q., Xu, R., Song, J., Bi, X., Zhang, H., Zhang, M.,
  Li, Y., Wu, Y., and Guo, D.
\newblock DeepSeekMath: Pushing the limits of mathematical reasoning in open
  language models.
\newblock \textit{arXiv preprint arXiv:2402.03300}, 2024.

\bibitem[Sobania et~al.(2023)]{sobania2023analysis}
Sobania, D., Briesch, M., Hanna, C., and Petke, J.
\newblock An analysis of the automatic bug fixing performance of ChatGPT.
\newblock In \textit{IEEE/ACM International Workshop on Automated Program
  Repair (APR)}, pp.\ 23--30, 2023.

\bibitem[Thrun and Pratt(1998)]{thrun1998learning}
Thrun, S. and Pratt, L.
\newblock \textit{Learning to Learn}.
\newblock Springer, 1998.

\bibitem[Wang et~al.(2022)]{wang2022self}
Wang, Y., Kordi, Y., Mishra, S., Liu, A., Smith, N.~A., Khashabi, D., and
  Hajishirzi, H.
\newblock Self-Instruct: Aligning language models with self-generated
  instructions.
\newblock In \textit{Proceedings of the 61st Annual Meeting of the Association
  for Computational Linguistics (ACL)}, pp.\ 13484--13508, 2023.

\bibitem[Wang et~al.(2023)]{wang2023voyager}
Wang, G., Xie, Y., Jiang, Y., Mandlekar, A., Xiao, C., Zhu, Y., Fan, L., and
  Anandkumar, A.
\newblock Voyager: An open-ended embodied agent with large language models.
\newblock In \textit{Advances in Neural Information Processing Systems
  (NeurIPS)}, 2023.

\bibitem[Yang et~al.(2024)]{yang2024sweagent}
Yang, J., Jimenez, C.~E., Wettig, A., Lieret, K., Yao, S., Narasimhan, K.,
  and Press, O.
\newblock SWE-agent: Agent-computer interfaces enable automated software
  engineering.
\newblock In \textit{Advances in Neural Information Processing Systems
  (NeurIPS)}, 2024.

\bibitem[Yuan et~al.(2023)]{yuan2023scaling}
Yuan, Z., Yuan, H., Li, C., Dong, G., Lu, K., Tan, C., Zhou, C., and Zhou, J.
\newblock Scaling relationship on learning mathematical reasoning with large
  language models.
\newblock \textit{arXiv preprint arXiv:2308.01825}, 2023.

\bibitem[Zelikman et~al.(2022)]{zelikman2022star}
Zelikman, E., Wu, Y., Mu, J., and Goodman, N.~D.
\newblock STaR: Bootstrapping reasoning with reasoning.
\newblock In \textit{Advances in Neural Information Processing Systems
  (NeurIPS)}, 2022.

\end{thebibliography}

