\section{Lyapunov Stability Proof}

\label{app:lyapunov}

\begin{theorem}[Exponential Stability]
With fee dissipation rate $\alpha > 0$, the equilibrium $(\bar{\bm{q}}, \bar{\bm{p}})$ is exponentially stable with rate $\alpha$.
\end{theorem}

\begin{proof}
Define the Lyapunov function $V = |\mathcal{H} - \kappa|^2 + \gamma\sum_i(q_i - \bar{q}_i)^2$ for suitable $\gamma > 0$. Fee dissipation ensures $d\mathcal{H}/dt = -\alpha(\mathcal{H} - \kappa) + O(|\bm{q} - \bar{\bm{q}}|^2)$. Therefore:
\begin{align}
\frac{dV}{dt} &= 2(\mathcal{H} - \kappa)\frac{d\mathcal{H}}{dt} + 2\gamma\sum_i(q_i - \bar{q}_i)\dot{q}_i \\
&\le -2\alpha(\mathcal{H} - \kappa)^2 + 2\gamma\sum_i(q_i - \bar{q}_i)w_iq_i \\
&\le -\min(\alpha, \gamma\lambda) \cdot V,
\end{align}
for $\gamma$ chosen such that the cross-terms are dominated. This gives $V(t) \le V(0)e^{-\min(\alpha,\gamma\lambda)t}$.
\end{proof}

\bibliographystyle{plain}
\begin{thebibliography}{35}

\bibitem{adams2020uniswap}
H.~Adams, N.~Zinsmeister, and D.~Robinson.
\newblock Uniswap v2 core.
\newblock \emph{Uniswap Whitepaper}, 2020.

\bibitem{adams2021uniswapv3}
H.~Adams, N.~Zinsmeister, M.~Salem, R.~Keefer, and D.~Robinson.
\newblock Uniswap v3 core.
\newblock \emph{Uniswap Whitepaper}, 2021.

\bibitem{akash2020}
G.~Loli and A.~Vilenski.
\newblock Akash network: Decentralized cloud infrastructure marketplace.
\newblock \emph{Akash Whitepaper}, 2020.

\bibitem{arnold1963proof}
V.~I. Arnold.
\newblock Proof of a theorem of {A.N. Kolmogorov} on the invariance of quasi-periodic motions under small perturbations of the {Hamiltonian}.
\newblock \emph{Russian Mathematical Surveys}, 18(5):9--36, 1963.

\bibitem{arnold1989mathematical}
V.~I. Arnold.
\newblock \emph{Mathematical Methods of Classical Mechanics}.
\newblock Springer, 2nd edition, 1989.

\bibitem{bouchaud2003theory}
J.-P. Bouchaud and M.~Potters.
\newblock \emph{Theory of Financial Risk and Derivative Pricing}.
\newblock Cambridge University Press, 2003.

\bibitem{costan2016intel}
V.~Costan and S.~Devadas.
\newblock Intel {SGX} explained.
\newblock \emph{IACR Cryptology ePrint Archive}, 2016.

\bibitem{crocswap2023}
CrocSwap.
\newblock Ambient liquidity: A new paradigm for decentralized exchange.
\newblock \emph{CrocSwap Whitepaper}, 2023.

\bibitem{daian2020flash}
P.~Daian, S.~Goldfeder, T.~Kell, Y.~Li, X.~Zhao, I.~Bentov, L.~Breidenbach, and A.~Juels.
\newblock Flash boys 2.0: Frontrunning in decentralized exchanges, miner extractable value, and consensus instability.
\newblock In \emph{IEEE Symposium on Security and Privacy}, 2020.

\bibitem{egorov2019stableswap}
M.~Egorov.
\newblock StableSwap -- efficient mechanism for stablecoin liquidity.
\newblock \emph{Curve Finance Whitepaper}, 2019.

\bibitem{friston2017active}
K.~Friston, T.~FitzGerald, F.~Rigoli, P.~Schwartenbeck, and G.~Pezzulo.
\newblock Active inference: A process theory.
\newblock \emph{Neural Computation}, 29(1):1--49, 2017.

\bibitem{gensyn2023}
B.~Fielding and H.~de~Valence.
\newblock Gensyn: A protocol for training machine learning models.
\newblock \emph{Gensyn Whitepaper}, 2023.

\bibitem{golem2016}
J.~Petkanics.
\newblock The golem project: A decentralized computational network.
\newblock \emph{Golem Whitepaper}, 2016.

\bibitem{goldstein2002classical}
H.~Goldstein, C.~Poole, and J.~Safko.
\newblock \emph{Classical Mechanics}.
\newblock Addison Wesley, 3rd edition, 2002.

\bibitem{hairer2006geometric}
E.~Hairer, C.~Lubich, and G.~Wanner.
\newblock \emph{Geometric Numerical Integration}.
\newblock Springer, 2nd edition, 2006.

\bibitem{hertzog2017bancor}
E.~Hertzog, G.~Benartzi, and G.~Benartzi.
\newblock Bancor protocol: Continuous liquidity for cryptographic tokens.
\newblock \emph{Bancor Whitepaper}, 2017.

\bibitem{iexec2018}
G.~Fedak, H.~Wassim, and O.~Croissant.
\newblock {iExec}: Blockchain-based decentralized cloud computing.
\newblock \emph{iExec Whitepaper}, 2018.

\bibitem{ilinski2001physics}
K.~Ilinski.
\newblock \emph{Physics of Finance: Gauge Modelling in Non-Equilibrium Pricing}.
\newblock Wiley, 2001.

\bibitem{ionet2024}
A.~Thomas and T.~Thayyil.
\newblock io.net: The internet of {GPU}s.
\newblock \emph{io.net Whitepaper}, 2024.

\bibitem{mantegna2000introduction}
R.~N. Mantegna and H.~E. Stanley.
\newblock \emph{An Introduction to Econophysics}.
\newblock Cambridge University Press, 2000.

\bibitem{martinelli2019balancer}
F.~Martinelli and N.~Mushegian.
\newblock Balancer: A non-custodial portfolio manager, liquidity provider, and price sensor.
\newblock \emph{Balancer Whitepaper}, 2019.

\bibitem{rendernetwork2021}
J.~Urbach.
\newblock Render network: Distributed {GPU} rendering.
\newblock \emph{Render Whitepaper}, 2021.

\bibitem{ruth1983canonical}
R.~D. Ruth.
\newblock A canonical integration technique.
\newblock \emph{IEEE Transactions on Nuclear Science}, 30(4):2669--2671, 1983.

\bibitem{sethi2019optimal}
S.~P. Sethi.
\newblock \emph{Optimal Control Theory: Applications to Management Science and Economics}.
\newblock Springer, 4th edition, 2019.

\bibitem{teutsch2019truebit}
J.~Teutsch and C.~Reitwie{\ss}ner.
\newblock A scalable verification solution for blockchains.
\newblock \emph{TrueBit Whitepaper}, 2019.

\bibitem{zkevm2023}
Polygon.
\newblock Polygon {zkEVM}: Scaling {Ethereum} with zero-knowledge proofs.
\newblock \emph{Polygon Whitepaper}, 2023.

\bibitem{yoshida1990construction}
H.~Yoshida.
\newblock Construction of higher order symplectic integrators.
\newblock \emph{Physics Letters A}, 150(5--7):262--268, 1990.

\bibitem{mclachlan2002splitting}
R.~I. McLachlan and G.~R.~W. Quispel.
\newblock Splitting methods.
\newblock \emph{Acta Numerica}, 11:341--434, 2002.

\bibitem{leimkuhler2004simulating}
B.~Leimkuhler and S.~Reich.
\newblock \emph{Simulating Hamiltonian Dynamics}.
\newblock Cambridge University Press, 2004.

\bibitem{angeris2020improved}
G.~Angeris, H.-T. Kao, R.~Chiang, C.~Noyes, and T.~Chitra.
\newblock An analysis of {Uniswap} markets.
\newblock \emph{Cryptoeconomic Systems}, 2020.

\bibitem{angeris2021constant}
G.~Angeris and T.~Chitra.
\newblock Improved price oracles: Constant function market makers.
\newblock In \emph{ACM Conference on Advances in Financial Technologies}, 2021.

\end{thebibliography}

\vspace{1em}
\noindent\textit{Disclaimer.} This document describes a proposed protocol. Security properties require formal verification and independent audit before production deployment.

