% Hanzo 3.0: AGI for All
\documentclass[11pt,twocolumn]{article}
\usepackage{shared/hanzocover}
\usepackage{amsmath, amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{float}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,urlcolor=blue}

\title{Hanzo 3.0: AGI for All --- Next-Generation Agent Management with Agent
Chaining and Vector Database Architecture}
\author{
    Zach Kelling \\
    \textit{CTO, Hanzo AI (Techstars '17)} \\
    \texttt{zach@hanzo.ai}
}
\date{2023}

\begin{document}
\hanzocoverpage

\twocolumn[
\begin{@twocolumnfalse}
\maketitle
\begin{abstract}
We present Hanzo~3.0, a platform for managing multiple AI agents at scale. The
system introduces \emph{Agent Chaining}, an architecture for composing
open-source AI protocols into pipelines; a \emph{Vector Database Backend} for
infinite-recall retrieval-augmented generation; \emph{4-bit quantization} for
efficient local and private model deployment; and \emph{multidimensional user
interfaces} spanning voice, text, desktop, mobile, wearable, and AR/VR
modalities. We describe the platform's architecture, its applications across
finance, healthcare, manufacturing, and other industries, and its implications
for the trajectory toward artificial general intelligence.
\end{abstract}
\end{@twocolumnfalse}
]

\section{Introduction}
\label{sec:intro}

As AI capabilities expand, organizations increasingly deploy not one model but
constellations of specialized agents---each handling a distinct task, domain, or
modality. Managing these agent fleets introduces operational challenges:
monitoring performance across agents, distributing tasks equitably, maintaining
consistent configuration, and providing a unified interface to end users who
neither know nor care which agent is responding.

Hanzo~3.0 addresses these challenges with a platform purpose-built for
multi-agent management. The system treats AI agents as first-class operational
entities with lifecycle management, versioning, monitoring, and collaboration
primitives.

\section{Architecture}
\label{sec:architecture}

\subsection{Agent Chaining}

Agent Chaining is the core abstraction. Rather than deploying agents as isolated
services, Hanzo composes them into directed acyclic graphs (DAGs) where the
output of one agent feeds the input of the next. This enables:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Protocol composition.} Open-source AI protocols (language
    models, vision encoders, speech recognizers) snap together via typed
    interfaces.
    \item \textbf{Conditional routing.} Chains branch based on confidence
    scores, input type, or domain classification.
    \item \textbf{Fallback cascades.} If a primary agent fails or returns
    low-confidence output, the chain automatically routes to a secondary.
\end{itemize}

\subsection{Vector Database Backend}

The platform stores all agent interactions, user context, and domain knowledge in
a vector database optimized for high-concurrency, low-latency similarity search.
This provides:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Infinite recall.} Agents retrieve relevant context from the
    full history of interactions, not just a fixed context window.
    \item \textbf{Retrieval-augmented generation (RAG).} Every agent query is
    enriched with semantically similar prior interactions and domain documents.
    \item \textbf{Real-time indexing.} New interactions are indexed immediately,
    ensuring the knowledge base is always current.
\end{itemize}

\subsection{4-Bit Quantization}

Hanzo supports 4-bit model quantization, reducing model size by approximately
$8\times$ compared to FP32 with minimal accuracy degradation. This enables:

\begin{itemize}[leftmargin=1.1em]
    \item Local deployment on consumer hardware (laptops, phones).
    \item Private inference without sending data to cloud endpoints.
    \item Hybrid architectures where lightweight local models handle routine
    queries and cloud models handle complex ones.
\end{itemize}

\subsection{Token Length Research}

The platform incorporates ongoing research into extending effective context
length. Longer token windows enable more detailed, contextually rich agent
responses---critical for domains like legal analysis, medical records review, and
financial modeling where relevant context spans thousands of tokens.

\section{Multi-Agent Dashboard}
\label{sec:dashboard}

The dashboard provides a unified operational view:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Real-time metrics.} Conversation count, response latency,
    accuracy, and throughput per agent.
    \item \textbf{Filtering and grouping.} Agents sorted by model type, domain,
    performance tier, or deployment location.
    \item \textbf{Configuration management.} Parameters and training data managed
    at individual or group level.
    \item \textbf{Versioning.} Save and restore agents at any version, enabling
    safe rollback.
    \item \textbf{Parallel training and deployment.} Train and deploy multiple
    agents concurrently; clone existing agents for rapid scaling.
\end{itemize}

\section{Task Distribution}
\label{sec:tasks}

The Task Distribution System ensures equitable and efficient workload allocation:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Round-robin.} Simple rotation across available agents.
    \item \textbf{Least-busy.} Route to the agent with the lowest current load.
    \item \textbf{Performance-based.} Route to the agent with the highest
    accuracy or lowest latency for the given task type.
\end{itemize}

Automated scaling adjusts the active agent count in response to workload
fluctuations.

\section{Multidimensional Interfaces}
\label{sec:interfaces}

Hanzo~3.0 supports interaction across every modality:

\begin{table}[H]
\centering
\small
\begin{tabular}{ll}
\toprule
\textbf{Modality} & \textbf{Capabilities} \\
\midrule
Voice & Context-aware conversation, commands \\
Text & Chat, structured queries, forms \\
Desktop/Mobile & Native applications, notifications \\
Watch/Wearable & Glanceable summaries, quick actions \\
AR contact lens & Ambient overlays, real-time data \\
HoloLens/VR & Spatial data exploration, collaboration \\
\bottomrule
\end{tabular}
\caption{Supported interface modalities.}
\label{tab:modalities}
\end{table}

A novel contribution is the treatment of \emph{time as an interface dimension}.
Users navigate their data timeline---reviewing past interactions, projecting
future scenarios---transforming time from passive flow to active exploration
tool. In 3D/4D environments, temporal navigation enables trend visualization
and predictive modeling.

\subsection{Ad Hoc AI-Generated Interfaces}

The platform can construct personalized interfaces in real time based on user
context and task requirements. These range from simple 2D displays to spatial
3D environments, generated on demand rather than pre-designed.

\section{Multi-Agent Collaboration}
\label{sec:collaboration}

For tasks requiring multiple perspectives, agents collaborate through:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Sequential handoff.} Agents take turns, each adding domain
    expertise.
    \item \textbf{Parallel consensus.} Multiple agents process the same input
    and the system synthesizes responses.
    \item \textbf{Feedback integration.} User feedback flows back through
    reinforcement learning, continuously improving agent behavior.
\end{itemize}

\section{Industry Applications}
\label{sec:industries}

\subsection{Finance}
A unified platform integrating bank accounts, crypto wallets, and investment
portfolios. Automated trading executes across asset classes based on user-defined
goals. Risk management algorithms balance portfolios for optimal risk-adjusted
returns. Real-time insights track performance, identify risks, and surface
opportunities.

\subsection{Healthcare}
AR-enhanced interfaces for clinicians overlay patient data and predictive models
during procedures. Agent chains coordinate across imaging analysis, records
review, and treatment recommendation.

\subsection{Manufacturing}
Real-time AR visualization of production lines highlights bottlenecks,
maintenance needs, and safety concerns. Predictive analytics forecast equipment
failure before it occurs.

\subsection{Mining and Commodities}
Agent chains span the full value chain: deposit forecasting, equipment
monitoring, refining optimization, logistics routing, market analysis, and
provenance tracking via blockchain.

\section{Human Agency}
\label{sec:agency}

The platform is designed to augment rather than replace human decision-making.
Users set parameters, manage training data, and provide feedback---ensuring AI
adapts to human preferences rather than the reverse. Voice recognition broadens
accessibility, and every interface is designed so that technology serves the user.

\section{Toward Singularity}
\label{sec:singularity}

Hanzo~3.0 represents a step toward artificial general intelligence through
automated decision-making, continuous learning via reinforcement feedback, and
increasingly autonomous agent coordination. The platform's ethical commitments
---user privacy, data security, regulatory compliance---ensure that this
trajectory respects human rights and societal norms.

\section{Roadmap}
\label{sec:roadmap}

\begin{description}[leftmargin=1.1em]
    \item[Phase~I.] Advanced machine learning algorithms, expanded third-party
    integrations.
    \item[Phase~II.] Broader language support, new domain specializations.
    \item[Phase~III.] Integral everyday companion---AI assistance spanning all
    aspects of daily life.
\end{description}

\section{Conclusion}
\label{sec:conclusion}

Hanzo~3.0 provides the operational infrastructure for the multi-agent era. Agent
Chaining composes specialized models into capable pipelines. Vector database
storage delivers infinite recall. Quantization enables private, local deployment.
Multidimensional interfaces meet users in whatever modality suits the task. The
result is a platform that makes advanced AI management accessible across
industries while maintaining the ethical foundations necessary for responsible
progress toward artificial general intelligence.

\begin{thebibliography}{99}

\bibitem{vaswani2017attention}
A.~Vaswani, N.~Shazeer, N.~Parmar, J.~Uszkoreit, L.~Jones, A.~N.~Gomez,
\L.~Kaiser, and I.~Polosukhin.
\newblock Attention is all you need.
\newblock In \emph{NeurIPS}, 2017.

\bibitem{ouyang2022rlhf}
L.~Ouyang, J.~Wu, X.~Jiang, D.~Almeida, C.~L.~Wainwright, P.~Mishkin,
C.~Zhang, S.~Agarwal, K.~Slama, A.~Ray, et~al.
\newblock Training language models to follow instructions with human feedback.
\newblock In \emph{NeurIPS}, 2022.

\bibitem{chase2022langchain}
H.~Chase.
\newblock {LangChain}: Building applications with {LLMs} through composability.
\newblock \url{https://github.com/langchain-ai/langchain}, 2022.

\bibitem{yang2023autogpt}
G.~Yang, R.~Nakano, and S.~Richards.
\newblock {Auto-GPT}: An autonomous {GPT-4} experiment.
\newblock \url{https://github.com/Significant-Gravitas/Auto-GPT}, 2023.

\bibitem{lewis2020rag}
P.~Lewis, E.~Perez, A.~Piktus, F.~Petroni, V.~Karpukhin, N.~Goyal,
H.~K\"uttler, M.~Lewis, W.-t.~Yih, T.~Rockt\"aschel, S.~Riedel, and
D.~Kiela.
\newblock Retrieval-augmented generation for knowledge-intensive {NLP} tasks.
\newblock In \emph{NeurIPS}, 2020.

\bibitem{dettmers2022gptq}
T.~Dettmers, M.~Lewis, Y.~Belkada, and L.~Zettlemoyer.
\newblock {GPT3.int8()}: 8-bit matrix multiplication for transformers at scale.
\newblock In \emph{NeurIPS}, 2022.

\bibitem{touvron2023llama}
H.~Touvron, T.~Lavril, G.~Izacard, X.~Martinet, M.-A.~Lachaux,
T.~Lacroix, B.~Rozi\`ere, N.~Goyal, E.~Hambro, F.~Azhar, et~al.
\newblock {LLaMA}: Open and efficient foundation language models.
\newblock \emph{arXiv preprint arXiv:2302.13971}, 2023.

\bibitem{yao2023react}
S.~Yao, J.~Zhao, D.~Yu, N.~Du, I.~Shafran, K.~Narasimhan, and Y.~Cao.
\newblock {ReAct}: Synergizing reasoning and acting in language models.
\newblock In \emph{ICLR}, 2023.

\end{thebibliography}

\end{document}
