\section{Introduction}

\label{sec:intro}
% ============================================================================

Software is not sovereign. The applications people depend on daily---messaging,
notes, banking, health records---exist as tenants on infrastructure controlled
by third parties. The user's data lives in a provider's database, encrypted
with the provider's keys, governed by the provider's terms. If the provider
disappears, so does the application. If the provider changes policy, the user
complies or loses access. This is not ownership. It is tenancy.

The problem is architectural. Cloud-native design treats the network as a given:
every read is a round-trip, every write is a permission request, every query is
metered. The application is a thin client to someone else's computer. The
alternative---self-hosting---requires expertise that excludes all but a small
technical minority. There is no middle ground.

Base is the middle ground.

\begin{definition}[Sovereign Application]
An application is sovereign if: (i) the user controls the encryption keys for
all stored data, (ii) the application functions without any network connection,
(iii) replication to any remote host is opt-in and reversible, and (iv) no
third party can unilaterally revoke access to the application or its data.
\end{definition}

Base achieves sovereignty by inverting the cloud model. Instead of a thin
client talking to a remote server, Base is the server---running locally, owning
its own database, generating its own keys. Remote services (cloud storage, chain
anchoring, peer sync) are optional extensions, never requirements.

\subsection{Contributions}

\begin{enumerate}[leftmargin=1.4em]
    \item A single-binary application runtime with embedded encrypted database,
    realtime APIs, authentication, and plugin system
    (\S\ref{sec:architecture}).
    \item A data plane architecture where local storage is primary and remote
    replication is provider-optional (\S\ref{sec:dataplane}).
    \item Base Functions: TypeScript execution in an embedded Goja VM with
    zero cold start and event-driven hooks (\S\ref{sec:functions}).
    \item Base Collections: schemaless records with row-level access rules and
    field-level KMS encryption (\S\ref{sec:collections}).
    \item Base Sync: CRDT-based conflict-free replication supporting
    offline-first, peer-to-peer, and server-relay topologies
    (\S\ref{sec:sync}).
    \item Integration with Lux Network for key policy, identity, and threshold
    controls (\S\ref{sec:lux}).
    \item Performance evaluation showing 10,400 encrypted writes/sec and
    sub-millisecond reads on commodity hardware (\S\ref{sec:performance}).
\end{enumerate}

% ============================================================================
