\section{Architecture}

\label{sec:architecture}
% ============================================================================

Base is a single statically-linked Go binary. It embeds all components: the
database engine, the HTTP server, the WebSocket server, the JavaScript VM, the
authentication system, the admin dashboard, and the migration runner. There is
no external dependency at runtime---no PostgreSQL, no Redis, no message broker,
no reverse proxy.

\subsection{System Components}

\begin{table}[H]
\centering
\small
\begin{tabular}{llp{5.5cm}}
\toprule
\textbf{Component} & \textbf{Implementation} & \textbf{Function} \\
\midrule
Database & SQLite + sqlcipher & Encrypted embedded storage \\
HTTP API & Go net/http & REST endpoints, admin UI \\
Realtime & WebSocket & Live subscriptions, change feeds \\
Plugin VM & Goja (ES2023) & TypeScript/JS function runtime \\
Auth & Hanzo IAM (OIDC) & Federated identity, org scoping \\
Admin & Embedded SPA & Collection manager, log viewer \\
Migration & SQL + Go & Schema versioning, auto-migration \\
KMS Client & hanzoai/kms & Field-level encryption keys \\
\bottomrule
\end{tabular}
\caption{Base runtime components. All are compiled into a single binary.}
\label{tab:components}
\end{table}

\subsection{Binary Size and Startup}

The Base binary is 28~MB (stripped, compressed with UPX: 12~MB). Startup time
from cold to serving HTTP requests is 45~ms on an M4 MacBook and 120~ms on a
Raspberry Pi~4. This is not a framework that boots a JVM, connects to three
external services, and runs database migrations. It is a process that opens a
file and starts listening.

\subsection{Deployment Modes}

Base supports three deployment modes with identical API surfaces:

\begin{enumerate}[leftmargin=1.4em]
    \item \textbf{Local}: Single binary on the user's machine. Data in
    \texttt{\textasciitilde/.base/data.db}. No network required.
    \item \textbf{Server}: Single binary on a VPS or bare-metal host. Exposed
    via DNS. Optional TLS termination via Hanzo Ingress.
    \item \textbf{Cluster}: Multiple Base instances behind a load balancer,
    sharing a PostgreSQL database for write coordination. SQLite remains the
    local cache on each node.
\end{enumerate}

The critical property: applications written for local mode work in cluster mode
without modification. The API is the same. The data model is the same. The
auth flow is the same. Only the storage backend changes, and that change is a
single environment variable.

% ============================================================================
