\section{Base as Data Plane}

\label{sec:dataplane}
% ============================================================================

Base treats local storage as the primary data plane. This is the opposite of
every cloud-native system, where the network database is primary and the client
is a cache. In Base, the local SQLite database is the source of truth. Remote
storage is a replica.

\subsection{Local Vault}

Every Base instance maintains a local encrypted database using SQLite compiled
with the sqlcipher extension. The encryption key is derived from the user's
master key via Argon2id:

\begin{equation}
k_{\text{db}} = \text{Argon2id}(k_{\text{master}}, \text{salt}, t=3, m=64\text{MB}, p=4)
\end{equation}

The database file is AES-256-CBC encrypted at the page level. An attacker who
obtains the file without the key sees random bytes. The key never leaves the
device unless the user explicitly exports it (for backup or device migration).

\subsection{Provider-Optional Replication}

Replication in Base is a directed graph of sync targets. Each target is a
remote Base instance, an S3-compatible bucket, or a Lux chain endpoint. The
user configures zero or more targets:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Zero targets}: Fully offline. Data exists only on the local
    device. Backup is the user's responsibility.
    \item \textbf{One target}: Simple backup. Changes replicate to a single
    remote. Recovery is possible from the remote.
    \item \textbf{Multiple targets}: Redundant replication. Changes fan out to
    multiple remotes. Availability survives any single target failure.
\end{itemize}

No target is privileged. There is no ``primary'' remote. The local database
is always authoritative. If a remote diverges, the CRDT merge
(\S\ref{sec:sync}) resolves the conflict deterministically.

\subsection{Chain Anchoring}

Base optionally anchors data integrity proofs to the Lux Network. Anchoring
does not store data on-chain---it stores a Merkle root of the local database
state, creating an immutable timestamp proof.

\begin{equation}
\text{anchor}(t) = \text{SHA3-256}\bigl(\text{root}(\text{MerkleTree}(D_t))\bigr)
\end{equation}

where $D_t$ is the set of all records at time $t$. The anchor transaction is
submitted to the Lux T-Chain (transaction chain), providing sub-second
finality and costing less than 0.001 LUX per anchor. Applications that require
tamper-evidence (medical records, legal documents, audit logs) enable anchoring.
Applications that do not need it pay nothing.

% ============================================================================
