\section{Base Collections}

\label{sec:collections}
% ============================================================================

Collections are Base's data model. A collection is a named set of records with
a schema that is enforced but not rigid. Fields can be added at any time.
Fields can have types, defaults, validation rules, and encryption directives.

\subsection{Schema Model}

Collections use a schema-on-write model. Every record in a collection shares
the same field set, but the field set can evolve:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Add field}: New fields are added with a default value. Existing
    records gain the field lazily on next read or write.
    \item \textbf{Remove field}: Fields are marked deleted. Data is retained
    but hidden from the API. Physical deletion is a separate admin action.
    \item \textbf{Rename field}: Renames are tracked in the migration log,
    preserving query compatibility via aliases.
    \item \textbf{Type change}: Prohibited. A new field must be created and the
    old field deprecated.
\end{itemize}

\subsection{Access Rules}

Every collection has three access rule sets: \texttt{listRule},
\texttt{viewRule}, and \texttt{createRule}/\texttt{updateRule}/\texttt{deleteRule}.
Rules are expressions evaluated against the request context:

\begin{lstlisting}[language=JavaScript,caption={Collection access rules.}]
// Only the record owner can view
viewRule: "@request.auth.id = id"

// Org members can list
listRule: "@request.auth.org = org_id"

// Admins can delete
deleteRule: "@request.auth.role = 'admin'"
\end{lstlisting}

Rules compose with boolean operators (\texttt{\&\&}, \texttt{||}). A missing
rule defaults to deny. There is no ambient read or write access.

\subsection{Field-Level KMS Encryption}

Individual fields can be marked as encrypted. Encrypted fields are stored as
ciphertext in the SQLite database and decrypted only when returned to an
authorized requester. The encryption key for each field is managed by
Hanzo KMS (kms.hanzo.ai):

\begin{equation}
c_f = \text{AES-256-GCM}(k_f, \text{nonce}, \text{plaintext}_f, \text{aad})
\end{equation}

where $k_f$ is the field-specific data encryption key (DEK), itself encrypted
by a key encryption key (KEK) stored in KMS. The AAD (additional authenticated
data) includes the record ID and collection name, binding the ciphertext to
its context.

Use cases: social security numbers, medical records, API keys, private notes.
The database administrator (or an attacker who compromises the database file)
sees only ciphertext for encrypted fields.

% ============================================================================
