\section{Base Sync}

\label{sec:sync}
% ============================================================================

Base Sync is the replication protocol. It enables multiple Base instances to
share data without a central coordinator. Sync is built on CRDTs
(Conflict-free Replicated Data Types), guaranteeing that concurrent
modifications converge to the same state without manual conflict resolution.

\subsection{CRDT Log}

Every mutation in Base produces a log entry:

\begin{equation}
e = (\text{id}, \text{collection}, \text{op}, \text{fields}, \text{hlc}, \text{node\_id})
\end{equation}

where $\text{hlc}$ is a hybrid logical clock~\cite{kulkarni2014hlc} and
$\text{node\_id}$ is the originating Base instance. The log is append-only.
Deletions are tombstones, not physical removals.

\subsection{Conflict-Free Merge}

Sync uses a last-writer-wins register (LWW) at the field level. Given two
concurrent writes to the same field:

\begin{equation}
\text{resolve}(e_1, e_2) =
\begin{cases}
e_1 & \text{if } \text{hlc}(e_1) > \text{hlc}(e_2) \\
e_2 & \text{if } \text{hlc}(e_2) > \text{hlc}(e_1) \\
e_{\max(\text{node\_id})} & \text{if } \text{hlc}(e_1) = \text{hlc}(e_2)
\end{cases}
\end{equation}

Field-level LWW means that two users editing different fields of the same
record will both see their changes preserved. Only truly conflicting writes
(same field, same record, same logical time) fall back to node ID ordering.

\subsection{Sync Topologies}

\begin{table}[H]
\centering
\small
\begin{tabular}{lp{6cm}}
\toprule
\textbf{Topology} & \textbf{Description} \\
\midrule
Offline-first & Single device. Sync log accumulates locally. Replayed on
next connection. \\
Peer-to-peer & Two or more Base instances sync directly over WebSocket or
WebRTC. No server. \\
Server-relay & Base instances sync through a central relay. The relay is a
Base instance with no special privileges. \\
Hub-and-spoke & One Base instance acts as the persistent hub. Edge instances
sync through it. The hub can be a cloud VM or a NAS on the local network. \\
\bottomrule
\end{tabular}
\caption{Supported sync topologies. All use the same CRDT merge protocol.}
\label{tab:topologies}
\end{table}

\subsection{Selective Sync}

Not all data needs to sync everywhere. Base supports collection-level and
rule-based sync filters:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Collection filter}: Sync only specified collections to a
    target. A mobile device might sync \texttt{notes} but not
    \texttt{analytics}.
    \item \textbf{Record filter}: Sync only records matching a predicate.
    A shared workspace might sync only records where
    \texttt{org\_id = \$current\_org}.
    \item \textbf{Field filter}: Exclude specific fields from sync. Encrypted
    fields can be excluded from remote targets to keep sensitive data local.
\end{itemize}

% ============================================================================
