\section{Cost Tracking and Attribution}

\label{sec:cost}

\subsection{Pricing Model Normalization}

Provider pricing models exhibit significant heterogeneity:

\begin{table}[t]
\centering
\caption{Provider pricing model categories}
\label{tab:pricing}
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Model} & \textbf{Providers} & \textbf{Unit} \\
\midrule
Per-token & OpenAI, Anthropic & \$/1M tokens \\
Per-character & Google & \$/1M chars \\
Per-second & Replicate & \$/GPU-sec \\
Reserved & Azure, Bedrock & \$/hour \\
Per-request & Some embeddings & \$/request \\
Tiered & Together AI & Rate-dependent \\
Free tier & Groq, Cerebras & Quota-limited \\
\bottomrule
\end{tabular}
\end{table}

The cost engine normalizes all pricing to a per-token basis using provider-specific conversion factors. For character-based pricing, we use the empirical ratio of 1 token $\approx$ 4 characters for English text, calibrated per-model using tokenizer analysis. For time-based pricing, we estimate cost as $c_{\text{time}} = \text{price\_per\_second} \times \text{generation\_time}$.

\subsection{Real-Time Cost Computation}

For each completed request, the cost engine computes:

\begin{equation}
\text{cost}(q) = c_{\text{in}}(p) \cdot n_{\text{in}} + c_{\text{out}}(p) \cdot n_{\text{out}} + c_{\text{fixed}}(p)
\end{equation}

where $n_{\text{in}}$ and $n_{\text{out}}$ are input and output token counts from the provider response's \texttt{usage} field, and $c_{\text{fixed}}$ captures any per-request fixed costs.

For streaming requests where usage is reported only at stream completion, the gateway estimates running cost from token-counting the streamed content using a provider-matched tokenizer.

\subsection{Budget Enforcement}

Organizations can set budget limits at multiple granularities:

\begin{itemize}[leftmargin=*]
    \item \textbf{Per-key daily/monthly limits}: Hard caps on API key spend.
    \item \textbf{Per-model limits}: Restrict spend on expensive models.
    \item \textbf{Alerting thresholds}: Notifications at 50\%, 80\%, 90\%, 100\% of budget.
    \item \textbf{Automatic downgrade}: When budget thresholds are crossed, automatically route to cheaper model alternatives.
\end{itemize}

\subsection{Cost Attribution}

The gateway attributes costs to organizational units through a hierarchical tagging system. Each request can carry tags for organization, project, team, user, and feature. Costs are aggregated along each dimension and exposed through a reporting API:

\begin{lstlisting}[language=Python, caption=Cost attribution query]
GET /v1/costs?start=2026-02-01
    &end=2026-02-28
    &group_by=project,model
    &filter=team:ml-platform

{
  "total_cost": 14823.47,
  "by_project": {
    "chat-prod": {
      "gpt-4o": 8234.12,
      "claude-sonnet": 3421.89
    },
    "embeddings": {
      "text-embedding-3-large": 1167.46
    }
  }
}
\end{lstlisting}

