\section{Conclusion}

\label{sec:conclusion}

The Hanzo LLM Gateway demonstrates that a well-designed proxy layer can substantially reduce the complexity, cost, and risk of multi-provider LLM deployments. By normalizing 100+ provider APIs behind a single interface, implementing intelligent routing with formal optimization objectives, and providing semantic caching with high precision, the gateway achieves 23\% latency reduction, 31\% cost savings, and 99.97\% availability in production. The open-source release enables organizations to deploy multi-provider LLM infrastructure without vendor lock-in.

\bibliographystyle{plain}
\begin{thebibliography}{30}

\bibitem{brown2020language}
T.~Brown, B.~Mann, N.~Ryder, et~al.
\newblock Language models are few-shot learners.
\newblock In \emph{NeurIPS}, 2020.

\bibitem{bommasani2022opportunities}
R.~Bommasani, D.~Hudson, E.~Adeli, et~al.
\newblock On the opportunities and risks of foundation models.
\newblock \emph{arXiv preprint arXiv:2108.07258}, 2022.

\bibitem{aiinfra2025survey}
AI Infrastructure Alliance.
\newblock State of AI infrastructure 2025: Multi-provider deployment patterns.
\newblock Technical report, 2025.

\bibitem{kwon2023vllm}
W.~Kwon, Z.~Li, S.~Zhuang, et~al.
\newblock Efficient memory management for large language model serving with {PagedAttention}.
\newblock In \emph{SOSP}, 2023.

\bibitem{tgi2023}
HuggingFace.
\newblock Text generation inference: A {Rust}, {Python} and {gRPC} server for text generation.
\newblock \url{https://github.com/huggingface/text-generation-inference}, 2023.

\bibitem{litellm2023}
BerriAI.
\newblock {LiteLLM}: Call all {LLM} {APIs} using the {OpenAI} format.
\newblock \url{https://github.com/BerriAI/litellm}, 2023.

\bibitem{portkey2024}
Portkey.
\newblock Portkey: Control panel for {AI} apps.
\newblock \url{https://portkey.ai}, 2024.

\bibitem{martian2024}
Martian.
\newblock Model router: Automatic model selection for optimal cost and quality.
\newblock Technical report, Martian AI, 2024.

\bibitem{indyk1998approximate}
P.~Indyk and R.~Motwani.
\newblock Approximate nearest neighbors: Towards removing the curse of dimensionality.
\newblock In \emph{STOC}, 1998.

\bibitem{charikar2002similarity}
M.~Charikar.
\newblock Similarity estimation techniques from rounding algorithms.
\newblock In \emph{STOC}, 2002.

\bibitem{perez2022ignore}
F.~Perez and I.~Ribas.
\newblock Ignore this title and {HackAPrompt}: Exposing systemic weaknesses of {LLMs} through a global-scale prompt hacking competition.
\newblock \emph{arXiv preprint arXiv:2311.16119}, 2022.

\bibitem{openai2024api}
OpenAI.
\newblock {OpenAI} {API} reference.
\newblock \url{https://platform.openai.com/docs/api-reference}, 2024.

\bibitem{anthropic2024api}
Anthropic.
\newblock {Anthropic} {API} reference.
\newblock \url{https://docs.anthropic.com/en/api}, 2024.

\bibitem{google2024gemini}
Google.
\newblock {Gemini} {API} documentation.
\newblock \url{https://ai.google.dev/docs}, 2024.

\bibitem{nygard2018release}
M.~Nygard.
\newblock \emph{Release It!: Design and Deploy Production-Ready Software}.
\newblock Pragmatic Bookshelf, 2nd edition, 2018.

\bibitem{burns2016design}
B.~Burns.
\newblock \emph{Designing Distributed Systems}.
\newblock O'Reilly Media, 2018.

\bibitem{kleppmann2017designing}
M.~Kleppmann.
\newblock \emph{Designing Data-Intensive Applications}.
\newblock O'Reilly Media, 2017.

\bibitem{vargaftik2022rade}
S.~Vargaftik, R.~Ben-Basat, A.~Szpigel, and G.~Mendelson.
\newblock {RADE}: Resource-efficient supervised anomaly detection using decision tree encoding.
\newblock In \emph{MLSys}, 2022.

\bibitem{chen2023frugalgpt}
L.~Chen, M.~Zaharia, and J.~Zou.
\newblock {FrugalGPT}: How to use large language models while reducing cost and improving performance.
\newblock \emph{arXiv preprint arXiv:2305.05176}, 2023.

\bibitem{jiang2024mixtral}
A.~Jiang, A.~Sablayrolles, A.~Roux, et~al.
\newblock Mixtral of experts.
\newblock \emph{arXiv preprint arXiv:2401.04088}, 2024.

\bibitem{touvron2023llama}
H.~Touvron, L.~Martin, K.~Stone, et~al.
\newblock Llama 2: Open foundation and fine-tuned chat models.
\newblock \emph{arXiv preprint arXiv:2307.09288}, 2023.

\bibitem{team2024gemma}
Gemma Team.
\newblock Gemma 2: Improving open language models at a practical size.
\newblock \emph{arXiv preprint arXiv:2408.00118}, 2024.

\bibitem{zheng2023judging}
L.~Zheng, W.-L. Chiang, Y.~Sheng, et~al.
\newblock Judging {LLM}-as-a-judge with {MT-Bench} and {Chatbot Arena}.
\newblock In \emph{NeurIPS}, 2023.

\bibitem{hendrycks2021measuring}
D.~Hendrycks, C.~Burns, S.~Basart, et~al.
\newblock Measuring massive multitask language understanding.
\newblock In \emph{ICLR}, 2021.

\bibitem{chen2021evaluating}
M.~Chen, J.~Tworek, H.~Jun, et~al.
\newblock Evaluating large language models trained on code.
\newblock \emph{arXiv preprint arXiv:2107.03374}, 2021.

\end{thebibliography}

