\section{System Architecture}

\label{sec:architecture}

\subsection{Overview}

Hanzo Chat comprises five layers:

\begin{enumerate}[leftmargin=1.4em]
    \item \textbf{Presentation Layer}: React-based UI with markdown rendering, code highlighting, file previews, and streaming response display.
    \item \textbf{Routing Layer}: Model selection, load balancing, and failover across providers.
    \item \textbf{Memory Layer}: Hierarchical storage and retrieval of conversational context.
    \item \textbf{Tool Layer}: MCP server management, tool discovery, and execution sandboxing.
    \item \textbf{Gateway Layer}: Unified API proxy (Hanzo LLM Gateway) translating between provider-specific formats.
\end{enumerate}

\subsection{Request Flow}

A user message $m_t$ at turn $t$ triggers the following pipeline:

\begin{algorithm}[H]
\caption{Hanzo Chat Request Pipeline}
\label{alg:pipeline}
\begin{algorithmic}[1]
\Require User message $m_t$, conversation history $H_{<t}$, user profile $U$
\State $\bm{e}_t \gets \text{Embed}(m_t)$ \Comment{Semantic embedding}
\State $\mathcal{R} \gets \text{RetrieveMemory}(\bm{e}_t, U)$ \Comment{Memory retrieval}
\State $C_t \gets \text{BuildContext}(H_{<t}, \mathcal{R}, m_t)$ \Comment{Context assembly}
\State $\hat{M} \gets \text{RouteModel}(C_t, U.\text{policy})$ \Comment{Model selection}
\State $\mathcal{T} \gets \text{DiscoverTools}(C_t, \hat{M})$ \Comment{Tool discovery}
\State $r_t \gets \text{Generate}(\hat{M}, C_t, \mathcal{T})$ \Comment{LLM generation}
\While{$r_t$ contains tool calls}
    \State $\text{results} \gets \text{ExecuteTools}(r_t.\text{tool\_calls})$
    \State $r_t \gets \text{Generate}(\hat{M}, C_t \cup \text{results}, \mathcal{T})$
\EndWhile
\State $\text{UpdateMemory}(m_t, r_t, \bm{e}_t)$
\State \Return $r_t$
\end{algorithmic}
\end{algorithm}

\subsection{Gateway Integration}

Hanzo Chat connects to LLM providers through the Hanzo LLM Gateway~\cite{hanzollm2025}, which provides:

\begin{itemize}[leftmargin=1.1em]
    \item Unified OpenAI-compatible API for 100+ models across providers (OpenAI, Anthropic, Google, Together, Ollama, etc.).
    \item Automatic retry with exponential backoff and provider failover.
    \item Request/response logging for analytics and debugging.
    \item Token counting and cost tracking per request.
    \item Rate limiting and quota management per user/organization.
\end{itemize}

