\section{Related Work}

\label{sec:related}

\subsection{Multi-Model Systems}

Mixture-of-experts architectures~\cite{shazeer2017outrageously,fedus2022switch} route within a single model. FrugalGPT~\cite{chen2023frugalgpt} chains LLM calls for cost optimization. RouterBench~\cite{hu2024routerbench} evaluates model routing strategies. Martian~\cite{martian2024} provides model routing as a service. Our approach differs by combining routing with memory and tool integration in a unified system.

\subsection{Conversational Memory}

MemoryBank~\cite{zhong2024memorybank} extends LLMs with long-term memory inspired by Ebbinghaus forgetting curves. Reflexion~\cite{shinn2023reflexion} uses self-reflection for task improvement. MemGPT~\cite{packer2023memgpt} implements virtual context management with paging. Our hierarchical memory architecture extends these ideas with three-tier scoping and automatic consolidation.

\subsection{Tool Use in LLMs}

Toolformer~\cite{schick2023toolformer} trains models to use tools via self-supervised learning. Gorilla~\cite{patil2023gorilla} fine-tunes for API call generation. TaskWeaver~\cite{qin2024taskweaver} provides a code-first agent framework. The Model Context Protocol~\cite{anthropic2024mcp} standardizes tool interfaces. Hanzo Chat builds on MCP but adds dynamic tool discovery, sandboxed execution, and multi-step chaining.

\subsection{Chat Interfaces}

ChatGPT~\cite{openai2022chatgpt} popularized conversational AI interfaces. Claude.ai~\cite{anthropic2024claude} introduced artifacts for rich content. Gemini~\cite{google2024gemini} integrates multimodal understanding. Poe~\cite{quora2023poe} provides multi-model access. Hanzo Chat differentiates through unified memory, tool integration, and intelligent routing across all models.

