\section{Conclusion}

\label{sec:conclusion}

We have presented Hanzo Chat, a multi-model conversational AI system that combines intelligent model routing, hierarchical persistent memory, and extensible tool integration through the Model Context Protocol. Our evaluation demonstrates that the combination of these three capabilities achieves performance exceeding any single model while significantly reducing costs. The hierarchical memory architecture enables conversations spanning hundreds of turns without degradation, addressing one of the most significant limitations of current LLM-based chat systems. Production deployment over 18 months with 500K+ conversations validates the practical viability of the approach. Hanzo Chat is available at \texttt{chat.hanzo.ai} and its architecture is documented for the research community.

\bibliographystyle{plain}
\begin{thebibliography}{32}

\bibitem{anthropic2024claude}
Anthropic.
\newblock Claude 3.5 technical report.
\newblock \emph{Anthropic Technical Reports}, 2024.

\bibitem{anthropic2024mcp}
Anthropic.
\newblock Model context protocol specification.
\newblock \emph{MCP Documentation}, 2024.

\bibitem{chen2023frugalgpt}
L.~Chen, M.~Zaharia, and J.~Zou.
\newblock {FrugalGPT}: How to use large language models while reducing cost and improving performance.
\newblock \emph{arXiv preprint arXiv:2305.05176}, 2023.

\bibitem{chiang2024chatbot}
W.-L. Chiang, L.~Zheng, Y.~Sheng, et~al.
\newblock Chatbot arena: An open platform for evaluating {LLM}s by human preference.
\newblock \emph{arXiv preprint arXiv:2403.04132}, 2024.

\bibitem{fedus2022switch}
W.~Fedus, B.~Zoph, and N.~Shazeer.
\newblock Switch transformers: Scaling to trillion parameter models with simple and efficient sparsity.
\newblock \emph{JMLR}, 23(120):1--39, 2022.

\bibitem{gal2016dropout}
Y.~Gal and Z.~Ghahramani.
\newblock Dropout as a {Bayesian} approximation: Representing model uncertainty in deep learning.
\newblock In \emph{ICML}, 2016.

\bibitem{google2024gemini}
Google.
\newblock Gemini: A family of highly capable multimodal models.
\newblock \emph{Google Technical Reports}, 2024.

\bibitem{hanzollm2025}
Hanzo~AI.
\newblock Hanzo {LLM} gateway: Unified proxy for 100+ {LLM} providers.
\newblock \emph{Hanzo Technical Documentation}, 2025.

\bibitem{hu2024routerbench}
X.~Hu, Z.~Li, and J.~Chen.
\newblock {RouterBench}: A benchmark for multi-{LLM} routing system.
\newblock \emph{arXiv preprint arXiv:2403.12031}, 2024.

\bibitem{li2010contextual}
L.~Li, W.~Chu, J.~Langford, and R.~E. Schapire.
\newblock A contextual-bandit approach to personalized news article recommendation.
\newblock In \emph{WWW}, 2010.

\bibitem{martian2024}
Martian.
\newblock Intelligent model routing for {LLM} applications.
\newblock \emph{Martian Documentation}, 2024.

\bibitem{mcclelland1995memory}
J.~L. McClelland, B.~L. McNaughton, and R.~C. O'Reilly.
\newblock Why there are complementary learning systems in the hippocampus and neocortex.
\newblock \emph{Psychological Review}, 102(3):419, 1995.

\bibitem{nogueira2019passage}
R.~Nogueira and K.~Cho.
\newblock Passage re-ranking with {BERT}.
\newblock \emph{arXiv preprint arXiv:1901.04085}, 2019.

\bibitem{openai2022chatgpt}
OpenAI.
\newblock Introducing {ChatGPT}.
\newblock \emph{OpenAI Blog}, 2022.

\bibitem{openai2023gpt4}
OpenAI.
\newblock {GPT-4} technical report.
\newblock \emph{arXiv preprint arXiv:2303.08774}, 2023.

\bibitem{packer2023memgpt}
C.~Packer, S.~Wooders, K.~Lin, V.~Fang, S.~G. Patil, I.~Stoica, and J.~E. Gonzalez.
\newblock {MemGPT}: Towards {LLM}s as operating systems.
\newblock \emph{arXiv preprint arXiv:2310.08560}, 2023.

\bibitem{patil2023gorilla}
S.~G. Patil, T.~Zhang, X.~Wang, and J.~E. Gonzalez.
\newblock Gorilla: Large language model connected with massive {API}s.
\newblock \emph{arXiv preprint arXiv:2305.15334}, 2023.

\bibitem{pgvector2023}
A.~Katz.
\newblock pgvector: Open-source vector similarity search for {PostgreSQL}.
\newblock \emph{GitHub Repository}, 2023.

\bibitem{qin2024taskweaver}
B.~Qin, D.~Liang, X.~Ye, et~al.
\newblock {TaskWeaver}: A code-first agent framework.
\newblock \emph{arXiv preprint arXiv:2311.17541}, 2024.

\bibitem{quora2023poe}
Quora.
\newblock Poe: Fast, helpful {AI} chat.
\newblock \emph{Quora Blog}, 2023.

\bibitem{schick2023toolformer}
T.~Schick, J.~Dwivedi-Yu, R.~Dess\'i, et~al.
\newblock Toolformer: Language models can teach themselves to use tools.
\newblock In \emph{NeurIPS}, 2023.

\bibitem{shazeer2017outrageously}
N.~Shazeer, A.~Mirhoseini, K.~Maziarz, et~al.
\newblock Outrageously large neural networks: The sparsely-gated mixture-of-experts layer.
\newblock In \emph{ICLR}, 2017.

\bibitem{shinn2023reflexion}
N.~Shinn, F.~Cassano, A.~Gopinath, K.~R. Narasimhan, and S.~Yao.
\newblock Reflexion: Language agents with verbal reinforcement learning.
\newblock In \emph{NeurIPS}, 2023.

\bibitem{wang2022text}
L.~Wang, N.~Yang, X.~Huang, et~al.
\newblock Text embeddings by weakly-supervised contrastive pre-training.
\newblock \emph{arXiv preprint arXiv:2212.03533}, 2022.

\bibitem{zheng2023judging}
L.~Zheng, W.-L. Chiang, Y.~Sheng, et~al.
\newblock Judging {LLM}-as-a-judge with {MT-Bench} and chatbot arena.
\newblock In \emph{NeurIPS}, 2023.

\bibitem{zhong2024memorybank}
W.~Zhong, L.~Guo, Q.~Gao, et~al.
\newblock {MemoryBank}: Enhancing large language models with long-term memory.
\newblock In \emph{AAAI}, 2024.

\end{thebibliography}

