\section{Semantic Memory Manager}

\label{sec:memory}

\subsection{Experience Representation}

\begin{definition}[Experience Entry]
\label{def:experience}
An experience entry $e \in \mathcal{E}$ is a tuple:
\begin{equation}
    e = \bigl(\texttt{id},\; \texttt{text},\; c,\; d,\; \bm{v},\; u,\; k\bigr),
    \label{eq:experience}
\end{equation}
where:
\begin{itemize}[leftmargin=1.5em]
    \item $\texttt{id} \in \mathbb{N}$ is a unique identifier,
    \item $\texttt{text} \in \Sigma^{\leq 32}$ is the natural-language experience
          (at most 32 words),
    \item $c \in (0, 1)$ is the confidence score (initialized to $c_0 = 0.5$
          for new entries, updated via online learning),
    \item $d \in \mathcal{D}$ is the domain label,
    \item $\bm{v} \in \mathbb{R}^{384}$ is the sentence embedding produced by
          \texttt{all-MiniLM-L6-v2}~\cite{reimers2019sentence},
    \item $u \in \mathbb{N}$ is the usage count (number of times retrieved and
          applied),
    \item $k \in \mathbb{N}$ is the epoch of creation (session index).
\end{itemize}
\end{definition}

The confidence score is updated via a Bayesian-flavored online rule after each
session in which experience $e$ was applied:
\begin{equation}
    c \leftarrow
    \begin{cases}
        \min(c + \gamma(1 - c),\; c_{\max}) & \text{if applying } e \text{ improved task reward,} \\
        \max(c - \gamma c,\; c_{\min}) & \text{if applying } e \text{ did not improve reward,}
    \end{cases}
    \label{eq:confidence-update}
\end{equation}
where $\gamma = 0.1$ is the learning rate, $c_{\min} = 0.05$, and
$c_{\max} = 0.95$.

\subsection{Retrieval}
\label{sec:retrieval}

At the start of each agent session with task $x$, the memory manager retrieves
the top-$K$ experiences most relevant to $x$.

\paragraph{Query Embedding.}
The task query is embedded: $\bm{q} = \text{Enc}(x) \in \mathbb{R}^{384}$.

\paragraph{Domain-Filtered Cosine Similarity.}
For each experience $e \in \mathcal{E}$, the retrieval score is:
\begin{equation}
    \text{score}(e, x) =
    \begin{cases}
        \dfrac{\bm{q} \cdot \bm{v}_e}{\|\bm{q}\|\,\|\bm{v}_e\|} & \text{if } d_e \in D_x \text{ or } D_x = \emptyset, \\
        -\infty & \text{otherwise,}
    \end{cases}
    \label{eq:retrieval-score}
\end{equation}
where $D_x \subseteq \mathcal{D}$ is the set of domains inferred from the task
(empty if domain is ambiguous, in which case all domains are eligible). The
top-$K$ experiences by score are returned.

\paragraph{Context Injection.}
Retrieved experiences are prepended to the agent's system prompt in ranked order:
\begin{quote}
\texttt{[G0] \{e[0].text\}}\\
\texttt{[G1] \{e[1].text\}}\\
$\vdots$\\
\texttt{[G\{K-1\}] \{e[K-1].text\}}
\end{quote}
This format signals to the model that the bracketed items are learned
guidelines (``G'' for guideline), distinguishing them from task instructions
and conversation history. Default $K = 5$ retrieved experiences per session.

\subsection{Confidence-Weighted Retrieval Score}

The pure cosine similarity score can be augmented by confidence weighting to
prioritize well-validated experiences:
\begin{equation}
    \tilde{\text{score}}(e, x) = (1 - \lambda)\,\frac{\bm{q} \cdot \bm{v}_e}{\|\bm{q}\|\,\|\bm{v}_e\|}
    + \lambda\, c_e,
    \label{eq:conf-weighted-score}
\end{equation}
with $\lambda = 0.2$. This blends semantic relevance with empirical reliability.

\subsection{Memory Compression}
\label{sec:compression}

As the experience library grows, unconstrained addition would eventually
degrade retrieval precision (too many near-duplicate or stale entries) and
increase context injection cost. The memory manager applies one of four
compression strategies when $|\mathcal{E}| > N_{\max}$ (default $N_{\max} = 500$).

\paragraph{Strategy 1: Diversity Compression.}
Apply $K$-means clustering to the embedding matrix
$V = [\bm{v}_1, \ldots, \bm{v}_{|\mathcal{E}|}]^\top$. Within each cluster
$\mathcal{C}_k$, retain only the experience with highest confidence:
\begin{equation}
    e^*_k = \argmax_{e \in \mathcal{C}_k} c_e.
    \label{eq:diversity-compression}
\end{equation}
This ensures the surviving library spans the full semantic space of learned
patterns.

\paragraph{Strategy 2: Importance Compression.}
Sort experiences by confidence and discard the bottom-$p$ fraction:
\begin{equation}
    \mathcal{E}' = \bigl\{e \in \mathcal{E} : c_e \geq Q_{1-p}(\{c_e : e \in \mathcal{E}\})\bigr\},
    \label{eq:importance-compression}
\end{equation}
where $Q_{1-p}$ is the $(1-p)$-quantile. Default $p = 0.2$ (discard lowest
20\% by confidence).

\paragraph{Strategy 3: Temporal Decay Compression.}
Assign each experience a time weight via exponential decay:
\begin{equation}
    w_t(e) = \exp\bigl(-\lambda(k_{\text{now}} - k_e)\bigr),
    \label{eq:temporal-decay}
\end{equation}
where $k_e$ is the epoch of creation, $k_{\text{now}}$ is the current epoch,
and $\lambda > 0$ is the decay constant (default $\lambda = 0.05$). Experiences
with $w_t(e) < w_{\min}$ are pruned. Temporal decay is appropriate in
fast-changing environments where older experiences may reflect outdated tool
behavior.

\paragraph{Strategy 4: Hybrid Compression.}
The hybrid strategy combines confidence, temporal weight, and a diversity
penalty into a single utility score:
\begin{equation}
    U(e) = \alpha \cdot c_e^{\,\beta} \cdot w_t(e)^{\,\gamma} \cdot d_{\text{div}}(e)^{\,\delta},
    \label{eq:hybrid-utility}
\end{equation}
where $d_{\text{div}}(e) = \min_{e' \neq e, e' \in \mathcal{E}}\|\bm{v}_e - \bm{v}_{e'}\|_2$
is a diversity distance that penalizes experiences that are too similar to
others in the library (encouraging diversity), and $\alpha, \beta, \gamma, \delta > 0$
are tunable exponents (defaults: $\beta = 1.0$, $\gamma = 0.5$, $\delta = 0.3$).
Experiences are pruned in ascending order of $U(e)$ until $|\mathcal{E}| \leq N_{\max}$.

\begin{remark}
In practice, we recommend hybrid compression for production deployments, as
it balances all three considerations. Diversity compression alone can
inadvertently discard valid experiences that happen to share semantic space
with a higher-confidence entry. Importance compression alone can cause the
library to become stale. Temporal decay alone discards valuable stable patterns.
\end{remark}

\subsection{Memory Manager Algorithm}

\begin{algorithm}[H]
\caption{SemanticMemoryManager.Update}
\label{alg:memory-update}
\begin{algorithmic}[1]
\Require Experience library $\mathcal{E}$, operation set $\Omega^*$,
         capacity $N_{\max}$, compression strategy $\mathcal{C}$
\Ensure Updated $\mathcal{E}'$

\State $\mathcal{E}' \gets \mathcal{E}$
\For{each $\omega \in \Omega^*$}
    \If{$\omega = \textsc{add}(e_{\text{new}})$}
        \State $e_{\text{new}}.\bm{v} \gets \text{Enc}(e_{\text{new}}.\texttt{text})$
        \State $\mathcal{E}' \gets \mathcal{E}' \cup \{e_{\text{new}}\}$
    \ElsIf{$\omega = \textsc{modify}(e_{\text{id}}, e_{\text{new}})$}
        \State Find $e$ with $e.\texttt{id} = e_{\text{id}}$ in $\mathcal{E}'$
        \State $e_{\text{new}}.\bm{v} \gets \text{Enc}(e_{\text{new}}.\texttt{text})$
        \State $e_{\text{new}}.c \gets e.c$ \Comment{Preserve confidence history}
        \State $\mathcal{E}' \gets (\mathcal{E}' \setminus \{e\}) \cup \{e_{\text{new}}\}$
    \ElsIf{$\omega = \textsc{delete}(e_{\text{id}})$}
        \State $\mathcal{E}' \gets \mathcal{E}' \setminus \{e : e.\texttt{id} = e_{\text{id}}\}$
    \ElsIf{$\omega = \textsc{merge}(e_{\text{id}_1}, e_{\text{id}_2}, e_{\text{new}})$}
        \State $c_{\text{merged}} \gets \max(c_{e_{\text{id}_1}}, c_{e_{\text{id}_2}})$
        \State $e_{\text{new}}.c \gets c_{\text{merged}}$
        \State $e_{\text{new}}.\bm{v} \gets \text{Enc}(e_{\text{new}}.\texttt{text})$
        \State $\mathcal{E}' \gets (\mathcal{E}' \setminus \{e_{\text{id}_1}, e_{\text{id}_2}\}) \cup \{e_{\text{new}}\}$
    \EndIf
\EndFor
\If{$|\mathcal{E}'| > N_{\max}$}
    \State $\mathcal{E}' \gets \text{Compress}(\mathcal{E}',\; \mathcal{C},\; N_{\max})$
\EndIf
\State \Return $\mathcal{E}'$
\end{algorithmic}
\end{algorithm}

% ============================================================================
% 5. AGENT-SPECIFIC ADAPTATIONS
% ============================================================================
