\section{Agent-Specific Adaptations}

\label{sec:adaptations}

Beyond the structural adaptations to the GRPO pipeline, Agent-GRPO incorporates
three classes of agent-specific experience types.

\subsection{Tool Selection Experiences}

Tool selection experiences encode when to prefer one tool over another for
a class of tasks. These are the most valuable experiences for reducing tool
call errors, as incorrect tool selection typically leads to cascading failures.

\paragraph{Examples of Tool Selection Experiences.}
\begin{itemize}[leftmargin=1.5em]
    \item \textit{For JSON extraction from \texttt{package.json}, prefer
          \texttt{read\_json} over \texttt{bash\,+\,jq}; 3$\times$ faster,
          no shell escaping required.}
    \item \textit{For Kubernetes pod logs, use \texttt{kubectl\_logs}
          with \texttt{--tail=100} before \texttt{kubectl\_exec}; avoids
          unnecessary shell sessions.}
    \item \textit{When searching code for function definitions, prefer
          \texttt{ast\_search} over \texttt{grep}; handles multiline
          definitions correctly.}
\end{itemize}

Tool selection experiences are tagged with domain $d \in \mathcal{D}$ and
assigned high initial confidence ($c_0 = 0.7$) when extracted from groups with
large positive advantages ($|A^{(i)}| > 1.5$).

\subsection{Failure Pattern Experiences}

Failure pattern experiences encode common failure modes and their remediation
strategies. They are derived primarily from Stage-2 analysis of negative
examples ($A^{(i)} < -0.5$).

\paragraph{Examples of Failure Pattern Experiences.}
\begin{itemize}[leftmargin=1.5em]
    \item \textit{Git push to \texttt{main} fails on protected repos; always
          create a PR branch first.}
    \item \textit{Docker build fails if \texttt{COPY} precedes \texttt{RUN
          apt-get}; reorder to install deps first.}
    \item \textit{Kubernetes \texttt{kubectl apply} without \texttt{--namespace}
          silently targets \texttt{default}; always specify namespace.}
\end{itemize}

Failure pattern experiences carry a \texttt{failure\_mode} annotation that
enables targeted retrieval when the agent encounters error messages matching
known patterns. The retrieval query is augmented with the error message:
$x' = \text{concat}(x, \text{error\_message})$ before embedding.

\subsection{Domain-Specific Convention Experiences}

Convention experiences encode task-domain norms that are not explicit in tool
documentation but emerge from observed agent behavior across many sessions.
These are partitioned by domain $d \in \mathcal{D}$.

\paragraph{Coding Domain ($d_{\text{code}}$).}
\begin{itemize}[leftmargin=1.5em]
    \item Commit message format: \texttt{type(scope): description}.
    \item Test before commit: always run \texttt{pytest} or \texttt{go test}
          before pushing.
    \item PR description template: Summary, Test Plan, Breaking Changes.
\end{itemize}

\paragraph{DevOps Domain ($d_{\text{devops}}$).}
\begin{itemize}[leftmargin=1.5em]
    \item Deployment sequence: build $\to$ push image $\to$ update manifest
          $\to$ apply $\to$ verify rollout.
    \item Rollback procedure: \texttt{kubectl rollout undo} before any
          destructive cleanup.
    \item Health check: \texttt{kubectl get pods} within 60\,s of deployment.
\end{itemize}

\paragraph{Data Science Domain ($d_{\text{ds}}$).}
\begin{itemize}[leftmargin=1.5em]
    \item Always set random seeds before experiments.
    \item Log hyperparameters before training starts.
    \item Validate data shapes immediately after loading.
\end{itemize}

Convention experiences are assigned domain $d$ during Stage-1 summarization
(the LLM is instructed to classify the domain of each trajectory). They are
retrieved with domain filtering active ($D_x = \{d\}$), preventing convention
experiences from one domain from polluting retrieval for another.

\subsection{Hierarchical Experience Structure}

The three experience types form a natural hierarchy:
\begin{equation}
    \text{Convention} \supseteq \text{Tool Selection} \supseteq \text{Failure Pattern}
\end{equation}
in terms of generality. Convention experiences apply broadly; tool selection
experiences apply to specific tool pairs; failure pattern experiences apply
to specific error conditions. At retrieval time, we sample experiences from
all three types, with a soft budget: at most $K_c = 2$ convention experiences,
$K_t = 2$ tool selection experiences, and $K_f = 1$ failure pattern experience
in the default $K = 5$ budget.

% ============================================================================
% 6. INTEGRATION WITH SELF-IMPROVEMENT SYSTEM
% ============================================================================
