\section{Proof of AI (PoAI)}

\label{sec:poai}

\subsection{Design Goals}

PoAI must satisfy four properties:

\begin{enumerate}[leftmargin=1.4em]
    \item \textbf{Correctness}: Honest computation is always accepted.
    \item \textbf{Soundness}: Incorrect computation is detected with high probability.
    \item \textbf{Efficiency}: Verification cost is a small fraction of computation cost.
    \item \textbf{Incentive compatibility}: Rational actors are incentivized to compute honestly.
\end{enumerate}

\subsection{TEE Attestation}

Workers execute AI jobs within Trusted Execution Environments (Intel SGX, AMD SEV, ARM TrustZone)~\cite{costan2016intel}. The TEE generates a hardware-signed attestation:

\begin{definition}[PoAI Attestation]
An attestation is a tuple $\sigma = (h_{\text{input}}, h_{\text{output}}, h_{\text{model}}, t, \text{metrics}, \text{sig}_{\text{TEE}})$ where:
\begin{itemize}[leftmargin=1.1em]
    \item $h_{\text{input}} = \text{SHA-256}(\text{input})$: Hash of job input.
    \item $h_{\text{output}} = \text{SHA-256}(\text{output})$: Hash of job output.
    \item $h_{\text{model}} = \text{SHA-256}(\text{model weights})$: Hash of model used.
    \item $t$: Execution timestamp and duration.
    \item $\text{metrics}$: GPU utilization, memory usage, floating-point operations count.
    \item $\text{sig}_{\text{TEE}}$: Hardware signature from the TEE enclave.
\end{itemize}
\end{definition}

\subsection{Statistical Verification}

TEE attestations can be forged if the TEE hardware is compromised. PoAI adds a statistical verification layer:

\begin{algorithm}[H]
\caption{PoAI Statistical Verification}
\label{alg:poai}
\begin{algorithmic}[1]
\Require Attestation $\sigma$, sampling rate $\alpha = 0.003$
\State \Comment{Level 1: Attestation validation}
\State Verify TEE signature $\text{sig}_{\text{TEE}}$
\State Verify hash chain: $h_{\text{input}}, h_{\text{output}}, h_{\text{model}}$
\State Verify metrics consistency (FLOPs vs. model size vs. duration)
\State \Comment{Level 2: Statistical sampling}
\State Draw $u \sim \text{Uniform}(0, 1)$
\If{$u < \alpha$} \Comment{Selected for full verification}
    \State Re-execute job on a verifier node
    \State Compare output hashes: $h_{\text{output}}' \stackrel{?}{=} h_{\text{output}}$
    \If{mismatch}
        \State Initiate dispute resolution
    \EndIf
\EndIf
\State \Comment{Level 3: Quality scoring}
\State If inference: compare output distribution with reference
\State Update worker quality score: $q_j \gets \alpha \cdot \text{result} + (1-\alpha) \cdot q_j$
\State \Return verification result
\end{algorithmic}
\end{algorithm}

\subsection{Handling Non-Determinism}

LLM inference is non-deterministic due to floating-point ordering and sampling. PoAI handles this via:

\begin{enumerate}[leftmargin=1.4em]
    \item \textbf{Seed pinning}: Workers use a deterministic seed derived from the job hash. Verification re-executes with the same seed.
    \item \textbf{Distribution matching}: For sampling-based generation, we compare the output distribution rather than exact tokens. The KL divergence between the worker's output distribution and the reference must be below a threshold $\epsilon_{\text{KL}}$.
    \item \textbf{Approximate matching}: For numerical outputs (embeddings, logits), we allow element-wise error up to $\epsilon_{\text{num}} = 10^{-4}$ to account for floating-point non-associativity.
\end{enumerate}

\begin{theorem}[PoAI Detection Rate]
\label{thm:detection}
With sampling rate $\alpha$ and $N$ jobs per verification epoch, the probability of detecting a worker who deviates on fraction $\delta$ of their jobs is:
\begin{equation}
P(\text{detect}) = 1 - (1 - \alpha)^{\delta N}.
\end{equation}
For $\alpha = 0.003$, $N = 10{,}000$, $\delta = 0.01$: $P(\text{detect}) = 1 - (1-0.003)^{100} = 0.259$. Over 10 epochs: $P = 1 - (1-0.259)^{10} = 0.953$.
\end{theorem}

The key insight is that workers who cheat on even 1\% of jobs are detected with 95.3\% probability within 10 epochs (10 hours), and the slashing penalty far exceeds the savings from cheating.

\subsection{Slashing Conditions}

Workers are slashed (lose staked collateral) for:

\begin{table}[H]
\centering
\small
\begin{tabular}{lcc}
\toprule
\textbf{Violation} & \textbf{Slash \%} & \textbf{Detection Method} \\
\midrule
Wrong output (verified) & 100\% & Statistical sampling \\
Invalid attestation & 50\% & Signature verification \\
Timeout ($> 2\times$ SLA) & 10\% & Deadline monitoring \\
Quality below threshold & 5\%/epoch & Quality score decay \\
Sybil detected & 100\% & Network analysis \\
\bottomrule
\end{tabular}
\caption{PoAI slashing conditions.}
\label{tab:slashing}
\end{table}

