\section{Multi-Consensus Architecture}

\label{sec:consensus}

\subsection{Dual-Track Design}

ACI uses two consensus mechanisms optimized for different latency requirements:

\begin{definition}[Dual-Track Consensus]
\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Fast-path} (DAG-based, $< 200$ms): Handles job matching, resource allocation, and real-time updates. Uses a Directed Acyclic Graph (DAG) consensus~\cite{danezis2022narwhal} where transactions are ordered by causal dependency rather than total order.
    \item \textbf{Settlement-path} (BFT, $< 2$s): Handles payment finalization, dispute resolution, and governance. Uses a classical BFT consensus~\cite{yin2019hotstuff} with $3f+1$ validators tolerating $f$ Byzantine faults.
\end{itemize}
\end{definition}

\subsection{Fast-Path Consensus}

The fast-path processes job matching as a DAG of causally related transactions:

\begin{algorithm}[H]
\caption{Fast-Path DAG Consensus}
\label{alg:fastpath}
\begin{algorithmic}[1]
\Require Transaction $\text{tx}$ from client/router
\State $\text{vertex} \gets (\text{tx}, \text{round}, \text{author}, \text{parents})$
\State Broadcast $\text{vertex}$ to all validators
\State \Comment{Each validator maintains a local DAG}
\State Add $\text{vertex}$ to local DAG
\State \Comment{Consensus by 2/3 acknowledgment}
\If{$\text{vertex}$ acknowledged by $\ge 2n/3$ validators}
    \State $\text{vertex}$ is committed
    \State Execute job matching transaction
\EndIf
\end{algorithmic}
\end{algorithm}

The DAG structure allows concurrent transaction processing (no sequential block ordering), achieving throughput proportional to network bandwidth rather than consensus rounds.

\subsection{Settlement-Path Consensus}

Payment settlement requires total ordering and finality. We use a modified HotStuff~\cite{yin2019hotstuff} protocol:

\begin{enumerate}[leftmargin=1.4em]
    \item \textbf{Prepare}: Leader proposes a block of settlement transactions.
    \item \textbf{Pre-commit}: Validators verify PoAI attestations and vote.
    \item \textbf{Commit}: Block is committed with $2f+1$ signatures.
    \item \textbf{Decide}: Block is finalized and irreversible.
\end{enumerate}

Settlement blocks are produced every 2 seconds, batching all verified jobs from the fast-path.

\subsection{Cross-Track Coordination}

The fast-path and settlement-path are coordinated via a checkpoint mechanism:

\begin{enumerate}[leftmargin=1.4em]
    \item Every 100 fast-path rounds, a checkpoint is submitted to the settlement-path.
    \item The checkpoint contains a Merkle root of all committed fast-path transactions.
    \item Settlement-path validators verify the checkpoint against their local DAG state.
    \item Once the checkpoint is settled, the corresponding fast-path transactions are considered final.
\end{enumerate}

\subsection{Throughput Analysis}

\begin{table}[H]
\centering
\small
\begin{tabular}{lccc}
\toprule
\textbf{Consensus} & \textbf{TPS} & \textbf{Finality} & \textbf{Validators} \\
\midrule
Ethereum L1 & 15 & 12 min & 900K+ \\
Solana & 3,000 & 12s & 2,000 \\
Sui (Narwhal) & 120,000 & 2s & 100 \\
ACI fast-path & 12,000 & 200ms & 100 \\
ACI settlement & 2,000 & 2s & 100 \\
\bottomrule
\end{tabular}
\caption{Consensus throughput comparison.}
\label{tab:consensus}
\end{table}

