\section{Identity: Hanzo IAM}

\label{sec:identity}
%% ============================================================================

\subsection{Overview}

Hanzo IAM is a multi-tenant identity and access management system implemented as a single Go binary with zero external dependencies beyond PostgreSQL. It supports OAuth~2.0, OpenID Connect, SAML~2.0, LDAP, WebAuthn/FIDO2, and direct username/password authentication. IAM integrates with over 40 social identity providers (Google, GitHub, Apple, Microsoft, Discord, Telegram, etc.) and implements Casbin-based RBAC with per-organization policy engines.

\subsection{Feature Matrix}

\begin{table}[H]
\centering
\caption{Identity provider feature comparison.}
\label{tab:identity}
\footnotesize
\begin{tabular}{@{}lccccc@{}}
\toprule
Feature & \rotatebox{60}{Hanzo IAM} & \rotatebox{60}{Auth0} & \rotatebox{60}{Clerk} & \rotatebox{60}{Supabase} & \rotatebox{60}{Firebase} \\
\midrule
Social providers & 40+ & 30+ & 20+ & 12 & 7 \\
OIDC server & \yes & \yes & \no & \no & \no \\
SAML 2.0 & \yes & \yes & \no & \no & \no \\
WebAuthn/FIDO2 & \yes & \yes & \yes & \no & \no \\
LDAP bridge & \yes & \yes & \no & \no & \no \\
SCIM 2.0 & \yes & \yes & \no & \no & \no \\
Casbin RBAC & \yes & \no & \no & \no & \no \\
Org hierarchy & \yes & \yes & \yes & \no & \no \\
Self-hostable & \yes & \no & \no & \yes & \no \\
Open source & \yes & \no & \no & \yes & \no \\
Argon2id hash & \yes & \no & \no & \yes & \no \\
MFA (TOTP+HW) & \yes & \yes & \yes & \yes & \yes \\
PQ-ready keys & \yes & \no & \no & \no & \no \\
K8s CRD & \yes & \no & \no & \no & \no \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Performance}

Benchmarked on a single 4-vCPU node (AMD EPYC 7R13, 8~GB RAM, PostgreSQL~16):

\begin{table}[H]
\centering
\caption{Authentication latency (median / p99, milliseconds).}
\footnotesize
\begin{tabular}{@{}lrr@{}}
\toprule
System & Median & p99 \\
\midrule
Hanzo IAM & 18 & 42 \\
Auth0 (managed) & 85 & 340 \\
Clerk (managed) & 62 & 210 \\
Supabase Auth & 35 & 120 \\
Firebase Auth & 95 & 450 \\
\bottomrule
\end{tabular}
\end{table}

Auth0 and Clerk latencies include network round-trip to their managed infrastructure. Hanzo IAM runs in-cluster, eliminating this overhead. Supabase Auth is also self-hostable and shows competitive latency; however, it lacks SAML, LDAP, SCIM, and Casbin RBAC.

\subsection{Integration Cost}

\begin{table}[H]
\centering
\caption{Integration effort for a typical multi-tenant SaaS.}
\footnotesize
\begin{tabular}{@{}lrr@{}}
\toprule
System & LoC (glue) & Setup time \\
\midrule
Hanzo IAM & 45 & 15 min \\
Auth0 & 180 & 2 hr \\
Clerk & 120 & 1 hr \\
Supabase Auth & 90 & 45 min \\
Firebase Auth & 150 & 1.5 hr \\
\bottomrule
\end{tabular}
\end{table}

Hanzo IAM integration consists of deploying the \texttt{HanzoIAM} CRD and adding a JWT validation middleware to the service. Auth0 requires configuring rules, actions, and custom database connections. Firebase requires installing the Firebase Admin SDK and implementing custom claims logic.

\subsection{Total Cost of Ownership}

At 100,000 monthly active users (MAU):

\begin{table}[H]
\centering
\caption{Monthly cost at 100K MAU (USD).}
\footnotesize
\begin{tabular}{@{}lr@{}}
\toprule
System & Monthly cost \\
\midrule
Hanzo IAM (self-hosted) & \$18 (compute) \\
Auth0 (B2B Essential) & \$1,300 \\
Clerk (Pro) & \$500 \\
Supabase (Pro) & \$25 + \$0.00325/MAU \\
Firebase (Blaze) & \$0.06/verification \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Key Differentiator}

Hanzo IAM is the only identity provider with integrated Casbin policy engines, SCIM~2.0, and post-quantum-ready key derivation (ML-KEM-768 for key encapsulation during session establishment). It runs as a single binary with a Kubernetes CRD for declarative management. No other system offers self-hosted, open-source identity with this breadth of enterprise protocol support.

%% ============================================================================
