\section{API Gateway: Hanzo Gateway}

\label{sec:gateway}
%% ============================================================================

\subsection{Overview}

Hanzo Gateway extends KrakenD, a stateless Go-based API gateway, with Hanzo-specific integrations: IAM JWT validation via JWKS endpoint auto-discovery, KMS-backed configuration secrets, Prometheus metrics aligned with the Hanzo observability stack, and declarative endpoint definitions across 147+ routes. The gateway requires no database and no distributed cache, enabling horizontal scaling by adding replicas behind a load balancer.

\subsection{Feature Matrix}

\begin{table}[H]
\centering
\caption{API gateway feature comparison.}
\label{tab:gateway}
\footnotesize
\begin{tabular}{@{}lccccc@{}}
\toprule
Feature & \rotatebox{60}{Hanzo} & \rotatebox{60}{Kong} & \rotatebox{60}{Envoy} & \rotatebox{60}{AWS APIGW} & \rotatebox{60}{Nginx} \\
\midrule
Stateless & \yes & \no & \yes & N/A & \yes \\
JWT validation & \yes & \yes & \yes & \yes & \pmark \\
Rate limiting & \yes & \yes & \yes & \yes & \pmark \\
Circuit breaker & \yes & \yes & \yes & \no & \no \\
Response merge & \yes & \no & \no & \no & \no \\
Backend agg. & \yes & \no & \no & \no & \no \\
IAM-native & \yes & \no & \no & \no & \no \\
KMS-native & \yes & \no & \no & \yes & \no \\
Declarative cfg & \yes & \yes & \yes & \yes & \no \\
Zero-downtime & \yes & \yes & \yes & \yes & \pmark \\
Self-hosted & \yes & \yes & \yes & \no & \yes \\
Open source & \yes & \yes & \yes & \no & \yes \\
Plugin system & \yes & \yes & \yes & \yes & \yes \\
K8s CRD & \yes & \yes & \yes & \no & \pmark \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Performance}

Benchmarked using \texttt{wrk2} with 256 concurrent connections, 30-second duration, targeting a pass-through proxy to a static JSON backend on the same host:

\begin{table}[H]
\centering
\caption{Gateway throughput and latency overhead.}
\footnotesize
\begin{tabular}{@{}lrrr@{}}
\toprule
System & RPS & Median (ms) & p99 (ms) \\
\midrule
Hanzo Gateway & 45,200 & 1.2 & 4.8 \\
Kong (dbless) & 28,400 & 2.1 & 12.3 \\
Envoy & 52,100 & 0.9 & 3.2 \\
Nginx (proxy) & 61,300 & 0.6 & 2.1 \\
AWS API GW & 10,000 & 8.5 & 45.0 \\
\bottomrule
\end{tabular}
\end{table}

Envoy and Nginx outperform Hanzo Gateway on raw proxy throughput. However, neither provides built-in response merging, backend aggregation, or IAM-native JWT validation. Kong's database dependency (even in dbless mode, it loads declarative config into an in-memory store) adds overhead. AWS API Gateway's 10,000 RPS limit is a hard regional default that requires quota increases.

\subsection{Key Differentiator}

Hanzo Gateway combines KrakenD's zero-allocation proxy pipeline with native Hanzo IAM JWKS validation and KMS secret injection. A single JSON configuration file defines all 147+ endpoints, including backend aggregation rules that merge responses from multiple microservices into a single client response. No competing gateway provides this combination without additional plugins or sidecars.

%% ============================================================================
