\section{Runtime: Hanzo Base}

\label{sec:runtime}
%% ============================================================================

\subsection{Overview}

Hanzo Base is a fork of PocketBase extended with a Goja-based JavaScript VM for server-side scripting, native Hanzo IAM integration, multi-tenant data isolation via organization-scoped collections, and S3-compatible storage backends. It provides a real-time database (SQLite or PostgreSQL), REST API auto-generation, file storage, and an admin dashboard in a single binary.

\subsection{Feature Matrix}

\begin{table}[H]
\centering
\caption{Application runtime comparison.}
\footnotesize
\begin{tabular}{@{}lccccc@{}}
\toprule
Feature & \rotatebox{60}{Base} & \rotatebox{60}{Supabase} & \rotatebox{60}{Firebase} & \rotatebox{60}{Hasura} & \rotatebox{60}{Appwrite} \\
\midrule
Single binary & \yes & \no & \no & \no & \no \\
JS scripting & \yes & \yes & \yes & \no & \yes \\
Real-time DB & \yes & \yes & \yes & \yes & \yes \\
Auto REST API & \yes & \yes & \no & \no & \yes \\
GraphQL & \no & \no & \no & \yes & \yes \\
File storage & \yes & \yes & \yes & \no & \yes \\
Admin UI & \yes & \yes & \yes & \yes & \yes \\
Multi-tenant & \yes & \pmark & \no & \yes & \pmark \\
IAM-native & \yes & \no & \no & \no & \no \\
Self-hosted & \yes & \yes & \no & \yes & \yes \\
Open source & \yes & \yes & \no & \yes & \yes \\
Goja VM & \yes & \no & \no & \no & \no \\
SQLite mode & \yes & \no & \no & \no & \no \\
K8s CRD & \yes & \no & \no & \no & \no \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Performance}

Benchmarked on a single 2-vCPU node (SQLite mode, 1000 records, 100 concurrent connections):

\begin{table}[H]
\centering
\caption{CRUD operation latency (median, milliseconds).}
\footnotesize
\begin{tabular}{@{}lrrrr@{}}
\toprule
System & Create & Read & Update & Delete \\
\midrule
Hanzo Base & 0.8 & 0.3 & 0.7 & 0.5 \\
Supabase & 3.2 & 1.1 & 2.8 & 2.1 \\
Firebase & 12.0 & 4.5 & 11.0 & 8.0 \\
Hasura & 2.5 & 0.9 & 2.2 & 1.8 \\
Appwrite & 4.1 & 1.5 & 3.6 & 2.9 \\
\bottomrule
\end{tabular}
\end{table}

Hanzo Base in SQLite mode achieves sub-millisecond reads because the database is embedded in the same process with zero network overhead. PostgreSQL mode adds approximately 1--2ms per operation for network round-trip. Supabase and Hasura require separate PostgreSQL instances, contributing to their higher latencies.

\subsection{Integration Cost}

Deploying a new Hanzo Base instance with IAM authentication, multi-tenant isolation, and S3-compatible storage requires 12 lines of Kubernetes YAML (one \texttt{HanzoBase} CRD). The equivalent Supabase deployment requires 4 separate services (PostgreSQL, GoTrue, PostgREST, and Storage), each with independent configuration.

\subsection{Key Differentiator}

The embedded Goja JavaScript VM enables server-side hooks, computed fields, and custom endpoints without deploying additional services. Combined with native IAM integration, Hanzo Base replaces what would otherwise be Supabase + Auth0 + a custom Node.js service for business logic.

%% ============================================================================
