\section{Secrets: Hanzo KMS}

\label{sec:kms}
%% ============================================================================

\subsection{Overview}

Hanzo KMS is a fork of Infisical adapted for the Hanzo ecosystem. It provides centralized secrets management with end-to-end encryption, environment-scoped secret injection, Kubernetes CRD-based synchronization, and audit logging. Secrets are encrypted at rest using AES-256-GCM with keys derived from a master key stored in a hardware security module (HSM) or cloud KMS (AWS KMS, GCP KMS).

\subsection{Feature Matrix}

\begin{table}[H]
\centering
\caption{Secrets management comparison.}
\footnotesize
\begin{tabular}{@{}lccccc@{}}
\toprule
Feature & \rotatebox{60}{Hanzo KMS} & \rotatebox{60}{Vault} & \rotatebox{60}{AWS KMS} & \rotatebox{60}{Doppler} & \rotatebox{60}{Infisical} \\
\midrule
E2E encryption & \yes & \yes & \yes & \yes & \yes \\
Env scoping & \yes & \yes & \no & \yes & \yes \\
K8s sync & \yes & \yes & \pmark & \yes & \yes \\
HSM backend & \yes & \yes & \yes & \no & \yes \\
Dynamic secrets & \no & \yes & \no & \no & \no \\
Secret rotation & \yes & \yes & \yes & \yes & \yes \\
Audit log & \yes & \yes & \yes & \yes & \yes \\
IAM-native & \yes & \no & \no & \no & \no \\
Self-hosted & \yes & \yes & \no & \no & \yes \\
Open source & \yes & \yes & \no & \no & \yes \\
Web UI & \yes & \yes & \yes & \yes & \yes \\
CLI & \yes & \yes & \yes & \yes & \yes \\
K8s CRD & \yes & \yes & \no & \no & \yes \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Performance and Operational Complexity}

HashiCorp Vault is the most feature-rich alternative, offering dynamic secrets (database credentials generated on demand), PKI certificate authority, and transit encryption. However, Vault requires an unsealing ceremony (Shamir's secret sharing or auto-unseal via cloud KMS), a storage backend (Consul, etcd, or integrated Raft), and significant operational expertise. A production Vault deployment typically requires 3--5 nodes and a dedicated operations team.

Hanzo KMS trades Vault's dynamic secret generation for operational simplicity: a single binary, PostgreSQL-backed storage, and automatic K8s secret synchronization via the \texttt{KMSSecret} CRD. For organizations that do not require dynamic secrets, this eliminates approximately 80\% of Vault's operational surface.

\begin{table}[H]
\centering
\caption{Operational complexity comparison.}
\footnotesize
\begin{tabular}{@{}lrr@{}}
\toprule
System & Min nodes & Config files \\
\midrule
Hanzo KMS & 1 & 1 \\
HashiCorp Vault & 3 & 5+ \\
AWS KMS & N/A & IAM policies \\
Doppler & N/A & 0 (managed) \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Key Differentiator}

Hanzo KMS authenticates against Hanzo IAM, meaning the same OIDC token that authenticates a developer to the platform also authorizes their access to secrets. No separate credential management. The \texttt{KMSSecret} CRD automatically syncs secrets to Kubernetes namespaces, eliminating manual \texttt{kubectl create secret} workflows.

%% ============================================================================
