\section{Tier Complexity Thresholds}

\label{app:thresholds}

The complexity thresholds $\theta_j$ used in Algorithm~\ref{alg:tier} were
tuned on a 2,000-task calibration set:

\begin{table}[H]
\centering
\caption{Tier Complexity Thresholds and Typical Task Types}
\label{tab:thresholds}
\begin{tabular}{clll}
\toprule
\textbf{Tier} & $\bm{\theta_j}$ & \textbf{Typical Task Type} &
\textbf{Example} \\
\midrule
$T_0$ & $[0.0, 0.3)$ & Q\&A, summarization, translation &
    ``What is RAG?'' \\
$T_1$ & $[0.3, 0.55)$ & Code gen, data analysis, scripting &
    ``Fix this bug'' \\
$T_2$ & $[0.55, 0.70)$ & Web automation, GUI workflows &
    ``Book a flight'' \\
$T_3$ & $[0.70, 0.88)$ & System admin, multi-service tasks &
    ``Deploy staging env'' \\
$T_4$ & $[0.88, 1.0]$ & Hardware access, GPU compute &
    ``Train classifier'' \\
\bottomrule
\end{tabular}
\end{table}

% ============================================================================
% REFERENCES
% ============================================================================
\bibliographystyle{plainnat}
\begin{thebibliography}{40}

\bibitem[Aider(2024)]{aider2024}
Aider.
\newblock Aider: AI pair programming in your terminal.
\newblock \url{https://aider.chat}, 2024.

\bibitem[Anthropic(2024a)]{anthropic2024claude}
Anthropic.
\newblock The Claude 3 model family: Opus, Sonnet, Haiku.
\newblock Technical report, Anthropic, 2024.

\bibitem[Anthropic(2024b)]{anthropic2024computer}
Anthropic.
\newblock Introducing computer use, a new Claude 3.5 Sonnet, and Claude 3.5 Haiku.
\newblock \url{https://www.anthropic.com/news/3-5-models-and-computer-use}, 2024.

\bibitem[AutoGPT(2023)]{autogpt2023}
AutoGPT Team.
\newblock AutoGPT: An autonomous GPT-4 experiment.
\newblock \url{https://github.com/Significant-Gravitas/AutoGPT}, 2023.

\bibitem[Brown et~al.(2020)]{brown2020language}
Brown, T., Mann, B., Ryder, N., Subbiah, M., et~al.
\newblock Language models are few-shot learners.
\newblock \emph{NeurIPS}, 33:1877--1901, 2020.

\bibitem[Casdoor(2023)]{casdoor2023}
Casbin.
\newblock Casdoor: An open-source identity and access management platform.
\newblock \url{https://casdoor.org}, 2023.

\bibitem[Clark et~al.(2005)]{clark2005live}
Clark, C., Fraser, K., Hand, S., Hansen, J.~G., Jul, E., Limpach, C.,
  Pratt, I., and Warfield, A.
\newblock Live migration of virtual machines.
\newblock \emph{NSDI}, 2:273--286, 2005.

\bibitem[Cognition(2024)]{cognition2024devin}
Cognition Labs.
\newblock Introducing Devin, the first AI software engineer.
\newblock \url{https://www.cognition-labs.com/blog}, 2024.

\bibitem[Cursor(2024)]{cursor2024}
Anysphere Inc.
\newblock Cursor: The AI-first code editor.
\newblock \url{https://cursor.com}, 2024.

\bibitem[Hanzo AI(2024a)]{hanzo2024dso}
Hanzo AI Research.
\newblock Decentralized semantic optimization: Collaborative knowledge
  accumulation for AI agents.
\newblock Technical report, Hanzo AI Inc, 2024.

\bibitem[Hanzo AI(2024b)]{hanzo2024operative}
Hanzo AI Research.
\newblock Operative: A computer use framework for autonomous AI agents.
\newblock Technical report, Hanzo AI Inc, 2024.

\bibitem[Hanzo AI(2024c)]{hanzo2024sdk}
Hanzo AI Research.
\newblock Hanzo agent SDK: Building AI agents with orchestration, tool use,
  and memory systems.
\newblock Technical report, Hanzo AI Inc, 2024.

\bibitem[Infisical(2023)]{infisical2023}
Infisical Inc.
\newblock Infisical: Open-source secret management platform.
\newblock \url{https://infisical.com}, 2023.

\bibitem[Jimenez et~al.(2024)]{jimenez2024swebench}
Jimenez, C.~E., Yang, J., Wettig, A., Yao, S., Pei, K., Press, O., and
  Narasimhan, K.
\newblock SWE-bench: Can language models resolve real-world GitHub issues?
\newblock \emph{ICLR}, 2024.

\bibitem[LangChain(2022)]{langchain2022}
Chase, H.
\newblock LangChain: Building applications with LLMs through composability.
\newblock \url{https://github.com/langchain-ai/langchain}, 2022.

\bibitem[McMahan et~al.(2017)]{mcmahan2017communication}
McMahan, B., Moore, E., Ramage, D., Hampson, S., and y~Arcas, B.~A.
\newblock Communication-efficient learning of deep networks from decentralized
  data.
\newblock \emph{AISTATS}, pp.\ 1273--1282, 2017.

\bibitem[Anthropic(2024c)]{mcp2024}
Anthropic.
\newblock Model Context Protocol: An open standard for connecting AI models
  to tools and data.
\newblock \url{https://modelcontextprotocol.io}, 2024.

\bibitem[NATS(2023)]{nats2023}
Synadia Communications.
\newblock NATS: Cloud native messaging.
\newblock \url{https://nats.io}, 2023.

\bibitem[OpenAI(2023)]{openai2023gpt4}
OpenAI.
\newblock GPT-4 technical report.
\newblock \emph{arXiv preprint arXiv:2303.08774}, 2023.

\bibitem[OpenAI(2025)]{openai2025operator}
OpenAI.
\newblock Introducing Operator.
\newblock \url{https://openai.com/blog/introducing-operator}, 2025.

\bibitem[OpenTelemetry(2023)]{opentelemetry2023}
Cloud Native Computing Foundation.
\newblock OpenTelemetry: High-quality, ubiquitous, and portable telemetry.
\newblock \url{https://opentelemetry.io}, 2023.

\bibitem[Park et~al.(2023)]{park2023generative}
Park, J.~S., O'Brien, J., Cai, C.~J., Morris, M.~R., Liang, P., and
  Bernstein, M.~S.
\newblock Generative agents: Interactive simulacra of human behavior.
\newblock \emph{UIST}, pp.\ 1--22, 2023.

\bibitem[Shapiro et~al.(2011)]{shapiro2011conflict}
Shapiro, M., Pregui{\c{c}}a, N., Baquero, C., and Zawirski, M.
\newblock Conflict-free replicated data types.
\newblock \emph{Symposium on Self-Stabilizing Systems}, pp.\ 386--400, 2011.

\bibitem[Soares et~al.(2015)]{soares2015corrigibility}
Soares, N., Fallenstein, B., Yudkowsky, E., and Armstrong, S.
\newblock Corrigibility.
\newblock \emph{AAAI Workshop on AI and Ethics}, 2015.

\bibitem[Tailscale(2023)]{tailscale2023}
Tailscale Inc.
\newblock Tailscale: Zero config VPN.
\newblock \url{https://tailscale.com}, 2023.

\bibitem[Temporal(2023)]{temporal2023}
Temporal Technologies.
\newblock Temporal: Open source durable execution system.
\newblock \url{https://temporal.io}, 2023.

\bibitem[Wu et~al.(2023)]{wu2023autogen}
Wu, Q., Bansal, G., Zhang, J., Wu, Y., Zhang, S., Zhu, E., Li, B., Jiang, L.,
  Zhang, X., and Wang, C.
\newblock AutoGen: Enabling next-gen LLM applications via multi-agent conversation.
\newblock \emph{arXiv preprint arXiv:2308.08155}, 2023.

\bibitem[Yang et~al.(2024)]{yang2024sweagent}
Yang, J., Jimenez, C.~E., Wettig, A., Lieret, K., Yao, S., Narasimhan, K.,
  and Press, O.
\newblock SWE-agent: Agent-computer interfaces enable automated software
  engineering.
\newblock \emph{NeurIPS}, 2024.

\end{thebibliography}

