\section{Adaptive Continuous Batching}

\label{sec:batching}
% ============================================================================

Static batching wastes GPU cycles: short sequences pad to the longest sequence
in the batch, and completed sequences block new admissions until the entire
batch finishes. Continuous batching~\cite{orca2022} solves this by admitting and
evicting sequences at each decode step.

\subsection{Iteration-Level Scheduling}

At each decode iteration $t$, the scheduler decides which sequences to include
in the batch $\mathcal{B}_t$:

\begin{equation}
\mathcal{B}_t = \arg\max_{\mathcal{B} \subseteq \mathcal{R}_t}
\sum_{r \in \mathcal{B}} w(r) \quad \text{s.t.} \quad
\sum_{r \in \mathcal{B}} k(r) \leq K_{\max},
\label{eq:batch}
\end{equation}
where $\mathcal{R}_t$ is the set of runnable sequences, $w(r)$ is priority
weight, and $k(r)$ is the KV-cache block count for sequence $r$.
$K_{\max}$ is the total available cache blocks.

\begin{proposition}[Throughput Bound]
Under continuous batching with $K_{\max}$ cache blocks and average sequence
length $\bar{n}$, the maximum sustainable throughput is:
\begin{equation}
\Theta_{\max} = \frac{K_{\max}}{\bar{n}} \cdot f_{\text{iter}},
\end{equation}
where $f_{\text{iter}}$ is the iteration frequency (decode steps per second).
\end{proposition}

\subsection{Priority-Aware Batching}

Not all requests are equal. Realtime requests (chat) must preempt batch
requests (document processing). We assign priority weights:
\begin{equation}
w(r) = \begin{cases}
    10.0 & \text{if } \tau_r = \texttt{realtime}, \\
    1.0 & \text{if } \tau_r = \texttt{standard}, \\
    0.1 & \text{if } \tau_r = \texttt{batch}.
\end{cases}
\end{equation}

When GPU memory is contiguous, realtime requests preempt batch sequences via
KV-cache swapping to CPU memory. The preempted sequence resumes when GPU memory
becomes available.

\subsection{Prefill--Decode Disaggregation}

Prefill (processing input tokens) and decode (generating output tokens) have
fundamentally different compute profiles. Prefill is compute-bound; decode is
memory-bandwidth-bound. We separate them:

\begin{enumerate}[leftmargin=1.4em]
    \item \textbf{Prefill phase}: Dedicated GPU SMs process the input sequence
    in a single forward pass, populating the KV-cache.
    \item \textbf{Decode phase}: The sequence joins the continuous batch for
    autoregressive generation.
\end{enumerate}

On H100 GPUs with 80GB HBM3, this disaggregation improves prefill throughput by
2.1$\times$ and reduces TTFT by 45\% compared to unified processing.

% ============================================================================
