\section{Multi-Model GPU Sharing}

\label{sec:sharing}
% ============================================================================

Dedicating one GPU per model wastes capacity. A 7B-parameter model in FP16
occupies 14GB; an 80GB H100 can host multiple models simultaneously.

\subsection{Spatial Multiplexing}

We partition GPU memory into fixed regions:
\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Model weights region}: Pinned allocations for model parameters.
    Multiple models co-reside if total weight size fits.
    \item \textbf{KV-cache pool}: Shared pool of cache blocks, allocated
    dynamically across models based on demand.
    \item \textbf{Workspace}: Scratch memory for intermediate activations,
    shared across all models (only one model computes at a time per SM group).
\end{itemize}

\begin{definition}[Memory Budget]
For a GPU with total memory $M_{\text{total}}$, the allocation is:
\begin{equation}
M_{\text{total}} = \sum_{i=1}^{N} M_{\text{weights}}^{(i)} +
M_{\text{kv}} + M_{\text{workspace}} + M_{\text{reserved}},
\end{equation}
where $N$ is the number of co-located models, $M_{\text{kv}}$ is the shared
KV-cache pool, $M_{\text{workspace}}$ is the activation workspace, and
$M_{\text{reserved}}$ is a safety margin (typically 2GB).
\end{definition}

\subsection{Temporal Multiplexing}

When models cannot co-reside (e.g., two 70B models on one GPU), we use temporal
multiplexing with fast model swapping:
\begin{enumerate}[leftmargin=1.4em]
    \item Model weights are kept in host pinned memory (PCIe-accessible).
    \item On model switch, weights are streamed via PCIe Gen5 at 64 GB/s.
    \item A 14GB model swap completes in $\sim$220ms. KV-cache is preserved
    across swaps for the active model's in-flight sequences.
\end{enumerate}

\begin{theorem}[Sharing Efficiency]
Let $\lambda_i$ be the request rate for model $i$ and $\mu_i$ be its service
rate on a dedicated GPU. Under spatial multiplexing with $N$ models, the
effective service rate for model $i$ is:
\begin{equation}
\mu_i^{\text{shared}} = \mu_i \cdot \frac{M_{\text{kv}}^{(i)}}{M_{\text{kv}}},
\end{equation}
where $M_{\text{kv}}^{(i)}$ is the KV-cache allocation for model $i$.
Sharing is efficient when $\sum_i \lambda_i / \mu_i^{\text{shared}} < 1$.
\end{theorem}

% ============================================================================
