\section{Conclusion}

\label{sec:conclusion}
% ============================================================================

We have described the GPU inference architecture of Hanzo Cloud, achieving
3.2$\times$ throughput improvement and p99 latency under 180ms through adaptive
continuous batching, multi-model GPU sharing, and GPU-aware cluster scheduling.
The system serves 47M+ monthly inference requests across 192 GPUs with 94\%
average utilization. All optimizations are deployed in production and integrated
with the Hanzo ACI compute market, where GPU operators earn HANZO tokens for
verified inference work.

% ============================================================================
% BIBLIOGRAPHY
% ============================================================================
\begin{thebibliography}{20}

\bibitem{hanzoaci}
{Hanzo AI Research}.
\newblock Hanzo {ACI}: {AI} Chain Infrastructure for decentralized compute markets.
\newblock \emph{Hanzo AI Research}, December 2024.

\bibitem{hanzocandle}
{Hanzo AI Research}.
\newblock Hanzo Candle: Rust-native {ML} framework for production inference.
\newblock \emph{Hanzo AI Research}, 2024.

\bibitem{hanzojin}
{Hanzo AI Research}.
\newblock Jin: Unified multimodal {AI} architecture.
\newblock \emph{Hanzo AI Research}, 2025.

\bibitem{hanzogateway}
{Hanzo AI Research}.
\newblock Hanzo Gateway: {API} gateway with {KrakenD} backend.
\newblock \emph{Hanzo AI Research}, 2024.

\bibitem{hanzoiam}
{Hanzo AI Research}.
\newblock Hanzo {IAM}: Identity and access management for {AI} platforms.
\newblock \emph{Hanzo AI Research}, 2024.

\bibitem{hanzoedge}
{Hanzo AI Research}.
\newblock Hanzo Edge Inference: On-device and edge {AI} serving.
\newblock \emph{Hanzo AI Research}, 2025.

\bibitem{luxwhitepaper}
{Lux Partners}.
\newblock Lux: A scalable multi-consensus blockchain with post-quantum cryptography.
\newblock \emph{Lux Network Technical Report}, 2024.

\bibitem{vllm2023}
W.~Kwon, Z.~Li, S.~Zhuang, Y.~Sheng, L.~Zheng, C.~Yu, J.~Gonzalez, H.~Zhang, and I.~Stoica.
\newblock Efficient memory management for large language model serving with {PagedAttention}.
\newblock In \emph{Proc.\ SOSP}, 2023.

\bibitem{orca2022}
G.~Yu, J.~Jeong, G.~Kim, S.~Kim, and B.~Chun.
\newblock Orca: A distributed serving system for {Transformer}-based generative models.
\newblock In \emph{Proc.\ OSDI}, 2022.

\bibitem{tensorrt}
{NVIDIA}.
\newblock {TensorRT-LLM}: High-performance inference for large language models.
\newblock \emph{NVIDIA Developer Documentation}, 2024.

\end{thebibliography}

