\section{Token Distribution}

\label{sec:distribution}
% ============================================================================

The genesis allocation distributes 1 billion HANZO tokens across five categories
designed to balance immediate liquidity, long-term alignment, and ecosystem
growth.

\begin{table}[H]
\centering
\begin{tabular}{lrl}
\toprule
\textbf{Category} & \textbf{Allocation} & \textbf{Tokens} \\
\midrule
Community \& Ecosystem & 40\% & 400,000,000 \\
Team \& Advisors & 20\% & 200,000,000 \\
Treasury (DAO) & 15\% & 150,000,000 \\
Infrastructure & 15\% & 150,000,000 \\
Liquidity & 10\% & 100,000,000 \\
\midrule
\textbf{Total} & \textbf{100\%} & \textbf{1,000,000,000} \\
\bottomrule
\end{tabular}
\caption{Genesis token distribution.}
\label{tab:distribution}
\end{table}

\subsection{Community \& Ecosystem (40\%)}

The largest allocation funds long-term ecosystem growth:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Compute mining rewards} (20\%): Distributed to inference node
    operators proportional to verified compute work (PoAI attestations). Emission
    follows a halving schedule: 50\% emitted in years 1--4, 25\% in years 5--8,
    12.5\% in years 9--12, remainder over years 13--20.
    \item \textbf{Developer grants} (10\%): Allocated by DAO governance to
    projects building on Hanzo infrastructure. Minimum 6-month milestone-based
    vesting.
    \item \textbf{User rewards} (5\%): Retroactive distribution to active
    platform users based on compute consumption, staking participation, and
    governance activity.
    \item \textbf{Ecosystem partnerships} (5\%): Strategic allocations for
    integrations with Lux Exchange, Zoo Labs, and third-party platforms.
\end{itemize}

\subsection{Team \& Advisors (20\%)}

Subject to a 4-year vesting schedule with a 1-year cliff:

\begin{itemize}[leftmargin=1.1em]
    \item No tokens transferable during the first 12 months (cliff period).
    \item After the cliff, tokens vest linearly over the remaining 36 months.
    \item Unvested tokens of departing team members return to the treasury.
\end{itemize}

\subsection{Treasury (15\%)}

Governed by the Hanzo DAO with a 7-day timelock and multi-sig execution:

\begin{itemize}[leftmargin=1.1em]
    \item Security audits and bug bounties.
    \item Operational expenses (infrastructure, legal, compliance).
    \item Emergency reserves for market stabilization.
    \item Proposals require $>$50\% quorum and $>$66\% approval.
\end{itemize}

\subsection{Infrastructure (15\%)}

Reserved for node operators and compute providers:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Validator incentives} (8\%): Rewards for validators maintaining
    consensus on Hanzo Network.
    \item \textbf{Compute provider bootstrap} (7\%): Initial subsidies for GPU
    providers joining the network, tapering over 3 years.
\end{itemize}

\subsection{Liquidity (10\%)}

Dedicated to DEX pool liquidity on the Lux Exchange:

\begin{itemize}[leftmargin=1.1em]
    \item HANZO/WLUX pool on Lux V3 (concentrated liquidity).
    \item HANZO/LUSDC pool on Lux V2 (full-range).
    \item Cross-chain HANZO/ETH pool via Teleport bridge.
    \item Protocol-owned liquidity (POL) managed by the DAO, never sold.
\end{itemize}

% ============================================================================
