\section{Zen Model Economics}

\label{sec:zen}
% ============================================================================

The Zen model family comprises 26 specialized models built on the Qwen3+
architecture. Each model has distinct compute costs and pricing:

\begin{table}[H]
\centering
\small
\begin{tabular}{llrr}
\toprule
\textbf{Model} & \textbf{Specialization} & \textbf{Params} &
\textbf{Cost/1K tokens} \\
\midrule
zen-coder & Code generation & 72B & 0.12 HANZO \\
zen-artist & Image generation & 14B & 0.45 HANZO \\
zen-scribe & Long-form writing & 32B & 0.08 HANZO \\
zen-designer & UI/UX generation & 14B & 0.35 HANZO \\
zen-analyst & Data analysis & 32B & 0.10 HANZO \\
zen-reason & Chain-of-thought & 72B & 0.15 HANZO \\
zen-chat & Conversational & 8B & 0.03 HANZO \\
zen-embed & Embedding & 1.5B & 0.01 HANZO \\
\bottomrule
\end{tabular}
\caption{Selected Zen models with per-1K-token pricing in HANZO. Full catalog of 26 models available at docs.hanzo.ai.}
\label{tab:zen}
\end{table}

\subsection{Model Provider Economics}

Third-party model providers can list models on the Hanzo Marketplace. The
economic structure for providers:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Listing fee}: One-time 1,000 HANZO to list a model (burned).
    \item \textbf{Revenue split}: Provider receives 70\%, platform receives 30\%.
    \item \textbf{Quality bond}: Providers stake 10,000 HANZO per model, slashable
    for SLA violations.
    \item \textbf{Governance approval}: New models require DAO vote for inclusion
    in the default catalog.
\end{itemize}

\subsection{Training-Free Adaptation}

Hanzo's Active Semantic Optimization (ASO) and Decentralized Semantic
Optimization (DSO) protocols~\cite{hanzoaso, hanzodso} enable zero-training model
adaptation via Product-of-Experts (PoE) decoding. Users pay for experiential
prior retrieval at a fraction of fine-tuning cost:

\begin{itemize}[leftmargin=1.1em]
    \item Prior retrieval: 0.001 HANZO per query (10$\times$ cheaper than
    inference).
    \item Prior submission: Free, but requires quality bond of 25 HANZO.
    \item Quality rewards: High-quality priors earn 0.01 HANZO per downstream
    usage.
\end{itemize}

% ============================================================================
