\section{Defense \& Enterprise}

\label{sec:defense}
% ============================================================================

Hanzo's defense division publishes research and provides enterprise-grade AI
infrastructure for government and defense applications. This creates a distinct
revenue stream that feeds back into the token economy.

\subsection{Research Portfolio}

Hanzo has published 10 defense-focused papers spanning:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Agentic AI}: Multi-agent coordination for autonomous systems,
    federated agent networks, self-improving agent architectures.
    \item \textbf{Formal verification}: Harness-based testing, unified test
    frameworks, provably correct AI system deployment.
    \item \textbf{Quantum sensing}: Post-quantum cryptographic integration,
    quantum-resistant key management.
\end{itemize}

\subsection{Enterprise Token Utility}

Enterprise customers interact with the HANZO economy through:

\begin{itemize}[leftmargin=1.1em]
    \item \textbf{Volume commitments}: Pre-purchasing HANZO at a discount for
    guaranteed compute capacity (1-year minimum term).
    \item \textbf{Dedicated staking}: Enterprise nodes with dedicated GPU
    allocation, requiring higher stake (500,000+ HANZO) but receiving
    priority scheduling.
    \item \textbf{Compliance features}: KMS-integrated key management, SOC 2 Type
    II audited infrastructure, ITAR-compliant compute isolation.
    \item \textbf{SLA guarantees}: On-chain SLA contracts with automatic
    compensation (in HANZO) for downtime or latency violations.
\end{itemize}

% ============================================================================
