\section{Comparison to Other AI Tokens}

\label{sec:comparison}
% ============================================================================

\begin{table}[H]
\centering
\small
\begin{tabular}{lccccc}
\toprule
\textbf{Feature} & \textbf{HANZO} & \textbf{FET} & \textbf{RNDR} &
\textbf{TAO} & \textbf{AKT} \\
\midrule
Fixed supply & \checkmark & \checkmark & \checkmark & $\times$ & \checkmark \\
Fee burn & \checkmark & $\times$ & $\times$ & $\times$ & $\times$ \\
Compute payment & \checkmark & $\times$ & \checkmark & $\times$ & \checkmark \\
Node staking & \checkmark & \checkmark & $\times$ & \checkmark & \checkmark \\
Governance & \checkmark & \checkmark & $\times$ & \checkmark & \checkmark \\
Fee reduction & \checkmark & $\times$ & $\times$ & $\times$ & $\times$ \\
Revenue share & \checkmark & $\times$ & $\times$ & $\times$ & $\times$ \\
Own L1 chain & \checkmark & \checkmark & $\times$ & \checkmark & \checkmark \\
Model serving & \checkmark & $\times$ & $\times$ & \checkmark & $\times$ \\
Full PaaS & \checkmark & $\times$ & $\times$ & $\times$ & $\times$ \\
\bottomrule
\end{tabular}
\caption{Feature comparison of AI infrastructure tokens.}
\label{tab:comparison}
\end{table}

\paragraph{Fetch.ai (FET).} Agent-focused with staking and governance, but no
direct compute payment utility. The token primarily coordinates autonomous
agents rather than pricing infrastructure.

\paragraph{Render (RNDR).} GPU rendering marketplace with direct compute
payment, but no staking, governance, or fee reduction. The token is purely
transactional, subject to the velocity problem.

\paragraph{Bittensor (TAO).} Decentralized model training with staking and
governance, but inflationary emissions with no burn. TAO's perpetual inflation
dilutes holders without bound.

\paragraph{Akash (AKT).} General-purpose compute marketplace with staking and
governance. Closest to HANZO in design, but lacks fee burn, fee reduction,
revenue sharing, and integrated model serving.

\paragraph{HANZO differentiation.} HANZO is the only AI token that combines
all five utility dimensions (compute payment, staking, governance, fee
reduction, revenue sharing) with deflationary burn, an integrated PaaS, and
native model serving. The closest competitor (AKT) lacks three of these five
dimensions.

% ============================================================================
