\section{Introduction}

\label{sec:intro}
% ============================================================================

Hanzo infrastructure depends on cryptographic signing for security-critical
operations:
\begin{enumerate}[leftmargin=1.4em]
    \item \textbf{IAM session tokens}: JWT access tokens signed with Ed25519,
    issued by Hanzo IAM~\cite{hanzoiam} and validated by every service in the
    platform.
    \item \textbf{KMS envelope encryption}: Master keys that encrypt all
    secrets in Hanzo KMS~\cite{hanzokms}. Compromise of a master key exposes
    every secret in the system.
    \item \textbf{Agent credentials}: AI agents~\cite{hanzoagentsdk} receive
    delegated credentials for tool access, API calls, and resource management.
    The signing key authorizing these delegations must not be compromised.
    \item \textbf{ACI compute attestations}: GPU nodes attest to inference
    results by signing output hashes. These attestations are verified on-chain
    for the Hanzo ACI compute market~\cite{hanzoaci}.
\end{enumerate}

In all cases, a single signing key stored on a single server is a single point
of failure. If that server is compromised, the attacker can forge IAM tokens,
decrypt all secrets, issue fraudulent agent credentials, or submit false compute
attestations.

Threshold signatures solve this by distributing the signing key across $n$
servers such that any $t$ must cooperate to sign, but fewer than $t$ servers
learn nothing about the key. The combined public key is a standard Ed25519 or
ECDSA public key---verifiers need no knowledge of the threshold scheme.

\subsection{Protocol Selection}

We use two protocols, chosen for specific properties:

\begin{table}[H]
\centering
\small
\begin{tabular}{lll}
\toprule
\textbf{Protocol} & \textbf{Curve} & \textbf{Use Case} \\
\midrule
FROST~\cite{frost} & Ed25519 & IAM tokens, agent credentials, API signing \\
CGGMP21~\cite{cggmp21} & secp256k1 & Blockchain transactions, ACI staking \\
\bottomrule
\end{tabular}
\caption{Protocol selection by use case.}
\label{tab:protocols}
\end{table}

FROST is preferred for Ed25519 operations due to its simplicity (2-round
signing, no zero-knowledge proofs during signing). CGGMP21 is required for
ECDSA because Schnorr-based threshold constructions do not produce valid ECDSA
signatures.

% ============================================================================
