\section{FROST for Ed25519}

\label{sec:frost}
% ============================================================================

\subsection{Distributed Key Generation}

FROST DKG produces key shares without any party learning the full secret key.
Each participant $P_i$ acts as both dealer and verifier in a Pedersen
DKG~\cite{pedersen1991}:

\begin{definition}[$(t, n)$-Secret Sharing]
A $(t, n)$-Shamir secret sharing of $s \in \mathbb{Z}_q$ distributes shares
$s_i = f(i)$ where $f(X) = s + \sum_{j=1}^{t-1} a_j X^j$ is a random
polynomial of degree $t-1$. Any $t$ shares determine $s$; fewer than $t$
shares reveal no information about $s$.
\end{definition}

\begin{algorithm}
\caption{FROST Distributed Key Generation}
\label{alg:dkg}
\begin{algorithmic}[1]
\Require Participants $P_1, \ldots, P_n$, threshold $t$
\Ensure Public key $Y$, shares $s_1, \ldots, s_n$
\For{each participant $P_i$}
    \State Sample random polynomial $f_i(X) = a_{i,0} + \sum_{j=1}^{t-1} a_{i,j} X^j$
    \State Broadcast commitments $C_{i,j} = a_{i,j} \cdot G$ for $j = 0, \ldots, t-1$
    \State Send share $f_i(k)$ privately to each $P_k$
\EndFor
\For{each participant $P_k$}
    \State Verify: $f_i(k) \cdot G \stackrel{?}{=} \sum_{j=0}^{t-1} C_{i,j} \cdot k^j$ for all $i$
    \State Compute combined share: $s_k = \sum_{i=1}^{n} f_i(k)$
\EndFor
\State $Y = \sum_{i=1}^{n} C_{i,0}$ \Comment{Combined public key}
\end{algorithmic}
\end{algorithm}

\begin{theorem}[DKG Security]
The FROST DKG protocol is secure against $t-1$ corrupted participants under
the discrete logarithm assumption on the Ed25519 curve. No coalition of fewer
than $t$ participants can reconstruct the secret key $s = \sum_i a_{i,0}$.
\end{theorem}

\subsection{Threshold Signing}

FROST signing requires two rounds. In the first round, each signer generates
and broadcasts a nonce commitment. In the second round, each signer produces a
partial signature using their key share and the aggregated nonce.

\begin{definition}[FROST Signature Share]
Given message $m$, group nonce $R = \sum_{i \in S} R_i$, and challenge
$c = H(R \| Y \| m)$, signer $P_i$ with share $s_i$ computes:
\begin{equation}
z_i = d_i + e_i \cdot \rho_i + \lambda_i \cdot s_i \cdot c,
\end{equation}
where $(d_i, e_i)$ are nonce secrets, $\rho_i = H(i, m, B)$ is the binding
factor, and $\lambda_i$ is the Lagrange interpolation coefficient for the
signing set $S$.
\end{definition}

The aggregated signature $\sigma = (R, z)$ where $z = \sum_{i \in S} z_i$ is a
standard Ed25519 signature verifiable with the public key $Y$.

\begin{proposition}[Unforgeability]
FROST signatures are existentially unforgeable under chosen-message attack
(EUF-CMA) in the random oracle model, assuming the hardness of the discrete
logarithm problem on Curve25519~\cite{frost}.
\end{proposition}

\subsection{Nonce Pre-computation}

To reduce signing latency, signers pre-compute nonce pairs during idle periods:
\begin{enumerate}[leftmargin=1.4em]
    \item Each signer generates a batch of $B$ nonce pairs $(d_j, e_j)$ and
    commitments $(D_j, E_j)$.
    \item Commitments are broadcast and stored by all participants.
    \item During signing, a pre-computed nonce pair is consumed. No
    communication is needed in the first round.
\end{enumerate}

With $B = 10000$ pre-computed nonces, signing reduces to a single round with
median latency of 8ms (vs.\ 12ms for 2-round signing).

% ============================================================================
