\section{Integration with Hanzo Services}

\label{sec:services}
% ============================================================================

\subsection{IAM Token Signing}

Hanzo IAM~\cite{hanzoiam} issues JWT tokens signed with Ed25519. The IAM
signing key is split across 5 threshold signing nodes in a $(3, 5)$
configuration:

\begin{enumerate}[leftmargin=1.4em]
    \item User authenticates via OIDC (password, passkey, or SSO).
    \item IAM constructs the JWT payload (claims, expiry, org scope).
    \item IAM sends the signing request to the threshold signing service.
    \item Three of five signing nodes produce partial signatures; the
    coordinator aggregates them into a standard Ed25519 signature.
    \item The JWT is returned to the user. Verifiers use the public key
    from the IAM JWKS endpoint---no knowledge of threshold signing needed.
\end{enumerate}

\subsection{KMS Master Key Protection}

Hanzo KMS~\cite{hanzokms} uses envelope encryption: data encryption keys (DEKs)
are encrypted with a master key (KEK). The KEK is threshold-shared:
\begin{itemize}[leftmargin=1.1em]
    \item KEK encryption uses threshold ECDH: each signing node holds a share
    of the ECDH private key, and $t$ nodes must cooperate to derive the shared
    secret used for AES-256-GCM encryption of the DEK.
    \item KEK rotation is a key refresh operation---existing DEKs encrypted
    under the old KEK are re-encrypted under the refreshed key without
    exposing plaintext DEKs.
\end{itemize}

\subsection{Agent Credential Signing}

AI agents~\cite{hanzoagentsdk} receive delegated credentials via certificate
chains:
\begin{enumerate}[leftmargin=1.4em]
    \item The platform root key (threshold-shared) signs an org-level
    intermediate certificate.
    \item The org intermediate signs an agent-specific credential with
    constrained permissions (allowed tools, API scopes, time bound).
    \item Agents present the credential chain when accessing resources. Each
    service verifies the chain back to the well-known root public key.
\end{enumerate}

\subsection{ACI Compute Attestation}

ACI GPU nodes~\cite{hanzoaci} sign inference attestations with threshold ECDSA:
\begin{itemize}[leftmargin=1.1em]
    \item Each GPU node in a verification group holds a share of the group
    signing key.
    \item After verifying an inference result, $t$ of $n$ verifiers produce a
    threshold ECDSA signature over the result hash.
    \item The signature is submitted on-chain as proof of verified computation.
\end{itemize}

% ============================================================================
