\section{Conclusion}

\label{sec:conclusion}
% ============================================================================

Hanzo Threshold Signing eliminates single points of compromise across Hanzo
infrastructure by distributing signing keys via $(3, 5)$ threshold schemes.
FROST for Ed25519 achieves 8--12ms signing latency for IAM tokens and agent
credentials. CGGMP21 for ECDSA achieves 45ms signing latency for blockchain
operations. Proactive key refresh every 24 hours limits exposure from share
compromise. The system processes 2M+ daily threshold signatures with 99.97\%
availability.

% ============================================================================
% BIBLIOGRAPHY
% ============================================================================
\begin{thebibliography}{15}

\bibitem{hanzoiam}
{Hanzo AI Research}.
\newblock Hanzo {IAM}: Identity and access management for {AI} platforms.
\newblock \emph{Hanzo AI Research}, 2024.

\bibitem{hanzokms}
{Hanzo AI Research}.
\newblock Hanzo {KMS}: Secrets management with per-environment rotation and audit trail.
\newblock \emph{Hanzo AI Research}, September 2021.

\bibitem{hanzoagentsdk}
{Hanzo AI Research}.
\newblock Hanzo Agent {SDK}: Framework for building {AI} agents.
\newblock \emph{Hanzo AI Research}, 2025.

\bibitem{hanzoaci}
{Hanzo AI Research}.
\newblock Hanzo {ACI}: {AI} Chain Infrastructure for decentralized compute markets.
\newblock \emph{Hanzo AI Research}, December 2024.

\bibitem{luxwhitepaper}
{Lux Partners}.
\newblock Lux: A scalable multi-consensus blockchain with post-quantum cryptography.
\newblock \emph{Lux Network Technical Report}, 2024.

\bibitem{frost}
C.~Komlo and I.~Goldberg.
\newblock {FROST}: Flexible round-optimized {Schnorr} threshold signatures.
\newblock In \emph{Proc.\ SAC}, 2020.

\bibitem{cggmp21}
R.~Canetti, R.~Gennaro, S.~Goldfeder, N.~Makriyannis, and U.~Peled.
\newblock {UC} non-interactive, proactive, threshold {ECDSA} with identifiable aborts.
\newblock In \emph{Proc.\ CCS}, 2020.

\bibitem{pedersen1991}
T.~Pedersen.
\newblock Non-interactive and information-theoretic secure verifiable secret sharing.
\newblock In \emph{Proc.\ CRYPTO}, 1991.

\bibitem{herzberg1995}
A.~Herzberg, S.~Jarecki, H.~Krawczyk, and M.~Yung.
\newblock Proactive secret sharing, or: How to cope with perpetual leakage.
\newblock In \emph{Proc.\ CRYPTO}, 1995.

\bibitem{gg20}
R.~Gennaro and S.~Goldfeder.
\newblock One round threshold {ECDSA} with identifiable abort.
\newblock \emph{IACR Cryptology ePrint Archive}, 2020/540.

\bibitem{fireblocks}
{Fireblocks}.
\newblock {MPC} wallet infrastructure for digital assets.
\newblock \emph{Fireblocks Documentation}, 2023.

\end{thebibliography}

